### 2019-10-17 scripts/01_clean_pheno.Rmd
scritps

---
title: "I mean really clean the pheno and get to know the hell of the samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(readxl)
library(RVenn)
```

## Valid patient samples

```{r}
# the newest pheno table from M
# this sheet have nice sampleids for the samples, and also this is the table has all of the info
phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10162019.xlsx',
                    sheet = 5)
```

```{r}
# selecting the columns relavant to this study
phe_M_clean <- phe_M %>% 
   rename_all(funs(str_replace_all(., ' \\(.+$','') %>% 
    str_replace_all(., ' ','_') %>% 
    str_replace(., ':','_'))) %>% 
  # select the columns that may be relavant
  select(Sampleid = Baseline_Sample_ID, 
         Censor_YN,
         Censor_reason,
         CR, 
         ORR, 
         Toxicity, 
         Disease_Status_CAR = Key__Patient_Disease_Status_at_CAR, 
         Disease_Class, 
         CAR_Group) %>% 
  filter(is.na(Censor_reason)) %>% 
  # remove this one record that doesn't seem to be a valid record at all
  filter(!is.na(CR))

nrow(phe_M_clean)

head(phe_M_clean)

# output this pheno table cuz its gonna be used downstream
phe_M_clean %>% 
  write_csv('../data/cleaned_pheno/current_pheno.csv')
```

## Valid shotgun data

```{r}
# loading the table from John to see how many samples we actually have shotgun data
have_sg <- read_csv('../data/baselinesforAngel10032019.csv') %>% 
  filter(shotgun_YN == 'yes') %>% 
  # removig 1943A cuz its low quality
  filter(sample_id != '1943A') %>% 
  # THERE ARE THREE THAT ARE CURRENTLY BEING SEQUENCED, WE DON'T HAVE YET!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
  filter(!sample_id %in% c('161','171','173'))

nrow(have_sg)

```

## Overlap: the real valid shotgun samples

```{r}
# visualize the difference 
two <- Venn(list(
  have_file = have_sg$sample_id,
  useable = phe_M_clean$Sampleid
))

ggvenn(two)      
```
```{r}
# write out the overlap sampleid cuz that's gonna be the sample_names in use on the cluster for normalization
intersect(have_sg$sample_id, phe_M_clean$Sampleid) %>% 
  write.table('../data/input_for_cluster_sampleid/valid_sample_names.txt', col.names = F, row.names = F, quote = F)
```

### 2019-10-17 scripts/02_renorm_and_other_cluster.Rmd
scritps

---
title: "renorm of the genefa and pathway abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is to document what happened.

This step is done at the cluster

For the given valid samples("1725E" "2222A" "2245A" removed)
* merge them into one table (note that humann2 join table with take all of the tsv files that are in the folder)
* renorm to relab
* split them into stratified and unstratified

```{bash}
scp -r  daia1@lilac.mskcc.org:/home/daia1/my_workdir/samples/melody/humann2_final_out .
```

### 2019-10-17 scripts/03_lefse_CR_tox.Rmd
scritps

---
title: "Format for lefse and run on the classes CR and tox"
output: html_document
---

Melody's project

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Run unstratified pathway abundance of Melody's project thru lefse
* CR & Toxicity

Now we have 3 subclasses and together with the 2 classes of 35 that gonna make a total combination of 6

## Sorting out the pheno table for the valid samples

```{r}
# loading the table with the valid sampleids
valid_s <- read_table('../data/input_for_cluster_sampleid/valid_sample_names.txt', col_names = F)

pheno <- read_csv('../data/cleaned_pheno/current_pheno.csv') %>% 
  filter(Sampleid %in% valid_s$X1) %>% 
  select(-Censor_YN, -Censor_reason ) 
```

## Creating the class+subclass combi

```{r}
# get different combination of the classes
classes <- c('CR', 'Toxicity')
subclass <- c('Disease_Status_CAR','Disease_Class','CAR_Group')

combi <- expand.grid(classes = classes,subclass = subclass)

# use each combi row in the above df to create the pheno table for the lefse input

apply(combi, 1, function(row) {
  pheno %>% 
    select(row[['classes']], row[['subclass']]) %>% 
    t() %>% 
    write.table(str_glue('../data/lefse_input/pheno__{row[["classes"]]}__{row[["subclass"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
})
```

## Sorting the relab abundance table so that sampleid would align

```{r}
# input: relative abundance unstratified for pathway abundance
relab <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_unstratified.tsv') %>% 
  rename(sampleid = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  # remove the unmapped and the unintegrated record
  filter(!sampleid %in% c('UNMAPPED','UNINTEGRATED'))

# see if the sampleid in the relab is clean(do I need to clean additionally)
all.equal(sort(pheno$Sampleid), sort(colnames(relab)[2:ncol(relab)]))

# reorder the relab column to be the same with pheno table
relab_or <- bind_cols(relab[,1],relab[,pheno$Sampleid])

# just to be sure
all.equal(colnames(relab_or)[2:ncol(relab_or)], pheno$Sampleid)

# write out the table
relab_or %>% 
  write_tsv('../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv')
```

## Create the bash commmand to run the lefse steps

```{r}
# create the commad in df with str_glue and then output to a bash script
pheno_combi <- data_frame(pheno_file = list.files('../data/lefse_input', pattern = '.txt$', full.names = T)) %>% 
  mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))

pheno_combi %>% 
  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_{combi}tsv')) %>% 
  select(cat_cmd) %>% 
  write_csv('cat_combi_6.sh', col_names = F)
```

```{r}
# create the cmd to do the format input stuff
pheno_combi %>% 
  mutate(input_tsv = str_glue('../data/lefse_input/pathway_abundace_{combi}tsv'),
         fmt_input = str_replace(input_tsv, 'tsv','in'),
         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -s 2 -u 3 -o 1000000')) %>% 
  select(fmt_cmd) %>% 
  write_csv('format_input_6.sh', col_names = F)
```

```{r}
# the actual step to run lefse
pheno_combi %>% 
  mutate(fmt_input = str_glue('../data/lefse_input/pathway_abundace_{combi}in'),
         lefse_res = str_replace(fmt_input, 'in$','res'),
         run_cmd = str_glue('run_lefse.py {fmt_input} {lefse_res}')) %>% 
  select(run_cmd) %>% 
  write_csv('run_lefse.sh', col_names = F)
```

```{r}
# plotting cmd 
pheno_combi %>% 
  mutate(lefse_res = str_glue('../data/lefse_input/pathway_abundace_{combi}res'),
         fea_png = str_glue('../figs/pathway_abundace_{combi}png'),
         cmd = str_glue('plot_res.py {lefse_res}  {fea_png} --feature_font_size 4 --width 10 --dpi 300 --title {combi}')) %>% 
  select(cmd) %>% 
  write_csv('plotting_de_feature.sh', col_names = F)
```

####################################### below is command line ##########################################

## Run in terminal the below bash script

```{bash}
# this is the bash script to do the actual concatenation 
bash cat_combi_6.sh
```

```{bash}
# run in the cmd line THIS HAS TO BE DONE IN THE CMD SINCE RSTUDIO DOES NOT RECOGNIZE THE LEFSE SCRIPT
bash format_input_6.sh
```

```{bash}
bash run_lefse.sh
```

```{bash}
bash plotting_de_feature.sh
```

### 2019-10-17 scripts/04_lefse_ORR.Rmd
scritps

---
title: "lefse with ORR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# because there are currently 36 valid ORR samples so it needs to be normalized separately.

* ORR & Toxicity

```{r}
# the two samples that are currenly NR in ORR
NRs <- read_csv('../data/current_pheno.csv') %>% 
  filter(ORR == "NR") %>% 
  pull(Sampleid)

# the current pathway abundance matrix that is unnormalized
abun <- read_tsv('../data/humann2_res/humann2_pathabundance.tsv') %>% 
  rename(sampleid = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  select(-NRs)

ncol(abun)

# write out to do the normalization
abun %>% 
  write_tsv('../data/humann2_res/humann2_pathabundance_36.tsv')

```

```{bash}
# renorm
humann2_renorm_table --input humann2_pathabundance_36.tsv -u relab --output humann2_pathabundance_36_relab.tsv
```

```{bash}
# split the table
humann2_split_stratified_table --input humann2_pathabundance_36_relab.tsv --output humann2_final_out_36
```

```{r}
# load the split table with 36 and align the sampleid with the pheno table
cts <- read_tsv('../data/humann2_res/humann2_final_out_36/humann2_pathabundance_36_relab_unstratified.tsv')

# ORR  there is gonna be 36 samples cuz 2 are NR
pheno_O <- read_csv('../data/current_pheno.csv') %>% 
  filter(ORR != "NR") %>% 
  select(-CR, -Toxicity )

# reorder the cts table
cts <- bind_cols(cts[,1],cts[,pheno_O$Sampleid])

# check if the sampleids really align now
all.equal(colnames(cts)[2:ncol(cts)], pheno_O$Sampleid)

# write out the cts table for later binding
cts %>% 
  write_tsv('../data/cts_36_reorder.tsv')

```

# Different combination of the classes + subclasses

```{r}
# get different combination of the classes
classes <- c('ORR')
subclass <- c('Disease_Status_CAR','Disease_Class','CAR_Group')

combi <- expand.grid(classes = classes,subclass = subclass)

# just to be sure
all.equal(pheno_O$Sampleid, colnames(cts)[2:ncol(cts)])

# use each combi row in the above df to create the pheno table for the lefse input

apply(combi, 1, function(row) {
  pheno_O %>% 
    select(row[['classes']], row[['subclass']]) %>% 
    t() %>% 
    write.table(str_glue('../data/lefse_input/pheno__{row[["classes"]]}__{row[["subclass"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
})
```

# Create the formatted input for lefse

```{r}
# create the commad in df with str_glue and then output to a bash script
pheno_combi <- data_frame(pheno_file = list.files('../data/lefse_input', pattern = 'pheno__ORR', full.names = T)) %>% 
  mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))

pheno_combi %>% 
  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_{combi}tsv')) %>% 
  select(cat_cmd) %>% 
  write_csv('cat_combi_3.sh', col_names = F)
```

```{r}
# create the cmd to do the format input stuff
pheno_combi %>% 
  mutate(input_tsv = str_glue('../data/lefse_input/pathway_abundace_{combi}tsv'),
         fmt_input = str_replace(input_tsv, 'tsv','in'),
         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -s 2 -u 3 -o 1000000')) %>% 
  select(fmt_cmd) %>% 
  write_csv('format_input_3.sh', col_names = F)
```

```{r}  
# the actual step to run lefse
pheno_combi %>%  
  mutate(fmt_input = str_glue('../data/lefse_input/pathway_abundace_{combi}in'),
         lefse_res = str_replace(fmt_input, 'in$','res'),
         run_cmd = str_glue('run_lefse.py {fmt_input} {lefse_res}')) %>% 
  select(run_cmd) %>% 
  write_csv('run_lefse_3.sh', col_names = F)
```

```{r}
# plotting cmd 
pheno_combi %>% 
  mutate(lefse_res = str_glue('../data/lefse_input/pathway_abundace_{combi}res'),
         fea_png = str_glue('../figs/pathway_abundace_{combi}png'),
         cmd = str_glue('plot_res.py {lefse_res}  {fea_png} --feature_font_size 4 --width 10 --dpi 300 --title {combi}')) %>% 
  select(cmd) %>% 
  write_csv('plottint_de_feature_3.sh', col_names = F)

# No differentially abundant features found  
```

### 2019-10-17 scripts/05_pathway_feature_cpm.Rmd
scritps

---
title: "Check the pathway cpm of the feature"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

Check the pathway abundance in cpm to see if there is a decent effect size

```{r}
fn <- read_tsv('../data/humann2_res/humann2_final_out_cpm/humann2_pathabundance_cpm_unstratified.tsv') %>%
  rename(pathway = names(.)[1]) %>% 
  # remove the things up till the colon
  mutate(pathway = str_replace_all(pathway, '^.+\\: ','')) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  # removing things in the parenthesis and remove the spaces in the pathway
  mutate(pathway = str_replace_all(pathway, ' \\(.+\\)',''),
         pathway = str_replace_all(pathway, ' ',''))
```

```{r}
# loading the enriched pathways from 6 combi and filter on the abundance matrix
feat_pw <- read_csv('../data/unique_enriched_pathways.csv')

cpms <- fn %>% 
  filter(pathway %in% feat_pw$pathway)

# only half is exactly the same

same_pw <- cpms %>% 
  pull(pathway)
```

```{r}
# plot the counts 
# get the pheno info of those samples
pheno <- read_csv('../data/current_pheno.csv')

res <- cpms %>% 
  gather(key = 'Sampleid', value = 'cpm', '1191V' : 'V0040A') %>% 
  left_join(pheno, by = 'Sampleid')

res
```

```{r}

same_pw %>% 
  map(~ res %>% 
          arrange(CR, desc(cpm)) %>% 
          filter(pathway == .) %>% 
          ggbarplot(x = 'Sampleid', y = 'cpm', fill = 'CR', color = 'CR') +
          scale_color_manual(values = c('#00468B', '#EC0000'))  +
          scale_fill_manual(values = c('#00468B', '#EC0000')) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
          labs(title = str_glue('Pathway: {.}')) +
          ggsave(str_glue('../figs/pathway_abundance_cpm_{.}.jpg'), width = 4, height = 3, dpi = 300))

for (i in same_pw) {
  res %>% 
          arrange(Toxicity, desc(cpm)) %>% 
          filter(pathway == i) %>% 
          ggbarplot(x = 'Sampleid', y = 'cpm', fill = 'Toxicity', color = 'Toxicity') +
          scale_color_manual(values = c('#00468B', '#EC0000'))  +
          scale_fill_manual(values = c('#00468B', '#EC0000')) +
          theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
          labs(title = str_glue('Pathway: {i}')) +
          ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/pathway_abundance_cpm/Toxicity/pathway_abundance_cpm_{i}.jpg'), width = 8, height = 6, dpi = 300)
}

```

```{r}
res %>% 
    arrange(CR, desc(cpm)) %>% 
    filter(pathway == 'pantothenateandcoenzymeAbiosynthesisI') %>% 
    ggbarplot(x = 'Sampleid', y = 'cpm', fill = 'CR', color = 'CR') +
    scale_color_manual(values = c('#00468B', '#EC0000'))  +
    scale_fill_manual(values = c('#00468B', '#EC0000')) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
    labs(title = str_glue('Pathway: pantothenateandcoenzymeAbiosynthesisI')) +
    ggsave(str_glue('../figs/tox/pathway_abundance_cpm_pantothenateandcoenzymeAbiosynthesisI.jpg'), width = 4, height = 3, dpi = 300)
```

### 2019-10-17 scripts/06_lefse_res_ggplot.Rmd
scritps

---
title: "Replot the lefse plots with ggplot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Replot the lefse plots with ggplot2 (with each class + subclass combi)

## Combine and clean the res table from lefse

```{r}
# the input : lefse res files
fns <- list.files('../data/lefse_input/', pattern = 'pathway_abundace_pheno__[CR|Toxicity].+res$', full.names = T)

fns_name <- list.files('../data/lefse_input/', pattern = 'pathway_abundace_pheno__[CR|Toxicity].+res$')

# join all of the tables feature together
feature <- fns %>% 
  set_names(fns_name) %>% 
  map(~ read_tsv(., col_names = c('pathway','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  bind_rows(.id = 'group') %>% 
  mutate(group = str_replace(group, 'pathway_abundace_pheno__','')) %>% 
  mutate(group = str_replace(group, '.res$','')) %>% 
  mutate(pathway = str_replace_all(pathway, '^f_','')) %>% 
  select(-xx, -pval)

# change the "N" direction to be minus score
feature <- bind_rows(
  feature %>% 
    split(.$direction) %>% 
    pluck('N') %>% 
    mutate(score = -score),
  feature %>% 
    split(.$direction) %>% 
    pluck('Y') 
) %>% 
  arrange(group, pathway, score)
```

```{r}
# output the sig unique pathway from the above table for later use in the bar plot
feature  %>% 
  distinct(pathway) %>% 
  write_csv('../data/unique_sig_pathway.csv')
```

## Recreate the lefse figure with ggplot

Each combi per figure

```{r}
feature %>% 
  split(.$group) %>% 
  map(~  ggplot(data = ., aes(x = reorder(pathway, score), y = score, fill = direction)) +
  geom_bar(stat='identity') +
  coord_flip() +
  scale_color_manual(values = c('#00468B', '#EC0000'))  +
  scale_fill_manual(values = c('#00468B', '#EC0000')) +
  theme_minimal() +
  theme(axis.title.y  = element_blank(), 
        legend.position="bottom") +
  labs(title = str_glue('Response to class (with subclass): {.$group}') ,
       y = 'Score')+ 
  ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to class (with subclass): {.$group}.pdf'), device = 'pdf', width = 15, height = 6, dpi = 300))

```

### 2019-10-17 scripts/07_pathway_feature_overview.Rmd
scritps

---
title: "Check the lefse feature overview"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(tidyverse)
library(kableExtra)
```

```{r}
# the input : lefse res files
fns <- list.files('../data/lefse_input/', pattern = 'pathway_abundace_pheno__[CR|Toxicity].+res$', full.names = T)

fns_name <- list.files('../data/lefse_input/', pattern = 'pathway_abundace_pheno__[CR|Toxicity].+res$')
```

```{r}
# join all of the tables feature together
feature <- fns %>% 
  set_names(fns_name) %>% 
  map(~ read_tsv(., col_names = c('pathway','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  bind_rows(.id = 'group') %>% 
  mutate(group = str_replace(group, 'pathway_abundace_pheno__',''))

feature %>% 
  distinct(pathway) %>% 
  write_csv('../data/unique_enriched_pathways.csv')
```

```{r}
# see the Y and N direction respectively
Y_response <- feature %>% 
  filter(direction == 'Y')

Y_response %>% 
  kable(caption = 'Yes CR and Yes tox') %>% 
  kable_styling(full_width = F)

```

Some keywords appeared in yes CR and yes tox:

* L-citrulline : boosts nitric oxide production in the body. Nitric oxide helps your arteries relax and work better, which improves blood flow throughout your body. This may be helpful for treating or preventing some diseases.
* urate biosynthesis: related to purine metabolism.
* About pantothenate and phosphopantothenate: Coenzyme A is synthesized in a five-step process from pantothenate. First, **pantothenate (Vitamin B5) is phosphorylated to 4'-phosphopantothenate by the enzyme pantothenate kinase,** next a cysteine is added to 4'-phosphopantothenate by the enzyme phosphopantothenoylcysteine synthetase to form 4'-phospho-N-pantothenoylcysteine (PPC). In the next step, PPC is decarboxylated to 4'-phosphopantetheine by phosphopantothenoylcysteine decarboxylase. In the fourth step, 4'-phosphopantetheine is adenylylated to form dephospho-CoA by the enzyme phosphopantetheine adenylyl transferase. Finally, dephospho-CoA is phosphorylated using ATP to coenzyme A by the enzyme dephosphocoenzyme A kinase.
* tetrahydrofolate (vitamin B9)

```{r}
N_response <- feature %>% 
  filter(direction == 'N')

N_response %>% 
  kable(caption = 'No CR and no tox') %>% 
  kable_styling(full_width = F)
```

```{r}
# it looks like pantothenate biosynthesis is associated with no CR and no Tox regardless what subclass you have
N_response %>% 
  mutate(vb5 = str_detect(pathway, '^pantothenate')) %>% 
  separate(group, into = c('class','subclass'), sep = '__') %>% 
  group_by(class, subclass, direction,  vb5) %>% 
  summarise(cnt = n()) %>% 
  filter(vb5 == 'TRUE') %>% 
  kable(caption = 'Pantothenate biosynthesis existence in all class + subclass combinations') %>% 
  kable_styling(full_width = F)
```

It looks like pantothenate biosynthesis is associated with no CR and no Tox regardless what subclass you have

### what about the yes to CR and no to tox

```{r}
ideal <- feature  %>% 
  separate(group , into = c('class','subclass'), sep = '__') %>% 
  mutate(together = str_glue('{class}_{direction}')) %>% 
  filter(together %in% c('CR_Y','Toxicity_N'))

noidea <- feature  %>% 
  separate(group , into = c('class','subclass'), sep = '__') %>% 
  mutate(together = str_glue('{class}_{direction}')) %>% 
  filter(together %in% c('CR_N','Toxicity_Y'))
```

### 2019-10-17 scripts/08_barplot_species_pathway.Rmd
scritps

---
title: "Bar plot species contribution to pathway with humann2 script"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(RColorBrewer)
```

humann2_barplot
Basic usage: *humann2_barplot --input TABLE.tsv --feature FEATURE --outfile $FIGURE*
$TABLE.tsv = a stratified HUMAnN2 output file
$FEATURE = Feature from the table to plot (defaults to first feature)
$FIGURE = Where to save the figure
Run with -h to see additional command line options
humann2_barplot produces plots of stratified HUMAnN2 features. Note: unlike many other HUMAnN2 utilities, humann2_barplot requires the Python scientific stack (notably matplotlib) to operate. humann2_barplot includes many options for sorting and scaling data, as detailed in the --help menu. humann2_barplot is also explored in greater detail in the HUMAnN2 Tutorial.

Here is an example of a HUMAnN2 barplot for a pathway (denitrification) that was preferentially enriched in Human Microbiome Project oral samples relative to other body sites. This figure uses many options from humann2_barplot, including *regrouping by genus*, *pseudolog scaling*, and *sorting samples by similarity and metadata*:

```{r}
# parse the distinct pathway feature in the lefse results from the class: CR and tox
pw <- read_csv('../data/unique_enriched_pathways.csv') %>% 
  # something has a f_ at the beginning need to trim that off too
  mutate(pathway = str_replace(pathway, '^f_','')) %>% 
  # extract the pathway id from the whole messy string
  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY\\d*_\\d{3,4}')) %>% 
  mutate(pwid = str_replace_all(pwid, '_','-'))

```

## Make species level bar plot

```{r}
# check one last time whether the damn sampleids align
cnts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

# omg, its really hard to sort the pheno table the way I wanted, I will have to resort the cnts table to match
# reorder the relab column to be the same with pheno table
cnts_or <- bind_cols(cnts[,1],cnts[,pheno$Sampleid])

# just to be sure
all.equal(colnames(cnts_or)[2:ncol(cnts_or)], pheno$Sampleid)

# write out the reordered cnts file
cnts_or %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv' )

# write out the pheno as metadata
pheno %>% 
  select(Toxicity) %>% 
  t() %>% 
  write.table(str_glue('../data/tox_pheno.tsv'), sep = '\t', quote = F, row.names = T, col.names = F)

cnts_or_ <- cnts_or %>% 
  column_to_rownames('# Pathway')

in_meta_t <- data.frame(t(pheno %>% 
  select(Toxicity)), stringsAsFactors = FALSE)

names(in_meta_t) <- names(cnts_or_)
cat_pw_table <- rbind(in_meta_t, cnts_or_ )

cat_pw_table %>% 
  write.table('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl',
              quote = F,
              col.names=FALSE, 
              row.names=TRUE, 
              sep="\t")

```

```{bash}
# guess we need to cat the pheno and cnts together?
cat ../data/tox_pheno.tsv \
    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv > \
    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl
```

```{bash}
# the actual cmd to plot 
humann2_barplot  \
  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl \
  --focal-feature PWY-4242 \
  --sort similarity metadata \
  --focal-metadatum Toxicity --last-metadatum Toxicity \
  --scaling pseudolog \
  --dimensions 10 5 \
  -o ../figs/PWY-4242_pw.png
```

```{bash}
# without metadata
humann2_barplot  \
  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv \
  --focal-feature PWY-4242 \
  --sort sum \
  --dimensions 10 5 \
  -o ../figs/PWY-4242_pw.png

  #   --scaling pseudolog \
```

```{bash}
# the actual cmd to plot 
`humann2_barplot  \
  --input ../data/humann2_res/test_ARO-PWY_pw_abundance.tsv \
  --focal-feature ARO-PWY \
  --sort metadata \
  --focal-metadatum Toxicity --last-metadatum Toxicity \
  --scaling pseudolog \
  --dimensions 10 5 \
  -o ../figs/ARO-PWY_pw.png`

```

## Make genus level bar plot

Genus level gene families and pathways
By default, the gene families and pathways output files from HUMAnN2 are species level. To obtain genus level gene families and pathways, follow these steps.

Create a genus level gene families file

$ humann2_gene_families_genus_level --input $SAMPLE_genefamilies.tsv --output $SAMPLE_genefamilies_genus_level.tsv
In this command, replace $SAMPLE_genefamilies.tsv with the species level gene families file created by default by HUMAnN2 and $SAMPLE_genefamilies_genus_level.tsv with the name of the gene families genus level file that will be created.
Run HUMAnN2, with the genus level gene families file as input, to get genus level pathways output files

$ humann2 --input $SAMPLE_genefamilies_genus_level.tsv --output humann2_genus_level_output
This run will be much faster and require less memory than the original run as HUMAnN2 is provided gene family abundances so it only needs to compute the pathways.

### 2019-10-17 scripts/09_barplot_myself.Rmd
scritps

---
title: "Humann2 bar plot by myself"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The stacked bar plot like humann2 did, but do it myself.

```{r}
# send the test file to google group
# read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl', col_names = F) %>% 
#   filter(str_detect(X1, 'ARO-PWY|Toxicity'))  %>% 
#   write_tsv('../data/humann2_res/test_ARO-PWY_pw_abundance.tsv', col_names = F)
```

## Reorder the stratified pathabundance table and align with the pheno sampleid

```{r}
# pheno table 
pheno <- read_csv('../data/cleaned_pheno/current_pheno.csv')%>% 
  filter(Sampleid %in% valid_s$X1) %>% 
  select(-Censor_YN, -Censor_reason )

# the stratified table is used cuz we wanna know the species info
stratified <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
  rename(pathway = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$','')))

stratified <- bind_cols(stratified[,1],stratified[,pheno$Sampleid])

all.equal(colnames(stratified)[2:ncol(stratified)], pheno$Sampleid)

# write out this reordered table for later use
stratified %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv')
```

### Starting from toxicity

```{r}
# get the pheno lable of the samples
phelabel <-  pheno %>% 
  select(Sampleid, Toxicity)

cnt <- stratified %>% 
  filter(str_detect(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III')) %>% 
  mutate(pathway = str_replace(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III\\|', ''))  %>% 
  # extract the species info
  mutate(pathway = str_extract(pathway, 's__.+$|unclassified'),
         pathway = str_replace(pathway, 's__','')) %>% 
  rename(species = pathway)

# get the top 7 taxa and the unclassfied and other for all the others 
# get the above table 
mean_s <- cnt %>% 
  column_to_rownames('species') %>% 
  as.matrix() %>% 
  apply(., 1, mean) %>% 
  sort(decreasing = T)

# top 8
n <- 8
toppw <- names(mean_s)[1:n]

# colors and dark gray and light gray
hex <- c(brewer.pal(n -1, name = "Dark2"), '#A9A9A9', '#696969')

# now reduce the abundance to 9 rows, the top 8 and "other"

# split on whether in 8 or in "other"
splitother <- cnt %>% 
  split(f = str_detect(.$species, paste0(toppw, collapse = '|')))

cleaned_cnt <- bind_rows(splitother %>% 
    pluck('TRUE'),
  splitother %>% 
    pluck('FALSE') %>% 
    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
    group_by(Sample) %>% 
    summarise(other = sum(relab)) %>% 
    spread(Sample , other) %>% 
    mutate(species = 'Other') %>% 
    select(species, pheno$Sampleid)) %>% 
  # rearrange the species order factor lever
  mutate(species = factor(species, levels = c(toppw[toppw != 'unclassified'] , 'Other', 'unclassified'))) %>% 
  # the long format is the one for plotting
  gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(phelabel %>% 
              rename(Sample = Sampleid), by = 'Sample') %>% 
  mutate(Toxicity = factor(Toxicity))

# to put different color to the background to the phenotype
# create a dataframe to show where the coloring start and end for each group
# how many phenotypes ?
num_pheno <- length(names(table(phelabel$Toxicity)))
# have pre-defined background color for the pheno (to better contrast them)
bg_col <- c('#00468B', '#EC0000')
# number of samples in different pheno 
# the first phenotype(appearing on the left most of the plot)
phe1_num <- phelabel %>% 
  mutate(Toxicity = factor(Toxicity)) %>% 
  filter(Toxicity == as.character(levels(.$Toxicity)[1])) %>% 
  nrow

background_col <- data_frame(
  col = bg_col,
  xstart = c(0, phe1_num) + 0.5,
  xend = c(phe1_num, nrow(phelabel))+ 0.5,
  toxicity = names(table(phelabel$Toxicity))
)

# the ggplot
ggplot() +
  geom_bar(data = cleaned_cnt, 
           aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species, color = Toxicity), stat = 'identity') +
  geom_rect(data = background_col, 
            aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf), 
            fill = background_col$col, 
            alpha = 0.2) +
  geom_bar(data = cleaned_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
  #guides(alpha = FALSE) +
  scale_fill_manual(values = hex) +
  labs(y = 'relative abundance',
       x = 'Sample',
       title = str_glue('')) +
  theme_light() +
  theme(
    axis.text.x = element_blank(),
    panel.grid.major = element_blank(), 
    axis.ticks.x = element_blank(),  
    panel.grid.minor = element_blank()
    ) +
  ggsave('../figs/species_pw/PWY-4242: pantothenate and coenzyme A biosynthesis III.jpg', width = 12, height = 8, dpi = 300)
```

## Find how many sig pathways from lefse I wanna plot

```{r}
# load the table that has the unique sig pathways
sigpw <- read_csv('../data/unique_sig_pathway.csv')

# find how many species total it has in all of the lefse sig pathways
siglefse <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv') %>% 
  filter(str_detect(pathway,  paste0(pw$pwid, collapse = '|'))) %>% 
  separate(X1, into = c('pwid','taxa'), sep = '\\|')

sig_num <- siglefse %>% 
  split(f = .$pwid) %>% 
  map_dfc(~ nrow(.)) %>% 
  gather(key = 'pw', value = 'taxa_num') %>% 
  arrange(desc(taxa_num)) %>% 
  separate(pw, into = c('pwid','pw'), sep = ': ')

```

### 2019-10-17 scripts/cat_combi_3.sh

@@ -1,0 +1,3 @@
+cat ../data/lefse_input/pheno__ORR__CAR_Group.txt ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.tsv
+cat ../data/lefse_input/pheno__ORR__Disease_Class.txt ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.tsv
+cat ../data/lefse_input/pheno__ORR__Disease_Status_CAR.txt ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.tsv

### 2019-10-17 scripts/cat_combi_6.sh

@@ -1,0 +1,6 @@
+cat ../data/lefse_input/pheno__CR__CAR_Group.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.tsv
+cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv
+cat ../data/lefse_input/pheno__CR__Disease_Status_CAR.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.tsv
+cat ../data/lefse_input/pheno__Toxicity__CAR_Group.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.tsv
+cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv
+cat ../data/lefse_input/pheno__Toxicity__Disease_Status_CAR.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.tsv

### 2019-10-17 scripts/format_input_3.sh

@@ -1,0 +1,3 @@
+format_input.py ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.tsv ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.tsv ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000

### 2019-10-17 scripts/format_input_6.sh

@@ -1,0 +1,6 @@
+format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.tsv ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.tsv ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000

### 2019-10-17 scripts/plotting_de_feature.sh

@@ -1,0 +1,6 @@
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.res  ../figs/pathway_abundace_pheno__CR__CAR_Group.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__CAR_Group.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res  ../figs/pathway_abundace_pheno__CR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Class.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.res  ../figs/pathway_abundace_pheno__CR__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Status_CAR.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.res  ../figs/pathway_abundace_pheno__Toxicity__CAR_Group.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__CAR_Group.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res  ../figs/pathway_abundace_pheno__Toxicity__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Class.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.res  ../figs/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Status_CAR.

### 2019-10-17 scripts/plottint_de_feature_3.sh

@@ -1,0 +1,3 @@
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.res  ../figs/pathway_abundace_pheno__ORR__CAR_Group.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__ORR__CAR_Group.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.res  ../figs/pathway_abundace_pheno__ORR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__ORR__Disease_Class.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.res  ../figs/pathway_abundace_pheno__ORR__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__ORR__Disease_Status_CAR.

### 2019-10-17 scripts/run_lefse.sh

@@ -1,0 +1,6 @@
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.in ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.in ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.res

### 2019-10-17 scripts/run_lefse_3.sh

@@ -1,0 +1,3 @@
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.in ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.in ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.res

### 2019-10-17 scripts/09_barplot_myself.Rmd
update

@@
 
-# top 8
-n <- 8
+# top n (variable) including unclassified
+n <- floor(length(mean_s)*0.15) 
 toppw <- names(mean_s)[1:n]
@@
 
-# now reduce the abundance to 9 rows, the top 8 and "other"
+# now reduce the abundance to 9 rows, the top 8 and "other" (the current will work if the df has >= 9 rows)
 
@@
 # load the table that has the unique sig pathways
-sigpw <- read_csv('../data/unique_sig_pathway.csv')
+sigpw <- read_csv('../data/unique_sig_pathway.csv') %>% 
+  # grab the pw id which is gonna be able to match the rownames in the stratified table
+  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
+  mutate(pwid = str_replace_all(pwid, '_', '-'))
 
 # find how many species total it has in all of the lefse sig pathways
-siglefse <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv') %>% 
-  filter(str_detect(pathway,  paste0(pw$pwid, collapse = '|'))) %>% 
-  separate(X1, into = c('pwid','taxa'), sep = '\\|')
+siglefse <- stratified %>% 
+  filter(str_detect(pathway,  paste0(sigpw$pwid, collapse = '|'))) %>% 
+  separate(pathway, into = c('pwid','taxa'), sep = '\\|')
 

### 2019-10-17 scripts/09_barplot_myself.Rmd
update on the freaking bar plot

--- 2019-10-17 scripts/09_barplot_myself.Rmd
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
+library(tidyverse)
 ```
@@
 stratified <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
-  rename(pathway = names(.)[1]) %>% 
+  rename(PATHWAYID = names(.)[1]) %>% 
   rename_all(funs(str_replace(., '_Abundance$','')))
@@
 
-### Starting from toxicity
+## Find how many sig PATHWAYIDs from lefse I wanna plot
 
 ```{r}
+# load the table that has the unique sig PATHWAYIDs
+sigpw <- read_csv('../data/unique_sig_PATHWAYID.csv') %>% 
+  # grab the pw id which is gonna be able to match the rownames in the stratified table
+  mutate(pwid = str_extract(PATHWAYID, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
+  mutate(pwid = str_replace_all(pwid, '_', '-'))
+
+# find how many species total it has in all of the lefse sig PATHWAYIDs
+siglefse <- stratified %>% 
+  filter(str_detect(PATHWAYID,  paste0(sigpw$pwid, collapse = '|'))) %>% 
+  separate(PATHWAYID, into = c('pwid','taxa'), sep = '\\|')
+
+sig_num <- siglefse %>% 
+  split(f = .$pwid) %>% 
+  map_dfc(~ nrow(.)) %>% 
+  gather(key = 'pw', value = 'taxa_num') %>% 
+  arrange(desc(taxa_num)) %>% 
+  separate(pw, into = c('pwid','pw'), sep = ': ') %>% 
+  # for plotting reason filer on >= 9 rows
+  filter(taxa_num >= 9)
+
+```
+
+## Starting from toxicity
+
+```{r}
+species_compo_barplot <- function(PHENOVAR, PATHWAYID) {
+  # TWO INPUTS: 
+  # The pheno variable: CR, Toxicity, ORR
+  # The full PATHWAYID
   # get the pheno lable of the samples
+
+  #PATHWAYID <- 'COA-PWY'
   phelabel <-  pheno %>% 
@@
   cnt <- stratified %>% 
-  filter(str_detect(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III')) %>% 
-  mutate(pathway = str_replace(pathway, 'PWY-4242: pantothenate and coenzyme A biosynthesis III\\|', ''))  %>% 
+    filter(str_detect(pathway, PATHWAYID)) %>% 
     # extract the species info
@@
   mean_s <- cnt %>% 
-  column_to_rownames('species') %>% 
-  as.matrix() %>% 
-  apply(., 1, mean) %>% 
-  sort(decreasing = T)
+    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
+    group_by(species) %>% 
+    summarise(species_sum = sum(relab)) %>% 
+    arrange(desc(species_sum)) %>% 
+    pull(species)
 
-# top 8
+  # top n (variable) including unclassified
   n <- 8
-toppw <- names(mean_s)[1:n]
+  toppw <- mean_s[1:n]
 
@@
 
-# now reduce the abundance to 9 rows, the top 8 and "other"
+  # now reduce the abundance to 9 rows, the top 8 and "other" (the current will work if the df has >= 9 rows)
 
@@
     # the long format is the one for plotting
-  gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
+    gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)])
+
+  # get the sum of relab per sample cuz we wanna order by this
+  sample_relab_sum <- cleaned_cnt %>% 
+    group_by(Sample) %>% 
+    summarise(sample_sum = sum(relab))
+
+  final_cnt <- cleaned_cnt %>% 
+    left_join(sample_relab_sum, by = 'Sample') %>% 
     left_join(phelabel %>% 
                 rename(Sample = Sampleid), by = 'Sample') %>% 
-  mutate(Toxicity = factor(Toxicity))
+    mutate(Toxicity = factor(Toxicity)) %>% 
+    arrange(Toxicity, sample_sum)
 
@@
     xstart = c(0, phe1_num) + 0.5,
-  xend = c(phe1_num, nrow(phelabel))+ 0.5,
-  toxicity = names(table(phelabel$Toxicity))
+    xend = c(phe1_num, nrow(phelabel))+ 0.5
   )
@@
   ggplot() +
-  geom_bar(data = cleaned_cnt, 
-           aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species, color = Toxicity), stat = 'identity') +
+    geom_bar(data = final_cnt, 
+             aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
     geom_rect(data = background_col, 
@@
               alpha = 0.2) +
-  geom_bar(data = cleaned_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
-  #guides(alpha = FALSE) +
-  scale_fill_manual(values = hex) +
+    geom_bar(data = final_cnt, aes(x = reorder(Sample, as.numeric(Toxicity)), y = relab, fill = species), stat = 'identity') +
+    scale_fill_manual(values = hex ) +
     labs(y = 'relative abundance',
          x = 'Sample',
-       title = str_glue('')) +
+         title = str_glue('Response to Toxicity NO and YES:\nspecies composition of the pathway abundance\n{PATHWAYID}')) +
     theme_light() +
     theme(
-    axis.text.x = element_blank(),
+      axis.text.x =  element_text(angle = 40, hjust = 1, size = 8),
       panel.grid.major = element_blank(), 
@@
       ) +
-  ggsave('../figs/species_pw/PWY-4242: pantothenate and coenzyme A biosynthesis III.jpg', width = 12, height = 8, dpi = 300)
-```
+    ggsave(str_glue('../figs/species_pw/{PATHWAYID}.jpg'), width = 12, height = 8, dpi = 300)
+}
 
-## Find how many sig pathways from lefse I wanna plot
-
-```{r}
-# load the table that has the unique sig pathways
-sigpw <- read_csv('../data/unique_sig_pathway.csv')
-
-# find how many species total it has in all of the lefse sig pathways
-siglefse <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv') %>% 
-  filter(str_detect(pathway,  paste0(pw$pwid, collapse = '|'))) %>% 
-  separate(X1, into = c('pwid','taxa'), sep = '\\|')
-
-sig_num <- siglefse %>% 
-  split(f = .$pwid) %>% 
-  map_dfc(~ nrow(.)) %>% 
-  gather(key = 'pw', value = 'taxa_num') %>% 
-  arrange(desc(taxa_num)) %>% 
-  separate(pw, into = c('pwid','pw'), sep = ': ')
+sig_num$pwid %>% 
+  map(~  species_compo_barplot('Toxicity', .))
 

### 2019-10-17 scripts/08_barplot_species_pathway.Rmd
that stupid humann2 bar plot script finally worked

@@
 
+## Make species level bar plot using humann2 script
+
+### Create the correct pheno and relab table format
+
 ```{r}
-# parse the distinct pathway feature in the lefse results from the class: CR and tox
-pw <- read_csv('../data/unique_enriched_pathways.csv') %>% 
-  # something has a f_ at the beginning need to trim that off too
-  mutate(pathway = str_replace(pathway, '^f_','')) %>% 
-  # extract the pathway id from the whole messy string
-  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY\\d*_\\d{3,4}')) %>% 
-  mutate(pwid = str_replace_all(pwid, '_','-'))
+# ok so the correct format should be header at the first row, and phenotype and then the abundance values 
+# pheno table 
+read_csv('../data/cleaned_pheno/current_pheno.csv') %>% 
+  filter(Sampleid %in% valid_s$X1) %>% 
+  select(-Censor_YN, -Censor_reason) %>% 
+  # USE THE Toxicity PHENO FIRST
+  select(Sampleid, Toxicity) %>% 
+  t() %>% 
+  write.table('../data/horizontal_pheno/tox_wi_sample_header.txt', quote = F, row.names = T, col.names = F, sep = '\t')
 
+# reordered relab table but without header
+read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv') %>% 
+  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_wo_header.tsv', col_names = F)
 ```
 
-## Make species level bar plot
-
-```{r}
-# check one last time whether the damn sampleids align
-cnts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified.tsv') %>% 
-  rename_all(funs(str_replace(., '_Abundance$','')))
-
-# omg, its really hard to sort the pheno table the way I wanted, I will have to resort the cnts table to match
-# reorder the relab column to be the same with pheno table
-cnts_or <- bind_cols(cnts[,1],cnts[,pheno$Sampleid])
-
-# just to be sure
-all.equal(colnames(cnts_or)[2:ncol(cnts_or)], pheno$Sampleid)
-
-# write out the reordered cnts file
-cnts_or %>% 
-  write_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv' )
-
-# write out the pheno as metadata
-pheno %>% 
-  select(Toxicity) %>% 
-  t() %>% 
-  write.table(str_glue('../data/tox_pheno.tsv'), sep = '\t', quote = F, row.names = T, col.names = F)
-
-cnts_or_ <- cnts_or %>% 
-  column_to_rownames('# Pathway')
-
-in_meta_t <- data.frame(t(pheno %>% 
-  select(Toxicity)), stringsAsFactors = FALSE)
-
-names(in_meta_t) <- names(cnts_or_)
-cat_pw_table <- rbind(in_meta_t, cnts_or_ )
-
-cat_pw_table %>% 
-  write.table('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl',
-              quote = F,
-              col.names=FALSE, 
-              row.names=TRUE, 
-              sep="\t")
-
-```
+### row bind them together
 
@@
 # guess we need to cat the pheno and cnts together?
-cat ../data/tox_pheno.tsv \
-    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv > \
-    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl
+cat ../data/horizontal_pheno/tox_wi_sample_header.txt \
+    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_wo_header.tsv > \
+    ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv
 ```
+
+FINALLY WORKED!!!!!!!!!!!!!!!!!!!!!!
 
@@
 humann2_barplot  \
-  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.pcl \
+  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv \
   --focal-feature PWY-4242 \
-  --sort similarity metadata \
+  --sort similarity sum metadata \
   --focal-metadatum Toxicity --last-metadatum Toxicity \
@@
   -o ../figs/PWY-4242_pw.png
-```
-
-```{bash}
-# without metadata
-humann2_barplot  \
-  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered.tsv \
-  --focal-feature PWY-4242 \
-  --sort sum \
-  --dimensions 10 5 \
-  -o ../figs/PWY-4242_pw.png
-
-  #   --scaling pseudolog \
-```
-
-```{bash}
-# the actual cmd to plot 
-`humann2_barplot  \
-  --input ../data/humann2_res/test_ARO-PWY_pw_abundance.tsv \
-  --focal-feature ARO-PWY \
-  --sort metadata \
-  --focal-metadatum Toxicity --last-metadatum Toxicity \
-  --scaling pseudolog \
-  --dimensions 10 5 \
-  -o ../figs/ARO-PWY_pw.png`
-
 ```

### 2019-10-17 scripts/09_barplot_myself.Rmd
that stupid humann2 bar plot script finally worked

@@
   arrange(desc(taxa_num)) %>% 
-  separate(pw, into = c('pwid','pw'), sep = ': ') %>% 
+  separate(pw, into = c('pwid','pw'), sep = ': ', remove = F) %>% 
   # for plotting reason filer on >= 9 rows
@@
 ```{r}
-species_compo_barplot <- function(PHENOVAR, PATHWAYID) {
+species_compo_barplot <- function(PATHWAYID, PWFULLNAME) {
   # TWO INPUTS: 
@@
          x = 'Sample',
-         title = str_glue('Response to Toxicity NO and YES:\nspecies composition of the pathway abundance\n{PATHWAYID}')) +
+         title = str_glue('Response to Toxicity NO (blue background) and YES (red background):\nspecies composition of the pathway abundance\n{PATHWAYID}: {PWFULLNAME}')) +
     theme_light() +
@@
 
-sig_num$pwid %>% 
-  map(~  species_compo_barplot('Toxicity', .))
+map2(sig_num$pwid, sig_num$pw, ~ species_compo_barplot(.x, .y))
 
+sig_num$pw
 ```

### 2019-10-17 scripts/command.sh

@@ -1,0 +1,9 @@
+humann2_barplot  \
+    --input $1 \
+    --focal-feature $2 \
+    --sort similarity metadata \
+    --focal-metadatum Toxicity \
+    --last-metadatum Toxicity \
+    --scaling pseudolog \
+    --dimensions 10 5 \
+    -o $2.png

### 2019-10-17 scripts/08_barplot_species_pathway.Rmd
finally able to produce the damn bar pltos

@@
 # ok so the correct format should be header at the first row, and phenotype and then the abundance values 
+valid_s <- read_table('../data/input_for_cluster_sampleid/valid_sample_names.txt', col_names = F)
 # pheno table 
@@
   --dimensions 10 5 \
-  -o ../figs/PWY-4242_pw.png
+  -o ../figs/barplot_humann2/PWY-4242_pw.png
+```
+
+### now make the cmd with str_glue
+
+```{r}
+# load the sig pathway table with pwid
+cmd_barplot <- read_csv('../data/unique_sig_pathway_with_clean_pwid.csv') %>% 
+  select(pwid) %>% 
+  mutate(barcmd = str_glue('humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature {pwid} --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/{pwid}_pw.png')) %>%
+  select(barcmd)
+
+cmd_barplot %>% 
+  write.table('humann2_bar_cmd.sh', quote = F, col.names = F, row.names = F )
+```
+
+```{bash}
+bash humann2_bar_cmd.sh
 ```

### 2019-10-17 scripts/09_barplot_myself.Rmd
finally able to produce the damn bar pltos

@@
 # load the table that has the unique sig PATHWAYIDs
-sigpw <- read_csv('../data/unique_sig_PATHWAYID.csv') %>% 
+sigpw <- read_csv('../data/unique_sig_pathway.csv') %>% 
   # grab the pw id which is gonna be able to match the rownames in the stratified table
-  mutate(pwid = str_extract(PATHWAYID, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
+  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
   mutate(pwid = str_replace_all(pwid, '_', '-'))
@@
 siglefse <- stratified %>% 
-  filter(str_detect(PATHWAYID,  paste0(sigpw$pwid, collapse = '|'))) %>% 
-  separate(PATHWAYID, into = c('pwid','taxa'), sep = '\\|')
+  filter(str_detect(pathway,  paste0(sigpw$pwid, collapse = '|'))) %>% 
+  separate(pathway, into = c('pwid','taxa'), sep = '\\|')
 
@@
 
+# aoutput the reformatted sigpw table
+sigpw %>% 
+  write_csv('../data/unique_sig_pathway_with_clean_pwid.csv')
 ```

### 2019-10-17 scripts/humann2_bar_cmd.sh

@@ -1,0 +1,23 @@
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature ARGDEG-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/ARGDEG-PWY_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature HOMOSER-METSYN-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/HOMOSER-METSYN-PWY_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature MET-SAM-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/MET-SAM-PWY_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature METSYN-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/METSYN-PWY_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature ORNARGDEG-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/ORNARGDEG-PWY_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature P125-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/P125-PWY_pw.png
[…]

### 2019-10-18 scripts/08_barplot_species_pathway.Rmd
update

@@
 ```{r}
+# highlight the bar plots of tox in pathways only come out tox lefse
+tox_pathway <- read_csv('../data/lefse_stats_all_sig_pathway.csv') %>% 
+  filter(str_detect(group, '^Toxicity')) %>% 
+  distinct(pathway, .keep_all = T) %>% 
+  select(pathway)  %>% 
+  mutate(pathway = str_replace(pathway, '^f_','')) %>% 
+  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
+  mutate(pwid = str_replace_all(pwid, '_', '-'))
+
+```
+
+# --exclude-unclassified
+
+```{r}
 # load the sig pathway table with pwid
-cmd_barplot <- read_csv('../data/unique_sig_pathway_with_clean_pwid.csv') %>% 
+cmd_barplot <- tox_pathway %>% 
   select(pwid) %>% 
-  mutate(barcmd = str_glue('humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature {pwid} --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/{pwid}_pw.png')) %>%
+  mutate(barcmd = str_glue('humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature {pwid} --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/{pwid}_pw_no_unclassfified.png')) %>%
   select(barcmd)
@@
 cmd_barplot %>% 
-  write.table('humann2_bar_cmd.sh', quote = F, col.names = F, row.names = F )
+  write.table('humann2_bar_cmd_no_unclassfified_tox.sh', quote = F, col.names = F, row.names = F )
 ```
 
-```{bash}
-bash humann2_bar_cmd.sh
+# has unclassified
+
+```{r}
+# with unclassfified
+cmd_barplot <- tox_pathway %>% 
+  select(pwid) %>% 
+  mutate(barcmd = str_glue('humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature {pwid} --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/{pwid}_pw.png')) %>%
+  select(barcmd)
+
+cmd_barplot %>% 
+  write.table('humann2_bar_cmd_tox.sh', quote = F, col.names = F, row.names = F )
 ```

### 2019-10-18 scripts/command.sh

@@ -1,9 +1,0 @@
-humann2_barplot  \
-    --input $1 \
-    --focal-feature $2 \
-    --sort similarity metadata \
-    --focal-metadatum Toxicity \
-    --last-metadatum Toxicity \
-    --scaling pseudolog \
-    --dimensions 10 5 \
-    -o $2.png

### 2019-10-18 scripts/humann2_bar_cmd.sh

@@ -1,23 +1,23 @@
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature ARGDEG-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/ARGDEG-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature HOMOSER-METSYN-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/HOMOSER-METSYN-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature MET-SAM-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/MET-SAM-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature METSYN-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/METSYN-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature ORNARGDEG-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/ORNARGDEG-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature P125-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o ../figs/barplot_humann2/P125-PWY_pw.png
[…]

### 2019-10-18 scripts/humann2_bar_cmd_no_unclassfified.sh

@@ -1,0 +1,23 @@
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature ARGDEG-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/ARGDEG-PWY_pw_no_unclassfified.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature HOMOSER-METSYN-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/HOMOSER-METSYN-PWY_pw_no_unclassfified.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature MET-SAM-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/MET-SAM-PWY_pw_no_unclassfified.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature METSYN-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/METSYN-PWY_pw_no_unclassfified.png
[…]

### 2019-10-18 scripts/humann2_bar_cmd_no_unclassfified_tox.sh

@@ -1,0 +1,10 @@
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature 7ALPHADEHYDROX-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/7ALPHADEHYDROX-PWY_pw_no_unclassfified.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-4242 --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/PWY-4242_pw_no_unclassfified.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature CITRULBIO-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/CITRULBIO-PWY_pw_no_unclassfified.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-5695 --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/PWY-5695_pw_no_unclassfified.png
[…]

### 2019-10-18 scripts/humann2_bar_cmd_tox.sh

@@ -1,0 +1,10 @@
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature 7ALPHADEHYDROX-PWY --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/7ALPHADEHYDROX-PWY_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-4242 --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/PWY-4242_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature CITRULBIO-PWY --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/CITRULBIO-PWY_pw.png
+humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-5695 --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/PWY-5695_pw.png
[…]

### 2019-10-18 scripts/01_clean_pheno.Rmd
update on the genefamily lefse

@@
 # this sheet have nice sampleids for the samples, and also this is the table has all of the info
-phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10162019.xlsx',
+phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10172019.xlsx',
                     sheet = 5)

### 2019-10-18 scripts/10_barplot_genefamilies.Rmd
update on the genefamily lefse

---
title: "Humann2 bar plot for gene family"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

######################################## Bash ########################################

```{bash}
# needs to download the utility mapping db first
humann2_databases --download utility_mapping full /Users/daia1/melodyP/data/humann2_res
```

```{bash}
# regroup to KEGG Orthogroups (KOs)
humann2_regroup_table \
            --input ../data/humann2_res/humann2_genefamilies.tsv  \
            --groups uniref90_ko  \
            --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
```

```{bash}
# renorm to relab
humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u relab --output ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv
```

```{bash}
# split into unstratified and ..
humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv  --output ../data/humann2_res/genefamily_split_KO
```

######################################## R ########################################

## What features(gene orthologs) do you wanna see? Do lefse first.

```{r}

```

### 2019-10-18 scripts/11_lefse_geneorthologs.Rmd
update on the genefamily lefse

---
title: "Lefse on gene orthologs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Make pheno table(with only classes)

```{r}
# loading the table with the valid sampleids
pheno <- read_csv('../data/current_pheno.csv') %>% 
  filter(! Sampleid %in% c("1725E", "2222A" ,"2245A"))

```

```{r}
# get different combination of the classes
classes <- c('CR', 'Toxicity')

#subclass <- c('Disease_Status_CAR','Disease_Class','CAR_Group')
combi <- expand.grid(classes = classes)

# use each combi row in the above df to create the pheno table for the lefse input
apply(combi, 1, function(row) {
  pheno %>% 
    mutate(CR = str_glue('CR_{CR}'),
           Toxicity = str_glue('Toxicity_{Toxicity}')) %>% 
    select(row[['classes']]) %>% 
    t() %>% 
    write.table(str_glue('../data/lefse_genefam/pheno__{row[["classes"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
})
```

## Reorder the GO stratified table

```{r}
stra_go <- read_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified.tsv') %>% 
  rename(sampleid = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance-RPKs$','')))

stra_go <- bind_cols(stra_go[,1],stra_go[,pheno$Sampleid])

all.equal(pheno$Sampleid, colnames(stra_go)[2:ncol(stra_go)])

# write out

stra_go %>% 
  write_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv')
```

## str_glue

```{r}
# create the commad in df with str_glue and then output to a bash script
pheno_combi <- data_frame(pheno_file = list.files('../data/lefse_genefam', pattern = '.txt$', full.names = T)) %>% 
  mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))

pheno_combi %>% 
  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv > ../data/lefse_genefam/genefam_abundace_{combi}tsv')) %>% 
  select(cat_cmd) %>% 
  write_csv('cat_combi_genef.sh', col_names = F)
```

```{r}
# create the cmd to do the format input stuff
pheno_combi %>% 
  mutate(input_tsv = str_glue('../data/lefse_genefam/genefam_abundace_{combi}tsv'),
         fmt_input = str_replace(input_tsv, 'tsv','in'),
         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -u 2 -o 1000000')) %>% 
  select(fmt_cmd) %>% 
  write_csv('format_input_genef.sh', col_names = F)
```

```{r}
# the actual step to run lefse
pheno_combi %>% 
  mutate(fmt_input = str_glue('../data/lefse_genefam/genefam_abundace_{combi}in'),
         lefse_res = str_replace(fmt_input, 'in$','res'),
         run_cmd = str_glue('run_lefse.py {fmt_input} {lefse_res}')) %>% 
  select(run_cmd) %>% 
  write_csv('run_lefse_genef.sh', col_names = F)
```

```{r}
# plotting cmd 
pheno_combi %>% 
  mutate(lefse_res = str_glue('../data/lefse_genefam/genefam_abundace_{combi}res'),
         fea_png = str_glue('../figs/genefam_abundace_{combi}png'),
         cmd = str_glue('plot_res.py {lefse_res}  {fea_png} --feature_font_size 4 --width 10 --dpi 300 --title {combi}')) %>% 
  select(cmd) %>% 
  write_csv('plotting_de_feature_genef.sh', col_names = F)
```

## bash

```{bash}
bash cat_combi_genef.sh
bash format_input_genef.sh
bash run_lefse_genef.sh
bash plotting_de_feature_genef.sh
```

### 2019-10-18 scripts/cat_combi_genef.sh

@@ -1,0 +1,2 @@
+cat ../data/lefse_genefam/pheno__CR.txt ../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv > ../data/lefse_genefam/genefam_abundace_pheno__CR.tsv
+cat ../data/lefse_genefam/pheno__Toxicity.txt ../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv > ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.tsv

### 2019-10-18 scripts/format_input_genef.sh

@@ -1,0 +1,2 @@
+format_input.py ../data/lefse_genefam/genefam_abundace_pheno__CR.tsv ../data/lefse_genefam/genefam_abundace_pheno__CR.in -c 1 -u 2 -o 1000000
+format_input.py ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.tsv ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.in -c 1 -u 2 -o 1000000

### 2019-10-18 scripts/plotting_de_feature_genef.sh

@@ -1,0 +1,2 @@
+plot_res.py ../data/lefse_genefam/genefam_abundace_pheno__CR.res  ../figs/genefam_abundace_pheno__CR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR.
+plot_res.py ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.res  ../figs/genefam_abundace_pheno__Toxicity.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity.

### 2019-10-18 scripts/run_lefse_genef.sh

@@ -1,0 +1,2 @@
+run_lefse.py ../data/lefse_genefam/genefam_abundace_pheno__CR.in ../data/lefse_genefam/genefam_abundace_pheno__CR.res
+run_lefse.py ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.in ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.res

### 2019-10-18 scripts/10_barplot_genefamilies.Rmd
update

@@
 ```{r}
+# load the lefse res of tox
+# the input : lefse res files
+fns <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$', full.names = T)
+fns_name <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$')
+
+# join all of the tables feature together
+feature <- fns %>% 
+  set_names(fns_name) %>% 
+  # when meeting parsing failures, you need to specify all column types.
+  map(~ read_tsv(., col_names = c('genefam','xx','direction','score','pval'),  col_types = 'ccccc') %>% 
+        filter(!is.na(score))) %>% 
+  bind_rows(.id = 'group') %>% 
+  mutate(group = str_replace(group, 'genefam_abundace_pheno__','')) %>% 
+  mutate(group = str_replace(group, '.res$','')) %>% 
+  select(-xx, -pval)
+
+# change the "N" direction to be minus score
+# feature <- bind_rows(
+#   feature %>% 
+#     split(.$direction) %>% 
+#     pluck('N') %>% 
+#     mutate(score = -score),
+#   feature %>% 
+#     split(.$direction) %>% 
+#     pluck('Y') 
+# ) %>% 
+#   arrange(group, pathway, score)
+
+feature %>% 
+  write_csv('../data/lefse_genefam/with_unmapped_sig_feature.csv')
 

### 2019-10-18 scripts/11_lefse_geneorthologs.Rmd
update

@@
   rename(sampleid = names(.)[1]) %>% 
-  rename_all(funs(str_replace(., '_Abundance-RPKs$','')))
+  rename_all(funs(str_replace(., '_Abundance-RPKs$',''))) %>% 
+  # remove the ungrouped since it's not really giving gene fam info
+  filter(! str_detect(sampleid, '^UNGROUPED'))
 
@@
 bash run_lefse_genef.sh
-bash plotting_de_feature_genef.sh
+
 ```

### 2019-10-20 scripts/12_lefse_genefam_term_ggplot.Rmd
haha able to retrieve the actual term of KO

--- 2019-10-18 scripts/10_barplot_genefamilies.Rmd
@@
 ---
-title: "Humann2 bar plot for gene family"
+title: ""
 output: html_document
@@
 library(tidyverse)
+library(KEGGREST)
 ```
 
-######################################## Bash ########################################
-
-```{bash}
-# needs to download the utility mapping db first
-humann2_databases --download utility_mapping full /Users/daia1/melodyP/data/humann2_res
-```
-
-```{bash}
-# regroup to KEGG Orthogroups (KOs)
-humann2_regroup_table \
-            --input ../data/humann2_res/humann2_genefamilies.tsv  \
-            --groups uniref90_ko  \
-            --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
-```
-
-```{bash}
-# renorm to relab
-humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u relab --output ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv
-```
-
-```{bash}
-# split into unstratified and ..
-humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv  --output ../data/humann2_res/genefamily_split_KO
-```
-
-######################################## R ########################################
-
-## What features(gene orthologs) do you wanna see? Do lefse first.
+## Load the lefse results
 
@@
   mutate(group = str_replace(group, '.res$','')) %>% 
-  select(-xx, -pval)
+  select(-xx, -pval) %>% 
+  mutate(KOID = str_replace(genefam, '\\.g.+$',''))
 
@@
 
-feature %>% 
-  write_csv('../data/lefse_genefam/with_unmapped_sig_feature.csv')
+# feature %>% 
+#   write_csv('../data/lefse_genefam/with_unmapped_sig_feature.csv')
 
 ```
+
+## Use KEGGREST pkg to get the actual term
+
+### work on the Toxicity first
+
+```{r}
+tox_df <-  feature %>% 
+  filter(group == 'Toxicity') %>% 
+  split(cut_width(1:nrow(.), 10, boundary=0)) %>% 
+  map(~ pull(.data = ., KOID) %>% 
+        map(~ list(KO_name = keggGet(.)[[1]]$NAME,
+             KO_def = keggGet(.)[[1]]$DEFINITION)) %>% 
+        bind_rows()) %>% 
+  bind_rows()
+
+```

### 2019-10-21 scripts/12_lefse_genefam_term_ggplot.Rmd
update

@@
 
-# change the "N" direction to be minus score
-# feature <- bind_rows(
-#   feature %>% 
-#     split(.$direction) %>% 
-#     pluck('N') %>% 
-#     mutate(score = -score),
-#   feature %>% 
-#     split(.$direction) %>% 
-#     pluck('Y') 
-# ) %>% 
-#   arrange(group, pathway, score)
-
 # feature %>% 
@@
 ```{r}
+# since this pkg only retrieve 10 records one time, so need to split the df.
 tox_df <- feature %>% 
-  filter(group == 'Toxicity') %>% 
+  filter(group == 'Toxicity')
+
+tox_ko <- tox_df %>% 
   split(cut_width(1:nrow(.), 10, boundary=0)) %>% 
   map(~ pull(.data = ., KOID) %>% 
-        map(~ list(KO_name = keggGet(.)[[1]]$NAME,
+        map(~ list(KOID = keggGet(.)[[1]]$ENTRY,
+                   KO_name = keggGet(.)[[1]]$NAME,
                    KO_def = keggGet(.)[[1]]$DEFINITION)) %>% 
         bind_rows()) %>% 
-  bind_rows()
+  bind_rows() %>% 
+  distinct(KOID, .keep_all = T)
+
+# in the table some have duplicated KOID since several terms have a trailing species name
+tox_lefse <- tox_df %>% 
+  left_join(tox_ko, by = 'KOID')
 
 ```
+
+```{r}
+# format to visualize in lefse style
+# change the "N" direction to be minus score
+tox_final <- bind_rows(
+  tox_lefse %>%
+    split(.$direction) %>%
+    pluck('Toxicity_N') %>%
+    mutate(score = -as.numeric(score)),
+  tox_lefse %>%
+    split(.$direction) %>%
+    pluck('Toxicity_Y') %>%
+    mutate(score = as.numeric(score))
+) %>% 
+  # wanna highlight there are some terms have trailing species
+  mutate(have_species = if_else(str_detect(genefam, '\\.g__.+$'), str_extract(genefam, 'g__.+$'), ''))   %>% 
+  mutate(have_FP = if_else(have_species == 'g__Faecalibacterium_s__Faecalibacterium_prausnitzii', 'YES', 'NO')) %>% 
+  # combine the name and def to be a new var
+  mutate(KO_term = str_glue('{KO_name}: {KO_def}'))
+
+# for plotting purpose plotting the unique KO_term rows 
+ggplot(data = tox_final %>% 
+         distinct(KO_term, .keep_all = T) , aes(x = reorder(KO_term, score), y = score, fill = direction, color = have_FP)) +
+  geom_bar(stat='identity') +
+  coord_flip() +
+  scale_color_manual(values = c('blue', 'yellow'))+
+  scale_fill_manual(values = c('#EC0000', '#42B440')) +
+  theme_minimal() +
+  theme(axis.title.y  = element_blank(), 
+        legend.position="bottom") +
+  labs(title = str_glue('Response to class : Toxicity') ,
+       y = 'Score') + 
+  ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/KEGG_ortholog_lefse_plot/Response to class: Toxicity.pdf'), device = 'pdf', width = 18, height = 10, dpi = 300)
+
+```
+
+```{r}
+
+```

### 2019-10-21 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
PCoA with KO abundance

---
title: "PCoA with bray curtis on KO abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vegan)
```

Attempt on making the figure 7 A and B of paper https://www.nature.com/articles/srep34826
Making PCoA plot with Bray-Curtis distance on the KO abundace matrix

```{r}
# using the unstratified cts first
pheno <- read_csv('../data/current_pheno.csv') %>% 
  filter(! Sampleid %in% c("1725E", "2222A" ,"2245A"))

ko <- read_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_unstratified.tsv') %>% 
  rename(KO = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance-RPKs$','')))

ko <- bind_cols(ko[,1],ko[,pheno$Sampleid]) 
all.equal(pheno$Sampleid, colnames(ko)[2:ncol(ko)])
```
```{r}

```

### 2019-10-21 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
no separation

@@
 all.equal(pheno$Sampleid, colnames(ko)[2:ncol(ko)])
+
+# turn the ko table in to a matrix
+ko_matrix <- ko %>% 
+  column_to_rownames('KO') %>% 
+  as.matrix()
 ```
+
 ```{r}
+pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_unstratified.tsv') %>% 
+  rename(pw = names(.)[1]) %>% 
+  rename_all(funs(str_replace(., '_Abundance$','')))
 
+pw <- bind_cols(pw[,1],pw[,pheno$Sampleid]) 
+all.equal(pheno$Sampleid, colnames(pw)[2:ncol(pw)])
+
+pw_matrix <- pw %>% 
+  column_to_rownames('pw') %>% 
+  as.matrix()
 ```
+
+```{r}
+# PCoA
+pheno_col <- pheno %>% 
+  mutate(tox_col = if_else(Toxicity == 'Y', 'red', 'green '),
+         CR_col = if_else(CR == 'Y', 'red', 'green '))
+
+bc_dist <-  vegdist(t(pw_matrix), "bray")
+pc_bc <- cmdscale(bc_dist, k = 2)
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$tox_col, main = 'Pathway abundance color by tox')
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$CR_col, main = 'Pathway abundance color by CR')
+```
+
+```{r}
+# do the PcoA with bray curtis distance  
+bc_dist <-  vegdist(t(ko_matrix), "bray")
+pc_bc <- cmdscale(bc_dist, k = 2)
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$tox_col, main = 'KO abundance color by tox')
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$CR_col, main = 'KO abundance color by CR')
+```
+
+It just doesn't separate whether using gene ortholog or pathway abundance.

### 2019-10-21 scripts/10_genefamilies_KO_regroup.Rmd
haha able to retrieve the actual term of KO
update

--- 2019-10-18 scripts/10_barplot_genefamilies.Rmd
@@
 
-######################################## R ########################################
-
-## What features(gene orthologs) do you wanna see? Do lefse first.
+######################################## R ############################################
 
@@
 
+# reorder the unstratified table
+genef_KO <- read_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_unstratified.tsv')%>% 
+  rename(KO = names(.)[1]) %>% 
+  rename_all(funs(str_replace(., '_Abundance-RPKs$','')))
+
+pheno <- read_csv('../data/current_pheno.csv') %>% 
+  filter(! Sampleid %in% c("1725E", "2222A" ,"2245A"))
+
+genef_KO <- bind_cols(genef_KO[,1],genef_KO[,pheno$Sampleid]) 
+all.equal(pheno$Sampleid, colnames(genef_KO)[2:ncol(genef_KO)])
+
 ```
+```{r}
+# look at the relab abundance of the pathways related to pantothenate for example
+patha <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_unstratified.tsv') %>% 
+  # choose the pantothenate and fol
+```

### 2019-10-21 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
update

@@
 ```
+
+## Looking at the whole table
 
@@
 It just doesn't separate whether using gene ortholog or pathway abundance.
+
+## Looking at subset of pathways
+
+```{r}
+vb_matrix <- pw %>% 
+  filter(str_detect(pw, 'folate|pantothenate|cobalamin|biotin')) %>% 
+  column_to_rownames('pw') %>% 
+  as.matrix()
+
+bc_dist <-  vegdist(t(vb_matrix), "bray")
+pc_bc <- cmdscale(bc_dist, k = 2)
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$tox_col, main = 'Pathway abundance color by tox')
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$CR_col, main = 'Pathway abundance color by CR')
+
+```

### 2019-10-21 scripts/14_pathway_KO_regroup.Rmd
update

---
title: "pathway KO regroup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{bash}
# regroup to KEGG Orthogroups (KOs)
humann2_regroup_table \
            --input ../data/humann2_res/humann2_pathabundance.tsv  \
            --groups uniref90_pfam  \
            --output ../data/humann2_res/humann2_pathabundance_pfam_regroup.tsv
```

```{bash}
# renorm to relab
<!-- humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u relab --output ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv -->
```

```{bash}
# split into unstratified and ..
<!-- humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv  --output ../data/humann2_res/genefamily_split_KO -->
```

### 2019-10-23 scripts/10_genefamilies_KO_regroup.Rmd
update

@@
 library(tidyverse)
+library(ggpubr)
 ```
@@
 ```{r}
+patha
+```
+
+```{r}
 # look at the relab abundance of the pathways related to pantothenate for example
 patha <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_unstratified.tsv') %>% 
-  # choose the pantothenate and fol
+  rename(pw = names(.)[1]) %>% 
+  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
+  filter(str_detect(pw, 'folate|pantothenate|cobalamin|biotin'))
+
+# gather and combine with pheno info
+patha <- patha %>% 
+  gather(key = 'Sample', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
+  left_join(pheno %>% 
+              select(Sample = Sampleid, CR, Toxicity), by = 'Sample')
+
+# plot
+
+stat.test <- compare_means(
+  relab ~ pw, data = patha, group.by = "Toxicity",
+  method = "wilcox.test", ref.group = 'PWY-2201: folate transformations I'
+)
+
+patha %>% 
+  mutate(pw = str_replace(pw, '.+:','')) %>% 
+  ggboxplot(x = 'pw', y = 'relab', color = 'Toxicity', palette  = 'lancet') +
+  theme(axis.text.x = element_text(angle = 15, hjust = 1)) + 
+  #stat_compare_means(aes(group = Toxicity), label = "p.signif") +
+  labs(x = 'Vitamin B related pathways',
+       y = 'relative abundance',
+       title = 'Relative abundance in different response to Toxicity in vb related pathways') +
+  ggsave('../figs/Relative abundance in different response to Toxicity in vb related pathways.jpg', width = 10, height = 7, dpi = 300)
+
 ```
+
+```{r}
+load('../data/kegg_lookup_tables.RData')
+```

### 2019-10-24 scripts/01_clean_pheno.Rmd
update

@@
 # this sheet have nice sampleids for the samples, and also this is the table has all of the info
-phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10172019.xlsx',
+phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10222019.xlsx',
                     sheet = 5)
@@
   select(Sampleid = Baseline_Sample_ID, 
-         Censor_YN,
          Censor_reason,
@@
          CAR_Group) %>% 
-  filter(is.na(Censor_reason)) %>% 
-  # remove this one record that doesn't seem to be a valid record at all
-  filter(!is.na(CR))
+  filter(is.na(Censor_reason))
 
@@
   filter(sample_id != '1943A') %>% 
-  # THERE ARE THREE THAT ARE CURRENTLY BEING SEQUENCED, WE DON'T HAVE YET!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-  filter(!sample_id %in% c('161','171','173'))
+  # 171 is back, removing the the other two that doesn't make to the final valid sample
+  filter(!sample_id %in% c('161','173'))
 
@@
 ```
+
 ```{r}

### 2019-10-24 scripts/06_lefse_res_ggplot.Rmd
update

@@
 library(tidyverse)
+library(grid)
+library(cowplot)
+library(ggpubr)
 ```
@@
 feature %>% 
+  filter(str_detect(group, pattern = 'Disease_Class$')) %>% 
   split(.$group) %>% 
   map(~  ggplot(data = ., aes(x = reorder(pathway, score), y = score, fill = direction)) +
-  geom_bar(stat='identity') +
+            geom_bar(position = position_dodge2(preserve = "single"), stat = 'identity') +
             coord_flip() +
@@
                   legend.position="bottom") +
-  labs(title = str_glue('Response to class (with subclass): {.$group}') ,
+            labs(title = str_glue('Response to class') ,
                  y = 'Score')+ 
-  ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to class (with subclass): {.$group}.pdf'), device = 'pdf', width = 15, height = 6, dpi = 300))
+            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to class {.$group}.pdf'), device = 'pdf', width = 15, dpi = 300))
 
 ```
+
+```{r}
+CR <- feature %>% 
+  filter(str_detect(group, pattern = 'Disease_Class$')) %>% 
+  filter(group == 'CR__Disease_Class') %>% 
+  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
+            geom_bar(position = position_dodge2(preserve = "single"), stat = 'identity') +
+            coord_flip() +
+            scale_color_manual(values = c('#00468B', '#EC0000'))  +
+            scale_fill_manual(values = c('#00468B', '#EC0000')) +
+            theme_minimal() +
+            theme(axis.title.y  = element_blank(), 
+                  legend.position="bottom") +
+            labs(title = str_glue('Response to Complete Response') ,
+                 y = 'Score')
+
+tox <- feature %>% 
+  filter(str_detect(group, pattern = 'Disease_Class$')) %>% 
+  filter(group == 'Toxicity__Disease_Class') %>% 
+  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
+            geom_bar(position = position_dodge2(preserve = "single"), stat = 'identity') +
+            coord_flip() +
+            scale_color_manual(values = c('#00468B', '#EC0000'))  +
+            scale_fill_manual(values = c('#00468B', '#EC0000')) +
+            theme_minimal() +
+            theme(axis.title.y  = element_blank(), 
+                  legend.position="bottom") +
+            labs(title = str_glue('Response to Toxicity') ,
+                 y = 'Score')
+
+g <- plot_grid(CR, tox,
+          nrow = 2, 
+          align = 'hv',
+          labels = 'AUTO',
+          axis = 'b')  +
+  ggsave('../figs/dtes.jpg', width = 13, height = 12, dpi = 300)
+```

### 2019-10-24 scripts/09_barplot_myself.Rmd
update

@@
 # pheno table 
+valid_s <- read_table('../data/input_for_cluster_sampleid/valid_sample_names.txt', col_names = F)
+
 pheno <- read_csv('../data/cleaned_pheno/current_pheno.csv')%>% 

### 2019-10-28 scripts/01_clean_pheno.Rmd
update the final sample name

@@
 # this sheet have nice sampleids for the samples, and also this is the table has all of the info
-phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10222019.xlsx',
+phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10232019 wo Censor .xlsx',
                     sheet = 5)
@@
 # output this pheno table cuz its gonna be used downstream
+# this is all of the baseline samples that has 16s.
 phe_M_clean %>% 
@@
   filter(sample_id != '1943A') %>% 
-  # 171 is back, removing the the other two that doesn't make to the final valid sample
-  filter(!sample_id %in% c('161','173'))
+  ### FUCK THIS TABLE OF ../data/baselinesforAngel10032019.csv IS WRONG IN 161, it didn’t pass QC because not enough DNA material was present. SO WE DON'T HAVE THE SHOTGUN DATA.
+  filter(!sample_id %in% c('161'))
 
@@
 ggvenn(two)
+
+# so the final shotgun samples number will be 33
+length(intersect(have_sg$sample_id, phe_M_clean$Sampleid))
+
+final_sample_id <- intersect(have_sg$sample_id, phe_M_clean$Sampleid)
+
+# so the final removed 2 samples are:
+setdiff(pheno_col$Sampleid, final_sample_id) 
 ```

### 2019-10-28 scripts/06_lefse_res_ggplot.Rmd
update the final sample name

@@
             labs(title = str_glue('Response to Complete Response') ,
-                 y = 'Score')
+                 y = 'Score')+ 
+            ylim(-4,4) +
+            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to CR.jpg'),  width = 12, dpi = 300, )
 
@@
             labs(title = str_glue('Response to Toxicity') ,
-                 y = 'Score')
+                 y = 'Score')+ 
+            ylim(-4,4) +
+            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to tox.jpg'),  width = 12, dpi = 300, height = 3.5)
 
-g <- plot_grid(CR, tox,
-          nrow = 2, 
-          align = 'hv',
-          labels = 'AUTO',
-          axis = 'b')  +
-  ggsave('../figs/dtes.jpg', width = 13, height = 12, dpi = 300)
 ```

### 2019-10-28 scripts/03_lefse_CR_tox.Rmd
update

@@
 
-Now we have 3 subclasses and together with the 2 classes of 35 that gonna make a total combination of 6
-
 ## Sorting out the pheno table for the valid samples
@@
 ```{r}
-# loading the table with the valid sampleids
-valid_s <- read_table('../data/input_for_cluster_sampleid/valid_sample_names.txt', col_names = F)
-
-pheno <- read_csv('../data/cleaned_pheno/current_pheno.csv') %>% 
-  filter(Sampleid %in% valid_s$X1) %>% 
-  select(-Censor_YN, -Censor_reason ) 
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
+  select(-CAR_Group, -Censor_reason ) 
 ```
@@
 classes <- c('CR', 'Toxicity')
-subclass <- c('Disease_Status_CAR','Disease_Class','CAR_Group')
+subclass <- c('Disease_Class')
 
@@
 # create the commad in df with str_glue and then output to a bash script
-pheno_combi <- data_frame(pheno_file = list.files('../data/lefse_input', pattern = '.txt$', full.names = T)) %>% 
+pheno_combi <- tibble(pheno_file = list.files('../data/lefse_input', pattern = '.txt$', full.names = T)) %>% 
   mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))
@@
   select(cat_cmd) %>% 
-  write_csv('cat_combi_6.sh', col_names = F)
+  write_csv('cat_combi_CT.sh', col_names = F)
 ```
@@
   select(fmt_cmd) %>% 
-  write_csv('format_input_6.sh', col_names = F)
+  write_csv('format_input_CT.sh', col_names = F)
 ```
@@
   select(run_cmd) %>% 
-  write_csv('run_lefse.sh', col_names = F)
+  write_csv('run_lefse_CT.sh', col_names = F)
 ```
@@
   select(cmd) %>% 
-  write_csv('plotting_de_feature.sh', col_names = F)
+  write_csv('plotting_de_feature_CT.sh', col_names = F)
 ```
@@
 # this is the bash script to do the actual concatenation 
-bash cat_combi_6.sh
+bash cat_combi_CT.sh
+bash format_input_CT.sh
+bash run_lefse_CT.sh
+bash plotting_de_feature_CT.sh
 ```
-
-```{bash}
-# run in the cmd line THIS HAS TO BE DONE IN THE CMD SINCE RSTUDIO DOES NOT RECOGNIZE THE LEFSE SCRIPT
-bash format_input_6.sh
-```
-
-```{bash}
-bash run_lefse.sh
-```
-
-```{bash}
-bash plotting_de_feature.sh
-```

### 2019-10-28 scripts/cat_combi_3.sh

@@ -1,3 +1,0 @@
-cat ../data/lefse_input/pheno__ORR__CAR_Group.txt ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.tsv
-cat ../data/lefse_input/pheno__ORR__Disease_Class.txt ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.tsv
-cat ../data/lefse_input/pheno__ORR__Disease_Status_CAR.txt ../data/cts_36_reorder.tsv > ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.tsv

### 2019-10-28 scripts/cat_combi_6.sh

@@ -1,6 +1,0 @@
-cat ../data/lefse_input/pheno__CR__CAR_Group.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.tsv
-cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv
-cat ../data/lefse_input/pheno__CR__Disease_Status_CAR.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.tsv
-cat ../data/lefse_input/pheno__Toxicity__CAR_Group.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.tsv
-cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv
-cat ../data/lefse_input/pheno__Toxicity__Disease_Status_CAR.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.tsv

### 2019-10-28 scripts/cat_combi_CT.sh

@@ -1,0 +1,2 @@
+cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv
+cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv

### 2019-10-28 scripts/format_input_3.sh

@@ -1,3 +1,0 @@
-format_input.py ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.tsv ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.tsv ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000

### 2019-10-28 scripts/format_input_6.sh

@@ -1,6 +1,0 @@
-format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.tsv ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.tsv ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000

### 2019-10-28 scripts/format_input_CT.sh

@@ -1,0 +1,2 @@
+format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000

### 2019-10-28 scripts/plotting_de_feature_CT.sh

@@ -1,0 +1,2 @@
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res  ../figs/pathway_abundace_pheno__CR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Class.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res  ../figs/pathway_abundace_pheno__Toxicity__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Class.

### 2019-10-28 scripts/plottint_de_feature_3.sh

@@ -1,3 +1,0 @@
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.res  ../figs/pathway_abundace_pheno__ORR__CAR_Group.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__ORR__CAR_Group.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.res  ../figs/pathway_abundace_pheno__ORR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__ORR__Disease_Class.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.res  ../figs/pathway_abundace_pheno__ORR__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__ORR__Disease_Status_CAR.

### 2019-10-28 scripts/run_lefse_3.sh

@@ -1,3 +1,0 @@
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.in ../data/lefse_input/pathway_abundace_pheno__ORR__CAR_Group.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Class.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.in ../data/lefse_input/pathway_abundace_pheno__ORR__Disease_Status_CAR.res

### 2019-10-28 scripts/run_lefse_CT.sh

@@ -1,0 +1,2 @@
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res

### 2019-10-29 scripts/06_lefse_res_ggplot.Rmd
update

--- 2019-10-24 scripts/06_lefse_res_ggplot.Rmd
@@
 CR <- feature %>% 
-  filter(str_detect(group, pattern = 'Disease_Class$')) %>% 
   filter(group == 'CR__Disease_Class') %>% 
@@
             coord_flip() +
-            scale_color_manual(values = c('#00468B', '#EC0000'))  +
-            scale_fill_manual(values = c('#00468B', '#EC0000')) +
+            scale_color_manual(values = c('#42B440', '#EC0000'))  +
+            scale_fill_manual(values = c('#42B440', '#EC0000')) +
             theme_minimal() +
@@
             labs(title = str_glue('Response to Complete Response') ,
-                 y = 'Score')
+                 y = 'Score')+ 
+            ylim(-4,4) +
+            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to CR.jpg'),  width = 12, dpi = 300 )
 
 tox <- feature %>% 
-  filter(str_detect(group, pattern = 'Disease_Class$')) %>% 
   filter(group == 'Toxicity__Disease_Class') %>% 

### 2019-10-30 scripts/03_lefse_CR_tox.Rmd
use cpm instead of relab

@@
 # input: relative abundance unstratified for pathway abundance
-relab <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_unstratified.tsv') %>% 
+relab <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
   rename(sampleid = names(.)[1]) %>% 
@@
 relab_or %>% 
-  write_tsv('../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv')
+  write_tsv('../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv')
 ```
@@
 pheno_combi %>% 
-  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_{combi}tsv')) %>% 
+  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_{combi}tsv')) %>% 
   select(cat_cmd) %>% 

### 2019-10-30 scripts/cat_combi_CT.sh

@@ -1,2 +1,2 @@
-cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv
-cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_relab_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv
+cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv
+cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv

### 2019-10-30 scripts/15_RF_find_top_features.Rmd
update on the random forest

---
title: "RF to find the top important features"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
```

```{r}
# the species table from metaphlan2
# the species abundance here are actually in percent
s_tb <- read_tsv('../data/metaphlan2_result/merged_abundance_table_species.txt') %>% 
  rename_all(funs(str_replace(., '_profile$','')))

s_tb_sum <- s_tb %>% 
  column_to_rownames('ID') %>% 
  as.matrix() %>% 
  apply(., 2, sum)

# load the pheno
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
  select(-CAR_Group, -Censor_reason )

# to rearrange the cts table 
s_tb <- bind_cols(s_tb[,1],s_tb[,pheno$Sampleid])

all.equal(colnames(s_tb)[2:ncol(s_tb)], pheno$Sampleid)
```

```{r}
# to put the feature table in the sample as rowname format to do the ML 
s_tb_rf <- s_tb %>% 
  column_to_rownames('ID') %>% 
  as.matrix() %>% 
  t

# check the sample name really match
all.equal(rownames(s_tb_rf), pheno$Sampleid)

tox_RF <- randomForest(s_tb_rf, factor(pheno$Toxicity), importance = T)

impor_feature <- importance(tox_RF) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'species') %>% 
  arrange(MeanDecreaseAccuracy)
```

Although Clostridium innocuum are often present and harmless in healthy people, they have been isolated in various infections and predominantly in patients that are immunocompromised as an opportunistic bacteria.

```{r}
CR_RF <- randomForest(s_tb_rf, factor(pheno$CR), importance = T)

impor_feature <- importance(CR_RF) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'species') %>% 
  arrange(MeanDecreaseAccuracy)
```

```{r}
# the pathway abundance table
pabun <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  rename(pathway = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  # remove the unmapped and the unintegrated record
  filter(!pathway %in% c('UNMAPPED','UNINTEGRATED'))

pabun <- bind_cols(pabun[,1],pabun[,pheno$Sampleid])

pabun_rf <- pabun %>% 
  column_to_rownames('pathway') %>% 
  as.matrix() %>% 
  t

all.equal(rownames(pabun_rf), pheno$Sampleid)

tox_RF_PW <- randomForest(pabun_rf, factor(pheno$Toxicity), importance = T)

impor_pw <- importance(tox_RF_PW) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'pathway') %>% 
  arrange(MeanDecreaseAccuracy)
```

### 2019-10-31 scripts/15_RF_find_top_features.Rmd
random forest

@@
 ---
-title: "RF to find the top important features"
-output: html_document
+title: "Random Forest to find the top important features"
+author: "Anqi Dai"
+date: "`r Sys.Date()`"
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    code_folding: hide
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
 library(randomForest)
+library(ggpubr)
+library(tidyverse)
+library(kableExtra)
 ```
+
+## Response to Toxicity
+
+### Glimpse at pheno table
 
@@
 pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
-  select(-CAR_Group, -Censor_reason )
+  select(-CAR_Group, -Censor_reason ) %>% 
+  mutate_at(vars(str_which(names(.),'Sampleid', negate = T)), factor)
 
@@
 
-all.equal(colnames(s_tb)[2:ncol(s_tb)], pheno$Sampleid)
-```
-
-```{r}
 # to put the feature table in the sample as rowname format to do the ML 
@@
 
+# pheno %>% 
+#   kable(caption = 'Phenotype info of the final 33 baseline sample patients') %>% 
+#   kable_styling(full_width = F)
+
+pheno  %>% 
+  summary 
+```
+
+### Use only the toxicity as the response variable
+
+```{r}
 tox_RF <- randomForest(s_tb_rf, factor(pheno$Toxicity), importance = T)
 
-impor_feature <- importance(tox_RF) %>% 
+impor_feature <- importance(tox_RF, type=1) %>% 
   as.data.frame() %>% 
@@
   arrange(MeanDecreaseAccuracy)
+
+# plot the err rate change
+tox_RF  %>% 
+  pluck('err.rate') %>% 
+  as.data.frame() %>% 
+  mutate(ntree = seq(1, nrow(.))) %>% 
+  gather(key = 'type', value = 'ERR', OOB:Y) %>% 
+  ggline(x = 'ntree', y = 'ERR', color = 'type', palette = 'lancet') +
+  labs(x = 'Number of trees', 
+       y = 'Error rate')
+
+tox_RF
+```
+
+There are much for Y patients than N. If all assigning to Y it will probably like this.
+
+### Use all of the variables in the pheno table to do RF
+
+```{r} 
+# bind the pheno with the transposed cts matrix
+combined <- cbind(s_tb_rf, pheno %>% 
+                    select(-Sampleid))
+
+combined_tox_RF <- randomForest(formula = Toxicity ~ ., data = combined,  importance = T )
+
+combined_tox_RF  %>% 
+  pluck('err.rate') %>% 
+  as.data.frame() %>% 
+  mutate(ntree = seq(1, nrow(.))) %>% 
+  gather(key = 'type', value = 'ERR', OOB:Y) %>% 
+  ggline(x = 'ntree', y = 'ERR', color = 'type', palette = 'lancet') +
+  labs(x = 'Number of trees', 
+       y = 'Error rate')
+
+combined_tox_RF
+```
+
+### Removing some not expressive features
+
+```{r}
+# check the dist of the species relab sum of the table
+s_sum <- s_tb_rf %>% 
+  apply(., 2, sum)
+
+data_frame(species = names(s_sum),
+           relab_sum = s_sum) %>% 
+  gghistogram(x = 'relab_sum', bins = 50)
+
+# remove the species that have less than 1(%) abundance 
+length(s_sum[s_sum > 1])
+
+combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > 1])], pheno %>% 
+                    select(-Sampleid))
+
+combined_tox_RF_fil <- randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T )
+
+combined_tox_RF_fil  %>% 
+  pluck('err.rate') %>% 
+  as.data.frame() %>% 
+  mutate(ntree = seq(1, nrow(.))) %>% 
+  gather(key = 'type', value = 'ERR', OOB:Y) %>% 
+  ggline(x = 'ntree', y = 'ERR', color = 'type', palette = 'lancet') +
+  labs(x = 'Number of trees', 
+       y = 'Error rate')
+
+combined_tox_RF_fil 
+```
+
+No improvements at all.
+
+### See the important features(species)
+
+```{r}
+top_feature <- importance(combined_tox_RF_fil, type=1, scale = F) %>% 
+  as.data.frame() %>% 
+  rownames_to_column(var = 'species') %>% 
+  arrange(desc(MeanDecreaseAccuracy))
+
+top_feature
 ```
@@
 Although Clostridium innocuum are often present and harmless in healthy people, they have been isolated in various infections and predominantly in patients that are immunocompromised as an opportunistic bacteria.
-
-```{r}
-CR_RF <- randomForest(s_tb_rf, factor(pheno$CR), importance = T)
-
-impor_feature <- importance(CR_RF) %>% 
-  as.data.frame() %>% 
-  rownames_to_column(var = 'species') %>% 
-  arrange(MeanDecreaseAccuracy)
-```
-
-```{r}
-# the pathway abundance table
-pabun <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
-  rename(pathway = names(.)[1]) %>% 
-  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
-  # remove the unmapped and the unintegrated record
-  filter(!pathway %in% c('UNMAPPED','UNINTEGRATED'))
-
-pabun <- bind_cols(pabun[,1],pabun[,pheno$Sampleid])
-
-pabun_rf <- pabun %>% 
-  column_to_rownames('pathway') %>% 
-  as.matrix() %>% 
-  t
-
-all.equal(rownames(pabun_rf), pheno$Sampleid)
-
-tox_RF_PW <- randomForest(pabun_rf, factor(pheno$Toxicity), importance = T)
-
-impor_pw <- importance(tox_RF_PW) %>% 
-  as.data.frame() %>% 
-  rownames_to_column(var = 'pathway') %>% 
-  arrange(MeanDecreaseAccuracy)
-```

### 2019-10-31 scripts/16_RF_pathway.Rmd
random forest

---
title: "RF pathway abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(randomForest)
library(ggpubr)
library(tidyverse)
library(kableExtra)
```

```{r}
# load the pheno
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
  select(-CAR_Group, -Censor_reason ) %>% 
  mutate_at(vars(str_which(names(.),'Sampleid', negate = T)), factor)

# the pathway abundance table
pabun <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  rename(pathway = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  # remove the unmapped and the unintegrated record
  filter(!pathway %in% c('UNMAPPED','UNINTEGRATED'))

pabun <- bind_cols(pabun[,1],pabun[,pheno$Sampleid])

pabun_rf <- pabun %>% 
  column_to_rownames('pathway') %>% 
  as.matrix() %>% 
  t

all.equal(rownames(pabun_rf), pheno$Sampleid)

tox_RF_PW <- randomForest(pabun_rf, factor(pheno$Toxicity), importance = T)

impor_pw <- importance(tox_RF_PW, type=1) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = 'pathway') %>% 
  arrange(MeanDecreaseAccuracy) 
```

### 2019-10-31 scripts/15_RF_find_top_features.Rmd
update

@@
     toc_depth: 3
-    code_folding: hide
+
 ---
@@
 
-### Use only the toxicity as the response variable
+### Use all features as predictors
 
@@
 
-### Use all of the variables in the pheno table to do RF
+### Use all features + all pheno var as predictors
 
@@
 
-### Removing some not expressive features
+### Use more abundant features(species abundance > 1%) + all pheno var as predictors
 
@@
 
-No improvements at all.
+No improvements at all. I think because the relatively small sample size and uneven distribution of the groups.
 
@@
 
+Using the third table predictors.
+
 ```{r}
-top_feature <- importance(combined_tox_RF_fil, type=1, scale = F) %>% 
+# repeat the RF M = 100 times
+# pull top 10 features
+N <- 10
+M <- 100 
+pull_top_features <- function(){
+  importance(randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T ) , type=1, scale = F) %>% 
     as.data.frame() %>% 
     rownames_to_column(var = 'species') %>% 
-  arrange(desc(MeanDecreaseAccuracy))
+    arrange(desc(MeanDecreaseAccuracy)) %>% 
+    top_n(N) %>% 
+    pull(species)
+}
 
-top_feature
+res <- replicate(M, pull_top_features())
+
+stable_top_fea <- table(res) %>% 
+  as.data.frame() %>% 
+  rename(species = names(.)[1]) %>% 
+  arrange(desc(Freq))
+
+stable_top_fea %>% 
+  kable(caption = 'Frequency of features shown as top 10 in 100 random forest runs') %>% 
+  kable_styling(full_width = F, position = 'left')
+
 ```
 
-Although Clostridium innocuum are often present and harmless in healthy people, they have been isolated in various infections and predominantly in patients that are immunocompromised as an opportunistic bacteria.
+What's interesting is this species *Ruminococcus gnavus*.
+[This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines, which could be hypothesized to cause toxicity too.

### 2019-11-01 scripts/17_RF_CR_find_top_features copy.Rmd
update

--- 2019-10-31 scripts/15_RF_find_top_features.Rmd
@@
 library(kableExtra)
+library(arrangements)
+library(caret)
 ```
 
-## Response to Toxicity
+## Response to CR
 
@@
 ```{r}
-tox_RF <- randomForest(s_tb_rf, factor(pheno$Toxicity), importance = T)
+tox_RF <- randomForest(s_tb_rf, factor(pheno$CR), importance = T)
 
@@
 
-combined_tox_RF <- randomForest(formula = Toxicity ~ ., data = combined,  importance = T )
+combined_tox_RF <- randomForest(formula = CR ~ ., data = combined,  importance = T )
 
@@
 
-combined_tox_RF_fil <- randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T )
+combined_tox_RF_fil <- randomForest(formula = CR ~ ., data = combined_fil,  importance = T , ntree = 1000 )
 
@@
 No improvements at all. I think because the relatively small sample size and uneven distribution of the groups.
+
+```{r}
+# to use the top N percent features
+
+# to test on different threshold of selecting most abundant species
+sum_thre <- seq(0.5, 1, 0.05)
+tree_num <- 1000
+
+res <- sum_thre %>% 
+  set_names(sum_thre) %>% 
+  map(function(thre) {
+    s_sum_top <- quantile(s_sum, thre)
+    combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
+                    select(-Sampleid))
+
+    tibble(
+      num_fea = length(s_sum[s_sum > s_sum_top]) ,
+      OOB_err_rate = randomForest(formula = CR ~ ., data = combined_fil,  importance = T , ntree = tree_num )[['err.rate']][tree_num, 1]
+
+    ) 
+  }) %>% 
+  bind_rows(.id = 'thre')
+```
+
+Interesting, using no feature info actually has smaller OOB error rate.
+
+```{r}
+# try different combination of pheno to predict. (doesn't seem they are correlated right?)
+# select 2 out of 4 pheno vars other than CR
+all_var <- names(pheno)
+combi <- combinations(all_var[!all_var %in% c('Sampleid','CR')], 2) %>% 
+  as.data.frame(stringsAsFactors = F)
+
+phe_table <- apply(combi, 1, function(Row) {
+  phe_fil = pheno %>% 
+    select(CR, Row[['V1']], Row[['V2']])
+})
+
+# use those tables to do a random forest
+phe_fil_res <- phe_table %>% 
+  map(function(tbl) {
+     tibble(
+      OOB_err_rate = randomForest(formula = CR ~ ., data = tbl,  importance = T , ntree = tree_num )[['err.rate']][tree_num, 1]
+    ) 
+  }) %>% 
+  bind_rows() %>% 
+  bind_cols(combi) %>% 
+  mutate(phe_combi = str_glue('CR_{V1}_{V2}')) %>% 
+  select(-V1 , -V2) %>% 
+  arrange(desc(OOB_err_rate))
+
+# so the combi with the smallest error rate is:
+phe_fil_res %>% 
+  slice(which.min(OOB_err_rate)) %>% 
+  pull(phe_combi)
+
+# Is there correlation in the pheno table between CR and Disease_Class?
+phe_2 <- pheno  %>% 
+  select(CR, Disease_Class) %>% 
+  mutate(Disease_Class = if_else(as.character(Disease_Class) == 'CD19 Malignancy', 'CD19', 'Multi')) %>% 
+  table()
+
+chi2 <-  chisq.test(phe_2)
+chi2$p.value # reject the null hypothesis of independence
+
+# CR and ORR
+phe_2 <- pheno  %>% 
+  select(CR, ORR)  %>% 
+  table()
+
+chi2 <-  chisq.test(phe_2)
+chi2$p.value # reject the null hypothesis of independence
+
+# so they are actually correlated
+```
 

### 2019-11-01 scripts/15_RF_find_top_features.Rmd
update

@@
     toc_depth: 3 
-
+    code_folding: hide
 ---
@@
 library(kableExtra)
+library(arrangements)
 ```
@@
 
+```{r}
+# check whether the pheno vars are correlated (in pairs)
+all_var <- names(pheno)
+combi <- combinations(all_var[all_var != 'Sampleid'], 2) %>% 
+  as.data.frame(stringsAsFactors = F)
+
+phe_cor <- apply(combi, 1, function(Row) {
+  phe_fil = pheno %>% 
+    select(Row[['V1']], Row[['V2']]) %>% 
+    table()
+
+  chisq.test(phe_fil)$p.value
+}) %>% 
+  as.data.frame() %>% 
+  bind_cols(combi) %>% 
+  mutate(phe_var = str_glue('{V1}__{V2}')) %>% 
+  select(-V1, -V2) %>% 
+  rename(corrected_pval = names(.)[1]) %>% 
+  arrange(corrected_pval) %>% 
+  mutate(CORR = if_else(corrected_pval < 0.1, 'Yes', 'No'))
+
+phe_cor %>% 
+  kable(caption = 'Test on independence between two of the pheno vars') %>% 
+  kable_styling(full_width = F, position = 'left')
+```
+
 ### Use all features as predictors
@@
 
-There are much for Y patients than N. If all assigning to Y it will probably like this.
+Basically assigning everything to Y.
 
-### Use all features + all pheno var as predictors
+### Use all features + uncorrelated pheno var as predictors
+
+Uncorrelated pheno var: ORR, Disease_Status_CAR for Toxicity
 
@@
 combined <- cbind(s_tb_rf, pheno %>% 
-                    select(-Sampleid))
+                    select(Toxicity, ORR, Disease_Status_CAR))
 
@@
 
-### Use more abundant features(species abundance > 1%) + all pheno var as predictors
+### Use more abundant features(species abundance > 1%) + uncorrelated pheno var as predictors
 
@@
 combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > 1])], pheno %>% 
-                    select(-Sampleid))
+                    select(Toxicity, ORR, Disease_Status_CAR))
 
-combined_tox_RF_fil <- randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T )
+combined_tox_RF_fil <- randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T , ntree = 1000 )
 
@@
 
-No improvements at all. I think because the relatively small sample size and uneven distribution of the groups.
+No improvements at all. I think because the relatively small sample size and the signal itself is weak. There's no guarantee that there is association between microbiome and the outcome to begin with.
+
+```{r}
+# to use the top N percent features
+
+# to test on different threshold of selecting most abundant species
+sum_thre <- seq(0.5, 1, 0.05)
+tree_num <- 1000
+
+res <- sum_thre %>% 
+  set_names(sum_thre) %>% 
+  map(function(thre) {
+    s_sum_top <- quantile(s_sum, thre)
+    combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
+                    select(Toxicity, ORR, Disease_Status_CAR))
+
+    tibble(
+      num_fea = length(s_sum[s_sum > s_sum_top]) + 2,
+      OOB_err_rate = randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T , ntree = tree_num )[['err.rate']][tree_num, 1]
+
+    ) 
+  }) %>% 
+  bind_rows(.id = 'thre')
+
+res %>% 
+  kable(caption = 'Top N fraction features and the OOB error rate') %>% 
+  kable_styling(full_width = F, position = 'left')
+```
 
@@
 
-Using the third table predictors.
+Using the table with features abundance > 1% and the two uncorrelated vars.
 
@@
 What's interesting is this species *Ruminococcus gnavus*.
-[This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines, which could be hypothesized to cause toxicity too.
+[This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines.
+
+And [this paper](https://www.ncbi.nlm.nih.gov/pubmed/29330424) mentions that "*Faecalibacterium prausnitzii* decrease and *Ruminococcus gnavus* increase were the essential characteristics in CHF patients' gut microbiota".

### 2019-11-01 scripts/151_RF_find_top_features_other.Rmd
update

--- 2019-11-01 scripts/15_RF_find_top_features.Rmd
@@
 # the species abundance here are actually in percent
-s_tb <- read_tsv('../data/metaphlan2_result/merged_abundance_table_species.txt') %>% 
+s_tb <- read_tsv('../data/metaphlan2_result/merged_abundance_table_phylum.txt') %>% 
   rename_all(funs(str_replace(., '_profile$','')))
@@
 
-### Use more abundant features(species abundance > 1%) + uncorrelated pheno var as predictors
+### Use more abundant features(phylum abundance > 1%) + uncorrelated pheno var as predictors
 
@@
 
-```{r}
-# to use the top N percent features
-
-# to test on different threshold of selecting most abundant species
-sum_thre <- seq(0.5, 1, 0.05)
-tree_num <- 1000
-
-res <- sum_thre %>% 
-  set_names(sum_thre) %>% 
-  map(function(thre) {
-    s_sum_top <- quantile(s_sum, thre)
-    combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
-                    select(Toxicity, ORR, Disease_Status_CAR))
-
-    tibble(
-      num_fea = length(s_sum[s_sum > s_sum_top]) + 2,
-      OOB_err_rate = randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T , ntree = tree_num )[['err.rate']][tree_num, 1]
-
-    ) 
-  }) %>% 
-  bind_rows(.id = 'thre')
-
-res %>% 
-  kable(caption = 'Top N fraction features and the OOB error rate') %>% 
-  kable_styling(full_width = F, position = 'left')
-```
-
-### See the important features(species)
+### See the important features(phylum)
 
@@
 # pull top 10 features
-N <- 10 
+N <- 3 
 M <- 100 
@@
 stable_top_fea %>% 
-  kable(caption = 'Frequency of features shown as top 10 in 100 random forest runs') %>% 
+  kable(caption = 'Frequency of features shown as top 3 in 100 random forest runs') %>% 
   kable_styling(full_width = F, position = 'left')
@@
 ```
-
-What's interesting is this species *Ruminococcus gnavus*.
-[This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines.
-
-And [this paper](https://www.ncbi.nlm.nih.gov/pubmed/29330424) mentions that "*Faecalibacterium prausnitzii* decrease and *Ruminococcus gnavus* increase were the essential characteristics in CHF patients' gut microbiota".

### 2019-11-01 scripts/16_RF_find_top_pathways.Rmd
update

--- 2019-11-01 scripts/15_RF_find_top_features.Rmd
@@
 # the species abundance here are actually in percent
-s_tb <- read_tsv('../data/metaphlan2_result/merged_abundance_table_species.txt') %>% 
-  rename_all(funs(str_replace(., '_profile$','')))
-
-s_tb_sum <- s_tb %>% 
-  column_to_rownames('ID') %>% 
-  as.matrix() %>% 
-  apply(., 2, sum)
+s_tb <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
+  rename(pathway = names(.)[1]) %>% 
+  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
+  # remove the unmapped and the unintegrated record
+  filter(!pathway %in% c('UNMAPPED','UNINTEGRATED')) %>% 
+  mutate(pathway = str_extract(pathway, '^.+:')) %>% 
+  mutate(pathway = str_replace_all(pathway, ':','')) %>% 
+  mutate(pathway = str_replace_all(pathway, '-','_'))%>% 
+  mutate(pathway = str_replace_all(pathway, '\\+','_'))
 
@@
 s_tb_rf <- s_tb %>% 
-  column_to_rownames('ID') %>% 
+  column_to_rownames('pathway') %>% 
   as.matrix() %>% 
@@
 tox_RF <- randomForest(s_tb_rf, factor(pheno$Toxicity), importance = T)
-
-impor_feature <- importance(tox_RF, type=1) %>% 
-  as.data.frame() %>% 
-  rownames_to_column(var = 'species') %>% 
-  arrange(MeanDecreaseAccuracy)
 
@@
 
+### Use top N fraction features + uncorrelated pheno var as predictors
+
 ```{r}

### 2019-11-01 scripts/17_RF_CR_find_top_features.Rmd
update

--- 2019-11-01 scripts/17_RF_CR_find_top_features copy.Rmd
@@
     toc_depth: 3
-
+    code_folding: hide
 ---
@@
 library(arrangements)
-library(caret)
 ```
@@
 
-impor_feature <- importance(tox_RF, type=1) %>% 
-  as.data.frame() %>% 
-  rownames_to_column(var = 'species') %>% 
-  arrange(MeanDecreaseAccuracy)
-
 # plot the err rate change
@@
 
-There are much for Y patients than N. If all assigning to Y it will probably like this.
+### Use all features + uncorrelated pheno var as predictors
 
-### Use all features + all pheno var as predictors
+uncorrelated pheno var: Disease_Status_CAR for CR
 
@@
 combined <- cbind(s_tb_rf, pheno %>% 
-                    select(-Sampleid))
+                    select(Disease_Status_CAR, CR))
 
@@
 combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > 1])], pheno %>% 
-                    select(-Sampleid))
+                    select(Disease_Status_CAR, CR))
 
@@
 
-No improvements at all. I think because the relatively small sample size and uneven distribution of the groups.
+### Use top N fraction features + uncorrelated pheno var as predictors
 
@@
     combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
-                    select(-Sampleid))
+                    select(Disease_Status_CAR, CR))
 
@@
   bind_rows(.id = 'thre')
+
+res %>% 
+  kable(caption = 'Top N fraction features and the OOB error rate') %>% 
+  kable_styling(full_width = F, position = 'left')
 ```
-
-Interesting, using no feature info actually has smaller OOB error rate.
 
@@
 
-### See the important features(species)
+### See the important features (species)
 
@@
 pull_top_features <- function(){
-  importance(randomForest(formula = Toxicity ~ ., data = combined_fil,  importance = T ) , type=1, scale = F) %>% 
+  importance(randomForest(formula = CR ~ ., data = combined_fil,  importance = T ) , type=1, scale = F) %>% 
     as.data.frame() %>% 
@@
 ```
-
-What's interesting is this species *Ruminococcus gnavus*.
-[This recent paper](https://www.pnas.org/content/116/26/12672) points out that *Ruminococcus gnavus* produces a polysaccharide molecule that induces the production of inflammatory cytokines, which could be hypothesized to cause toxicity too.

### 2019-11-03 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
update

@@
 ```
+
+#############################################################################
+
+```{r}
+# the PCoA of the species abundance table
+spp <- read_tsv('../data/metaphlan2_result/merged_abundance_table_species.txt') %>% 
+  rename_all(funs(str_replace(., '_profile$','')))
+
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
+  select(-CAR_Group, -Censor_reason ) %>% 
+  mutate_at(vars(str_which(names(.),'Sampleid', negate = T)), factor)
+
+# to rearrange the cts table 
+spp <- bind_cols(spp[,1],spp[,pheno$Sampleid])
+
+# to put the feature table in the sample as rowname format to do the ML 
+spp <- spp %>% 
+  column_to_rownames('ID') %>% 
+  as.matrix()
+
+all.equal(colnames(spp), pheno$Sampleid)
+```
+
+```{r}
+pheno_col <- pheno %>% 
+  mutate(tox_col = if_else(Toxicity == 'Y', 'red', 'green '),
+         CR_col = if_else(CR == 'Y', 'red', 'green '))
+
+bc_dist <-  vegdist(t(spp), "bray")
+pc_bc <- cmdscale(bc_dist, k = 2) 
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$tox_col, main = 'species abundance color by tox')
+plot(pc_bc[,1], pc_bc[,2], col = pheno_col$CR_col, main = 'species abundance color by CR')
+```

### 2019-11-03 scripts/01_clean_pheno.Rmd
update
clean the fucking toxicity grade var

@@
 ```{r}
+# the final pheno table which has 33 rows
+final_pheno <- phe_M_clean %>% 
+  filter(Sampleid %in% have_sg$sample_id)
+
+final_pheno %>% 
+  write_csv('../data/cleaned_pheno/final_pheno.csv')
+```
+
+```{r}
 # write out the overlap sampleid cuz that's gonna be the sample_names in use on the cluster for normalization
@@
 ```
+
+```{r}
+# retrieve the toxicity grade info from the event table
+events <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10232019 wo Censor .xlsx',
+                    sheet = 3) %>% 
+  select(MRN, Event, `Event Description`, `Event Stop`) %>% 
+  rename(descr = names(.)[3],
+         stop = names(.)[4]) %>% 
+  filter(Event %in% c('CRS','ICANS')) %>% 
+  mutate(grade = if_else(str_detect(descr, '^No'), '0', str_extract(descr, '[1|2|3]'))) %>% 
+  select(-descr)
+
+table(events$Event)
+
+# 35402481 patient ICANS record missing.
+
+# try with the CRS first
+tox_CRS <- events %>% 
+  filter(Event == 'CRS')
+
+table(tox_CRS$grade)
+
+# try to get the sample id from MRN
+# this part was manually done by me cleaning M's table
+
+# below is for CRS only
+samps <- read_csv('../data/cleaned_pheno/MRN_sampleid.txt', col_types = 'cc') %>% 
+  left_join(tox_CRS, by  = 'MRN') %>% 
+  filter(sampleid %in% pheno$Sampleid) %>% 
+  # really mad that i have to remove some records in this way
+  mutate(stop = replace(stop, MRN == '35546332' & grade == '0', '2018-01-30')) %>% 
+  arrange(desc(stop)) %>% 
+  distinct(sampleid, .keep_all = T)
+
+# 35546332 has both None and grade 1 for CRS
+# based on the date choose the 0
+# choose the latest date grade for 00332772 : 1
+# choose the latest date grade for 35546332 : 0
+```

### 2019-11-03 scripts/01_clean_pheno.Rmd
update

@@
   arrange(desc(stop)) %>% 
-  distinct(sampleid, .keep_all = T)
+  distinct(sampleid, .keep_all = T) %>% 
+  rename(CRS_grade = grade) %>% 
+  select(-stop)
 
@@
 # choose the latest date grade for 35546332 : 0
+
+samps %>% 
+  write_csv('../data/cleaned_pheno/cleaned_CRS.csv')
 ```

### 2019-11-03 scripts/15_RF_find_top_features.Rmd
clean the fucking toxicity grade var
update

@@
 And [this paper](https://www.ncbi.nlm.nih.gov/pubmed/29330424) mentions that "*Faecalibacterium prausnitzii* decrease and *Ruminococcus gnavus* increase were the essential characteristics in CHF patients' gut microbiota".
+
+### Use the toxicity grade
+
+```{r}
+crs <- read_csv('../data/cleaned_pheno/cleaned_CRS.csv')
+
+pheno <- pheno %>% 
+  full_join(crs %>% 
+              select(Sampleid  = sampleid, CRS_grade), by = 'Sampleid') %>% 
+  mutate(CRS_grade = factor(CRS_grade))
+
+all.equal(rownames(s_tb_rf), pheno$Sampleid)
+
+########################################################################################
+
+# what if just do as Tsoni says that classify 0 and 1 as N the other as Y
+pheno <- pheno %>% 
+  mutate(tox_crs = factor(if_else(as.character(CRS_grade) %in% c('0','1'), 'NO', 'YES')))
+
+# select the 18 features that have > 95% relab
+s_sum_top <- quantile(s_sum, 0.95)
+combined_fil <- cbind(s_tb_rf[,names(s_sum[s_sum > s_sum_top])], pheno %>% 
+                    select(tox_crs, ORR, Disease_Status_CAR))
+
+dim(combined_fil)
+
+CRS_RF <- randomForest(formula = tox_crs ~ ., data = combined_fil,  importance = T )
+
+CRS_RF  %>% 
+  pluck('err.rate') %>% 
+  as.data.frame() %>% 
+  mutate(ntree = seq(1, nrow(.))) %>% 
+  gather(key = 'type', value = 'ERR', 'OOB':'YES') %>% 
+  ggline(x = 'ntree', y = 'ERR', color = 'type', palette = 'lancet') +
+  labs(x = 'Number of trees', 
+       y = 'Error rate')
+
+CRS_RF    # very high error rate. not working. 
+```

### 2019-11-03 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
the tox are not separating the sample at all for spp table

@@
 library(tidyverse)
+library(ggpubr)
 library(vegan)
@@
 plot(pc_bc[,1], pc_bc[,2], col = pheno_col$CR_col, main = 'species abundance color by CR')
+
+# updated pheno
+uppheno <- read_csv('../data/cleaned_pheno/pheno_w_expanded_tox.csv', col_types = 'cfffffff')
+
+# ggplot 
+pc_bc %>% 
+  as.data.frame() %>% 
+  rownames_to_column('Sampleid') %>% 
+  full_join(uppheno, by = 'Sampleid') %>% 
+  ggdotplot(x  = 'V1', y = 'V2', color =  'Toxicity', fill = 'Toxicity', 
+            palette = 'lancet')
+
 ```

### 2019-11-03 scripts/13_PCoA_brayCurtis_KO_abundance.Rmd
update

@@
   full_join(uppheno, by = 'Sampleid') %>% 
-  ggdotplot(x  = 'V1', y = 'V2', color =  'Toxicity', fill = 'Toxicity', 
-            palette = 'lancet')
+  ggplot(aes(x  = V1, y = V2, color =  Toxicity, fill = Toxicity)) +
+  geom_point() +
+  labs(x = 'PCo 1', y = 'PCo 2', title = 'PCoA plot for response to Toxicity') +
+  theme_classic() +
+  scale_color_manual(values = c('#42B440', '#EC0000'))  +
+  scale_fill_manual(values = c('#42B440', '#EC0000')) +
+  ggsave('../figs/PCoA/response_tox.jpg', dpi = 300, width = 6, height = 5)
 
+pc_bc %>% 
+  as.data.frame() %>% 
+  rownames_to_column('Sampleid') %>% 
+  full_join(uppheno, by = 'Sampleid') %>% 
+  ggplot(aes(x  = V1, y = V2, color =  CR, fill = CR)) +
+  geom_point() +
+  labs(x = 'PCo 1', y = 'PCo 2', title = 'PCoA plot for response to CR') +
+  theme_classic() +
+  scale_color_manual(values = c('#42B440', '#EC0000'))  +
+  scale_fill_manual(values = c('#42B440', '#EC0000')) +
+  ggsave('../figs/PCoA/response_CR.jpg', dpi = 300, width = 6, height = 5)
+
+pc_bc %>% 
+  as.data.frame() %>% 
+  rownames_to_column('Sampleid') %>% 
+  full_join(uppheno, by = 'Sampleid') %>% 
+  ggplot(aes(x  = V1, y = V2, color =  ORR, fill = ORR)) +
+  geom_point() +
+  labs(x = 'PCo 1', y = 'PCo 2', title = 'PCoA plot for response to ORR') +
+  theme_classic() +
+  scale_color_manual(values = c('#42B440', '#EC0000'))  +
+  scale_fill_manual(values = c('#42B440', '#EC0000')) +
+  ggsave('../figs/PCoA/response_ORR.jpg', dpi = 300, width = 6, height = 5) 
 ```

### 2019-11-03 scripts/18_lefse_metaphlan2_CR_tox.Rmd
update

--- 2019-10-30 scripts/03_lefse_CR_tox.Rmd
@@
 classes <- c('CR', 'Toxicity')
-subclass <- c('Disease_Class')
+subclass <- c('Disease_Status_CAR')
 
@@
 ```{r}
-# input: relative abundance unstratified for pathway abundance
-relab <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
-  rename(sampleid = names(.)[1]) %>% 
-  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
-  # remove the unmapped and the unintegrated record
-  filter(!sampleid %in% c('UNMAPPED','UNINTEGRATED'))
+spp <- read_csv('../data/metaphlan2_result/merged_abundance_table_species_reordered.txt')
 
-# see if the sampleid in the relab is clean(do I need to clean additionally)
-all.equal(sort(pheno$Sampleid), sort(colnames(relab)[2:ncol(relab)]))
-
-# reorder the relab column to be the same with pheno table
-relab_or <- bind_cols(relab[,1],relab[,pheno$Sampleid])
-
-# just to be sure
-all.equal(colnames(relab_or)[2:ncol(relab_or)], pheno$Sampleid)
-
-# write out the table
-relab_or %>% 
-  write_tsv('../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv')
+all.equal(colnames(spp)[2:ncol(spp)], pheno$Sampleid)
 ```
@@
 # create the commad in df with str_glue and then output to a bash script
-pheno_combi <- tibble(pheno_file = list.files('../data/lefse_input', pattern = '.txt$', full.names = T)) %>% 
+pheno_combi <- tibble(pheno_file = list.files('../data/lefse_input', pattern = 'Disease_Status_CAR.txt$', full.names = T)) %>% 
   mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))

### 2019-11-03 scripts/cat_combi_spp.sh

@@ -1,0 +1,2 @@
+cat ../data/lefse_input/pheno__CR__Disease_Status_CAR.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.txt > ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.tsv
+cat ../data/lefse_input/pheno__Toxicity__Disease_Status_CAR.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.txt > ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.tsv

### 2019-11-03 scripts/format_input_spp.sh

@@ -1,0 +1,2 @@
+format_input.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.tsv ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.tsv ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000

### 2019-11-03 scripts/plotting_de_feature_spp.sh

@@ -1,0 +1,2 @@
+plot_res.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.res  ../figs/spp_abundace_pheno__CR__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Status_CAR.
+plot_res.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.res  ../figs/spp_abundace_pheno__Toxicity__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Status_CAR.

### 2019-11-03 scripts/run_lefse_spp.sh

@@ -1,0 +1,2 @@
+run_lefse.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.in ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.res
+run_lefse.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.in ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.res

### 2019-11-03 scripts/18_lefse_metaphlan2_CR_tox.Rmd
update

@@
 classes <- c('CR', 'Toxicity')
-subclass <- c('Disease_Status_CAR')
+subclass <- c('Disease_Class')
 
@@
 # create the commad in df with str_glue and then output to a bash script
-pheno_combi <- tibble(pheno_file = list.files('../data/lefse_input', pattern = 'Disease_Status_CAR.txt$', full.names = T)) %>% 
+pheno_combi <- tibble(pheno_file = list.files('../data/lefse_input', pattern = 'Disease_Class.txt$', full.names = T)) %>% 
   mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))
@@
 pheno_combi %>% 
-  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_{combi}tsv')) %>% 
+  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/metaphlan2_result/merged_abundance_table_species_reordered.tsv > ../data/lefse_input/spp_abundace_{combi}tsv')) %>% 
   select(cat_cmd) %>% 
-  write_csv('cat_combi_CT.sh', col_names = F)
+  write_csv('cat_combi_spp.sh', col_names = F)
 ```
@@
 pheno_combi %>% 
-  mutate(input_tsv = str_glue('../data/lefse_input/pathway_abundace_{combi}tsv'),
+  mutate(input_tsv = str_glue('../data/lefse_input/spp_abundace_{combi}tsv'),
          fmt_input = str_replace(input_tsv, 'tsv','in'),
@@
   select(fmt_cmd) %>% 
-  write_csv('format_input_CT.sh', col_names = F)
+  write_csv('format_input_spp.sh', col_names = F)
 ```
@@
 pheno_combi %>% 
-  mutate(fmt_input = str_glue('../data/lefse_input/pathway_abundace_{combi}in'),
+  mutate(fmt_input = str_glue('../data/lefse_input/spp_abundace_{combi}in'),
          lefse_res = str_replace(fmt_input, 'in$','res'),
@@
   select(run_cmd) %>% 
-  write_csv('run_lefse_CT.sh', col_names = F)
+  write_csv('run_lefse_spp.sh', col_names = F)
 ```
@@
 pheno_combi %>% 
-  mutate(lefse_res = str_glue('../data/lefse_input/pathway_abundace_{combi}res'),
-         fea_png = str_glue('../figs/pathway_abundace_{combi}png'),
+  mutate(lefse_res = str_glue('../data/lefse_input/spp_abundace_{combi}res'),
+         fea_png = str_glue('../figs/spp_abundace_{combi}png'),
          cmd = str_glue('plot_res.py {lefse_res}  {fea_png} --feature_font_size 4 --width 10 --dpi 300 --title {combi}')) %>% 
   select(cmd) %>% 
-  write_csv('plotting_de_feature_CT.sh', col_names = F)
+  write_csv('plotting_de_feature_spp.sh', col_names = F)
 ```
@@
 # this is the bash script to do the actual concatenation 
-bash cat_combi_CT.sh
-bash format_input_CT.sh
-bash run_lefse_CT.sh
-bash plotting_de_feature_CT.sh
+bash cat_combi_spp.sh
+bash format_input_spp.sh
+bash run_lefse_spp.sh
+bash plotting_de_feature_spp.sh
 ```

### 2019-11-03 scripts/cat_combi_spp.sh

@@ -1,2 +1,2 @@
-cat ../data/lefse_input/pheno__CR__Disease_Status_CAR.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.txt > ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.tsv
-cat ../data/lefse_input/pheno__Toxicity__Disease_Status_CAR.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.txt > ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.tsv
+cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.tsv
+cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.tsv

### 2019-11-03 scripts/format_input_spp.sh

@@ -1,2 +1,2 @@
-format_input.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.tsv ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.tsv ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.tsv ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.tsv ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000

### 2019-11-03 scripts/plotting_de_feature_spp.sh

@@ -1,2 +1,2 @@
-plot_res.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.res  ../figs/spp_abundace_pheno__CR__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Status_CAR.
-plot_res.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.res  ../figs/spp_abundace_pheno__Toxicity__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Status_CAR.
+plot_res.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.res  ../figs/spp_abundace_pheno__CR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Class.
+plot_res.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.res  ../figs/spp_abundace_pheno__Toxicity__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Class.

### 2019-11-03 scripts/run_lefse_spp.sh

@@ -1,2 +1,2 @@
-run_lefse.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.in ../data/lefse_input/spp_abundace_pheno__CR__Disease_Status_CAR.res
-run_lefse.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.in ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Status_CAR.res
+run_lefse.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.in ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.res
+run_lefse.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.in ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.res

### 2019-11-04 scripts/18_lefse_metaphlan2_CR_tox.Rmd
cladogram

@@
 # get different combination of the classes
+# using no subclass (11-4-2019)
+
 classes <- c('CR', 'Toxicity')
-subclass <- c('Disease_Class')
 
@@
   pheno %>% 
-    select(row[['classes']], row[['subclass']]) %>% 
+    select(row[['classes']]) %>% 
     t() %>% 
-    write.table(str_glue('../data/lefse_input/pheno__{row[["classes"]]}__{row[["subclass"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
+    write.table(str_glue('../data/lefse_input/pheno__{row[["classes"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
 })
@@
 ```{r}
-spp <- read_csv('../data/metaphlan2_result/merged_abundance_table_species_reordered.txt')
+# keeping the leading taxa info 
+spp <- read_tsv('../data/metaphlan2_result/merged_abundance_table.txt') %>% 
+  # remove the rows that have strain info
+  filter(!str_detect(ID, 't__')) %>% 
+  # keep the rows that have species info and the header
+  filter(str_detect(ID, 's__|^ID')) %>% 
+  rename_all(funs(str_replace(., '_profile$','')))
+
+spp <- bind_cols(spp[,1],spp[,pheno$Sampleid])
 
 all.equal(colnames(spp)[2:ncol(spp)], pheno$Sampleid)
+
+# write out this table 
+spp %>% 
+  write_tsv('../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv')
 ```
@@
 # create the commad in df with str_glue and then output to a bash script
-pheno_combi <- tibble(pheno_file = list.files('../data/lefse_input', pattern = 'Disease_Class.txt$', full.names = T)) %>% 
+pheno_combi <- tibble(pheno_file = list.files('../data/lefse_input', pattern = '[Toxicity|CR].txt$', full.names = T)) %>% 
   mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))
@@
 pheno_combi %>% 
-  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/metaphlan2_result/merged_abundance_table_species_reordered.tsv > ../data/lefse_input/spp_abundace_{combi}tsv')) %>% 
+  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv > ../data/lefse_input/spp_abundace_{combi}tsv')) %>% 
   select(cat_cmd) %>% 
@@
          fmt_input = str_replace(input_tsv, 'tsv','in'),
-         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -s 2 -u 3 -o 1000000')) %>% 
+         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -u 2 -o 1000000')) %>% 
   select(fmt_cmd) %>% 
@@
 
+```{r}
+# plot the cladogram   
+pheno_combi %>% 
+  mutate(lefse_res = str_glue('../data/lefse_input/spp_abundace_{combi}res'),
+         cla_png = str_glue('../figs/cladogram/spp_abundace_lefse_{combi}png'),
+         cmd = str_glue('plot_cladogram.py {lefse_res}  {cla_png}  --dpi 300 --format png --title {combi} --label_font_size 3')) %>% 
+  select(cmd) %>% 
+  write_csv('plotting_cladogram_spp.sh', col_names = F)
+```
+
 ####################################### below is command line ##########################################
@@
 bash plotting_de_feature_spp.sh
+plotting_cladogram_spp.sh
 ```

### 2019-11-04 scripts/cat_combi_spp.sh

@@ -1,2 +1,2 @@
-cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.tsv
-cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.tsv
+cat ../data/lefse_input/pheno__CR.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv > ../data/lefse_input/spp_abundace_pheno__CR.tsv
+cat ../data/lefse_input/pheno__Toxicity.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv > ../data/lefse_input/spp_abundace_pheno__Toxicity.tsv

### 2019-11-04 scripts/format_input_spp.sh

@@ -1,2 +1,2 @@
-format_input.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.tsv ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.tsv ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/spp_abundace_pheno__CR.tsv ../data/lefse_input/spp_abundace_pheno__CR.in -c 1 -u 2 -o 1000000
+format_input.py ../data/lefse_input/spp_abundace_pheno__Toxicity.tsv ../data/lefse_input/spp_abundace_pheno__Toxicity.in -c 1 -u 2 -o 1000000

### 2019-11-04 scripts/plotting_cladogram_spp.sh

@@ -1,0 +1,2 @@
+plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__CR.res  ../figs/cladogram/spp_abundace_lefse_pheno__CR.png  --dpi 300 --format png --title pheno__CR. --label_font_size 3
+plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__Toxicity.res  ../figs/cladogram/spp_abundace_lefse_pheno__Toxicity.png  --dpi 300 --format png --title pheno__Toxicity. --label_font_size 3

### 2019-11-04 scripts/plotting_de_feature_spp.sh

@@ -1,2 +1,2 @@
-plot_res.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.res  ../figs/spp_abundace_pheno__CR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Class.
-plot_res.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.res  ../figs/spp_abundace_pheno__Toxicity__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Class.
+plot_res.py ../data/lefse_input/spp_abundace_pheno__CR.res  ../figs/spp_abundace_pheno__CR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR.
+plot_res.py ../data/lefse_input/spp_abundace_pheno__Toxicity.res  ../figs/spp_abundace_pheno__Toxicity.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity.

### 2019-11-04 scripts/run_lefse_spp.sh

@@ -1,2 +1,2 @@
-run_lefse.py ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.in ../data/lefse_input/spp_abundace_pheno__CR__Disease_Class.res
-run_lefse.py ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.in ../data/lefse_input/spp_abundace_pheno__Toxicity__Disease_Class.res
+run_lefse.py ../data/lefse_input/spp_abundace_pheno__CR.in ../data/lefse_input/spp_abundace_pheno__CR.res
+run_lefse.py ../data/lefse_input/spp_abundace_pheno__Toxicity.in ../data/lefse_input/spp_abundace_pheno__Toxicity.res

### 2019-11-04 scripts/18_lefse_metaphlan2_CR_tox.Rmd
update

@@
          cla_png = str_glue('../figs/cladogram/spp_abundace_lefse_{combi}png'),
-         cmd = str_glue('plot_cladogram.py {lefse_res}  {cla_png}  --dpi 300 --format png --title {combi} --label_font_size 3')) %>% 
+         cmd = str_glue('plot_cladogram.py {lefse_res}  {cla_png}  --dpi 300 --format png --title {combi} --label_font_size 4')) %>% 
   select(cmd) %>% 
@@
 bash plotting_de_feature_spp.sh
-plotting_cladogram_spp.sh
+bash plotting_cladogram_spp.sh
 ```

### 2019-11-04 scripts/plotting_cladogram_spp.sh

@@ -1,2 +1,2 @@
-plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__CR.res  ../figs/cladogram/spp_abundace_lefse_pheno__CR.png  --dpi 300 --format png --title pheno__CR. --label_font_size 3
-plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__Toxicity.res  ../figs/cladogram/spp_abundace_lefse_pheno__Toxicity.png  --dpi 300 --format png --title pheno__Toxicity. --label_font_size 3
+plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__CR.res  ../figs/cladogram/spp_abundace_lefse_pheno__CR.png  --dpi 300 --format png --title pheno__CR. --label_font_size 4
+plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__Toxicity.res  ../figs/cladogram/spp_abundace_lefse_pheno__Toxicity.png  --dpi 300 --format png --title pheno__Toxicity. --label_font_size 4

### 2019-11-04 scripts/19_phylum_table.Rmd
update

---
title: "Look at the phylum level"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

```{r}
phy <- read_tsv('../data/metaphlan2_result/merged_abundance_table_phylum.txt') %>% 
  rename_all(funs(str_replace(., '_profile$','')))

pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
  select(-CAR_Group, -Censor_reason )

#mutate_at(vars(str_which(names(.),'Sampleid', negate = T)), factor)

phy <- bind_cols(phy[,1],phy[,pheno$Sampleid])

all.equal(colnames(phy)[2:ncol(phy)], pheno$Sampleid)

```

```{r}
phy %>% 
  gather(key = 'Sampleid', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(pheno, by = 'Sampleid') %>% 
  ggboxplot(x = 'ID', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')  +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) 
```

```{r}
phy %>% 
  gather(key = 'Sampleid', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(pheno, by = 'Sampleid') %>% 
  filter(ID == 'Bacteroidetes') %>% 
  select(ID, relab, Toxicity) %>% 
  ggboxplot(x = 'Toxicity', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')   + 
  stat_compare_means()
```

```{r}
phy %>% 
  gather(key = 'Sampleid', value = 'relab', names(.)[2]: names(.)[ncol(.)]) %>% 
  left_join(pheno, by = 'Sampleid') %>% 
  filter(ID == 'Firmicutes') %>% 
  select(ID, relab, Toxicity) %>% 
  ggboxplot(x = 'Toxicity', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')   + 
  stat_compare_means()  
```

### 2019-11-05 scripts/19_phylum_table.Rmd
the correlation

@@
   ggboxplot(x = 'Toxicity', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')   + 
-  stat_compare_means()
+  stat_compare_means()+
+  ggsave('../figs/phylum_abundance_Bacteroidetes.jpg')
 ```
@@
   ggboxplot(x = 'Toxicity', y = 'relab', color = 'Toxicity', palette  = 'lancet', add = 'jitter')   + 
-  stat_compare_means()  
+  stat_compare_means()  +
+  ggsave('../figs/phylum_abundance_Firmicutes.jpg')
 ```

### 2019-11-05 scripts/20_correlation_bw_16S_shotgun.Rmd
the correlation

---
title: "Correlation between 16S and shotgun"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

## Load the two datasets

```{r}
# shotgun
gun <- read_tsv('../data/metaphlan2_result/merged_abundance_table_phylum.txt') %>% 
  rename_all(funs(str_replace(., '_profile$',''))) %>% 
  rename(phylum = ID)

pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
  select(-CAR_Group, -Censor_reason )
```

```{r}
dada2 <- read_rds('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/data/asv_set.RDS')

# how many samples are in the dada2 data
table(dada2 %>% 
  pluck('asv_counts') %>% 
    pull(sampleid)) %>% 
  names %>% 
  length

# get the relab at phylum level
sixteen <- dada2 %>% 
  pluck('asv_counts') %>% 
  select(ASV, count_relative, sampleid) %>% 
  left_join(dada2 %>% 
    pluck('asv_annotation') %>% 
    select(ASV, phylum ), by = 'ASV') %>% 
  filter(!is.na(phylum)) %>% 
  # select the 33 that are in the shotgun
  filter(sampleid %in% pheno$Sampleid) %>% 
  group_by(sampleid, phylum) %>% 
  summarise(relab = sum(count_relative)) %>% 
  spread(key = 'sampleid', value = 'relab', fill = 0)

```

```{r}
# same phylum term
ol_phylum <- intersect(sixteen$phylum, gun$phylum)

# have an idea that i better put the two of them table in the same table with overlapped phylum terms

two <- list(shotgun = gun,
            s16S = sixteen) %>% 
  imap(~  ..1 %>% 
      filter(phylum %in% ol_phylum) %>% 
      gather(key = 'sampleid', value = 'relab', names(.)[2] : names(.)[ncol(.)]) %>% 
      mutate(Group = ..2) %>% 
      mutate_at(vars(str_which(names(.),'relab', negate = T)), factor)
  )

Two <- two %>% 
  pluck('shotgun') %>% 
  rename(shotgun_relab = relab) %>% 
  select(-Group) %>% 
    full_join(two %>% 
    pluck('s16S') %>% 
    rename(s16S_relab = relab) %>% 
    select(-Group))

```

## Compare with Spearman correlation for each sample

```{r}
Two %>% 
  ggplot(aes(x = shotgun_relab, y = s16S_relab)) +
    geom_point() +
    facet_grid(sampleid~.) +
  ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)
```

### 2019-11-05 scripts/20_correlation_phylum_bw_16S_shotgun.Rmd
update

--- 2019-11-05 scripts/20_correlation_bw_16S_shotgun.Rmd
@@
 title: "Correlation between 16S and shotgun"
-output: html_document
+author: "Anqi Dai"
+date: "`r Sys.Date()`"
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    code_folding: hide
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
 library(tidyverse)
 library(ggpubr)
+library(RColorBrewer)
+library(kableExtra)
 ```
@@
 # how many samples are in the dada2 data
-table(dada2 %>% 
+dada2_nsamps <- table(dada2 %>% 
   pluck('asv_counts') %>% 
@@
     rename(s16S_relab = relab) %>% 
-    select(-Group))
+    select(-Group)) %>%  
+  # to uniform the two data
+  mutate(s16S_relab = s16S_relab*100)
 
@@
 Two %>% 
-  ggplot(aes(x = shotgun_relab, y = s16S_relab)) +
+  ggplot(aes(x = shotgun_relab, y = s16S_relab, col = phylum)) +
     geom_point() +
-    facet_grid(sampleid~.) +
+    facet_wrap(~sampleid) +
+  scale_color_manual(values = brewer.pal(n = length(ol_phylum), name = "Dark2")) +
+  labs(x = 'Shotgun data',
+       y = '16S data',
+       title = 'Relative abundance comparison between shotgun and 16S data\nat phylum level') +
   ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)
+
+spearman_cor <- Two %>% 
+  split(.$sampleid) %>% 
+  map_dfr(~ round(cor.test(.$shotgun_relab, 
+                           .$s16S_relab, 
+                           method = 'spearman', 
+                           exact = F)$estimate, 2)) %>% 
+  gather(key = 'sampleid', value = 'rho') %>% 
+  arrange(rho)
+
+spearman_cor %>% 
+  kable(caption = 'Spearman correlation at phylum level') %>% 
+  kable_styling(full_width = T, position = 'left')
+
+spearman_cor %>% 
+  write_csv('../data/spearman_cor_phylum.csv')
 ```

### 2019-11-05 scripts/21_correlation_genus_bw_16S_shotgun.Rmd
update

--- 2019-11-05 scripts/20_correlation_phylum_bw_16S_shotgun.Rmd
@@
 
+# At the genus level
+
 ## Load the two datasets
@@
 # shotgun
-gun <- read_tsv('../data/metaphlan2_result/merged_abundance_table_phylum.txt') %>% 
+gun <- read_tsv('../data/metaphlan2_result/merged_abundance_table_genus.txt') %>% 
   rename_all(funs(str_replace(., '_profile$',''))) %>% 
-  rename(phylum = ID)
+  rename(genus = ID)
 
@@
     pluck('asv_annotation') %>% 
-    select(ASV, phylum ), by = 'ASV') %>% 
-  filter(!is.na(phylum)) %>% 
+    select(ASV, genus ), by = 'ASV') %>% 
+  filter(!is.na(genus)) %>% 
   # select the 33 that are in the shotgun
   filter(sampleid %in% pheno$Sampleid) %>% 
-  group_by(sampleid, phylum) %>% 
+  group_by(sampleid, genus) %>% 
   summarise(relab = sum(count_relative)) %>% 
@@
 ```{r}
-# same phylum term
-ol_phylum <- intersect(sixteen$phylum, gun$phylum)
+# same genus term
+ol_genus <- intersect(sixteen$genus, gun$genus)
 
@@
   imap(~  ..1 %>% 
-      filter(phylum %in% ol_phylum) %>% 
+      filter(genus %in% ol_genus) %>% 
       gather(key = 'sampleid', value = 'relab', names(.)[2] : names(.)[ncol(.)]) %>% 
@@
 Two %>% 
-  ggplot(aes(x = shotgun_relab, y = s16S_relab, col = phylum)) +
+  ggplot(aes(x = shotgun_relab, y = s16S_relab)) +
     geom_point() +
     facet_wrap(~sampleid) +
-  scale_color_manual(values = brewer.pal(n = length(ol_phylum), name = "Dark2")) +
   labs(x = 'Shotgun data',
        y = '16S data',
-       title = 'Relative abundance comparison between shotgun and 16S data\nat phylum level') +
-  ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)
+       title = 'Relative abundance comparison between shotgun and 16S data\nat genus level') +
+  ggsave('../figs/correlation/cor_between_16_and_shotgun_genus.jpg', width = 10, height = 10, dpi = 300)
 
@@
 spearman_cor %>% 
-  kable(caption = 'Spearman correlation at phylum level') %>% 
+  kable(caption = 'Spearman correlation at genus level') %>% 
   kable_styling(full_width = T, position = 'left')
 
-spearman_cor %>% 
-  write_csv('../data/spearman_cor_phylum.csv')
 ```
+
+```{r}
+py_cor <- read_csv('../data/spearman_cor_phylum.csv') %>% 
+  rename(pylum = rho)
+
+py_cor %>% 
+  full_join(spearman_cor %>% 
+              rename(genus = rho), by = 'sampleid') %>% 
+  gather(key = 'Group', value = 'Cor', names(.)[2]: names(.)[ncol(.)]) %>% 
+  ggboxplot(x = 'Group', y = 'Cor',  title = 'Comparison of spearman correlation') +
+  stat_compare_means() +
+  ggsave('../figs/pylum_genus_correlation.jpg')
+```

### 2019-11-06 scripts/01_clean_pheno.Rmd
update

@@
   write_csv('../data/cleaned_pheno/final_pheno.csv')
+
+# write out the meta data info with tox and CR in each of the table
+final_pheno %>% 
+  select(Sampleid, Toxicity) %>% 
+  write_csv('../data/cleaned_pheno/tox_metadata.csv')
 ```

### 2019-11-06 scripts/01_clean_pheno.Rmd
update

--- 2019-11-03 scripts/01_clean_pheno.Rmd
@@
   write_csv('../data/cleaned_pheno/final_pheno.csv')
+
+# write out the meta data info with tox and CR in each of the table
+final_pheno %>% 
+  select(Sampleid, Toxicity) %>% 
+  write_csv('../data/cleaned_pheno/melody_tox_metadata.csv')
 ```

### 2019-11-06 scripts/18_lefse_metaphlan2_CR_tox.Rmd
update

@@
 spp <- read_tsv('../data/metaphlan2_result/merged_abundance_table.txt') %>% 
-  # remove the rows that have strain info
-  filter(!str_detect(ID, 't__')) %>% 
-  # keep the rows that have species info and the header
-  filter(str_detect(ID, 's__|^ID')) %>% 
   rename_all(funs(str_replace(., '_profile$','')))
@@
 spp %>% 
-  write_tsv('../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv')
+  write_tsv('../data/metaphlan2_result/merged_abundance_table_WHOLE_reordered.tsv')
 ```
@@
 pheno_combi %>% 
-  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv > ../data/lefse_input/spp_abundace_{combi}tsv')) %>% 
+  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/metaphlan2_result/merged_abundance_table_WHOLE_reordered.tsv > ../data/lefse_input/spp_abundace_{combi}tsv')) %>% 
   select(cat_cmd) %>% 

### 2019-11-06 scripts/cat_combi_spp.sh

@@ -1,2 +1,2 @@
-cat ../data/lefse_input/pheno__CR.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv > ../data/lefse_input/spp_abundace_pheno__CR.tsv
-cat ../data/lefse_input/pheno__Toxicity.txt ../data/metaphlan2_result/merged_abundance_table_species_reordered_w_taxa.tsv > ../data/lefse_input/spp_abundace_pheno__Toxicity.tsv
+cat ../data/lefse_input/pheno__CR.txt ../data/metaphlan2_result/merged_abundance_table_WHOLE_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__CR.tsv
+cat ../data/lefse_input/pheno__Toxicity.txt ../data/metaphlan2_result/merged_abundance_table_WHOLE_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__Toxicity.tsv

### 2019-11-08 scripts/20_correlation_phylum_bw_16S_shotgun.Rmd
log transform the relab

@@
   # to uniform the two data
-  mutate(s16S_relab = s16S_relab*100)
+  mutate(s16S_relab = s16S_relab*100) %>% 
+  # log10 transform the data add pseudocounto f 0.000001
+  mutate(shotgun_relab = log10(shotgun_relab + 0.0001),
+         s16S_relab = log10(s16S_relab + 0.0001))
 
@@
   facet_wrap(~sampleid) +
+  geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
+  geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
+  geom_hline(yintercept = 0, col= 'forestgreen', linetype = 'dashed') +
   scale_color_manual(values = brewer.pal(n = length(ol_phylum), name = "Dark2")) +
-  labs(x = 'Shotgun data',
-       y = '16S data',
-       title = 'Relative abundance comparison between shotgun and 16S data\nat phylum level') +
+  labs(x = 'log10(Shotgun data)',
+       y = 'log10(16S data)',
+       title = 'log10 transformed relative abundance comparison between shotgun and 16S data\nat phylum level') +
   ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)
@@
   kable(caption = 'Spearman correlation at phylum level') %>% 
-  kable_styling(full_width = T, position = 'left')
+  kable_styling(full_width = F, position = 'left')
 

### 2019-11-08 scripts/21_correlation_genus_bw_16S_shotgun.Rmd
log transform the relab

@@
     facet_wrap(~sampleid) +
-  labs(x = 'Shotgun data',
-       y = '16S data',
-       title = 'Relative abundance comparison between shotgun and 16S data\nat genus level') +
+  geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
+  geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
+  geom_hline(yintercept = 0, col= 'forestgreen', linetype = 'dashed') +
+  labs(x = 'log10(Shotgun data)',
+       y = 'log10(16S data)',
+       title = 'log10 transformed relative abundance comparison between shotgun and 16S data\nat genus level') +
+  theme_light()+
   ggsave('../figs/correlation/cor_between_16_and_shotgun_genus.jpg', width = 10, height = 10, dpi = 300)
 
+# plot a single sample's plot to show the details
+Two %>% 
+  filter(sampleid == '275G') %>% 
+  ggplot(aes(x = shotgun_relab, y = s16S_relab)) +
+    geom_point() +
+    geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
+    geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
+    geom_hline(yintercept = 0, col= 'forestgreen', linetype = 'dashed') +
+    labs(x = 'log10(Shotgun data)',
+         y = 'log10(16S data)',
+         title = 'log10(relative abundance) comparison (genus): 275G') +
+    theme_light()+
+    ggsave('../figs/correlation/275G_cor_between_16_and_shotgun_genus.jpg', width = 5, height = 5, dpi = 300)
+
+# spearman_cor
 spearman_cor <- Two %>% 
@@
   kable(caption = 'Spearman correlation at genus level') %>% 
-  kable_styling(full_width = T, position = 'left')
+  kable_styling(full_width = F, position = 'left')
 

### 2019-11-09 scripts/20_correlation_phylum_bw_16S_shotgun.Rmd
update

@@
   ggplot(aes(x = shotgun_relab, y = s16S_relab, col = phylum)) +
-  geom_point() +
+  geom_point(size = 3) +
   facet_wrap(~sampleid) +
-  geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
-  geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
-  geom_hline(yintercept = 0, col= 'forestgreen', linetype = 'dashed') +
+  geom_abline(intercept = 0, slope = 1, col = 'pink', linetype = 'dashed') +
+  geom_vline(xintercept = 0, col= 'lightgreen', linetype = 'dashed') +
+  geom_hline(yintercept = 0, col= 'lightgreen', linetype = 'dashed') +
   scale_color_manual(values = brewer.pal(n = length(ol_phylum), name = "Dark2")) +
@@
        title = 'log10 transformed relative abundance comparison between shotgun and 16S data\nat phylum level') +
+  theme_light()+
   ggsave('../figs/correlation/cor_between_16_and_shotgun.jpg', width = 10, height = 10, dpi = 300)

### 2019-11-09 scripts/21_correlation_genus_bw_16S_shotgun.Rmd
update

@@
     facet_wrap(~sampleid) +
-  geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
+  geom_abline(intercept = 0, slope = 1, col = 'pink', linetype = 'dashed') +
   geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
@@
     geom_point() +
-    geom_abline(intercept = 0, slope = 1, col = 'red', linetype = 'dashed') +
+    geom_abline(intercept = 0, slope = 1, col = 'pink', linetype = 'dashed') +
     geom_vline(xintercept = 0, col= 'forestgreen', linetype = 'dashed') +
@@
 ```
+
+```{bash}
+# create the shotgun table for the class order and family level
+2
+```
+
+```{r}
+compare_each_taxa_level <- function(){
+
+}
+```

### 2019-11-12 scripts/22_shortbred.Rmd
update

---
title: "Shortbred results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(ggpubr)
```

```{r}
# create symlinks for the fastq files to work in the quant step
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')  %>% 
  select(Sampleid)  %>% 
  mutate(fastq = str_glue('{Sampleid}.fastq'),
         idx = seq(1, nrow(.)),
         idx_fn = str_glue('WGS_{idx}.fastq'),
         cmd = str_glue('ln -s /home/daia1/my_workdir/samples/melody/{fastq} /home/daia1/my_workdir/samples/melody/{idx_fn}'))

pheno %>% 
  select(cmd) %>% 
  write_tsv('../scripts/link_wgs_fastq.sh', col_names = F)
```

### 2019-11-12 scripts/link_wgs_fastq.sh

@@ -1,0 +1,33 @@
+ln -s /home/daia1/my_workdir/samples/melody/1191V.fastq /home/daia1/my_workdir/samples/melody/WGS_1.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1388A.fastq /home/daia1/my_workdir/samples/melody/WGS_2.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1409B.fastq /home/daia1/my_workdir/samples/melody/WGS_3.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1652A.fastq /home/daia1/my_workdir/samples/melody/WGS_4.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1661A.fastq /home/daia1/my_workdir/samples/melody/WGS_5.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1779A.fastq /home/daia1/my_workdir/samples/melody/WGS_6.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1742A.fastq /home/daia1/my_workdir/samples/melody/WGS_7.fastq
+ln -s /home/daia1/my_workdir/samples/melody/V0040A.fastq /home/daia1/my_workdir/samples/melody/WGS_8.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1765A.fastq /home/daia1/my_workdir/samples/melody/WGS_9.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1787A.fastq /home/daia1/my_workdir/samples/melody/WGS_10.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1715A.fastq /home/daia1/my_workdir/samples/melody/WGS_11.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1511N.fastq /home/daia1/my_workdir/samples/melody/WGS_12.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1783C.fastq /home/daia1/my_workdir/samples/melody/WGS_13.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1638A.fastq /home/daia1/my_workdir/samples/melody/WGS_14.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1508A.fastq /home/daia1/my_workdir/samples/melody/WGS_15.fastq
+ln -s /home/daia1/my_workdir/samples/melody/1450A.fastq /home/daia1/my_workdir/samples/melody/WGS_16.fastq
[…]

### 2019-11-13 scripts/20_correlation_phylum_bw_16S_shotgun.Rmd
update

@@
   write_csv('../data/spearman_cor_phylum.csv')
+
+pearson_cor <- Two %>% 
+  split(.$sampleid) %>% 
+  map_dfr(~ round(cor.test(.$shotgun_relab, 
+                           .$s16S_relab, 
+                           method = 'pearson', 
+                           exact = F)$estimate, 2)) %>% 
+  gather(key = 'sampleid', value = 'cor') %>% 
+  arrange(cor)
+
+pearson_cor %>% 
+  write_csv('../data/pearson_cor_phylum.csv')
 ```

### 2019-11-13 scripts/21_correlation_genus_bw_16S_shotgun.Rmd
update

@@
 
+pearson_cor <- Two %>% 
+  split(.$sampleid) %>% 
+  map_dfr(~ round(cor.test(.$shotgun_relab, 
+                           .$s16S_relab, 
+                           method = 'pearson', 
+                           exact = F)$estimate, 2)) %>% 
+  gather(key = 'sampleid', value = 'cor') %>% 
+  arrange(cor)
+
 ```
@@
   ggsave('../figs/pylum_genus_correlation.jpg')
-```
 
-```{bash}
-# create the shotgun table for the class order and family level
-2
 ```
@@
 ```{r}
-compare_each_taxa_level <- function(){
+py_cor <- read_csv('../data/pearson_cor_phylum.csv') %>% 
+  rename(pylum = cor)
 
-}
+py_cor %>% 
+  full_join(pearson_cor %>% 
+              rename(genus = cor), by = 'sampleid') %>% 
+  gather(key = 'Group', value = 'Cor', names(.)[2]: names(.)[ncol(.)]) %>% 
+  ggboxplot(x = 'Group', y = 'Cor',  title = 'Comparison of Pearson correlation') +
+  stat_compare_means() +
+  ggsave('../figs/pylum_genus_correlation_pearson.jpg')
 ```

### 2019-11-14 scripts/23_unifrac.Rmd
update

---
title: "Unifrac tree"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(seqinr)
source('~/MSK/work/microbiome_db/SQL/scripts/get_data_from_query_OTU.R')
```

```{r}
# load the table
annot <- get_data_from_query_OTU(0,'asv_annotation_blast_ag')  
seqs <- get_data_from_query_OTU(0,'asv_sequences_ag') 
```

```{r}
# create fasta file and write out to disk which will do the msa later
res <- apply(seqs, 1, function(Row){
  write.fasta(sequences = as.list(Row[['asv_sequence']]), 
              names = Row[['key']],
              file.out = str_glue('../output/ASV_fa/{Row[["key"]]}_ASV_seq.fa'))
})

# should be a fasta file with 80,000 sequences 
```

```{bash}
# do the msa
```

### 2019-11-15 scripts/23_unifrac.Rmd
update

@@
 library(seqinr)
+library("seqRFLP")
 source('~/MSK/work/microbiome_db/SQL/scripts/get_data_from_query_OTU.R')
@@
 ```{r}
-# create fasta file and write out to disk which will do the msa later
-res <- apply(seqs, 1, function(Row){
-  write.fasta(sequences = as.list(Row[['asv_sequence']]), 
-              names = Row[['key']],
-              file.out = str_glue('../output/ASV_fa/{Row[["key"]]}_ASV_seq.fa'))
-})
-
 # should be a fasta file with 80,000 sequences 
+res <- dataframe2fas(seqs %>% 
+                       select(key, asv_sequence) %>% 
+                       sample_n(20000), file="../output/all.fa")
 ```
@@
 # do the msa
+# Fastest possible (nucleotides)
+~/Downloads/SDE2/analysis/14_Day3/analysis/19_MSA/output/muscle -in ../output/all.fa -out ../output/all_msa_fast.fa -maxiters 1 -diags  
 ```
+
+## Use QIIME to build the tree

### 2019-11-17 scripts/06_lefse_res_ggplot.Rmd
update

--- 2019-10-24 scripts/06_lefse_res_ggplot.Rmd
@@
             coord_flip() +
-            scale_color_manual(values = c('#00468B', '#EC0000'))  +
-            scale_fill_manual(values = c('#00468B', '#EC0000')) +
+            scale_color_manual(values = c('#ED0000', '#00468B'))  +
+            scale_fill_manual(values = c('#ED0000', '#00468B')) +
             theme_minimal() +
@@
             coord_flip() +
-            scale_color_manual(values = c('#00468B', '#EC0000'))  +
-            scale_fill_manual(values = c('#00468B', '#EC0000')) +
+            scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            scale_fill_manual(values = c('#0099B4', '#AD002A')) +
             theme_minimal() +

### 2019-11-17 scripts/16_RF_pathway.Rmd
update

@@
 ```{r}
-# load the pheno
-pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
-  select(-CAR_Group, -Censor_reason ) %>% 
-  mutate_at(vars(str_which(names(.),'Sampleid', negate = T)), factor)
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')
 
-# the pathway abundance table
-pabun <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
+pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
   rename(pathway = names(.)[1]) %>% 
@@
 
-pabun <- bind_cols(pabun[,1],pabun[,pheno$Sampleid])
+pw <- bind_cols(pw[,1],pw[,pheno$Sampleid])
 
-pabun_rf <- pabun %>% 
+pw_rf <- pw %>% 
   column_to_rownames('pathway') %>% 
@@
 
-all.equal(rownames(pabun_rf), pheno$Sampleid)
+all.equal(rownames(pw_rf), pheno$Sampleid)
 
-tox_RF_PW <- randomForest(pabun_rf, factor(pheno$Toxicity), importance = T)
-
-impor_pw <- importance(tox_RF_PW, type=1) %>% 
-  as.data.frame() %>% 
-  rownames_to_column(var = 'pathway') %>% 
-  arrange(MeanDecreaseAccuracy) 
+tox_RF_PW <- randomForest(pw_rf, factor(pheno$Toxicity), importance = T)
 ```

### 2019-11-18 scripts/z01_kate_sample_sort.Rmd
update

---
title: "Kate chronic samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
source('~/MSK/work/microbiome_db/SQL/scripts/get_data_from_query_OTU.R')
```

```{r}
chronic <- get_data_from_query_OTU(0,'frozen_set_cgvhd_ag')
```

```{r}
# We care about 3 time periods:
# PreHCT: (day_relative_to_hct >= -30 & day_relative_to_hct <= -6)
# PeriHCT: (day_relative_to_hct >= 7 & day_relative_to_hct <= 21)
# PeriD100: (day_relative_to_hct >= 70 & day_relative_to_hct <= 130)
# 
# We selected only one sample per patient. In case of multiple samples for the same patient, we selected the first baseline sample for PreHCT (closest to day –30) 
# or the sample closest to day 14 or day 100 in case of PeriHCT and PeriD100.

period_list <- chronic %>% 
  select(patient_id, sampleid, case_status, day_relative_to_hct) %>% 
  mutate(period = case_when(
    day_relative_to_hct >= -30 & day_relative_to_hct <= -6 ~ 'PreHCT',
    day_relative_to_hct >= 7 & day_relative_to_hct <= 21 ~ 'PeriHCT',
    day_relative_to_hct >= 70 & day_relative_to_hct <= 130 ~ 'PeriD100'
  )) %>% 
  # keep the records that are in the above periods
  filter(!is.na(period)) %>% 
  split(.$period)

# now get the time difference between the actual time and the ideal date we emphasize on
period_diff <- bind_rows(
  period_list %>% 
    pluck('PreHCT') %>% 
    mutate(tdiff = abs(day_relative_to_hct - (-30))),
  period_list %>% 
    pluck('PeriHCT') %>% 
    mutate(tdiff = abs(day_relative_to_hct - (14))),
  period_list %>% 
    pluck('PeriD100') %>% 
    mutate(tdiff = abs(day_relative_to_hct - (100)))
)

# target samples
target <- period_diff %>% 
  group_by(patient_id, period) %>% 
  arrange(tdiff, .by_group = T) %>% 
  slice(1)  %>% 
  arrange(period, tdiff, sampleid)

# how many unique patients
period_diff %>% 
   group_by(patient_id) %>% 
   summarise(cnt = n())
```

```{r}
# how many we actually have shotgun
catalog <- read_csv('/Volumes/vandenBrinkLab/Angel_Dai/Full_human_shotgun_catalog/full_human_shotgun_catalog_updated.csv')

target_have <- target %>% 
  inner_join(catalog, by = 'sampleid') %>% 
  # only looking at the PeriD100 time
  filter(period == 'PeriD100')

current <- read_csv('~/projects/MSS_pipeline-/output/current_240_sample_preprocessing_stats.csv')

# how many we actually have 
length(intersect(current$sampleid, target_have$sampleid))

target_have %>% 
  group_by(period, case_status) %>% 
  summarise(cnt = n())

# how many need to be uploaded (35)
target_have %>% 
  filter(!sampleid %in% current$sampleid) %>% 
  select(directory) %>% 
  mutate(cmd = str_glue('rsync --progress --partial -avz {directory} daia1@lilac.mskcc.org:~/my_workdir/samples/kate')) %>% 
  ungroup() %>%
  select(cmd) %>%  
  write_csv('../data/upload_kate_70.sh', col_names = F)
```

### 2019-11-19 scripts/z01_kate_sample_sort.Rmd
update

@@
 library(tidyverse)
+library(ggvenn)
 source('~/MSK/work/microbiome_db/SQL/scripts/get_data_from_query_OTU.R')
@@
 
+**So the key is that first you filter on the samples that have shotgun data!!!! **
+**And then select the ones that are the closest !!! so that you don't waste any shotgun data!!!!!**
+
 ```{r}
 chronic <- get_data_from_query_OTU(0,'frozen_set_cgvhd_ag')
+
+# how many we actually have shotgun
+catalog <- read_csv('/Volumes/vandenBrinkLab/Angel_Dai/Full_human_shotgun_catalog/full_human_shotgun_catalog_updated.csv')
+
+chronic_shot <- chronic %>% 
+  inner_join(catalog, by = 'sampleid')
 ```
@@
 
-period_list <- chronic %>% 
-  select(patient_id, sampleid, case_status, day_relative_to_hct) %>% 
+period_list <- chronic_shot %>% 
+  select(patient_id, sampleid, case_status, day_relative_to_hct, directory) %>% 
   mutate(period = case_when(
@@
 target <- period_diff %>% 
-  group_by(patient_id, period) %>% 
+  filter(period == 'PeriD100') %>% 
+  group_by(patient_id) %>% 
   arrange(tdiff, .by_group = T) %>% 
   slice(1)  %>% 
-  arrange(period, tdiff, sampleid)
+  arrange(patient_id, tdiff, sampleid) %>% 
+  distinct(sampleid, .keep_all = T)
 
-# how many unique patients
-period_diff %>% 
-   group_by(patient_id) %>% 
-   summarise(cnt = n())
+target %>% 
+  write_csv('../data/kate/final_p100_sample.csv')
+
 ```
@@
 ```{r}
-# how many we actually have shotgun
-catalog <- read_csv('/Volumes/vandenBrinkLab/Angel_Dai/Full_human_shotgun_catalog/full_human_shotgun_catalog_updated.csv')
-
-target_have <- target %>% 
-  inner_join(catalog, by = 'sampleid') %>% 
-  # only looking at the PeriD100 time
-  filter(period == 'PeriD100')
-
 current <- read_csv('~/projects/MSS_pipeline-/output/current_240_sample_preprocessing_stats.csv')
@@
 # how many we actually have 
-length(intersect(current$sampleid, target_have$sampleid))
+length(intersect(target$sampleid, current$sampleid))
 
@@
 # how many need to be uploaded (35)
-target_have %>% 
+target %>% 
   filter(!sampleid %in% current$sampleid) %>% 
@@
   select(cmd) %>%  
-  write_csv('../data/upload_kate_70.sh', col_names = F)
+  write_csv('../data/upload_kate_43.sh', col_names = F)
 ```

### 2019-11-21 scripts/03_lefse_CR_tox.Rmd
no subclass for the pathway lefse

@@
          fmt_input = str_replace(input_tsv, 'tsv','in'),
-         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -s 2 -u 3 -o 1000000')) %>% 
+         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1  -u 2 -o 1000000')) %>% 
   select(fmt_cmd) %>% 

### 2019-11-21 scripts/06_lefse_res_ggplot.Rmd
no subclass for the pathway lefse

@@
 
-Replot the lefse plots with ggplot2 (with each class + subclass combi)
+Replot the lefse plots with ggplot2 (lefse with only class )
 
@@
 # the input : lefse res files
-fns <- list.files('../data/lefse_input/', pattern = 'pathway_abundace_pheno__[CR|Toxicity].+res$', full.names = T)
+fns <- list.files('../data/lefse_input/', pattern = 'pathway_.+res$', full.names = T)
 
-fns_name <- list.files('../data/lefse_input/', pattern = 'pathway_abundace_pheno__[CR|Toxicity].+res$')
+fns_name <- list.files('../data/lefse_input/', pattern = 'pathway_.+res$')
 
@@
   mutate(group = str_replace(group, '.res$','')) %>% 
-  mutate(pathway = str_replace_all(pathway, '^f_','')) %>% 
   select(-xx, -pval)
@@
 ```{r}
-feature %>% 
-  filter(str_detect(group, pattern = 'Disease_Class$')) %>% 
-  split(.$group) %>% 
-  map(~  ggplot(data = ., aes(x = reorder(pathway, score), y = score, fill = direction)) +
-            geom_bar(position = position_dodge2(preserve = "single"), stat = 'identity') +
+all_title_fs <- 20
+axis_text_fs <- 16
+
+CR <- feature %>% 
+  filter(group == 'CR') %>% 
+  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
+            geom_bar( stat = 'identity') +
             coord_flip() +
-            scale_color_manual(values = c('#00468B', '#EC0000'))  +
-            scale_fill_manual(values = c('#00468B', '#EC0000')) +
-            theme_minimal() +
+            scale_color_manual(values = c('#ED0000', '#00468B'))  +
+            scale_fill_manual(values = c('#ED0000', '#00468B')) +
+            theme_classic() +
             theme(axis.title.y  = element_blank(), 
+                  plot.title = element_text(size=all_title_fs),
+                  axis.title.x = element_text(size=axis_text_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position='bottom') +
+            labs(title = str_glue('Response to CR') ,
+                 y = 'Score')+ 
+            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to CR.pdf'),  width = 12, device = 'pdf' )
+
+tox <- feature %>% 
+  filter(group == 'Toxicity') %>% 
+  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
+            geom_bar(stat = 'identity') +
+            coord_flip() +
+            scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  axis.title.x = element_text(size=axis_text_fs),
+                  plot.title = element_text(size=all_title_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
                   legend.position="bottom") +
-            labs(title = str_glue('Response to class') ,
+            labs(title = str_glue('Response to Toxicity') ,
                  y = 'Score')+ 
-            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to class {.$group}.pdf'), device = 'pdf', width = 15, dpi = 300))
+            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to tox.pdf'), device = 'pdf', width = 12)
 
@@
 ```{r}
-CR <- feature %>% 
-  filter(group == 'CR__Disease_Class') %>% 
-  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
-            geom_bar(position = position_dodge2(preserve = "single"), stat = 'identity') +
-            coord_flip() +
-            scale_color_manual(values = c('#ED0000', '#00468B'))  +
-            scale_fill_manual(values = c('#ED0000', '#00468B')) +
-            theme_minimal() +
-            theme(axis.title.y  = element_blank(), 
-                  legend.position="bottom") +
-            labs(title = str_glue('Response to Complete Response') ,
-                 y = 'Score')+ 
-            ylim(-4,4) +
-            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to CR.jpg'),  width = 12, dpi = 300 )
-
-tox <- feature %>% 
-  filter(group == 'Toxicity__Disease_Class') %>% 
-  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
-            geom_bar(position = position_dodge2(preserve = "single"), stat = 'identity') +
-            coord_flip() +
-            scale_color_manual(values = c('#0099B4', '#AD002A'))  +
-            scale_fill_manual(values = c('#0099B4', '#AD002A')) +
-            theme_minimal() +
-            theme(axis.title.y  = element_blank(), 
-                  legend.position="bottom") +
-            labs(title = str_glue('Response to Toxicity') ,
-                 y = 'Score')+ 
-            ylim(-4,4) +
-            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/Response to tox.jpg'),  width = 12, dpi = 300, height = 3.5)
-
+# stack the above two plots vertically
+g <- plot_grid(CR,tox,  
+          nrow = 2, 
+          align = 'hv',
+          axis = 'b')  +
+  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/lefse_results_ggplot_recreate/pathway_lefse_combined.pdf', device = 'pdf', height = 15, width = 15)
 ```

### 2019-11-21 scripts/cat_combi_CT.sh

@@ -1,2 +1,2 @@
-cat ../data/lefse_input/pheno__CR__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv
-cat ../data/lefse_input/pheno__Toxicity__Disease_Class.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv
+cat ../data/lefse_input/pheno__CR.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR.tsv
+cat ../data/lefse_input/pheno__Toxicity.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity.tsv

### 2019-11-21 scripts/format_input_CT.sh

@@ -1,2 +1,2 @@
-format_input.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in -c 1 -s 2 -u 3 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__CR.tsv ../data/lefse_input/pathway_abundace_pheno__CR.in -c 1  -u 2 -o 1000000
+format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity.in -c 1  -u 2 -o 1000000

### 2019-11-21 scripts/plotting_de_feature_CT.sh

@@ -1,2 +1,2 @@
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res  ../figs/pathway_abundace_pheno__CR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Class.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res  ../figs/pathway_abundace_pheno__Toxicity__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Class.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR.res  ../figs/pathway_abundace_pheno__CR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR.
+plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity.res  ../figs/pathway_abundace_pheno__Toxicity.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity.

### 2019-11-21 scripts/run_lefse_CT.sh

@@ -1,2 +1,2 @@
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR.in ../data/lefse_input/pathway_abundace_pheno__CR.res
+run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity.in ../data/lefse_input/pathway_abundace_pheno__Toxicity.res

### 2019-11-26 scripts/03_lefse_CR_tox.Rmd
remove the special features when doing renorm

@@
   rename(sampleid = names(.)[1]) %>% 
-  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
-  # remove the unmapped and the unintegrated record
-  filter(!sampleid %in% c('UNMAPPED','UNINTEGRATED'))
+  rename_all(funs(str_replace(., '_Abundance$','')))
 

### 2019-11-26 scripts/24_pathcoverage.Rmd
remove the special features when doing renorm

---
title: "pathway coverage"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
pwc <- read_tsv('~/melodyP/data/humann2_res/humann2_final_out/humann2_pathcoverage_unstratified.tsv')
```

### 2019-11-26 scripts/z01_kate_sample_sort.Rmd
remove the special features when doing renorm

@@
 ```
+
+### Select the target samples that go into this comparison
 
@@
 
-target_have %>% 
-  group_by(period, case_status) %>% 
-  summarise(cnt = n())
-
-# how many need to be uploaded (35)
+# how many need to be uploaded 
 target %>% 
@@
 ```
+
+### Create the pheno table and make sure the input two tables are correctly sorted
+
+```{r}
+pheno <- target  %>% 
+  ungroup %>% 
+  mutate(fid = str_extract(directory, 'Sample.+IGO'),
+         fid = str_replace(fid, 'Sample_',''),
+         fid = str_replace(fid, '_IGO',''))  %>% 
+  select(sampleid, case_status, fid)
+
+# the cts table 
+cts <- read_tsv('../data/kate/pwabun/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
+  rename(sampleid = names(.)[1]) %>% 
+  rename_all(funs(str_replace(., '_Abundance$',''))) 
+cts <- bind_cols(cts[,1],cts[,pheno$fid])
+
+cts %>% 
+  write_tsv('../data/kate/input_lefse_cts.tsv')
+
+pheno_col <- pheno %>% 
+  select(case_status) %>% 
+  t %>% 
+  write.table('../data/kate/input_lefse_pheno.txt', sep = '\t', quote = F, row.names = T, col.names = F)
+
+```

### 2019-11-27 scripts/06_lefse_res_ggplot.Rmd
update

@@
 feature  %>% 
-  distinct(pathway) %>% 
+  distinct(pathway, .keep_all = T) %>% 
   write_csv('../data/unique_sig_pathway.csv')

### 2019-11-27 scripts/24_pathcoverage.Rmd
update

@@
 title: "pathway coverage"
-output: html_document
+author: "Anqi Dai"
+date: "`r Sys.Date()`"
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
+    code_folding: hide
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
 library(tidyverse)
+library(kableExtra)
 ```
@@
 ```{r}
-pwc <- read_tsv('~/melodyP/data/humann2_res/humann2_final_out/humann2_pathcoverage_unstratified.tsv')
+pwc <- read_tsv('~/melodyP/data/humann2_res/humann2_final_out/humann2_pathcoverage_unstratified.tsv') %>% 
+  rename(pw = names(.)[1]) %>% 
+  rename_all(funs(str_replace(., '_Coverage$',''))) %>% 
+  # create a new column just to be pwid 
+  mutate(pwid = str_extract(pw, '^.+:')) %>% 
+  mutate(pwid = str_replace(pwid, ':',''))
+
+# the pheno info
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
+  select(sampleid = Sampleid, 
+         Toxicity,
+         CR)
 ```
+
+```{r}
+# get the sig pathways in Melodys project to those two outcomes
+sig <- read_csv('../data/unique_sig_pathway.csv') %>% 
+  # extract the pw id part
+  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY.+\\d{3,4}'))  %>% 
+  mutate(pwid = str_replace_all(pwid, '_', '-'))
+
+# subset the pwc to be the sig ones only
+pwc_sig <- pwc %>% 
+  filter(pwid %in% sig$pwid) %>% 
+  select(-pwid) %>% 
+  gather(key = 'sampleid', value = 'cov', names(.)[2]:names(.)[ncol(.)]) %>% 
+  left_join(pheno, by  = 'sampleid')
+
+pwc %>% 
+  filter(pwid %in% sig$pwid) %>% 
+  select(-pwid) %>% 
+  kable(caption = 'Look at the table of pathway coverage in each sample') %>% 
+  kable_styling()
+```
+
+```{r out.width='120%'}
+pwc_sig_tox <- pwc_sig %>% 
+  group_by(pw, Toxicity) %>% 
+  summarise(cov_med = median(cov)) %>% 
+  arrange(desc(cov_med)) %>% 
+  filter(cov_med > 0)
+
+all_title_fs <- 20
+axis_text_fs <- 16
+
+pwc_sig_tox %>% 
+  ggplot(aes(x = pw, y = cov_med, fill = Toxicity)) +
+  geom_bar( stat = 'identity' , position="dodge") +
+  coord_flip() +
+  scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+  scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+  theme_classic() +
+  theme(axis.title.y  = element_blank(), 
+        axis.title.x = element_text(size=axis_text_fs),
+        plot.title = element_text(size=all_title_fs),
+        axis.text.x = element_text(size=axis_text_fs),
+        legend.position="bottom") +
+  labs(title = str_glue('Response to Toxicity') ,
+       y = 'Median coverage in each group') + 
+  ggsave('../figs/path_coverage_tox.jpg', width = 15, height = 8, dpi = 300)
+
+```
+
+```{r out.width='120%'}
+
+pwc_sig_cr <- pwc_sig %>% 
+  group_by(pw, CR) %>% 
+  summarise(cov_med = median(cov)) %>% 
+  arrange(desc(cov_med)) %>% 
+  filter(cov_med > 0)
+
+pwc_sig_cr %>% 
+  ggplot(aes(x = pw, y = cov_med, fill = CR)) +
+  geom_bar( stat = 'identity' , position="dodge") +
+  coord_flip() +
+  scale_color_manual(values = c('#ED0000', '#00468B'))  +
+  scale_fill_manual(values = c('#ED0000', '#00468B')) +
+  theme_classic() +
+  theme(axis.title.y  = element_blank(), 
+        axis.title.x = element_text(size=axis_text_fs),
+        plot.title = element_text(size=all_title_fs),
+        axis.text.x = element_text(size=axis_text_fs),
+        legend.position="bottom") +
+  labs(title = str_glue('Response to CR') ,
+       y = 'Median coverage in each group') + 
+  ggsave('../figs/path_coverage_CR.jpg', width = 15, height = 8, dpi = 300)
+```

### 2019-11-27 scripts/24_pathcoverage.Rmd
path coverage heatmap

@@
 library(kableExtra)
+library(pheatmap)
 ```
@@
   kable_styling()
+
+pwc_wide <-   pwc %>% 
+  filter(pwid %in% sig$pwid) %>% 
+  select(-pwid) %>% 
+  column_to_rownames('pw')
 ```
@@
 ```
+
+```{r}
+# try to do a heatmap to see if there is some sample that is interesting to look at 
+pheatmap(
+  pwc_wide,
+  #color=bwrPalette, 
+  #annotation_col = annot,
+  show_rownames = T,
+  show_colnames = TRUE,
+  filename = '../figs/pcov_hm.jpg',
+  height = 10,
+  width = 20,
+  cluster_rows =T,
+  cluster_cols=T
+)
+```

### 2019-11-27 scripts/111_genefam_regroup_go.Rmd
the regroup

---
title: "genefamily GO regroup"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GO.db)
```

```{bash}
# go regroup
 humann2_regroup_table             --input ../data/humann2_res/humann2_genefamilies.tsv              --groups uniref90_go             --output ../data/humann2_res/humann2_genefamilies_uniref90_go.tsv
# gropued 1+ more times:  ko 8%  eggnog 19.4% go 56.8%

# renorm to cpm
humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_uniref90_go.tsv -u cpm -s n --output ../data/humann2_res/humann2_genefamilies_uniref90_go_cpm.tsv
#Excluding special feature: UNGROUPED)

# split table
humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_uniref90_go_cpm.tsv --output ../data/humann2_res/humann2_final_out
```

```{r}
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

# reorder the cloumn names 
source('~/pipeline/scripts/lefse/01_prepare_pheno_cts_input_table_for_lefse_run.R')

cts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_go_cpm_unstratified.tsv')

cts <- prepare_lefse_input_tables(pheno, cts)

cts %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_go_cpm_unstratified_reordered.tsv')
```

```{bash}
# now use my run lefse bash script since I have already have the pheno table out 
mkdir ../data/humann2_res/lefse_GO
bash ~/projects/MSS_pipeline-/scripts/lefse/run_lefse.sh \
  ../data/lefse_input/pheno__Toxicity.txt \
  ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_go_cpm_unstratified_reordered.tsv \
  ../data/humann2_res/lefse_GO

bash ~/projects/MSS_pipeline-/scripts/lefse/run_lefse.sh \
  ../data/lefse_input/pheno__CR.txt \
  ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_go_cpm_unstratified_reordered.tsv \
  ../data/humann2_res/lefse_GO 
```

### 2019-11-27 scripts/112_genefam_eggnog.Rmd
the regroup

---
title: "genefam_eggnog"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GO.db)
```

```{bash}
# eggnog regroup
humann2_regroup_table             --input ../data/humann2_res/humann2_genefamilies.tsv              --groups uniref90_eggnog             --output ../data/humann2_res/humann2_genefamilies_uniref90_eggnog_regroup.tsv
# gropued 1+ more times:  ko 8%  eggnog 19.4% go 56.8%

# renorm to cpm
humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_uniref90_eggnog_regroup.tsv -u cpm -s n --output ../data/humann2_res/humann2_genefamilies_uniref90_eggnog_regroup_cpm.tsv
#Excluding special feature: UNGROUPED)

# split table
humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_uniref90_eggnog_regroup_cpm.tsv --output ../data/humann2_res/humann2_final_out
```

```{r}
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

# reorder the cloumn names 
source('~/pipeline/scripts/lefse/01_prepare_pheno_cts_input_table_for_lefse_run.R')

cts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_eggnog_regroup_cpm_unstratified.tsv')

cts <- prepare_lefse_input_tables(pheno, cts)

cts %>% 
  write_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_eggnog_regroup_cpm_unstratified_reordered.tsv')
```

```{bash}
# now use my run lefse bash script since I have already have the pheno table out 
mkdir ../data/humann2_res/lefse_eggnog

bash ~/projects/MSS_pipeline-/scripts/lefse/run_lefse.sh \
  ../data/lefse_input/pheno__Toxicity.txt \
  ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_eggnog_regroup_cpm_unstratified_reordered.tsv \
  ../data/humann2_res/lefse_eggnog

bash ~/projects/MSS_pipeline-/scripts/lefse/run_lefse.sh \
  ../data/lefse_input/pheno__CR.txt \
  ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_eggnog_regroup_cpm_unstratified_reordered.tsv \
  ../data/humann2_res/lefse_eggnog 
```

### Mapping the COG term to its name

### 2019-11-27 scripts/11_lefse_geneorthologs.Rmd
the regroup

@@
 # loading the table with the valid sampleids
-pheno <- read_csv('../data/current_pheno.csv') %>% 
-  filter(! Sampleid %in% c("1725E", "2222A" ,"2245A"))
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')
 
@@
 
-## Reorder the GO stratified table
+## Reorder the KO stratified table
+
+```{bash}
+# get the KO regrouped table
+humann2_regroup_table \
+            --input ../data/humann2_res/humann2_genefamilies.tsv  \
+            --groups uniref90_ko  \
+            --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
+```
 

### 2019-11-27 scripts/24_pathcoverage.Rmd
the regroup

@@
 
-pwc_sig_tox %>% 
+tox <- pwc_sig_tox %>% 
   ggplot(aes(x = pw, y = cov_med, fill = Toxicity)) +
@@
 
-pwc_sig_cr %>% 
+cr <- pwc_sig_cr %>% 
   ggplot(aes(x = pw, y = cov_med, fill = CR)) +
@@
 # try to do a heatmap to see if there is some sample that is interesting to look at 
+annot <- pheno %>% 
+  column_to_rownames('sampleid')
+
 pheatmap(
@@
   #color=bwrPalette, 
-  #annotation_col = annot,
+  annotation_col = annot,
   show_rownames = T,

### 2019-11-27 scripts/111_genefam_regroup_go.Rmd
update

@@
 mkdir ../data/humann2_res/lefse_GO
+mkdir ../data/humann2_res/lefse_GO/tox
+mkdir ../data/humann2_res/lefse_GO/CR
+
 bash ~/projects/MSS_pipeline-/scripts/lefse/run_lefse.sh \
@@
   ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_go_cpm_unstratified_reordered.tsv \
-  ../data/humann2_res/lefse_GO
+  ../data/humann2_res/lefse_GO/tox
 
@@
   ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_go_cpm_unstratified_reordered.tsv \
-  ../data/humann2_res/lefse_GO 
+  ../data/humann2_res/lefse_GO/CR
 ```

### 2019-11-27 scripts/112_genefam_eggnog.Rmd
update

@@
 # now use my run lefse bash script since I have already have the pheno table out 
-mkdir ../data/humann2_res/lefse_eggnog
+mkdir ../data/humann2_res/lefse_eggnog/tox
+mkdir ../data/humann2_res/lefse_eggnog/CR
 
@@
   ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_eggnog_regroup_cpm_unstratified_reordered.tsv \
-  ../data/humann2_res/lefse_eggnog
+  ../data/humann2_res/lefse_eggnog/tox
 
@@
   ../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_eggnog_regroup_cpm_unstratified_reordered.tsv \
-  ../data/humann2_res/lefse_eggnog 
+  ../data/humann2_res/lefse_eggnog/CR
 ```

### 2019-11-27 scripts/11_lefse_geneorthologs.Rmd
update

--- 2019-10-18 scripts/11_lefse_geneorthologs.Rmd
@@
 library(tidyverse)
+library(GO.db)
 ```
 
-## Make pheno table(with only classes)
+```{bash}
+# ko regroup
+humann2_regroup_table             --input ../data/humann2_res/humann2_genefamilies.tsv              --groups uniref90_ko              --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
+# gropued 1+ more times:  ko 8%  eggnog 19.4% go 56.8%
 
-```{r}
-# loading the table with the valid sampleids
-pheno <- read_csv('../data/current_pheno.csv') %>% 
-  filter(! Sampleid %in% c("1725E", "2222A" ,"2245A"))
+# renorm to cpm
+humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u cpm -s n --output ../data/humann2_res/humann2_genefamilies_KO_regroup_cpm.tsv
+#Excluding special feature: UNGROUPED)
 
+# split table
+humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_cpm.tsv --output ../data/humann2_res/humann2_final_out
 ```
@@
 ```{r}
-# get different combination of the classes
-classes <- c('CR', 'Toxicity')
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')
 
-#subclass <- c('Disease_Status_CAR','Disease_Class','CAR_Group')
-combi <- expand.grid(classes = classes)
+# reorder the cloumn names 
+source('~/pipeline/scripts/lefse/01_prepare_pheno_cts_input_table_for_lefse_run.R')
 
-# use each combi row in the above df to create the pheno table for the lefse input
-apply(combi, 1, function(row) {
-  pheno %>% 
-    mutate(CR = str_glue('CR_{CR}'),
-           Toxicity = str_glue('Toxicity_{Toxicity}')) %>% 
-    select(row[['classes']]) %>% 
-    t() %>% 
-    write.table(str_glue('../data/lefse_genefam/pheno__{row[["classes"]]}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
-})
+cts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_KO_regroup_cpm_unstratified.tsv')
+
+cts <- prepare_lefse_input_tables(pheno, cts)
+
+cts %>% 
+  write_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_KO_regroup_cpm_unstratified_reordered.tsv')
 ```
 
-## Reorder the GO stratified table
+```{bash}
+# now use my run lefse bash script since I have already have the pheno table out 
+mkdir ../data/humann2_res/lefse_ko
+mkdir ../data/humann2_res/lefse_ko/tox
+mkdir ../data/humann2_res/lefse_ko/CR
 
-```{r}
-stra_go <- read_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified.tsv') %>% 
-  rename(sampleid = names(.)[1]) %>% 
-  rename_all(funs(str_replace(., '_Abundance-RPKs$','')))
+bash ~/projects/MSS_pipeline-/scripts/lefse/run_lefse.sh \
+  ../data/lefse_input/pheno__Toxicity.txt \
+  ../data/humann2_res/humann2_final_out/humann2_genefamilies_KO_regroup_cpm_unstratified_reordered.tsv \
+  ../data/humann2_res/lefse_ko/tox
 
-stra_go <- bind_cols(stra_go[,1],stra_go[,pheno$Sampleid])
-
-all.equal(pheno$Sampleid, colnames(stra_go)[2:ncol(stra_go)])
-
-# write out
-
-stra_go %>% 
-  write_tsv('../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv')
+bash ~/projects/MSS_pipeline-/scripts/lefse/run_lefse.sh \
+  ../data/lefse_input/pheno__CR.txt \
+  ../data/humann2_res/humann2_final_out/humann2_genefamilies_KO_regroup_cpm_unstratified_reordered.tsv \
+  ../data/humann2_res/lefse_ko/CR
 ```
-
-## str_glue
-
-```{r}
-# create the commad in df with str_glue and then output to a bash script
-pheno_combi <- data_frame(pheno_file = list.files('../data/lefse_genefam', pattern = '.txt$', full.names = T)) %>% 
-  mutate(combi = str_extract(pheno_file, 'pheno.+\\.'))
-
-pheno_combi %>% 
-  mutate(cat_cmd = str_glue('cat {pheno_file} ../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv > ../data/lefse_genefam/genefam_abundace_{combi}tsv')) %>%
-  select(cat_cmd) %>% 
-  write_csv('cat_combi_genef.sh', col_names = F)
-```
-
-```{r}
-# create the cmd to do the format input stuff
-pheno_combi %>% 
-  mutate(input_tsv = str_glue('../data/lefse_genefam/genefam_abundace_{combi}tsv'),
-         fmt_input = str_replace(input_tsv, 'tsv','in'),
-         fmt_cmd = str_glue('format_input.py {input_tsv} {fmt_input} -c 1 -u 2 -o 1000000')) %>% 
-  select(fmt_cmd) %>% 
-  write_csv('format_input_genef.sh', col_names = F)
-```
-
-```{r}
-# the actual step to run lefse
-pheno_combi %>% 
-  mutate(fmt_input = str_glue('../data/lefse_genefam/genefam_abundace_{combi}in'),
-         lefse_res = str_replace(fmt_input, 'in$','res'),
-         run_cmd = str_glue('run_lefse.py {fmt_input} {lefse_res}')) %>% 
-  select(run_cmd) %>% 
-  write_csv('run_lefse_genef.sh', col_names = F)
-```
-
-```{r}
-# plotting cmd 
-pheno_combi %>% 
-  mutate(lefse_res = str_glue('../data/lefse_genefam/genefam_abundace_{combi}res'),
-         fea_png = str_glue('../figs/genefam_abundace_{combi}png'),
-         cmd = str_glue('plot_res.py {lefse_res}  {fea_png} --feature_font_size 4 --width 10 --dpi 300 --title {combi}')) %>% 
-  select(cmd) %>% 
-  write_csv('plotting_de_feature_genef.sh', col_names = F)
-```
-
-## bash
-
-```{bash}
-bash cat_combi_genef.sh
-bash format_input_genef.sh
-bash run_lefse_genef.sh
-bash plotting_de_feature_genef.sh
-```

### 2019-11-27 scripts/12_lefse_genefam_term_ggplot.Rmd
PCoA with KO abundance
update

@@
 ---
-title: ""
+title: "KO & EGGNOG"
 output: html_document
@@
 # the input : lefse res files
-fns <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$', full.names = T)
+fns <- list.files('../data/humann2_res/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$', full.names = T)
 fns_name <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$')
@@
 ```
-
-```{r}
-
-```

### 2019-11-27 scripts/05_pathway_feature_cpm.Rmd
update

@@
 ```{r}
-fn <- read_tsv('../data/humann2_res/humann2_final_out_cpm/humann2_pathabundance_cpm_unstratified.tsv') %>%
-  rename(pathway = names(.)[1]) %>% 
-  # remove the things up till the colon
-  mutate(pathway = str_replace_all(pathway, '^.+\\: ','')) %>% 
+pwa <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>%
+  dplyr::rename(pw = names(.)[1]) %>% 
   rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
+  # create a new column just to be pwid 
+  mutate(pwid = str_extract(pw, '^.+:')) %>% 
+  mutate(pwid = str_replace(pwid, ':',''))
+
   # removing things in the parenthesis and remove the spaces in the pathway
-  mutate(pathway = str_replace_all(pathway, ' \\(.+\\)',''),
-         pathway = str_replace_all(pathway, ' ',''))
+  # mutate(pathway = str_replace_all(pathway, ' \\(.+\\)',''),
+  #        pathway = str_replace_all(pathway, ' ',''))
+
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
+  dplyr::select(sampleid = Sampleid, 
+         Toxicity,
+         CR)
 ```
@@
 ```{r}
-# loading the enriched pathways from 6 combi and filter on the abundance matrix
-feat_pw <- read_csv('../data/unique_enriched_pathways.csv')
+sig <- read_csv('../data/unique_sig_pathway.csv') %>% 
+  # extract the pw id part
+  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY.+\\d{3,4}'))  %>% 
+  mutate(pwid = str_replace_all(pwid, '_', '-'))
 
-cpms <- fn %>% 
-  filter(pathway %in% feat_pw$pathway)
-
-# only half is exactly the same
-
-same_pw <- cpms %>% 
-  pull(pathway)
+pwa_sig <- pwa %>% 
+  filter(pwid %in% sig$pwid) %>% 
+  dplyr::select(-pwid) %>% 
+  gather(key = 'sampleid', value = 'abun', names(.)[2]:names(.)[ncol(.)]) %>% 
+  left_join(pheno, by  = 'sampleid')
 ```
@@
 ```{r}
-# plot the counts 
-# get the pheno info of those samples
-pheno <- read_csv('../data/current_pheno.csv')
+# bar plot to show the cpm in each sample for each sig pathway
+pw_list <- pwa_sig %>% 
+  distinct(pw) %>% 
+  pull(pw)
 
-res <- cpms %>% 
-  gather(key = 'Sampleid', value = 'cpm', '1191V' : 'V0040A') %>% 
-  left_join(pheno, by = 'Sampleid')
-
-res
+# tox
+for (p in pw_list) {
+  pwa_sig %>% 
+    arrange(Toxicity, desc(abun)) %>% 
+        filter(pw == p) %>% 
+        ggbarplot(x = 'sampleid', y = 'abun', fill = 'Toxicity', color = 'Toxicity') +
+          scale_color_manual(values = c('#00468B', '#EC0000'))  +
+          scale_fill_manual(values = c('#00468B', '#EC0000')) +
+          theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
+          labs(title = str_glue('Pathway: {p}')) +
+          ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/pathway_abundance_cpm/Toxicity/pathway_abundance_cpm_{p}.jpg'), width = 12, height = 10, dpi = 300)
+}
 ```
@@
 ```{r}
-
-same_pw %>% 
-  map(~ res %>% 
-          arrange(CR, desc(cpm)) %>% 
-          filter(pathway == .) %>% 
-          ggbarplot(x = 'Sampleid', y = 'cpm', fill = 'CR', color = 'CR') +
+# CR
+for (p in pw_list) {
+  pwa_sig %>% 
+    arrange(CR, desc(abun)) %>% 
+        filter(pw == p) %>% 
+        ggbarplot(x = 'sampleid', y = 'abun', fill = 'CR', color = 'CR') +
           scale_color_manual(values = c('#00468B', '#EC0000'))  +
@@
           theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
-          labs(title = str_glue('Pathway: {.}')) +
-          ggsave(str_glue('../figs/pathway_abundance_cpm_{.}.jpg'), width = 4, height = 3, dpi = 300))
-
-for (i in same_pw) {
-  res %>% 
-          arrange(Toxicity, desc(cpm)) %>% 
-          filter(pathway == i) %>% 
-          ggbarplot(x = 'Sampleid', y = 'cpm', fill = 'Toxicity', color = 'Toxicity') +
-          scale_color_manual(values = c('#00468B', '#EC0000'))  +
-          scale_fill_manual(values = c('#00468B', '#EC0000')) +
-          theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
-          labs(title = str_glue('Pathway: {i}')) +
-          ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/pathway_abundance_cpm/Toxicity/pathway_abundance_cpm_{i}.jpg'), width = 8, height = 6, dpi = 300)
+          labs(title = str_glue('Pathway: {p}')) +
+          ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/pathway_abundance_cpm/CR/pathway_abundance_cpm_{p}.jpg'), width = 12, height = 10, dpi = 300)
 }
-
 ```
-
-```{r}
-res %>% 
-    arrange(CR, desc(cpm)) %>% 
-    filter(pathway == 'pantothenateandcoenzymeAbiosynthesisI') %>% 
-    ggbarplot(x = 'Sampleid', y = 'cpm', fill = 'CR', color = 'CR') +
-    scale_color_manual(values = c('#00468B', '#EC0000'))  +
-    scale_fill_manual(values = c('#00468B', '#EC0000')) +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) +
-    labs(title = str_glue('Pathway: pantothenateandcoenzymeAbiosynthesisI')) +
-    ggsave(str_glue('../figs/tox/pathway_abundance_cpm_pantothenateandcoenzymeAbiosynthesisI.jpg'), width = 4, height = 3, dpi = 300)
-```

### 2019-11-27 scripts/12_lefse_genefam_term_ggplot.Rmd
update

@@
 # the input : lefse res files
-fns <- list.files('../data/humann2_res/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$', full.names = T)
-fns_name <- list.files('../data/lefse_genefam/', pattern = 'genefam_abundace_pheno__[CR|Toxicity].+res$')
+fns <- c(list.files('../data/humann2_res/lefse_ko/', pattern = '*/*.res$', full.names = T),
+         list.files('../data/humann2_res/lefse_ko/', pattern = '*/*.res$', full.names = T))
 

### 2019-11-27 scripts/25_followups.Rmd
update

---
title: "Two patient follow up"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
target <- read_csv('/Volumes/vandenBrinkLab/Angel_Dai/Full_human_shotgun_catalog/full_human_shotgun_catalog_updated.csv') %>% 
  # the patient 1487 and 2050
  filter(str_detect(sampleid, '1487|2050')) %>% 
  filter(!str_detect(directory, 'A_IGO'))

target %>% 
  dplyr::select(directory) %>% 
  mutate(cmd = str_glue('rsync --progress --partial -avz {directory} daia1@lilac.mskcc.org:~/my_workdir/samples/melody')) %>% 
  dplyr::select(cmd) %>%   
  write_csv('../data/upload_M_5.sh', col_names = F)
```

### 2019-12-03 scripts/26_unpack_pw.Rmd
update

---
title: "Find gene fam contribution to pw"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{bash}
# renorm the gene family table
humann2_renorm_table  --input humann2_genefamilies.tsv -u cpm -s n --output humann2_genefamilies_cpm.tsv

# unpack the pw using humann2 script
humann2_unpack_pathways --input-genes humann2_genefamilies_cpm.tsv --input-pathways humann2_pathabundance_cpm.tsv \
  --remove-taxonomy --output humann2_pwa_unpacked.tsv 
```

```{r}
# look at the results
unpacked <- read_tsv('../data/humann2_res/humann2_pwa_unpacked.tsv', skip = 1, col_names = c('pw','cpm'))

# the pantothenate (PWY-4242)
vb5 <- unpacked %>% 
  filter(str_detect(pw, '4242'))
```

### 2019-12-03 scripts/27_maASlin2.Rmd
try new method

---
title: "maASlin2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Maaslin2)
```

```{r}
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv', col_types = 'ccffffff') %>% 
  select(-Censor_reason, -CAR_Group) %>% 
  column_to_rownames('Sampleid')

pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  rename(pw = names(.)[1])  %>% 
  column_to_rownames('pw') %>% 
  as.matrix() %>% 
  t() %>% 
  as.data.frame()

fit_data <- Maaslin2(
    pw, pheno, '../data/maASlin2/output',normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
    standardize = FALSE)
```

### 2019-12-03 scripts/26_unpack_pw.Rmd
unpacked

@@
 # the pantothenate (PWY-4242)
+
+# get the top N results
+N <- 20
+
 vb5 <- unpacked %>% 
-  filter(str_detect(pw, '4242'))
+  filter(str_detect(pw, '4242')) %>% 
+  filter(str_detect(pw, '\\|')) %>% 
+  arrange(desc(cpm)) %>% 
+  separate(pw, into = c('name','uniref90_ID'), '\\|') %>% 
+  top_n(N)
+
+vb5 %>% 
+  ggbarplot(x = 'uniref90_ID', y = 'cpm', color = 'white', fill = 'forestgreen', title = 'PWY-4242')  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggsave('../figs/unpack/PWY-4242.jpg', width = 8, height = 7, dpi = 300)
+
+# COA 
+unpacked %>% 
+  filter(str_detect(pw, 'COA-PWY\\|')) %>% 
+  filter(str_detect(pw, '\\|')) %>% 
+  arrange(desc(cpm)) %>% 
+  separate(pw, into = c('name','uniref90_ID'), '\\|') %>% 
+  top_n(N) %>% 
+  ggbarplot(x = 'uniref90_ID', y = 'cpm', color = 'white', fill = 'forestgreen', title = 'COA-PWY')  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggsave('../figs/unpack/COA-PWY.jpg', width = 8, height = 7, dpi = 300)
+
+# PANTOSYN-PWY
+unpacked %>% 
+  filter(str_detect(pw, 'PANTOSYN-PWY\\|')) %>% 
+  filter(str_detect(pw, '\\|')) %>% 
+  arrange(desc(cpm)) %>% 
+  separate(pw, into = c('name','uniref90_ID'), '\\|') %>% 
+  top_n(N) %>% 
+  ggbarplot(x = 'uniref90_ID', y = 'cpm', color = 'white', fill = 'forestgreen', title = 'PANTOSYN-PWY')  +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  ggsave('../figs/unpack/PANTOSYN-PWY.jpg', width = 8, height = 7, dpi = 300)
 ```

### 2019-12-03 scripts/27_maASlin2.Rmd
unpacked

@@
 fit_data <- Maaslin2(
-    pw, pheno, '../data/maASlin2/output',normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
+    pw, pheno, '../data/maASlin2/output', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
     standardize = FALSE)

### 2019-12-05 scripts/10_genefamilies_KO_regroup.Rmd
update

@@
 
+### PCoA
+
 ```{r}
-load('../data/kegg_lookup_tables.RData')
+cts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_KO_regroup_cpm_unstratified_reordered.tsv')
+
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
+  dplyr::rename(sampleid = Sampleid)
+
+plot_BC_PCoA(cts, pheno, 'Toxicity')
 ```

### 2019-12-05 scripts/111_genefam_regroup_go.Rmd
update

@@
 library(GO.db)
+source('~/pipeline/scripts/Computation/BC_distance_PCoA.R')
 ```
@@
 ```
+
+### The PCoA of the GO regroup
+
+```{r}
+cts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_go_cpm_unstratified_reordered.tsv')
+
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
+  dplyr::rename(sampleid = Sampleid)
+
+plot_BC_PCoA(cts, pheno, 'CR')
+```

### 2019-12-05 scripts/112_genefam_eggnog.Rmd
update

@@
 humann2_regroup_table             --input ../data/humann2_res/humann2_genefamilies.tsv              --groups uniref90_eggnog             --output ../data/humann2_res/humann2_genefamilies_uniref90_eggnog_regroup.tsv
-# gropued 1+ more times:  ko 8%  eggnog 19.4% go 56.8%
+# gropued 1+ more times:  ko 8%  eggnog 19.4% go 56.8%  level4ec 9.8%
 
@@
 
-### Mapping the COG term to its name
+### PCoA
+
+```{r}
+cts <- read_tsv('../data/humann2_res/humann2_final_out/humann2_genefamilies_uniref90_eggnog_regroup_cpm_unstratified_reordered.tsv')
+
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') %>% 
+  dplyr::rename(sampleid = Sampleid)
+
+plot_BC_PCoA(cts, pheno, 'CR')
+```

### 2019-12-05 scripts/27_maASlin2.Rmd
update

@@
 library(Maaslin2)
+```
+
+```{r}
+# format the whole table column and keep the samples in the final cohort 33
+have_sg <- read_csv('../data/baselinesforAngel10032019.csv') %>% 
+  filter(shotgun_YN == 'yes') %>% 
+  # removig 1943A cuz its low quality
+  filter(sample_id != '1943A') %>% 
+  ### FUCK THIS TABLE OF ../data/baselinesforAngel10032019.csv IS WRONG IN 161, it didn’t pass QC because not enough DNA material was present. SO WE DON'T HAVE THE SHOTGUN DATA.
+  filter(!sample_id %in% c('161'))
+
+phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
+                    sheet = 5) %>% 
+  rename_all(funs(str_replace_all(., ' \\(.+$','') %>% 
+    str_replace_all(., ' ','_') %>% 
+    str_replace(., ':|-','_'))) %>% 
+  filter(is.na(Censor_reason)) %>% 
+  filter(Baseline_Sample_ID %in% have_sg$sample_id)
+
+```
+
+```{r}
+pheno <- phe_M %>% 
+  dplyr::select(sampleid = Baseline_Sample_ID,
+                CR, 
+                ORR, 
+                Toxicity, 
+                Cytokine_Release_Syndrome,
+                ICANS,
+                Antibiotics_4_weeks_prior_to_CAR_T_cells,
+                Antibiotics_4_weeks_post_CAR_T_cells,
+                Disease_Class,
+                Key__Patient_Disease_Status_at_CAR
+                ) %>% 
+  # change all the pheno vars to factors
+  mutate_if(is.character, as.factor) %>% 
+  column_to_rownames('sampleid') 
 ```

### 2019-12-05 scripts/27_maASlin2.Rmd
update on MaAsLin2

--- 2019-12-03 scripts/27_maASlin2.Rmd
@@
 library(Maaslin2) 
+library(eeptools)
+library(tidyverse)
 ```
@@
 ```{r}
-pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv', col_types = 'ccffffff') %>% 
-  select(-Censor_reason, -CAR_Group) %>% 
-  column_to_rownames('Sampleid')
+# get the MRN of my target samples 
+link <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
+                    sheet = 4) %>% 
+  dplyr::select(pt_id, MRN)
 
+# format the whole table column and keep the samples in the final cohort 33
+have_sg <- read_csv('../data/baselinesforAngel10032019.csv') %>% 
+  filter(shotgun_YN == 'yes') %>% 
+  # removig 1943A cuz its low quality
+  filter(sample_id != '1943A') %>% 
+  ### FUCK THIS TABLE OF ../data/baselinesforAngel10032019.csv IS WRONG IN 161, it didn’t pass QC because not enough DNA material was present. SO WE DON'T HAVE THE SHOTGUN DATA.
+  filter(!sample_id %in% c('161'))
+
+# now I have MRN in the below tble
+phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
+                    sheet = 5) %>% 
+  rename_all(funs(str_replace_all(., ' \\(.+$','') %>% 
+  str_replace_all(., ' ','_') %>% 
+  str_replace(., ':|-','_'))) %>% 
+  filter(is.na(Censor_reason)) %>% 
+  filter(Baseline_Sample_ID %in% have_sg$sample_id) %>% 
+  left_join(link, by = 'pt_id')
+
+# sub selecting the vars I'm interested in 
+pheno <- phe_M %>% 
+  dplyr::select(MRN,
+                sampleid = Baseline_Sample_ID,
+                CR, 
+                ORR, 
+                Toxicity, 
+                Cytokine_Release_Syndrome,
+                ICANS,
+                Antibiotics_4_weeks_prior_to_CAR_T_cells,
+                #Antibiotics_4_weeks_post_CAR_T_cells,
+                Disease_Class,
+                Key__Patient_Disease_Status_at_CAR
+                )
+```
+
+```{r}
+# get the clinical data info  from ying pulled data
+load('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/YingRCode/ctc_melody.RData')
+```
+
+```{r}
+# get the specifically the three ILs, ferriton and crp
+
+IL_table <- ctc.interleukin %>% 
+  # remove that one record that has serum in subtest_name
+  filter(!str_detect(subtest_name, 'Serum')) %>% 
+  # format the 1,000 such value to remove the comma
+  mutate(result_text = str_replace(result_text, ',',''),
+         result_text = if_else(str_detect(result_text, '^<'), '0', result_text)) %>% 
+  mutate(result_text = as.numeric(result_text)) %>% 
+  group_by(MRN, subtest_name) %>% 
+  summarise(IL =  max(result_text)) %>% 
+  mutate(subtest_name = str_replace(subtest_name, 'Interleukin','IL'),
+         subtest_name= str_replace_all(subtest_name, ' ','_')) %>% 
+  spread(key = 'subtest_name' , value = 'IL')
+
+# the three ILs, select the highest value if there are multiple 
+pheno <- pheno %>% 
+  # get patient info such as age and sex (using MRN)
+  left_join(ctc.patients %>% 
+              dplyr::select(MRN, sex, dob), by = 'MRN') %>% 
+  # get the hightest ferritin value 
+  left_join(ctc.ferritin %>% 
+              dplyr::select(MRN, result_text) %>% 
+              group_by(MRN) %>% 
+              summarise(ferritin = max(result_text)), by = 'MRN') %>% 
+  left_join(ctc.crp %>% 
+              dplyr::select(MRN, result_text) %>% 
+              mutate(result_text = if_else(str_detect(result_text, '^<'), '0', result_text),
+                     result_text = as.numeric(result_text)) %>% 
+              group_by(MRN) %>% 
+              summarise(crp  = max(result_text, na.rm = T)), by = 'MRN') %>% 
+  left_join(IL_table, by  ='MRN')
+
+pheno <- pheno %>% 
+  mutate(age = round(age_calc(pheno$dob, enddate = Sys.Date(), units = "years", precise = TRUE),0)) %>% 
+  dplyr::select(-dob)
+
+```
+
+```{r}
+# now change the columns to different type: dbl or factor
+pheno_final  <- pheno %>%  
+  # change the ferritin to dbl
+  mutate(ferritin = as.numeric(ferritin),
+         Key__Patient_Disease_Status_at_CAR = as.character(Key__Patient_Disease_Status_at_CAR)) %>% 
+  # change all the pheno vars to factors
+  mutate_if(is.character, as.factor) %>% 
+  mutate(MRN = as.character(MRN),
+         sampleid = as.character(sampleid)) %>% 
+  column_to_rownames('sampleid')
+
+colnames(pheno_final)
+```
+
+```{r}
 pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
-  rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
-  rename(pw = names(.)[1])  %>% 
+  dplyr::rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
+  dplyr::rename(pw = names(.)[1])  %>% 
   column_to_rownames('pw') %>% 
@@
 fit_data <- Maaslin2(
-    pw, pheno, '../data/maASlin2/output',normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
+    pw, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
+    fixed_effects = c(
+
+                      'Toxicity', 
+                      'ICANS',
+                      'Cytokine_Release_Syndrome',
+                      'Disease_Class',
+                      'Key__Patient_Disease_Status_at_CAR',
+                      'ferritin',
+                      'crp',
+                      'IL_1_beta',
+                      'IL_10',
+                      'IL_6'),
+    random_effects = c('MRN',
+                       'CR',
+                       'ORR', 
+                       'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
+                       'sex',
+                       'age'
+                       ),
+    standardize = FALSE)
+
+fit_data <- Maaslin2(
+    pw, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
+    fixed_effects = c(  'sex',
+                       'age'),
+
     standardize = FALSE)

### 2019-12-05 scripts/28_MaAsLin2_spp.Rmd
update on MaAsLin2

---
title: "MaAsLin2 species"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Maaslin2)
```

```{r}
# format the whole table column and keep the samples in the final cohort 33
have_sg <- read_csv('../data/baselinesforAngel10032019.csv') %>% 
  filter(shotgun_YN == 'yes') %>% 
  # removig 1943A cuz its low quality
  filter(sample_id != '1943A') %>% 
  ### FUCK THIS TABLE OF ../data/baselinesforAngel10032019.csv IS WRONG IN 161, it didn’t pass QC because not enough DNA material was present. SO WE DON'T HAVE THE SHOTGUN DATA.
  filter(!sample_id %in% c('161'))

phe_M <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
                    sheet = 5) %>% 
  rename_all(funs(str_replace_all(., ' \\(.+$','') %>% 
    str_replace_all(., ' ','_') %>% 
    str_replace(., ':|-','_'))) %>% 
  filter(is.na(Censor_reason)) %>% 
  filter(Baseline_Sample_ID %in% have_sg$sample_id)

```

```{r}
pheno <- phe_M %>% 
  dplyr::select(sampleid = Baseline_Sample_ID,
                CR, 
                #ORR, 
                Toxicity, 
                Cytokine_Release_Syndrome,   
                ICANS,
                Antibiotics_4_weeks_prior_to_CAR_T_cells,
                #Antibiotics_4_weeks_post_CAR_T_cells,
                Disease_Class,
                Key__Patient_Disease_Status_at_CAR
                ) %>% 
  # change all the pheno vars to factors
  mutate_if(is.character, as.factor) %>% 
  column_to_rownames('sampleid') 
```

```{r}
spp <- read_tsv('../data/metaphlan2_result/merged_abundance_table_genus.txt') %>% 
  dplyr::rename_all(funs(str_replace(., '_profile$',''))) %>% 
  dplyr::rename(spp = names(.)[1])  %>% 
  column_to_rownames('spp') %>% 
  as.matrix() %>% 
  t() %>% 
  as.data.frame()

fit_data <- Maaslin2(
    spp, pheno, '../data/maASlin2/genus', normalization  = 'NONE', transform = "AST",  cores = 4, plot_heatmap = F,
    standardize = FALSE)
```

### 2019-12-06 scripts/06_lefse_res_ggplot.Rmd
update

@@
   mutate(group = str_replace(group, '.res$','')) %>% 
-  select(-xx, -pval)
+  dplyr::select(-xx, -pval)
 
@@
             coord_flip() +
-            scale_color_manual(values = c('#ED0000', '#00468B'))  +
-            scale_fill_manual(values = c('#ED0000', '#00468B')) +
+            scale_color_manual(values = c('#925E9F', '#42B540'))  +
+            scale_fill_manual(values = c('#925E9F', '#42B540')) +
             theme_classic() +

### 2019-12-06 scripts/27_maASlin2.Rmd
update

@@
   dplyr::rename(pw = names(.)[1])  %>% 
+  # remove the things after the : so that its easier to see on the plot(the pathway name)
+  mutate(pw = str_replace(pw, ':.+$','')) %>% 
   column_to_rownames('pw') %>% 
@@
 fit_data <- Maaslin2(
-    pw, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
-    fixed_effects = c(
-
-                      'Toxicity', 
+    pw, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = T,
+    fixed_effects = c('Toxicity', 
+                      'Cytokine_Release_Syndrome',
                       'ICANS',
-                      'Cytokine_Release_Syndrome',
+                      'CR',
+                      'ORR', 
                       'Disease_Class',
                       'Key__Patient_Disease_Status_at_CAR',
+                      'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
                       'ferritin',
@@
                       'IL_6'),
-    random_effects = c('MRN',
-                       'CR',
-                       'ORR', 
-                       'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
-                       'sex',
-                       'age'
-                       ),
+    # random_effects = c('MRN',
+    #                    'sex',
+    #                    'age'),
     standardize = FALSE)
 
-fit_data <- Maaslin2(
-    pw, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = F,
-    fixed_effects = c(  'sex',
-                       'age'),
+```
 
-    standardize = FALSE)
+```{r}
+trans <- log(pw[,'PWY-7456'] + 1)
 ```

### 2019-12-09 scripts/27_maASlin2.Rmd
updte

@@
          sampleid = as.character(sampleid)) %>% 
+  mutate(IL_6 = log(IL_6),
+         IL_10 = log(IL_10)) %>% 
   column_to_rownames('sampleid')
@@
 
+# filter on the above matrix to include pathways existing in at least 20% of the samples (>=4)
+samp_num <- nrow(pw)
+
+ok_samps <- colSums(pw != 0)
+
+# to see how many samps pass the number of zeros threshold
+pers <- seq(0.1, 0.99, 0.1)
+
+pers %>% 
+  set_names(pers) %>% 
+  map_dfc(~ pw[,names(ok_samps[ok_samps >= ceiling(samp_num * .)])] %>% 
+            ncol) %>% 
+  gather( key = 'per', value = 'num')
+
+# temporarily decides to choose 
+per_thre <- 0.5
+
+pw_fil <- pw[,names(ok_samps[ok_samps >= ceiling(samp_num * per_thre)])]
+
+```
+
+```{r}
+
 fit_data <- Maaslin2(
-    pw, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = T,
+    pw_fil, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = T,
     fixed_effects = c('Toxicity', 
@@
                       'IL_6'),
-    # random_effects = c('MRN',
-    #                    'sex',
-    #                    'age'),
+    random_effects = c('sex',
+                       'age'),
     standardize = FALSE)

### 2019-12-11 scripts/27_maASlin2.Rmd
update

@@
 library(tidyverse)
+library(arrangements)
 ```
@@
 ```{r}
+# add the lines of therapy var
+lines <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
+                    sheet = 8)  %>% 
+  rename_all(funs(str_replace_all(., ' ','_'))) %>% 
+  dplyr::select(MRN, Prior_lines_of_therapy)
+```
+
+```{r}
 # get the clinical data info  from ying pulled data
@@
               summarise(crp  = max(result_text, na.rm = T)), by = 'MRN') %>% 
-  left_join(IL_table, by  ='MRN')
+  left_join(IL_table, by  ='MRN') %>% 
+  left_join(lines, by = 'MRN')
 
@@
 
-```{r}
+```{r pheno_final}
 # now change the columns to different type: dbl or factor
@@
          sampleid = as.character(sampleid)) %>% 
-  mutate(IL_6 = log(IL_6),
-         IL_10 = log(IL_10)) %>% 
+  #mutate(IL_6 = log(IL_6),
+  #       IL_10 = log(IL_10)) %>% 
   column_to_rownames('sampleid')
@@
 colnames(pheno_final)
+
+pheno_final %>% 
+  summary
+```
+
+```{r}
+# check the correlation between the pheno vars that are factors
+all_var <- pheno_final %>% 
+  select_if(is.factor) %>% 
+  names
+
+combi <- combinations(all_var, 2) %>% 
+  as.data.frame(stringsAsFactors = F)
+
+phe_cor <- apply(combi, 1, function(Row) {
+  phe_fil = pheno_final %>% 
+    dplyr::select(Row[['V1']], Row[['V2']])
+
+  round(cor.test(pheno_final[,Row[['V1']]],  pheno_final[,Row[['V2']]],method = 'spearman', 
+                           exact = F)$estimate, 2)
+  #chisq.test(phe_fil)$p.value
+}) %>% 
+  as.data.frame() %>% 
+  bind_cols(combi) %>% 
+  mutate(phe_var = str_glue('{V1}__{V2}')) %>% 
+  dplyr::select(-V1, -V2)
+
 ```
@@
 fit_data <- Maaslin2(
-    pw_fil, pheno_final, '../data/maASlin2/pw', normalization  = 'NONE', transform = "LOG",  cores = 4, plot_heatmap = T,
+    pw, pheno_final, '../data/maASlin2/pw', 
+    normalization  = 'NONE', 
+    transform = "LOG",  
+    cores = 4, 
+    plot_heatmap = T,
+    min_prevalence = 0.5,
     fixed_effects = c('Toxicity', 
-                      'Cytokine_Release_Syndrome',
-                      'ICANS',
                       'CR',
-                      'ORR', 
                       'Disease_Class',
@@
                       'IL_10',
-                      'IL_6'),
+                      'IL_6',
+                      'Prior_lines_of_therapy'),
     random_effects = c('sex',
                        'age'),
-    standardize = FALSE)
+    standardize = FALSE,
+    max_significance = 0.1)
 
 ```
-
-```{r}
-trans <- log(pw[,'PWY-7456'] + 1)
-```

### 2019-12-11 scripts/29_abx_usage.Rmd
update

--- 2019-10-18 scripts/10_barplot_genefamilies.Rmd
@@
 ---
-title: "Humann2 bar plot for gene family"
+title: "ABX usage"
 output: html_document
@@
 
-######################################## Bash ########################################
-
-```{bash}
-# needs to download the utility mapping db first
-humann2_databases --download utility_mapping full /Users/daia1/melodyP/data/humann2_res
-```
-
-```{bash}
-# regroup to KEGG Orthogroups (KOs)
-humann2_regroup_table \
-            --input ../data/humann2_res/humann2_genefamilies.tsv  \
-            --groups uniref90_ko  \
-            --output ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv
-```
-
-```{bash}
-# renorm to relab
-humann2_renorm_table  --input ../data/humann2_res/humann2_genefamilies_KO_regroup.tsv -u relab --output ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv
-```
-
-```{bash}
-# split into unstratified and ..
-humann2_split_stratified_table --input ../data/humann2_res/humann2_genefamilies_KO_regroup_relab.tsv  --output ../data/humann2_res/genefamily_split_KO
-```
-
-######################################## R ########################################
-
-## What features(gene orthologs) do you wanna see? Do lefse first.
-
 ```{r}

### 2019-12-13 scripts/29_abx_usage.Rmd
update

@@
 library(tidyverse)
+library(readxl)
+library(lubridate)
 ```
@@
 ```{r}
+# the current shotgun data
+sg <- read_csv('../data/cleaned_pheno/final_pheno.csv')
+
+# the updated baseline sample collection date
+bcd <- read_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/SampleDates12-12-2019.csv', 
+                col_types = cols(.default = col_character())) %>% 
+  mutate(datecollection = mdy(datecollection)) %>% 
+  filter(str_detect(window, 'Baseline')) %>% 
+  filter(sampleid %in% sg$sampleid)
+
+# load the updated abx usage table
+abx <- read_csv('../data/cleaned_pheno/copy12_12_abx.csv', col_types = cols(.default = col_character())) %>% 
+  filter(MRN %in% bcd$mrn) %>% 
+  dplyr::select(MRN, 
+                Antibiotics, 
+                start = `Start date`,
+                end = `End Dates`) %>% 
+  # there is one patient has None abx
+  filter(!is.na(start)) %>% 
+  mutate(start = mdy(start),
+         end = mdy(end)) %>% 
+  # join in the date of baseline sample collection
+  left_join(bcd %>% 
+              dplyr::select(mrn, datecollection), by = c('MRN'='mrn')) %>% 
+  mutate(startday = start - datecollection,
+         endday = end - datecollection)
+```
+
+```{r}
+# create a table that has Y and N for abx usage 4 weeks prior to the baseline sample collection date
+fil <- abx %>% 
+  filter(startday <= 0 ) %>% 
+  distinct(MRN)
 

### 2019-12-13 scripts/27_maASlin2.Rmd
update

@@
 # get the MRN of my target samples 
-link <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
+link <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 12122019.xlsx',
                     sheet = 4) %>% 
-  dplyr::select(pt_id, MRN)
+  dplyr::select(pt_id, 
+                MRN,
+                )
 
@@
 
+pheno %>% 
+  write_csv('../data/cleaned_pheno/final_pheno.csv')
 ```
@@
 ```
+
+```{r}
+source('~/MSK/work/microbiome_db/SQL/scripts/get_data_from_query_OTU.R')
+get_data_from_query_OTU(0,'antibiotics_ag') 
+dbListTables(con) %>% data.frame() %>% filter(grepl("_ag",.)) 
+```

### 2019-12-13 scripts/29_abx_usage.Rmd
update

@@
 library(lubridate)
+```
+
+```{r}
+### Using the clinical data info from ctc.medications
+ctc.medications.srx %>% 
+  filter(MRN == '38009710' & drug_name == 'vancomycin') %>% 
+  dplyr::select(MRN, drug_name, start_datetime, stop_datetime)
+
+ctc.medications.srx %>% 
+  filter(MRN == '35399272') %>% 
+  dplyr::select(MRN, drug_name, start_datetime, stop_datetime) %>% 
+  filter(str_detect(drug_name, 'cephalexin'))
+
 ```
@@
 # create a table that has Y and N for abx usage 4 weeks prior to the baseline sample collection date
+# only keep the MRN that have Y for abx 4 weeks prior to the baseline sample collection date
 fil <- abx %>% 

### 2019-12-16 scripts/29_abx_usage.Rmd
update

@@
                 col_types = cols(.default = col_character())) %>% 
+  # restore the MRN and pad necessary 0 in front of it
+  mutate(mrn = str_pad(mrn, 8, side = 'left', pad = '0')) %>% 
   mutate(datecollection = mdy(datecollection)) %>% 
@@
 # create a table that has Y and N for abx usage 4 weeks prior to the baseline sample collection date
-# only keep the MRN that have Y for abx 4 weeks prior to the baseline sample collection date
+# only keep the MRN that have Y for abx 4 weeks prior to the baseline sample collection date (in the current TJ antibiotics table)
 fil <- abx %>% 
   filter(startday <= 0 ) %>% 
-  distinct(MRN)
+  distinct(MRN) %>% 
+  pull(MRN)
 
+# the remaining MRNs that need to validate whether they have any abx
+other <- setdiff(sg$MRN, fil)
 ```
+
+## Start with Ying table and find out the abx medications
+
+```{r}
+# the below could only be run in the console but not a rmd trunk.
+source('~/MSK/work/microbiome_db/SQL/scripts/get_data_from_query_OTU.R')
+dbListTables(con) %>% data.frame() %>% filter(grepl("_ag",.)) 
+abxtb <- get_data_from_query_OTU(0,'antibiotics_antimicrobial_ordername_map_ag') 
+```
+
+```{r}
+# extract the unique abx names from the above table
+abxu <- abxtb %>% 
+  distinct(order_name) %>% 
+  mutate(abx_name = str_to_lower(str_replace(order_name, ' .+$',''))) %>% 
+  distinct(abx_name) %>% 
+  pull(abx_name)
+```
+
+```{r}
+# all medications of those 14 patients
+othermed <- ctc.medications.srx %>% 
+  filter(MRN %in% other) %>% 
+  dplyr::select(MRN, med_name, start_date, stop_date) %>% 
+  mutate(med_name = str_to_lower(str_replace(med_name, ' .+$',''))) %>% 
+  filter(med_name %in% abxu) %>%  
+  left_join(bcd %>% 
+              dplyr::select(mrn, datecollection), by = c('MRN'='mrn')) %>% 
+  mutate(startday = start_date - datecollection,
+         endday = stop_date - datecollection) %>% 
+  # selecting the records that are the proof of abx
+  filter(startday >= -28 & startday <= 0)
+
+# the final MRN of the yes exposure to abx in the 4 w prior to baseline window
+yexpo <- c(othermed %>% 
+             distinct(MRN) %>% 
+             pull(MRN), fil)
+```

### 2019-12-17 scripts/27_maASlin2.Rmd
update

--- 2019-12-11 scripts/27_maASlin2.Rmd
@@
 # get the MRN of my target samples 
-link <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 11252019 wo Censor .xlsx',
+link <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 12122019.xlsx',
                     sheet = 4) %>% 
-  dplyr::select(pt_id, MRN)
+  dplyr::select(pt_id, 
+                MRN,
+                disease_sta = `Key: Patient Disease Status at CAR`)
 
@@
                 ICANS,
-                Antibiotics_4_weeks_prior_to_CAR_T_cells,
-                #Antibiotics_4_weeks_post_CAR_T_cells,
+                Antibiotics_4_weeks_post_CAR_T_cells,
                 Disease_Class,
-                Key__Patient_Disease_Status_at_CAR
+                disease_sta
                 )
@@
 load('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/YingRCode/ctc_melody.RData')
+
+# load the table I made for the abx exposure
+abx <- read_csv('../data/cleaned_pheno/shotgun_cohort_abx_4w_baseline_expo_table.csv')
 ```
@@
   left_join(IL_table, by  ='MRN') %>% 
-  left_join(lines, by = 'MRN')
+  left_join(lines, by = 'MRN') %>% 
+  left_join(abx, by = 'MRN')
 
@@
   mutate(ferritin = as.numeric(ferritin),
-         Key__Patient_Disease_Status_at_CAR = as.character(Key__Patient_Disease_Status_at_CAR)) %>% 
+         disease_sta = as.character(disease_sta)) %>% 
   # change all the pheno vars to factors
@@
   dplyr::select(-V1, -V2) 
-
 ```
@@
 ```{r}
+# a look at the final pheno data
+colnames(pheno_final)
+
+cor.test(pheno_final$crp, pheno_final$IL_6, method = 'pearson')
+```
+
+```{r}
+# remove everything already in this folder
+#system('rm ../data/maASlin2/pw/*')
 
 fit_data <- Maaslin2(
-    pw, pheno_final, '../data/maASlin2/pw', 
+    pw, pheno_final, '../data/maASlin2/pw/onlyFixed', 
     normalization  = 'NONE', 
@@
     cores = 4, 
-    plot_heatmap = T,
-    min_prevalence = 0.5,
+    plot_heatmap = F,
+    min_abundance = 0.05,
+    min_prevalence = 0.1,
     fixed_effects = c('Toxicity', 
@@
                       'Disease_Class',
-                      'Key__Patient_Disease_Status_at_CAR',
-                      'Antibiotics_4_weeks_prior_to_CAR_T_cells', 
+                      'disease_sta',
+                      'abx_expo', 
                       'ferritin',
@@
                       'IL_1_beta',
-                      'IL_10',
+                      #'IL_10',
                       'IL_6',
-                      'Prior_lines_of_therapy'),
+                      'Prior_lines_of_therapy',
+                      'Antibiotics_4_weeks_post_CAR_T_cells'),
+    # random_effects = c('sex',
+    #                   'age'),
+    standardize = FALSE,
+    max_significance = 0.25)
+
+```
+
+```{r}
+system('rm ../data/maASlin2/pw/both/*')
+
+fit_data <- Maaslin2(
+    pw, pheno_final, '../data/maASlin2/pw/both', 
+    normalization  = 'NONE', 
+    transform = "LOG",  
+    cores = 4, 
+    plot_heatmap = F,
+    min_abundance = 0.05,
+    min_prevalence = 0.1,
+    fixed_effects = c('Toxicity', 
+                      'CR',
+                      'Disease_Class',
+                      'disease_sta',
+                      'abx_expo', 
+                      'ferritin',
+                      'crp',
+                      #'IL_1_beta',
+                      #'IL_10',
+                      #'IL_6',
+                      'Prior_lines_of_therapy',
+                      'Antibiotics_4_weeks_post_CAR_T_cells'),
     random_effects = c('sex',
@@
     standardize = FALSE,
-    max_significance = 0.1)
-
+    max_significance = 0.25)
 ```

### 2019-12-19 scripts/27_maASlin2.Rmd
update

@@
 # remove everything already in this folder
-#system('rm ../data/maASlin2/pw/*')
+system('rm ../data/maASlin2/pw/onlyFixed/*')
 
@@
                       'Prior_lines_of_therapy',
+                      'sex',
+                      'age',
                       'Antibiotics_4_weeks_post_CAR_T_cells'),
-    # random_effects = c('sex',
-    #                   'age'),
     standardize = FALSE,
@@
                       'crp',
-                      #'IL_1_beta',
+                      'IL_1_beta',
                       #'IL_10',
-                      #'IL_6',
+                      'IL_6',
                       'Prior_lines_of_therapy',

### 2020-01-08 scripts/30_univariate_lm.Rmd
update

---
title: "Univariate lm model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
final_pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  dplyr::rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  dplyr::rename(pw = names(.)[1])  %>% 
  # remove the things after the : so that its easier to see on the plot(the pathway name)
  mutate(pw = str_replace(pw, ':.+$',''))

pw <- bind_cols(pw[,1],pw[,final_pheno$sampleid])

all.equal(colnames(pw)[2:ncol(pw)], final_pheno$sampleid)
```

```{r}
# filter to have each pathway with at least 30% of nonzero (zero <= 50%)
pw_fil <- pw %>% 
  column_to_rownames('pw') %>% 
  as.matrix()

num_zero_thre <- floor(ncol(pw_fil) * 0.5)

pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]
nrow(pw_fil)

```

```{r}
# select the most abundant N pathways

N <- 10
topN_pw <- sort(rowSums(pw_fil), decreasing = T)[1:N]
pw_Fil <- pw_fil[names(topN_pw),]

all.equal(colnames(pw_Fil), final_pheno$sampleid)

# do a linear regression of each rows and then do a FDR correction later
# the pheno in the same order as the current sample names 
phe_tox <- final_pheno$Toxicity
phe_CR <- final_pheno$CR

res_tox <- apply(pw_Fil, 1, function(r){
  fit = summary(lm(r ~ phe_tox))$coefficients
  return(pval = fit[nrow(fit),ncol(fit)])
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))

res_CR <- apply(pw_Fil, 1, function(r){
  fit = summary(lm(r ~ phe_CR))$coefficients
  return(pval = fit[nrow(fit),ncol(fit)])
}) %>% 
  data.frame() %>% 
  rename(pval = names(.)[1]) %>% 
  rownames_to_column('pw') %>% 
  mutate(padj = p.adjust(pval, method = 'BH')) %>% 
  arrange(desc(padj))
```

### 2020-01-08 scripts/30_univariate_lm.Rmd
update

@@
 ---
-title: "Univariate lm model"
-output: html_document
+title: "Univariate linear regression"
+author: "Anqi Dai"
+date: "`r Sys.Date()`"
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
 library(tidyverse)
+library(kableExtra)
 ```
+
+Univariate linear regression to see if there is significant pathways from the humann2 normalized data
+
+## Load pheno and count data
 
@@
   mutate(pw = str_replace(pw, ':.+$','')) 
+```
 
+##### Reorder the column and make sure the sample names in the pheno table align with the sample name in the count table
+
+```{r}
 pw <- bind_cols(pw[,1],pw[,final_pheno$sampleid])
@@
 
+## filter to have each pathway with at least 50% of nonzero (zero <= 50%)
+
 ```{r}
-# filter to have each pathway with at least 30% of nonzero (zero <= 50%)
+# filter to have each pathway with at least 50% of nonzero (zero <= 50%)
 pw_fil <- pw %>% 
@@
 pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]
-nrow(pw_fil)
 
@@
 
+##### How many pathways left
+
 ```{r}
-# select the most abundant N pathways
+nrow(pw_fil)
+```
+
+## Select the most abundant 10 pathways and log2 transform the data
+
+```{r}
+# select the most abundant N pathways and log2 transform the data
 
@@
 
-all.equal(colnames(pw_Fil), final_pheno$sampleid)
+pw_Fil <- log2(pw_Fil)
 
+pw_Fil %>% 
+  kable(caption = 'log2 transformed count table with top 10 most abundant pathways') %>% 
+  kable_styling(full_width = F, position = 'left')
+```
+
+## Linear regression for each pathway with Toxicity and CR as response respectively, followed by FDR correction
+
+```{r}
 # do a linear regression of each rows and then do a FDR correction later
@@
   arrange(desc(padj))
+
+res_tox %>% 
+  kable(caption = 'Response to Toxicity') %>% 
+  kable_styling(full_width = F, position = 'left')
+
+res_CR %>% 
+  kable(caption = 'Response to CR') %>% 
+  kable_styling(full_width = F, position = 'left')
 ```
+
+## How many pathways are significant with FDR < 0.1
+
+```{r}
+# Toxicity
+res_tox %>% 
+  filter(padj < 0.1) %>% 
+  nrow
+
+# CR
+res_CR %>% 
+  filter(padj < 0.1) %>% 
+  nrow
+
+```
+
+##### I have tested: no sig pathways even select the top 200 most abundant ones.

### 2020-01-09 scripts/31_univariate_lm_wilcox.Rmd
update

--- 2020-01-08 scripts/30_univariate_lm.Rmd
@@
 ---
-title: "Univariate linear regression"
+title: "Univariate model using Wilcox Rank Sum test"
 author: "Anqi Dai"
@@
 
-Univariate linear regression to see if there is significant pathways from the humann2 normalized data
+Univariate Wilcox test to see if there is significant pathways from the humann2 normalized data
 
@@
 all.equal(colnames(pw)[2:ncol(pw)], final_pheno$sampleid)
+
+pw <- pw %>% 
+  column_to_rownames('pw') %>% 
+  as.matrix() 
 ```
 
-## filter to have each pathway with at least 50% of nonzero (zero <= 50%)
+```{r}
+## Filter to have each pathway with at least 10% of nonzero (zero <= 90%)
+
+#num_zero_thre <- floor(ncol(pw_fil) * 0.9)
+
+#pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]
+
+##### How many pathways left
+#nrow(pw_fil) 
+```
+
+## Log2 transform the data
 
 ```{r}
-# filter to have each pathway with at least 50% of nonzero (zero <= 50%)
-pw_fil <- pw %>% 
-  column_to_rownames('pw') %>% 
-  as.matrix()
+#pw_Fil <- log2(pw_fil + 0.01)
 
-num_zero_thre <- floor(ncol(pw_fil) * 0.5)
-
-pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]
-
+# use the whole table (not filter anything)
+pw_log2 <- log2(pw + 0.01)
 ```
 
-##### How many pathways left
+How many pathways went into the test
 
 ```{r}
-nrow(pw_fil)
+nrow(pw_log2)
 ```
 
-## Select the most abundant 10 pathways and log2 transform the data
+## Wilcox for each pathway with Toxicity and CR as response respectively, followed by FDR correction
 
 ```{r}
-# select the most abundant N pathways and log2 transform the data
-
-N <- 10
-topN_pw <- sort(rowSums(pw_fil), decreasing = T)[1:N]
-pw_Fil <- pw_fil[names(topN_pw),]
-
-pw_Fil <- log2(pw_Fil)
-
-pw_Fil %>% 
-  kable(caption = 'log2 transformed count table with top 10 most abundant pathways') %>% 
-  kable_styling(full_width = F, position = 'left')
-```
-
-## Linear regression for each pathway with Toxicity and CR as response respectively, followed by FDR correction
-
-```{r}
-# do a linear regression of each rows and then do a FDR correction later
-# the pheno in the same order as the current sample names 
 phe_tox <- final_pheno$Toxicity
@@
 
-res_tox <- apply(pw_Fil, 1, function(r){
-  fit = summary(lm(r ~ phe_tox))$coefficients
-  return(pval = fit[nrow(fit),ncol(fit)])
+res_tox <- apply(pw_log2, 1, function(r){
+  res = wilcox.test(r ~ phe_tox)
+  return(pval = res$p.value)
 }) %>% 
@@
 
-res_CR <- apply(pw_Fil, 1, function(r){
-  fit = summary(lm(r ~ phe_CR))$coefficients
-  return(pval = fit[nrow(fit),ncol(fit)])
+res_CR <- apply(pw_log2, 1, function(r){
+  res = wilcox.test(r ~ phe_CR)
+  return(pval = res$p.value)
 }) %>% 
@@
 ```
-
-##### I have tested: no sig pathways even select the top 200 most abundant ones.

### 2020-01-10 scripts/31_univariate_lm_wilcox.Rmd
update

@@
 
+phe_tox <- final_pheno$Toxicity
+phe_CR <- final_pheno$CR
+
 pw <- pw %>% 
@@
 
-```{r}
-## Filter to have each pathway with at least 10% of nonzero (zero <= 90%)
-
-#num_zero_thre <- floor(ncol(pw_fil) * 0.9)
-
-#pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]
-
-##### How many pathways left
-#nrow(pw_fil) 
-```
-
-## Log2 transform the data
+## Use different prevalence filtering threshold (percentage of zero count samples) and do the log2 transform and wilcox test
 
 ```{r}
-#pw_Fil <- log2(pw_fil + 0.01)
+num_zero_thre <- seq(0, 0.99, 0.05)
 
-# use the whole table (not filter anything)
-pw_log2 <- log2(pw + 0.01)
-```
+test_diff_zero_thre_filter <- function(fil_thre){
 
-How many pathways went into the test
+  num_zero_thre <- floor(ncol(pw_fil) * fil_thre)
 
-```{r}
-nrow(pw_log2)
-```
+  pw_fil <- pw[rowSums(pw == 0) <= num_zero_thre, ]
 
-## Wilcox for each pathway with Toxicity and CR as response respectively, followed by FDR correction
+  pw_log2 <- log2(pw_fil + 0.01)
 
-```{r}
-phe_tox <- final_pheno$Toxicity
-phe_CR <- final_pheno$CR
-
-res_tox <- apply(pw_log2, 1, function(r){
-  res = wilcox.test(r ~ phe_tox)
+  sig_tox <- apply(pw_log2, 1, function(r){
+      res = wilcox.test(r ~ phe_tox, exact = F)
       return(pval = res$p.value)
@@
       mutate(padj = p.adjust(pval, method = 'BH')) %>% 
-  arrange(desc(padj))
+      arrange(desc(padj)) %>% 
+      filter(padj < 0.25) %>% 
+      nrow
 
-res_CR <- apply(pw_log2, 1, function(r){
-  res = wilcox.test(r ~ phe_CR)
+  sig_CR <- apply(pw_log2, 1, function(r){
+      res = wilcox.test(r ~ phe_CR, exact = F)
       return(pval = res$p.value)
@@
       mutate(padj = p.adjust(pval, method = 'BH')) %>% 
-  arrange(desc(padj))
+      arrange(desc(padj)) %>% 
+      filter(padj < 0.25) %>% 
+      nrow
 
-res_tox %>% 
-  kable(caption = 'Response to Toxicity') %>% 
-  kable_styling(full_width = F, position = 'left')
+  return(list(sig_tox_num = sig_tox,
+         sig_CR_num = sig_CR,
+         pw_num = nrow(pw_log2)))
+}
 
-res_CR %>% 
-  kable(caption = 'Response to CR') %>% 
+res <- list()
+for (i in 1:length(num_zero_thre)){
+  res[[i]] = test_diff_zero_thre_filter(num_zero_thre[i])
+}
+
+final <- bind_rows(res) %>% 
+  mutate(zero_perc = num_zero_thre) %>% 
+  select(zero_perc, pw_num, sig_tox_num, sig_CR_num)
+
+final %>% 
+  kable(caption = 'Number of significant results (FDR < 0.25) in different zero percentage threshold') %>% 
   kable_styling(full_width = F, position = 'left')
 ```
-
-## How many pathways are significant with FDR < 0.1
-
-```{r}
-# Toxicity
-res_tox %>% 
-  filter(padj < 0.1) %>% 
-  nrow
-
-# CR
-res_CR %>% 
-  filter(padj < 0.1) %>% 
-  nrow
-
-```

### 2020-01-16 scripts/32_two_genera.Rmd
update

---
title: "Two genera pathway association"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
library(tidyverse)
library(pheatmap)
library(kableExtra)
```

Load the temporary table of humann2 gene fam in genus level results

Renorm

Split

# The genus of Enterococcus

Get the Enterococcus count of the pathway

```{r}
pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

genus <- read_tsv('../data/humann2_res/genus_level/humann2_pathabundance_genus_level_cpm_stratified.tsv') %>% 
  dplyr::rename(pw = names(.)[1]) %>% 
  rename_all(funs(str_replace(., '_genefamilies_.+$','')))

genus_entero <- genus %>% 
  filter(str_detect(pw, 'g__Enterococcus')) 
```

```{r}
# a heatmap to see the counts distribution

annot <- data.frame(row.names = pheno$sampleid, tox = pheno$Toxicity, CR = pheno$CR)

pheatmap(
  genus_entero %>% 
    column_to_rownames('pw') %>% 
    select(-`1409B`) %>% 
    as.matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = TRUE,
  show_colnames = TRUE,
  filename = '../figs/Enterococcus_heatmap.jpg',
  height = 18,
  width = 18,
  cluster_rows = F,
  cluster_cols = F
)
```

Many of them are zeros that we can see, I don't think it makes a lot of sense to do a linear regression.

## Do a wilcox in relation to Toxicity

```{r}
# ehh do it anyway
entero_matrix <- genus_entero %>% 
    column_to_rownames('pw') %>% 
    as.matrix

# rearrange the column names
entero_matrix <- entero_matrix[,pheno$sampleid]

all.equal(colnames(entero_matrix), pheno$sampleid)

# a linear regression of CR and tox respectively
sig_tox <- apply(entero_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$Toxicity, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_tox %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)

sig_CR <- apply(entero_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$CR, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_CR %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

# Gather all the genera in Bacteroidetes

```{r}
bac_genera_g <- c('Bacteroides',
                'Prevotella',
                'Parabacteroides',
                'Porphyromonas',
                'Flavobacterium',
                'Cytophaga',
                'Alistipes',
                'Flexibacter',
                'Capnocytophaga',
                'Dyadobacter',
                'Hymenobacter',
                'Barnesiella',
                'Dorea',
                'Flectobacillus',
                'Arcicella',
                'Chryseobacterium',
                'Salinibacter',
                'Polaribacter',
                'Cellulophaga',
                'Blattabacterium',
                'Runella',
                'Flammeovirga',
                'Tenacibaculum',
                'Sphingobacterium',
                'Pedobacter',
                'Leeuwenhoekiella',
                'Butyricimonas',
                'Bergeyella',
                'Robiginitalea',
                'Croceibacter',
                'Dysgonomonas',
                'Algoriphagus',
                'Thermonema',
                'Gelidibacter',
                'Arenibacter',
                'Alkaliflexus',
                'Aequorivita',
                'Muricauda',
                'Chitinophaga',
                'Epilithonimonas',
                'Paludibacter',
                'Belliella',
                'Butyricicoccus',
                'Megamonas',
                'Lewinella',
                'Niabella',
                'Vitellibacter',
                'Fabibacter',
                'Pontibacter',
                'Proteiniphilum',
                'Mucilaginibacter')

bac_genera <- read_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/data/Bacteroidetes_genus.csv') %>% 
  pull(genus)

intersect(bac_genera, bac_genera_g)
```

```{r}
# find out all the rows that have bac genera
genus_bact <- genus %>% 
  separate(pw, into = c('pathway','genus'), '\\|') %>% 
  mutate(genus = str_replace(genus, 'g__','')) %>% 
  filter(genus %in% bac_genera)

genus_bact  %>% 
  count(pathway)

# group by the pathway and get the summed count of all the bact genera
bact_matrix <- genus_bact %>% 
  select(-genus) %>% 
  gather(key = 'sampleid', value = 'CPM', names(.)[2]:names(.)[ncol(.)]) %>% 
  group_by(pathway, sampleid) %>% 
  summarise(cpm = sum(CPM)) %>% 
  spread(key = 'sampleid', value = 'cpm') %>% 
  column_to_rownames('pathway') %>% 
  as.matrix

bact_matrix <- bact_matrix[,pheno$sampleid]

all.equal(colnames(bact_matrix), pheno$sampleid)
```
## a wilcox of CR and tox respectively

```{r}
# a wilcox of CR and tox respectively
sig_tox <- apply(bact_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$Toxicity, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_tox %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)

sig_CR <- apply(bact_matrix, 1, function(r){
    res = wilcox.test(r ~ pheno$CR, exact = F)
    return(pval = res$p.value)
  }) %>% 
    data.frame() %>% 
    dplyr::rename(pval = names(.)[1]) %>% 
    rownames_to_column('pw') %>% 
    mutate(padj = p.adjust(pval, method = 'BH')) %>% 
    arrange(padj)

sig_CR %>% 
  head() %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

### 2020-01-17 scripts/32_two_genera.Rmd
update

@@
 
+genus_no_bact <- genus %>% 
+  separate(pw, into = c('pathway','genus'), '\\|') %>% 
+  mutate(genus = str_replace(genus, 'g__','')) %>% 
+  filter(!genus %in% bac_genera) %>% 
+  filter(genus != 'unclassified')
+
+genus_no_bact_matrix <- genus_no_bact %>% 
+  select(-genus) %>% 
+  gather(key = 'sampleid', value = 'CPM', names(.)[2]:names(.)[ncol(.)]) %>% 
+  group_by(pathway, sampleid) %>% 
+  summarise(cpm = sum(CPM)) %>% 
+  spread(key = 'sampleid', value = 'cpm') %>% 
+  column_to_rownames('pathway') %>% 
+  as.matrix
+
 genus_bact  %>% 

### 2020-01-17 scripts/33_distance_bar.Rmd
update

---
title: "Plot the within and between group distance in a bar plot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

```{r}
# make a matrix of filtered 50% cpm reordered counts
final_pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')

pw <- read_tsv('../data/humann2_res/humann2_final_out/humann2_pathabundance_cpm_unstratified.tsv') %>% 
  dplyr::rename_all(funs(str_replace(., '_Abundance$',''))) %>% 
  dplyr::rename(pw = names(.)[1])  %>% 
  # remove the things after the : so that its easier to see on the plot(the pathway name)
  mutate(pw = str_replace(pw, ':.+$',''))

pw <- bind_cols(pw[,1],pw[,final_pheno$sampleid])

all.equal(colnames(pw)[2:ncol(pw)], final_pheno$sampleid)

# filter to have each pathway with at least 50% of nonzero (zero <= 50%)
pw_fil <- pw %>% 
  column_to_rownames('pw') %>% 
  as.matrix()

num_zero_thre <- floor(ncol(pw_fil) * 0.5)

pw_fil <- pw_fil[rowSums(pw_fil == 0) <= num_zero_thre, ]

pw_fil %>% 
  write.csv('../data/00_filtered_matrix/humann2_pathabundance_cpm_unstratified_filtered50_reorder_matrix.csv', 
              row.names = T, col.names = T, quote = F)
```

You make a barplot with three bars.
- mean distance among samples from The “N” group and another (within distance group N)
- mean distance among samples from The “Y” group and another (within distance group Y)
- mean distance among samples from group “N” to group “Y” (between groups distance)

- For within distance in group “N”, you take the values from rows and columns that belong to samples of group “N”, excluding the diagonal. The matrix is symmetric and will give duplicate values, so take only one of them.
- Do the same for within distance in group “Y”.
- For between distance, you take the rows from group “N” and columns from group “Y”. The matrix is not symmetric in this case and give a single value for each possible pair.

```{r}
# calculate distance between N samples
N_samps <- final_pheno %>% 
  filter(Toxicity == 'N') %>% 
  pull(sampleid)

Y_samps <- final_pheno %>% 
  filter(Toxicity == 'Y') %>% 
  pull(sampleid)

N_mtx <- pw_fil[,N_samps]

Y_mtx <- pw_fil[,Y_samps]

# calculate distance between samples

N_dist <- stats::dist(t(N_mtx), method = "euclidean")

Y_dist <- stats::dist(t(Y_mtx), method = "euclidean")

N_dist_mean <- mean(N_dist)

Y_dist_mean <- mean(Y_dist)

# calcualte distance between the pairs in N and Y group
library(pdist)

dists <- pdist(t(N_mtx), t(Y_mtx))

res <- as.matrix(dists)

NY_dist_mean <- mean(res)

tibble(
  N_dist_mean = N_dist_mean,
  Y_dist_mean = Y_dist_mean,
  NY_dist_mean =  NY_dist_mean
) %>% 
  gather(key = 'group', value  = 'mean_dist') %>% 
  mutate(mean_dist = round(mean_dist, 0)) %>% 
  ggbarplot(x = 'group', y = 'mean_dist', label = T,
            color = 'white', fill = 'forestgreen',
            title = 'euclidean distance within and between group')
```

### 2020-01-17 scripts/33_distance_bar.Rmd
update

@@
 ```
+
+```{r}
+all <- bind_rows(data_frame(group = 'NY',distance = dists@dist),
+    data_frame(group = 'N',distance = N_dist),
+    data_frame(group = 'Y',distance = Y_dist))
+
+all %>% 
+  ggboxplot(x = 'group', y = 'distance', add = 'jitter') +
+  stat_compare_means(
+    comparisons = list(c('NY', 'N'), c('NY','Y'), c('Y','N')),
+    label = "p.signif",
+    method = 'wilcox.test')
+```

### 2020-01-19 scripts/34_secondary_bile.Rmd
update

---
title: "secondary bile acid biosynthesis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

```{r}
pheno <- read_csv('../../CART_microbiome/data/cleaned_pheno/final_pheno.csv')
mbile <- list.files('../../CART_microbiome/data/bile/', full.names = T)

all <- mbile %>% 
  set_names(mbile) %>% 
  map(~ read_tsv(.)) %>% 
  bind_rows(.id = 'sampleid') %>% 
  group_by(sampleid) %>% 
  summarise(RPKM = sum(Count))

all %>% 
  mutate(sampleid = str_replace(sampleid, '.+//',''),
         sampleid = str_replace(sampleid, '_short.+$','')) %>% 
  full_join(pheno, by = 'sampleid') %>% 
  ggboxplot(x = 'CR', y = 'RPKM', color = 'CR', palette = 'lancet', add = 'dotplot')  +
  stat_compare_means(comparisons = list(c('Y', 'N')), 
                     label = "p.signif",
                     method = 'wilcox.test',
                     correct=FALSE)

all %>% 
  mutate(sampleid = str_replace(sampleid, '.+//',''),
         sampleid = str_replace(sampleid, '_short.+$','')) %>% 
  full_join(pheno, by = 'sampleid') %>% 
  ggboxplot(x = 'Toxicity', y = 'RPKM', color = 'Toxicity', palette = 'lancet', add = 'dotplot',
            title = 'CART samples secondary bile acid')  +
  stat_compare_means(comparisons = list(c('Y', 'N')), 
                     label = "p.signif",
                     method = 'wilcox.test',
                     correct = FALSE) 
```

### 2020-01-28 scripts/33_distance_bar.Rmd
update

--- 2020-01-17 scripts/33_distance_bar.Rmd
@@
               row.names = T, col.names = T, quote = F)
+```
+
+```{r}
+# do the same thing as below but with matrix that only have enterococcos and ...
+pw_fil <- bact_matrix
 ```
@@
 
-tibble(
-  N_dist_mean = N_dist_mean,
-  Y_dist_mean = Y_dist_mean,
-  NY_dist_mean =  NY_dist_mean
-) %>% 
-  gather(key = 'group', value  = 'mean_dist') %>% 
-  mutate(mean_dist = round(mean_dist, 0)) %>% 
-  ggbarplot(x = 'group', y = 'mean_dist', label = T,
-            color = 'white', fill = 'forestgreen',
-            title = 'euclidean distance within and between group')
+# tibble(
+#   N_dist_mean = N_dist_mean,
+#   Y_dist_mean = Y_dist_mean,
+#   NY_dist_mean =  NY_dist_mean
+# ) %>% 
+#   gather(key = 'group', value  = 'mean_dist') %>% 
+#   mutate(mean_dist = round(mean_dist, 0)) %>% 
+#   ggbarplot(x = 'group', y = 'mean_dist', label = T,
+#             color = 'white', fill = 'forestgreen',
+#             title = 'euclidean distance within and between group')
+
+all <- bind_rows(data_frame(group = 'NY',distance = dists@dist),
+    data_frame(group = 'N',distance = N_dist),
+    data_frame(group = 'Y',distance = Y_dist))
+
+all %>% 
+  ggboxplot(x = 'group', y = 'distance', add = 'jitter') +
+  stat_compare_means(
+    comparisons = list(c('NY', 'N'), c('NY','Y'), c('Y','N')),
+    label = "p.signif",
+    method = 'wilcox.test') +
+  ggsave('../figs/distance/bact.jpg', dpi = 300, width = 6 , height = 5)
 ```

### 2020-01-31 scripts/35_Braycurtis_distance_two_genera.Rmd
update

---
title: "Bray curtis distance to the group centroid"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(pdist)
```

# entero_matrix

```{r}
sum_0_sample <- names(apply(entero_matrix, 2, sum)[apply(entero_matrix, 2, sum) == 0])
entero_matrix_fil <- entero_matrix[,!colnames(entero_matrix) %in% sum_0_sample]

# the N matrix and Y matrix
N_mat <- entero_matrix_fil[,colnames(entero_matrix_fil) %in% N_samps]
Y_mat <- entero_matrix_fil[,!colnames(entero_matrix_fil) %in% N_samps]

N_dist <- vegdist(t(N_mat), "bray")
Y_dist <- vegdist(t(Y_mat), "bray")

#dists <- pdist(t(N_mat), t(Y_mat))

all <- bind_rows(
    data_frame(group = 'N',distance = N_dist),
    data_frame(group = 'Y',distance = Y_dist)
    )

all %>% 
  ggboxplot(x = 'group', y = 'distance', add = 'jitter', title = 'entero_matrix') +
  stat_compare_means(
    comparisons = list(c('Y','N')),
    label = "p.signif",
    method = 'wilcox.test')
```

# bact_matrix

```{r}
sum_0_sample <- names(apply(bact_matrix, 2, sum)[apply(bact_matrix, 2, sum) == 0])

bact_matrix_fil <- bact_matrix[,!colnames(bact_matrix) %in% sum_0_sample]

# the N matrix and Y matrix
N_mat <- bact_matrix_fil[,colnames(bact_matrix_fil) %in% N_samps]
Y_mat <- bact_matrix_fil[,!colnames(bact_matrix_fil) %in% N_samps]

N_dist <- vegdist(t(N_mat), "bray")
Y_dist <- vegdist(t(Y_mat), "bray")

all <- bind_rows(
    data_frame(group = 'N',distance = N_dist),
    data_frame(group = 'Y',distance = Y_dist)
    )

all %>% 
  ggboxplot(x = 'group', y = 'distance', add = 'jitter', title = 'bact_matrix') +
  stat_compare_means(
    comparisons = list(c('Y','N')),
    label = "p.signif",
    method = 'wilcox.test')
```

### 2020-01-31 scripts/06_lefse_res_ggplot.Rmd
update

@@
   mutate(group = str_replace(group, 'pathway_abundace_pheno__','')) %>% 
-  mutate(group = str_replace(group, '.res$','')) %>% 
-  dplyr::select(-xx, -pval)
+  mutate(group = str_replace(group, '.res$',''))
 
@@
   arrange(group, pathway, score)
+
+feature %>% 
+  write_csv('../data/lefse_results_sig_pathway.csv')
 ```

### 2020-02-05 scripts/32_two_genera.Rmd
update

@@
     as.matrix
+
+genus_entero %>% 
+  write_tsv('../data/two_genera/genus_entero_pw_df.tsv')
 
@@
 
-bac_genera <- read_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/data/Bacteroidetes_genus.csv') %>% 
+bac_genera <- read_csv('../data/two_genera/Bacteroidetes_genus.csv') %>% 
   pull(genus)
@@
 ```
+```{r}
+
+pheatmap(
+  bact_matrix[,colnames(bact_matrix) != "2098A"],
+  #color=bwrPalette, 
+  annotation_col = annot,
+  show_rownames = T,
+  show_colnames = TRUE,
+  filename = '../figs/heatmap/bact_heatmap.jpg',
+  height = 18,
+  width = 28,
+  cluster_rows = T,
+  cluster_cols = T
+)
+```
+
+```{r}
+# because of the sparsity that we turn the count into 0 and 1
+
+entero_matrix[entero_matrix > 0] = 1
+
+bact_matrix[bact_matrix > 0] = 1
+
+pheatmap(
+  bact_matrix,
+  #color=bwrPalette, 
+  annotation_col = annot,
+  show_rownames = F,
+  show_colnames = TRUE,
+  filename = '../figs/heatmap/bact_heatmap_binary.jpg',
+  height = 18,
+  width = 20,
+  cluster_rows = T,
+  cluster_cols = T
+)
+
+pheatmap(
+  entero_matrix,
+  #color=bwrPalette, 
+  annotation_col = annot,
+  show_rownames = F,
+  show_colnames = TRUE,
+  filename = '../figs/heatmap/entero_heatmap_binary.jpg',
+  height = 18,
+  width = 20,
+  cluster_rows = T,
+  cluster_cols = T,
+  main = 'entero genus'
+)
+```

### 2020-02-05 scripts/36_two_heatmap.Rmd
update

---
title: "Enterococcus and Bacteroidetes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pheatmap)
library(tidyverse)
library(LaplacesDemon)
```

Investigating the contribution to the pathway abundances from the Bacteroidetes phylum and Enterococcus genus

```{r}
# load the normalized CPM count matrix with the stratified contribution from the above listed pylum and genus

entero_matrix <- read_tsv('../data/two_genera/genus_entero_pw_df.tsv') %>% 
    column_to_rownames('pw') %>% 
    as.matrix

bact_matrix <-  read_tsv('../data/two_genera/genus_bact_pw_df.tsv') %>% 
    column_to_rownames('pathway') %>% 
    as.matrix

pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv')
```

## Visualize normalized counts in Bacteroidetes phylum

```{r out.width='200%', out.height='200%'}
annot <- data.frame(row.names = pheno$sampleid, tox = pheno$Toxicity, CR = pheno$CR)

pheatmap(
  bact_matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = F,
  show_colnames = TRUE,
  #filename = '../figs/heatmap/bact_heatmap.jpg',
  height = 18,
  width = 28,
  cluster_rows = T,
  cluster_cols = T,
  main = 'All samples in Bacteroidetes'
)
```

Since 2098A seems to have too high counts and are making other samples counts ignorable. Try removing that one

```{r out.width='200%', out.height='200%'}
pheatmap(
  bact_matrix[,colnames(bact_matrix) != "2098A"],
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = F,
  show_colnames = TRUE,
  #filename = '../figs/heatmap/bact_heatmap.jpg',
  height = 18,
  width = 28,
  cluster_rows = T,
  cluster_cols = T,
  main = 'Excluding sample 2098A in Bacteroidetes'
) 
```

It doesn't suggest any association between counts of certain pathways with the phenotype eg. toxicity.

## Binarize Enterococcus counts and visualize

Since the Enterococcus is only a genus and the counts are more sparse for that, binarize the matrix to presense -- 1, absense -- 0

```{r}
entero_matrix[entero_matrix > 0] = 1

pheatmap(
  entero_matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = F,
  show_colnames = TRUE,
  #filename = '../figs/heatmap/entero_heatmap_binary.jpg',
  height = 18,
  width = 20,
  cluster_rows = T,
  cluster_cols = T,
  main = 'entero genus'
)
```

Again no differentiating between the groups

## Pairwise KL divergence

```{r}
library(arrangements)

combi <- combinations(colnames(bact_matrix), 2) %>% 
  as.data.frame(stringsAsFactors = F)

KLD_res <- apply(combi, 1, function(Row) {

  samp1_dens = density(bact_matrix[,Row[['V1']]])$y
  samp2_dens = density(bact_matrix[,Row[['V2']]])$y

  return(pairwise_KLD = KLD(samp1_dens, samp2_dens)$mean.sum.KLD	)

}) %>% 
  as.data.frame() %>% 
  bind_cols(combi) %>% 
  rename(KLD = names(.)[1])

KLD_res_matrix <- KLD_res %>% 
  spread(key = 'V2', value = 'KLD') %>% 
  column_to_rownames('V1') %>% 
  as.matrix()

pheatmap(
  KLD_res_matrix,
  #color=bwrPalette, 
  annotation_col = annot,
  show_rownames = T,
  show_colnames = T,
  #filename = '../figs/heatmap/entero_heatmap_binary.jpg',
  height = 18,
  width = 18,
  cluster_rows = T,
  cluster_cols = T,
  main = 'pairwise KL divergence'
)

```

### 2020-02-07 scripts/10_genefamilies_KO_regroup.Rmd
udpate

@@
 ```
+
+```{r}
+# look at the KOs that is involved in the secondary bile acid biosynthesis
+
+bile_KOs <- c('K01442','K15868','K15869','K15870','K15871','K15872','K15873','K15874')
+bile_KOs <- paste(bile_KOs, collapse = '|')
+
+ko_genef <- read_tsv('../data/humann2_res/humann2_genefamilies_KO_regroup.tsv') %>% 
+  filter(str_detect(`# Gene Family`, bile_KOs)) %>% 
+  filter(str_detect(`# Gene Family`, '\\|')) %>% 
+  distinct(`# Gene Family`) %>% 
+  rename(genefam = `# Gene Family`) %>% 
+  separate(genefam, into = c('KO_ID','species') ,sep = '\\|'  )
+
+```

### 2020-04-27 scripts/31_univariate_lm_wilcox.Rmd
update

@@
 knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
+```
+
+```{r}
 library(tidyverse)
@@
 
-  num_zero_thre <- floor(ncol(pw_fil) * fil_thre)
+  num_zero_thre <- floor(ncol(pw) * fil_thre)
 
@@
 ```
+
+```{r}
+final
+```

### 2020-11-09 scripts/37_clean_pheno_refreshed.Rmd
picking up this project

---
title: "cleaning the pheno table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## use Antonio's script to clean the 16s data meta info

### msk

```{r}
#Oct/16/2019
#
#Loading patient info for Melody's project.
#
#

library('openxlsx');
library(data.table);

if(!exists("dt_carT_patient_info_with_sample_and_oligoID_time_stamp")){

  #Load patient info
  # dt_carT_patient_info = read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table 10162019.xlsx",
  #                                  sheetName = "Summary Table with Some Oligos");
  # dt_carT_patient_info = data.table(dt_carT_patient_info);
  # dt_carT_patient_info = dt_carT_patient_info[!is.na(pt_id)];
  # 
  #Convert sampleid to oligo
  dt_carT_sampleid_info = openxlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table01132020.xlsx",#CTC Event Table 11252019 wo Censor .xlsx",
                                              sheet = "Sample IDs only ");
                                    # sheetName = "Sample IDs only ");
  dt_carT_sampleid_info = data.table(dt_carT_sampleid_info);

  dt_carT_sampleid_info2 = openxlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table01132020.xlsx",#CTC Event Table 11252019 wo Censor .xlsx",
                                              sheet = "Summary Table with Some Oligos");
  dt_carT_sampleid_info2 = data.table(dt_carT_sampleid_info2);
  pt_id_2_mrn = dt_carT_sampleid_info2[!duplicated(pt_id)]$MRN
  names(pt_id_2_mrn) = dt_carT_sampleid_info2[!duplicated(pt_id)]$pt_id

  dt_carT_sampleid_info[,c("mrn_str","mrn"):=.(pt_id_2_mrn[pt_id], as.numeric(pt_id_2_mrn[pt_id])) ]

  #dt_carT_sampleid_info = dt_carT_sampleid_info[!is.na(p)]

  dt_carT_sampleid_info_m = melt.data.table(dt_carT_sampleid_info,id.vars = c("pt_id"),
                                            measure.vars = c("Baseline.Sample.ID",
                                                             "Week.1.Sample.ID",
                                                             "Week.2.Sample.ID",
                                                             "Week.3.Sample.ID",
                                                             "Week.4.Sample.ID"),
                                            variable.name = "time_group",
                                            value.name = "sampleid",
                                            na.rm = T);

  #sample ids with no time assignment:
  #d_set_tsne2[!sampleid %in% dt_carT_sampleid_info_m$sampleid ]$sampleid

    dt_carT_patient_info_with_sample_and_oligoID = merge(dt_carT_sampleid_info[is.na(Censor_reason), #Censor_YN=="N",
                                                                            .(pt_id,
                                                                              mrn,
                                                                              mrn_str,
                                                                              CR=`CR.(Y/N)`,
                                                                              ORR=`ORR.(Y/N)`,
                                                                              Toxicity=`Toxicity.(Y/N)`,
                                                                              #CR=CR..Y.N.,
                                                                              #ORR=ORR..Y.N..,
                                                                              #Toxicity=Toxicity..Y.N.,
                                                                              Cytokine.Release.Syndrome=`Cytokine.Release.Syndrome.(Y/N)`,
                                                                              #Cytokine.Release.Syndrome..Y.N..,
                                                                              ICANS_neurotoxicity = `ICANS.(Neurotoxicity)(Y/N)`,
                                                                              #ICANS_neurotoxicity = ICANS..Neurotoxicity..Y.N..,
                                                                              Disease.Class,
                                                                              Disease,
                                                                              antibiotic_prior=`Antibiotics.4.weeks.prior.to.CAR.T.cells.(Y/N)`,
                                                                              antibiotic_post=`Antibiotics.4.weeks.post-CAR.T.cells.(Y/N)`,
                                                                              Patient.Disease.Status.at.CAR,
                                                                              CAR.Group)],
                                                       dt_carT_sampleid_info_m[,.(pt_id,
                                                                                  sampleid,
                                                                                  time_group)])

  dt_carT_patient_info_with_sample_and_oligoID[, CR_or_ORR := factor(ifelse(CR=="Y" | ORR=="Y","Y","N"),levels=c("Y","N"))]

  dt_carT_patient_info_with_sample_and_oligoID[,
                                               time_group_simple :=factor(gsub("Sample.ID","",time_group))]

#   dt_carT_patient_info_with_sample_and_oligoID_time_stamp=Sys.time();
#   cat("patient info loaded in variable:\n\tdt_carT_patient_info_with_sample_and_oligoID\n")
# }else{
#   if(as.numeric(difftime(Sys.time(), dt_carT_patient_info_with_sample_and_oligoID_time_stamp, units="min")) > 1 ){
#     rm("dt_carT_patient_info_with_sample_and_oligoID_time_stamp");
#     source('~/projects/human_microbiota/library/Melody/load_carT_patient_info.R');
#   }else{
#     cat("patient info already loaded in variable:\n\tdt_carT_patient_info_with_sample_and_oligoID\n") 
#   }
  cat("patient info loaded in variable:\n\tdt_carT_patient_info_with_sample_and_oligoID\n")

}
```

### upenn

```{r}
#Nov/04/2020
#
#Load UPenn. CART patients clinical data for patients with stool samples.
#
#

if(!exists("dt_cart_stool_upen_samples_and_outcome")){

  dt_cart_stool_upen_clinical = openxlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/Microbiota DB_commercial CART_to send.xlsx",
                                                    sheet= 1)

  dt_cart_stool_upen_clinical = data.table(dt_cart_stool_upen_clinical);

  dt_cart_stool_upen_samples = xlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/Project 11198 Sample Key.xlsx",
                                               sheetIndex = 1);
  dt_cart_stool_upen_samples = data.table(dt_cart_stool_upen_samples)
  dt_cart_stool_upen_samples = dt_cart_stool_upen_samples[!is.na(Subject.ID)]
  dt_cart_stool_upen_samples$NA. = NULL;
  dt_cart_stool_upen_samples$NA..1 = NULL;

  dt_cart_stool_upen_samples_and_outcome = merge(dt_cart_stool_upen_samples[,.(sampleid=OligoID,pt_id=Subject.ID)],
                                                 dt_cart_stool_upen_clinical[,.(pt_id=`pt_id.(de-identified)`,
                                                                                CR=toupper(`CR(Y/N)`),
                                                                                Tox=`Toxicity(Y/N)`)])

}

print("Upen clincal+stool set is loaded in dt_cart_stool_upen_samples_and_outcome")

```

### italy

```{r}
#Oct/14/2020
#
#This scripts load Italian data, organizes it in antibiotics and clinical side. 
#
#

if(!(exists("dt_CART_italy_antibiotics") & exists("dt_CART_italy_clinical") & exists("dt_CART_italy_samples_and_outcome"))){

  dt_CART_italy = xlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/DATABASE CAR T 091020.xlsx",
                                 sheetIndex = 1);
  dt_CART_italy = data.table(dt_CART_italy);
  setnames(dt_CART_italy,"pt_id.de.identified.","pid");
  dt_CART_italy = dt_CART_italy[!is.na(pid)]
  dt_CART_italy[, CR := ifelse(toupper(CR.Y.N.)=="Y",1,
                               ifelse(toupper(CR.Y.N.)=="N",0,
                               NA))]
  dt_CART_italy[, Tox := ifelse(toupper(Toxicity.Y.N.)=="Y",1,
                               ifelse(toupper(Toxicity.Y.N.)=="N",0,
                                      NA))]

  dt_CART_italy_clinical = dt_CART_italy[,infusion_date:=Data.infusione..dd.mm.yy.]

  dt_CART_italy_antibiotics = fread("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/abx_data_john10-19-2020.csv")
  setnames(dt_CART_italy_antibiotics,"P_ID","pid");

  dt_CART_italy_antibiotics[, drug_name := tolower(drug_name)]
  dt_CART_italy_antibiotics[, drug_name := gsub("-","_",drug_name)]
  dt_CART_italy_antibiotics[, drug_name := gsub("/","_",drug_name)]

  dt_CART_italy_antibiotics[, start_date:=as.Date(start_date,"%m/%d/%y")]
  dt_CART_italy_antibiotics[, stop_date:=as.Date(stop_date,"%m/%d/%y")]

  # x=sort(tolower(unique(dt_CART_italy_antibiotics$drug_name)))
  # x=gsub("-","_",x)
  # x=unique(gsub("/","_",x))

  drug_name_clean_map = c("amoxicilina_ac.clavulanico"="amoxicillin",
                          "amoxicillin clavulanate"="amoxicillin",
                          "amoxicillin clavulanate _"="amoxicillin",
                          "augmentin"="amoxicillin",
                          "bactrim"="sulfamethoxazole_trimethoprim",
                          "cefixima"="cefixime",
                          "cefixoral"="cefixime",
                          "ceftazidima_avibactam"="ceftazidime",
                          "ceftriaxone vii somministrazione al"="ceftriaxone",
                          "clarithromycin"="clarithromycin",
                          "fosfomicina"="Fosfomycin",
                          "levofloxacina"="levofloxacin",
                          "teicoplanin"="teicoplanin",
                          "tigeciclina"="tigecycline",
                          "trimethoprim_sulfamethoxazole non_sop prophylaxis since diagnosis"="sulfamethoxazole_trimethoprim",
                          "vancomicina"= "vancomycin.pending")

  for(i in 1:length(drug_name_clean_map)){
    drug_name_cur = names(drug_name_clean_map)[i];
    drug_name_clean_cur = drug_name_clean_map[i];

    dt_CART_italy_antibiotics[drug_name %in% drug_name_cur, drug_name := drug_name_clean_cur];
  }

  #Adding sampleid and pid link here

  dt_CART_italy_samples = fread("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/Set 81 Samples.txt")
  setnames(dt_CART_italy_samples,"pt_id","pid")
  setnames(dt_CART_italy_samples,"SampleID","sampleid")
  setnames(dt_CART_italy_samples,"OligoID","oligoid")

  dt_CART_italy_samples_and_outcome = merge(dt_CART_italy_samples[,.(pid,sampleid,oligoid,timepoint)],
                                            dt_CART_italy_clinical[,.(pid,CR,Tox)],
                                            all.x = T,
                                            by="pid")
  dt_CART_italy_samples_and_outcome[, CR:=.(ifelse(CR==1,"Y","N"))]
  dt_CART_italy_samples_and_outcome[, Tox:=.(ifelse(Tox==1,"Y","N"))]

}

print("dt_CART_italy_clinical, dt_CART_italy_antibiotics and dt_CART_italy_samples_and_outcome are loaded!")

```

## only look at the baseline samples for now (and we also need to have the CR out)

```{r}
# the msk table 
msk <- dt_carT_patient_info_with_sample_and_oligoID %>% 
  filter(time_group_simple == 'Baseline.')

upenn <- dt_cart_stool_upen_samples_and_outcome %>% 
  filter(!str_detect(sampleid, 'blank')) %>%  # two failed to amplify
  filter(!is.na(CR)) # one has not reached the CR record point

italy <- dt_CART_italy_samples_and_outcome %>% 
  filter(timepoint == 'PRE INFUSION') %>% 
  filter(!is.na(CR))  
```

### 2020-11-10 scripts/37_clean_pheno_refreshed.Rmd
update

@@
 
+This table has the CR_d100
+
 ```{r}
@@
 #
-
 library('openxlsx');
 library(data.table);
-
 if(!exists("dt_carT_patient_info_with_sample_and_oligoID_time_stamp")){
@@
 ```
+
+CR:
+day 30: msk and italy
+day 100: upenn
+
+Tox:
+day 30: all of them
+
+## getting the asv counts of those samples
+
+```{r}
+source('~/db_connect_simple.R')
+connect_database(config_file = '~/dbConfig.txt')
+
+get_table_from_database_predefined_filter('asv_counts_ag')
+get_table_from_database('asv_sequences_ag')
+```
+
+```{r}
+msk_cts <- asv_counts_ag %>% 
+  filter(sampleid %in% msk$sampleid) %>% 
+  select(asv_key, sampleid, count) %>% 
+  spread(key = 'sampleid' , value = 'count', fill = 0)
+
+# italy only has 3 in the db
+italy_cts <- asv_counts_ag %>% 
+  filter(sampleid %in% italy$oligoid) %>% 
+  select(asv_key, sampleid, count) %>% 
+  spread(key = 'sampleid' , value = 'count', fill = 0)
+
+upenn_cts <- asv_counts_ag %>% 
+  filter(sampleid %in% upenn$sampleid) %>% 
+  select(asv_key, sampleid, count) %>% 
+  spread(key = 'sampleid' , value = 'count', fill = 0)
+
+# update italy pheno table
+italy <- italy %>% 
+  filter(oligoid %in% asv_counts_ag$sampleid)
+
+```
+
+```{r}
+# write out the asv seq and cts to corresponding files
+library(seqinr)
+cts_fns <- list(msk = msk_cts,
+                italy = italy_cts,
+                upenn = upenn_cts)
+
+cts_fns %>% 
+  imap(function(.x, .y){
+
+    # the cts for the asv
+    .x %>% 
+      write_tsv(str_glue('../data/picrust/input/{.y}_cts.tsv'))
+
+    # the seqs for the asv
+    seqs = asv_sequences_ag %>%
+        filter(asv_key %in% .x$asv_key)
+
+    write.fasta(sequences = as.list(seqs$asv_sequence),
+              names = seqs$key,
+              file.out = str_glue('../data/picrust/input/{.y}_seqs.tsv'))
+  })
+```

### 2020-11-10 scripts/load_MSK_patients.R

@@ -1,0 +1,59 @@
+#Oct/05/2020
+#
+#This scripts load MSK patient data (clinical and antibiotic side),
+#It organizes it in antibiotics and clinical side. 
+#
+#
+
+library(data.table);
+
+
+if(!(exists("dt_CART_MSK_antibiotic") & exists("dt_CART_MSK_clinical"))){
+
+  
+  # dt_carT_sampleid_info = openxlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/CTC Event Table01132020.xlsx",#CTC Event Table 11252019 wo Censor .xlsx",
+  #                                             sheet = "Sample IDs only ");
+  
+  #dt_CART_MSK_clinical = openxlsx::read.xlsx("/Volumes/Attending queries/Smith. M/CAR Intestinal Microbiome (Jain; 16-834)/CAR Microbiome_2020Jan25v2.xlsx",sheet=1);
+  dt_CART_MSK_clinical = openxlsx::read.xlsx("/Volumes/Attending queries/Smith. M/CAR Intestinal Microbiome (Jain; 16-834)/CAR Microbiome_2020Oct22.xlsx",sheet=1);
+  dt_CART_MSK_clinical = data.table(dt_CART_MSK_clinical)
+  dt_CART_MSK_clinical$CR = ifelse(dt_CART_MSK_clinical$Day.30.Response=="CR",1,0)#"Y","N");
+  dt_CART_MSK_clinical$CR_d100 = ifelse(dt_CART_MSK_clinical$Day.100.Response=="CR",1,0)#"Y","N");
+  dt_CART_MSK_clinical$Tox_Neuro_and_CRS = ifelse(dt_CART_MSK_clinical$Neuro.tox=="Y" | dt_CART_MSK_clinical$Any.Grade.CRS=="Y",1,0)#"Y","N");
+  dt_CART_MSK_clinical$Tox_CRS = ifelse(dt_CART_MSK_clinical$Any.Grade.CRS=="Y" | dt_CART_MSK_clinical$Any.Grade.CRS=="Y",1,0)#"Y","N");
+  
+  
+  dt_CART_MSK_clinical[, infusion_date:= as.Date((IEC.Date),format = "%Y-%m-%d",origin="1899-12-30")]
+  
+
[…]

### 2020-11-10 scripts/37_clean_pheno_refreshed.Rmd
update

@@
 
+## write out the asv counts and seqs for picrust2
+
 ```{r}
 # write out the asv seq and cts to corresponding files
+# ITS NOT FILTERED BUT TO ALIGN WITH THE SNAKEFILE FILE NAMES
 library(seqinr)
@@
     .x %>% 
-      write_tsv(str_glue('../data/picrust/input/{.y}_cts.tsv'))
+      write_tsv(str_glue('../data/picrust/input/filtered_counts_{.y}.tsv'))
 
@@
     write.fasta(sequences = as.list(seqs$asv_sequence),
-              names = seqs$key,
-              file.out = str_glue('../data/picrust/input/{.y}_seqs.tsv'))
+              names = seqs$asv_key,
+              file.out = str_glue('../data/picrust/input/filtered_counts_asv_sequence_{.y}.tsv'))
   })
 ```
+
+## load the downloaded results from picrust2
+
+```{r}
+src <- tibble(
+  id = c('upenn','msk','italy'),
+  local_dir = str_glue('/Users/daia1/projects/CART_microbiome/data/picrust/output/{id}'),
+  pw_file = str_glue('daia1@lilac.mskcc.org:~/my_workdir/other_pipeline/picrust2/output/picrust2_out_{id}/pathways_out/path_abun_unstrat.tsv.gz'),
+  dl_pw = str_glue('scp {pw_file} {local_dir}'),
+  unzip_pw = str_glue('gunzip /Users/daia1/projects/CART_microbiome/data/picrust/output/{id}/path_abun_unstrat.tsv.gz')
+)
+# create local dir first
+src %>% 
+  mutate(mkdir_cmd = str_glue('mkdir {local_dir}')) %>% 
+  select(mkdir_cmd) %>% 
+  write_tsv('../data/mkdir.sh', col_names = F)
+
+src %>% 
+  select(dl_pw,  unzip_pw) %>% 
+  gather('fn', 'cmd') %>% 
+  select(cmd) %>% 
+  write_tsv('../data/downloding_pw.sh', col_names = F)
+```
+
+## comparison together and individually
+
+Lump italy and msk together since there are only three baseline samples from italy, and also italy and msk observe CR at day 30.
+
+```{r}
+pw_msk <- read_tsv('../data/picrust/output/msk/path_abun_unstrat.tsv')
+pw_italy <- read_tsv('../data/picrust/output/italy/path_abun_unstrat.tsv')
+pw_upenn <- read_tsv('../data/picrust/output/upenn/path_abun_unstrat.tsv')
+
+# msk and italy
+pheno_mitaly <- bind_rows(
+  msk %>% 
+    select(sampleid, CR, Toxicity),
+  italy %>% 
+    select(sampleid = oligoid,
+           CR,
+           Toxicity = Tox)
+)
+
+pw_mitaly <- pw_msk %>% 
+  inner_join(pw_italy)
+
+pw_mitaly_or <- bind_cols(pw_mitaly[,1],pw_mitaly[,pheno_mitaly$sampleid])
+
+all.equal(colnames(pw_mitaly_or)[2:ncol(pw_mitaly_or)], pheno_mitaly$sampleid)
+
+# upenn
+pheno_upenn <- upenn %>% 
+  mutate(sampleid = as.character(sampleid)) %>% 
+  select(sampleid, CR, Toxicity = Tox)
+
+pw_upenn_or <- bind_cols(pw_upenn[,1],pw_upenn[,pheno_upenn$sampleid])
+
+all.equal(colnames(pw_upenn_or)[2:ncol(pw_upenn_or)], pheno_upenn$sampleid)
+
+# all of them
+pheno_all <- bind_rows(
+  pheno_mitaly,
+  pheno_upenn
+)
+
+pw_all <- pw_mitaly %>% 
+  inner_join(pw_upenn)
+
+pw_all_or <- bind_cols(pw_all[,1],pw_all[,pheno_all$sampleid])
+
+all.equal(colnames(pw_all_or)[2:ncol(pw_all_or)], pheno_all$sampleid)
+```
+```{r}
+pheno <- list(mitaly = pheno_mitaly,
+              upenn = pheno_upenn,
+              all = pheno_all)
+
+pheno %>% 
+  imap(~ select(.data = .x, CR) %>% 
+            t %>% 
+            write.table(str_glue('../data/lefse/{.y}__pheno_CR.txt'), sep = '\t', quote = F, row.names = T, col.names = F))
+
+pheno %>% 
+  imap(~ select(.data = .x, Toxicity) %>% 
+            t %>% 
+            write.table(str_glue('../data/lefse/{.y}__pheno_Toxicity.txt'), sep = '\t', quote = F, row.names = T, col.names = F))
+
+cts_or <- list(
+              mitaly = pw_mitaly_or,
+              upenn = pw_upenn_or,
+              all = pw_all_or
+)
+
+cts_or %>% 
+  imap(~ .x %>% write_tsv(str_glue('../data/lefse/{.y}_cts.tsv')))
+
+```
+
+```{r}
+# a script to bind those things and produce the correct format for the galaxy portal
+cts <- list.files('../data/lefse/', pattern = 'cts')
+pheno <- list.files('../data/lefse/', pattern = 'pheno')
+combi <- expand.grid(cts_ = cts,pheno_ = pheno) %>% 
+  mutate(cts_id = str_replace(cts_, '_cts.tsv','')) %>% 
+  mutate(pheno_id = str_replace(pheno_, '__.+$','')) %>% 
+  filter(cts_id == pheno_id) %>% 
+  mutate(grp = if_else(str_detect(pheno_, 'CR'), 'CR', 'tox')) %>% 
+  mutate(cmd = str_glue('cat {pheno_} {cts_} > {pheno_id}_{grp}.tsv'))
+
+combi %>% select(cmd) %>% 
+  write_tsv('../data/lefse/cat_table.sh', col_names = F)
+
+```

### 2020-11-12 scripts/37_clean_pheno_refreshed.Rmd
jesus how much time I spent on this

@@
 library(tidyverse)
+library(data.table)
 ```
@@
 
-  dt_CART_italy_samples = fread("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/Set 81 Samples.txt")
+  dt_CART_italy_samples = fread("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/samples/opbg_sampleskey_11092020_with_oligoid.csv")
   setnames(dt_CART_italy_samples,"pt_id","pid")
-  setnames(dt_CART_italy_samples,"SampleID","sampleid")
-  setnames(dt_CART_italy_samples,"OligoID","oligoid")
+  #setnames(dt_CART_italy_samples,"SampleID","sampleid")
+  #setnames(dt_CART_italy_samples,"OligoID","oligoid")
 
@@
 all.equal(colnames(pw_all_or)[2:ncol(pw_all_or)], pheno_all$sampleid)
+
+# only msk data
+pheno_msk <-   msk %>% 
+    select(sampleid, CR, Toxicity)
+
+pheno_msk %>% 
+  select( Toxicity) %>% 
+            t %>% 
+            write.table(str_glue('../data/lefse/msk__pheno_Toxicity.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
+
+pw_msk_or <- bind_cols(pw_msk[,1],pw_msk[,pheno_msk$sampleid]) 
+all.equal(colnames(pw_msk_or)[2:ncol(pw_msk_or)], pheno_msk$sampleid)
+
+pw_msk_or %>% 
+  write_tsv(str_glue('../data/lefse/msk_cts.tsv'))
 ```

### 2020-11-12 scripts/38_link_shotgun_fid_with_pheno.Rmd
jesus how much time I spent on this

---
title: "link the shotgun fid and the actual sample and outcome table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

Actually the most important info with the sample: timepoint (so I know whether it is baseline) and the CR , Tox

```{r}
# the full italy samples and outcome joined from Antonio's two tables
italy_samp_outcome <- dt_CART_italy_samples %>% 
  full_join(dt_CART_italy_samples_and_outcome) %>% 
  distinct()

italy_samp_outcome %>% 
  write_csv('../data/shotgun_input/shotgun_italy_samp_outcome.csv')

# italy shotgun files
italy_fn <- read_csv('../../Catalog/data/full_project_11308_files.csv') %>% 
  distinct(sample_folder, .keep_all = T)

# there are 4 dups, and need to be concatenated to run
```

```{r}
samp_italy <- italy_fn %>% 
  left_join(italy_samp_outcome %>% mutate(fid = str_glue('{pid}{timepoint}')))

found1 <- samp_italy %>% 
  filter(!is.na(sampleid))

found2_df <- samp_italy %>% 
  filter(is.na(sampleid)) %>% 
  select(directory:fid) %>% 
  left_join(  italy_samp_outcome %>% 
    mutate(fid = str_replace_all(sampleid_raw, '\\+| |\\(|\\)|\\*','')  )
) %>% 
  distinct(fid, .keep_all = T) # there are two 16s samples CMT006 d172 and d168 both record as 180*

found2 <- found2_df %>% 
  filter(!is.na(sampleid))

found3_df <- found2_df %>% 
  filter(is.na(sampleid)) %>% 
  select(directory:fid) %>% 
  left_join(italy_samp_outcome %>% 
  filter(pid == 'CMT006') %>% 
  mutate(fid = str_replace_all(sampleid_raw, '\\+|\\(|\\)|\\*','')) %>% 
  filter(timepoint %in% c(90, 14))  %>% 
  mutate(fid = str_replace(fid, ' ','-'))) %>% 
  # there is one sample CMT006 that missed the datepoint that is actually day -3
  mutate(timepoint = if_else(fid == 'CMT006', -3.0, as.double(timepoint)),
         CR = if_else(fid == 'CMT006', 'Y', CR),
         Tox = if_else(fid == 'CMT006', 'N', Tox))

```

```{r}
# finally all of them matched ...

italy_pheno <- bind_rows(
  found1,
  found2,
  found3_df
) %>% 
  select(-sample_folder)

italy_pheno %>% 
  write_csv('../data/shotgun_pheno/italy_pheno_earlier.csv')
```

There are sevreral NA in the CR and tox in the baseline samples

Can you please clarify what “-B” when it is noted on the patient ID (pid)?
The “-B” for Patient ID CMT001B (sample ID “CMT001 PRE-BMT”) refers to a subsequent infusion for patient CMT001. Patient ID CMT001 after 42 days from his CAR T infusion received a BMT and it was recorded as Patient ID CMT001B and sampleID CMT001 PRE-BMT . Do not include Patient ID CMT001B and sampleID CMT001 PRE-BMT in the analysis

Can you please note once more the patients that should be excluded?
Exclude the patients in sponsored trial: CMT009; CMT011N; CMT013-B
Exclude the patients CMT001B for the reason at point “1”

so I think, CMT009; and ; CMT013-B ignore
what about CMT012 and CMT004B

### 2020-11-12 scripts/load_italian_patients.R

@@ -1,0 +1,79 @@
+#Oct/14/2020
+#
+#This scripts load Italian data, organizes it in antibiotics and clinical side. 
+#
+#
+
+if(!(exists("dt_CART_italy_antibiotics") & exists("dt_CART_italy_clinical") & exists("dt_CART_italy_samples_and_outcome"))){
+
+
+  dt_CART_italy = xlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/DATABASE CAR T 091020.xlsx",
+                                 sheetIndex = 1);
+  dt_CART_italy = data.table(dt_CART_italy);
+  setnames(dt_CART_italy,"pt_id.de.identified.","pid");
+  dt_CART_italy = dt_CART_italy[!is.na(pid)]
+  dt_CART_italy[, CR := ifelse(toupper(CR.Y.N.)=="Y",1,
+                               ifelse(toupper(CR.Y.N.)=="N",0,
+                               NA))]
+  dt_CART_italy[, Tox := ifelse(toupper(Toxicity.Y.N.)=="Y",1,
+                               ifelse(toupper(Toxicity.Y.N.)=="N",0,
+                                      NA))]
+  
+  dt_CART_italy_clinical = dt_CART_italy[,infusion_date:=Data.infusione..dd.mm.yy.]
+  
+  dt_CART_italy_antibiotics = fread("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/abx_data_john10-19-2020.csv")
+  setnames(dt_CART_italy_antibiotics,"P_ID","pid");
+  
+  dt_CART_italy_antibiotics[, drug_name := tolower(drug_name)]
+  dt_CART_italy_antibiotics[, drug_name := gsub("-","_",drug_name)]
+  dt_CART_italy_antibiotics[, drug_name := gsub("/","_",drug_name)]
+  
+  dt_CART_italy_antibiotics[, start_date:=as.Date(start_date,"%m/%d/%y")]
+  dt_CART_italy_antibiotics[, stop_date:=as.Date(stop_date,"%m/%d/%y")]
+  
+  # x=sort(tolower(unique(dt_CART_italy_antibiotics$drug_name)))
+  # x=gsub("-","_",x)
+  # x=unique(gsub("/","_",x))
+  
+  drug_name_clean_map = c("amoxicilina_ac.clavulanico"="amoxicillin",
+                          "amoxicillin clavulanate"="amoxicillin",
[…]

### 2020-11-24 scripts/37_clean_pheno_refreshed.Rmd
was very upset about the messiness but luckily John is such a reliable guy

@@
 
-  dt_cart_stool_upen_samples = xlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/Project 11198 Sample Key.xlsx",
+  dt_cart_stool_upen_samples = xlsx::read.xlsx("/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/samples/archive/Project 11198 Sample Key.xlsx",
                                                sheetIndex = 1);

### 2020-11-24 scripts/38_link_shotgun_fid_with_pheno.Rmd
was very upset about the messiness but luckily John is such a reliable guy

@@
 
+## current shotgun lookup table
+
+```{r}
+source('~/db_connect_simple.R')
+connect_database(config_file = '~/dbConfig.txt')
+get_table_from_database('shotgun_lookup_ad')
+
+list_table_from_database('bile')
+```
+
 Actually the most important info with the sample: timepoint (so I know whether it is baseline) and the CR , Tox
+
+## Italy data
 
@@
 
-There are sevreral NA in the CR and tox in the baseline samples
+```{r}
+# fix some more problems in the table including the duplicated sampleids
 
-Can you please clarify what “-B” when it is noted on the patient ID (pid)?
-The “-B” for Patient ID CMT001B (sample ID “CMT001 PRE-BMT”) refers to a subsequent infusion for patient CMT001.
-Patient ID CMT001 after 42 days from his CAR T infusion received a BMT and it was recorded as Patient ID CMT001B and sampleID CMT001 PRE-BMT .
-Do not include Patient ID CMT001B and sampleID CMT001 PRE-BMT in the analysis
+# now John has cleaned table and added a tab about ineligible patients 
+ineligible <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/Pt info/italy_essential_data_v4.xlsx', sheet = 2)
 
-Can you please note once more the patients that should be excluded?
-Exclude the patients in sponsored trial: CMT009; CMT011N; CMT013-B
-Exclude the patients CMT001B for the reason at point “1”
+italy <- read_csv('../data/shotgun_pheno/italy_pheno_earlier.csv') %>% 
+  # remove the row with many NA since it is a duplicate of CMT006-dm3
+  filter(!is.na(sampleid)) %>% 
+  # remove the ones that are ineligible
+  filter(!pid %in% ineligible$pt_id) %>% 
+  select(directory:fid, pid:sampleid, oligoid)
 
-so I think, CMT009; and ; CMT013-B ignore
-what about CMT012 and CMT004B
+# now load the outcome for each eligible patient
+eligible <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/Pt info/italy_essential_data_v4.xlsx', sheet = 1) %>% 
+  rename(pid = pt_id)
+
+italy_all <- italy %>% 
+  left_join(eligible) %>% 
+  filter(!is.na(disease)) # remove the ones that have NA since they are not in the eligible patients tab either
+
+# the sub table to be updated to shotgun lookup table
+italy_up <- italy_all %>% 
+  select(directory, projectID, sampleid, fid)
+
+```
+
+```{r}
+
+df %>% 
+  write_csv('/Volumes/vandenbrinklab/Angel_Dai/Full_human_shotgun_catalog/shotgun_italy_earlier.csv')
+```
+
+## Upenn data
+
+```{r}
+
+```
+
+## Update the new shotgun samples to the database table
+
+```{r}
+source('../../../microbiome_db/SQL/scripts/upload_data_from_query_OTU_AD.R')
+upload_data_from_query_OTU(1) # changed the file path of the table getting uploaded 
+```

### 2020-11-24 scripts/38_link_shotgun_fid_with_pheno.Rmd
yeah finally did this .... and now AG can help me upload

@@
 
+italy_all %>% 
+  write_csv('../data/shotgun_pheno/italy_all_earlier.csv')
+
 # the sub table to be updated to shotgun lookup table
@@
 
-```{r}
-
-df %>% 
-  write_csv('/Volumes/vandenbrinklab/Angel_Dai/Full_human_shotgun_catalog/shotgun_italy_earlier.csv')
-```
-
 ## Upenn data
@@
 ```{r}
+# get the file path of the shotgun files 
+upenn <- c('Project_11341')
+upenn2 <- tibble(
+  folder = '/Volumes/vandenBrinkLab/deep_sequencing'
+)
 
+upenn2_files <- upenn2 %>% 
+  pull(folder) %>% 
+  map_dfr(function(f){
+    upenn %>% 
+      set_names(upenn) %>% 
+      imap_dfr( ~ Sys.glob(str_glue('{f}/{..1}/*/*')) %>% 
+             data_frame %>% 
+              mutate(..2)) %>% 
+      rename(directory = names(.)[1],
+             projectID = names(.)[2])
+  }) %>% 
+  mutate(sample_folder = str_extract(directory, 'Sample_.+$'),
+         fid = str_replace(sample_folder, 'Sample_',''),
+         fid = str_replace(fid, '_IGO_.+$','')) %>% 
+  mutate(pid = str_replace(fid, 'MS_UPENN_','')) %>% 
+  mutate(pid = str_pad(pid, 2, side = 'left', pad = '0'))
+
+# the pheno of the ptb from John
+upenn <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 2) %>% 
+  filter(!is.na(pt_id)) %>% 
+  mutate(pt_id = as.character(pt_id)) %>% 
+  mutate(pid = str_replace(pt_id, '^80',''))
+
+upenn_samp <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/samples/sample_mapping_UPenn.xlsx') %>% 
+  select(pt_id, OligoID) %>% 
+  mutate(pt_id = as.character(pt_id))
+
+upenn_all <- upenn2_files %>% 
+  left_join(upenn) %>% 
+  full_join(upenn_samp)
+
+upenn_all %>% 
+  write_csv('../data/shotgun_pheno/upenn_all_earlier.csv')
+
+upenn_up <- upenn_all %>% 
+  select(directory, projectID, OligoID, fid) %>% 
+  rename(sampleid = OligoID)
+```
+
+```{r}
+# 11-24 to be uploaded 
+nov24_up <- bind_rows(upenn_up, italy_up)
+
+nov24_up %>% 
+  write_csv('/Volumes/vandenbrinklab/Angel_Dai/Full_human_shotgun_catalog/shotgun_cart_earlier.csv')
 ```
@@
 
+sometimes it doesn't work you just need to restart R and run it again
+
 ```{r}
+# change the d_set_input in the upload_data_from_query_OTU_AD.R script 
 source('../../../microbiome_db/SQL/scripts/upload_data_from_query_OTU_AD.R')

### 2020-12-02 scripts/39_new_MSK_shotgun.Rmd
udpate

---
title: "New MSK samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# check the sequencing files returned, how the fid is matched to sampleid, is there any duplicates etc
proj <- c('Project_11454') # there are three failed QC so this is the correct number

folder <- '/Volumes/vandenBrinkLab/deep_sequencing'

files <- folder %>% 
  map_dfr(function(f){
    proj %>% 
      set_names(proj) %>% 
      imap_dfr( ~ Sys.glob(str_glue('{f}/{..1}/*/*')) %>% 
             as.data.frame %>% 
              mutate(..2)) %>% 
      rename(directory = names(.)[1],
             projectID = names(.)[2])
  }) %>% 
  mutate(sample_folder = str_extract(directory, 'Sample_.+$'),
         fid = str_replace(sample_folder, 'Sample_',''),
         fid = str_replace(fid, '_IGO_.+$','')) %>% 
  mutate(sampleid = fid) # the fid looks clean should be the sampleid

files %>% 
  distinct(sampleid) %>% 
  nrow
```

```{r}
# check the sample patient pheno table organized by John
# the first sheet is the eligible samples
stb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 1)

length(intersect(stb$sampleid, files$sampleid))

files %>% 
  write_csv('../data/shotgun_input/Project_11454_shotgun_samples.csv')
```

### 2020-12-07 scripts/39_new_MSK_shotgun.Rmd
update
good i finally figured out the pheno for the msk patients ...

@@
   nrow
+
+files %>% 
+  write_csv('../data/shotgun_input/Project_11454_shotgun_samples.csv')
+
+# although 20 returned but few can be used, the non-useable ones are from ineligible pts
 ```
@@
 # the first sheet is the eligible samples
+# the below table has the most updated info organized by John
 stb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 1) 
+ineligible <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 2)
 
-length(intersect(stb$sampleid, files$sampleid))
+pheno <- read_csv('../data/cleaned_pheno/final_pheno.csv') # the 33 used last year
+intersect(pheno$sampleid, ineligible$sampleid)  # 6 has become ineligible 
+```
 
-files %>% 
-  write_csv('../data/shotgun_input/Project_11454_shotgun_samples.csv')
+## all MSK eligible baseline shotgun samples
+
+2020-12-7   32 total for eligible samples in the baseline time point from MSK patients
+
+```{r}
+current_Baseline <- stb %>% 
+  inner_join(shotgun_lookup_ad) %>% 
+  filter(str_detect(time_group, '^Baseline')) %>% 
+  distinct(sampleid, .keep_all = T)
+
+ptb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx') %>% 
+  mutate(mrn = as.numeric(mrn))
+
+pheno_msk <- current_Baseline %>% 
+  select(mrn:sampleid, fid) %>% 
+  left_join(ptb)
+
+pheno_msk %>% 
+  write_csv('../data/shotgun_pheno/msk_all_pheno.csv')
 ```

### 2020-12-07 scripts/39_new_MSK_shotgun.Rmd
update

@@
 intersect(pheno$sampleid, ineligible$sampleid)  # 6 has become ineligible
+
+setdiff(current_Baseline$sampleid, pheno$sampleid)
+
+missing <- c("2460A", "2563A", "1918C" ,"315"  )
 ```
@@
 pheno_msk <- current_Baseline %>% 
-  select(mrn:sampleid, fid) %>% 
+  select(mrn:sampleid, fid, directory) %>% 
   left_join(ptb)
@@
   write_csv('../data/shotgun_pheno/msk_all_pheno.csv')
+
+pheno_msk %>% 
+  filter(sampleid %in% missing)
 ```

### 2020-12-07 scripts/40_combine_all_samp_pheno.Rmd
update

---
title: "combine all centers pheno"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
# all centers pheno file 
# to get the baseline sampleid 
I <- read_csv('../data/shotgun_pheno/italy_all_earlier.csv') %>% 
  filter(timepoint < 0) %>% 
  ungroup() %>% 
  select(fid)

P <- read_csv('../data/shotgun_pheno/upenn_all_earlier.csv') %>% select(fid)

M <- read_csv('../data/shotgun_pheno/msk_all_pheno.csv') %>% select(fid)

all_baseline <- bind_rows(I, M, P)
```

### 2020-12-08 scripts/41_new_I_E_shotgun.Rmd
update

---
title: "NEW Italy and Upenn samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
source('~/db_connect_simple.R')
connect_database(config_file = '~/dbConfig.txt')
get_table_from_database('shotgun_lookup_ad')
```

## Italy ones

```{r}
ineligible <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/samples/opbg_sampleskey_with_oligoid_v2.xlsx', sheet = 2)

fntb_I <- shotgun_lookup_ad %>% 
  filter(projectid == 'Project_11453' & str_detect(fid, "^CMT")) %>% 
  # first see if that patient is eligible or not 
  mutate(pid = str_extract(fid, 'CMT...')) %>% 
  filter(! pid %in% ineligible$pt_id)  %>% 
  filter(!str_detect(fid, '_$')) %>% # the dups
  separate(fid, into = c('pre','day'), remove = F, convert = T, sep = '_') %>% 
  filter(is.na(day)) %>% # remove the ones I'm sure to be after infusion
  filter(fid != 'CMT029G0') %>%  # CMT029G0 is actually day 1 sample
  select(-pre, -day)
```
```{r}
fntb_P <- shotgun_lookup_ad %>%
  filter(projectid == 'Project_11453' & !str_detect(fid, "^CMT"))
```

### 2020-12-12 scripts/40_combine_all_samp_pheno.Rmd
aaaaahh i need to give a lab meeting presentation again on Dec 21
update
good Im making progess and continue sorting things out

@@
 
+I2 <- read_csv('../data/shotgun_pheno/italy_all_later.csv') %>% 
+  select(fid)
+
 P <- read_csv('../data/shotgun_pheno/upenn_all_earlier.csv') %>% select(fid)
+
+P2 <- read_csv('../data/shotgun_pheno/upenn_all_later.csv') %>% 
+  select(fid)
 
@@
 
-all_baseline <- bind_rows(I, M, P)
+all_baseline <- bind_rows(I,I2, M, P, P2)
+
+all_baseline %>% 
+  write_csv('../data/shotgun_input/samples_cart_all_msk_patial_italy_upenn.csv', col_names = F)
 ```
+
+## the pheno label for the samples
+
+```{r}
+I1 <- read_csv('../data/shotgun_pheno/italy_all_earlier.csv') %>% 
+  filter(timepoint < 0) %>% 
+  ungroup() %>% 
+  select(fid, cr_d30:toxicity)
+
+# the pheno lable for the I2!
+iptb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/patients/italy_essential_data_v4.xlsx', sheet = 1)
+I2_pheno <- I2 %>% 
+  rename(pt_id = pid) %>% 
+  left_join(iptb) %>% 
+  select(fid, cr_d30:toxicity)
+
+I_phe <- bind_rows(
+  I1, I2_pheno
+) %>% 
+  mutate(center = 'I')
+
+# upenn
+P_ptb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 1)
+
+P_ptb_ineligible <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 2)
+
+P_ <- bind_rows(
+  P %>% 
+    mutate(pt_id = str_replace(fid, 'MS_UPENN_','')) %>% select(fid, pt_id) %>% mutate(pt_id = as.numeric(pt_id)),
+  P2 %>% 
+    mutate(pt_id = as.numeric(fid)) %>% select(fid, pt_id)
+) %>% 
+  filter(! pt_id %in% P_ptb_ineligible$pt_id)
+
+P_phe <- P_ %>% 
+  left_join(P_ptb)  %>% 
+  select(fid, cr_d30:toxicity) %>% 
+  select(fid, cr_d30, cr_d100, cr_overall, toxicity)  %>% 
+  mutate(center = 'P')
+
+# MSK
+M_phe <- M %>% 
+  select(fid, cr_d30:toxicity) %>% 
+  mutate(center = 'M')
+
+# combined
+
+all <- bind_rows(
+  M_phe, P_phe, I_phe
+)
+
+# a tally table
+all %>% 
+  count(center, toxicity)
+
+all %>% 
+  count(center, cr_d30)
+```

### 2020-12-12 scripts/41_new_I_E_shotgun.Rmd
good Im making progess and continue sorting things out

@@
 ```
+
+```{r}
+fntb_I %>% 
+  write_csv('../data/shotgun_pheno/italy_all_later.csv')
+```
+
+## Upenn ones
+
 ```{r}
@@
 ```
+
+```{r}
+fntb_P %>% 
+  write_csv('../data/shotgun_pheno/upenn_all_later.csv')
+```

### 2020-12-12 scripts/42_humann3.Rmd
good Im making progess and continue sorting things out

---
title: "humann2 VS humann3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
hm3 <- list.files('../data/log/', full.names = T, pattern = '_humann3.log$')

hm3_metric <- hm3 %>% 
  set_names(hm3) %>% 
  purrr::map(~ read_table(., col_names = F, col_types = 'c') %>% 
  rename(content = names(.)[1]) %>% 
  filter(str_detect(content, 'INFO: Unaligned')) %>% 
  transmute(hm_perc = str_extract(content, 'after.+$')) %>% 
  mutate(hm_perc = str_replace(hm_perc, 'after ','')) %>% 
  separate(hm_perc, into = c('type','perc'), sep = ': ') %>% 
  mutate(type = str_replace(type, ' ','_'),
         perc = as.numeric(str_replace(perc, ' %',''))) %>% 
  spread(key = 'type', value = 'perc')) %>% 
  bind_rows(.id = 'sampleid') %>% 
  # filter out the rows that have NA in translated column
  filter(!is.na(translated_alignment)) %>% 
  arrange(nucleotide_alignment) %>% 
  mutate(grp = 'hm3')

```

```{r}
hm2 <- hm3_metric %>% 
  mutate(sampleid = str_replace(sampleid, 'humann3','humann2')) %>% 
  pull(sampleid)

hm2_metric <- hm2 %>% 
  set_names(hm2) %>% 
  purrr::map(~ read_table(., col_names = F, col_types = 'c') %>% 
  rename(content = names(.)[1]) %>% 
  filter(str_detect(content, 'INFO: Unaligned')) %>% 
  transmute(hm_perc = str_extract(content, 'after.+$')) %>% 
  mutate(hm_perc = str_replace(hm_perc, 'after ','')) %>% 
  separate(hm_perc, into = c('type','perc'), sep = ': ') %>% 
  mutate(type = str_replace(type, ' ','_'),
         perc = as.numeric(str_replace(perc, ' %',''))) %>% 
  spread(key = 'type', value = 'perc')) %>% 
  bind_rows(.id = 'sampleid') %>% 
  # filter out the rows that have NA in translated column
  filter(!is.na(translated_alignment)) %>% 
  arrange(nucleotide_alignment) %>% 
  mutate(grp = 'hm2')
```

```{r}
all <- bind_rows(hm2_metric, hm3_metric) %>% 
  gather('type', 'value', nucleotide_alignment:translated_alignment) %>% 
  mutate(samples = str_replace(sampleid, '_humann.+$','')) %>% 
  mutate(samples = as.factor(samples)) %>% 
  mutate(perc_aligned = 100-value)

all %>% 
  ggboxplot(x = 'grp', y = 'perc_aligned', color = 'grp', facet.by = 'type', add = 'jitter', line.color = "gray", palette = 'lancet')

```

### 2020-12-12 scripts/43_comparisons_3.Rmd
going slowly but moving forward....

---
title: "Comparisons in three center"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

USING THE HUMANN2 RESULTS FOR NOW
SUBSTITUTE WITH HUMANN3 RESULTS WHEN I CAN

```{r}
pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
  gather('pheno', 'value', cr_d30:toxicity)

sub_phenos <- pheno %>% 
  split(., list(.$center, .$pheno)) %>% 
  purrr::imap(~ filter(.data = ., value != 'not_assessed'))
```

```{r}
sub_phenos %>% 
  imap(function(.x, .y){
    select(.data = .x, value) %>% 
      t() %>% 
      write.table(str_glue('../data/shotgun_output/humann2/{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
  })
```

### 2020-12-13 scripts/43_comparisons_3.Rmd
jesus christ making tremendous advancements here

@@
 ```{r}
+# the pheno for the different center and tox/CR_d30
 sub_phenos %>% 
@@
 ```
+
+```{r}
+# the pathway counts table
+full <- read_tsv('../data/shotgun_output/humann2/humann2_pathabundance_CART.tsv') %>% 
+  select(!ends_with('_1'), MS_UPENN_1)
+
+pcts <- sub_phenos %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ full %>% select(`# Pathway`, matches(.x))) 
+```

### 2020-12-13 scripts/43_comparisons_3.Rmd
udpate

@@
   map(~ pull(.data = ., fid) ) %>% 
-  imap(~ full %>% select(`# Pathway`, matches(.x))) 
+  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann2/{.y}_pcts.tsv'))) 
 ```
+
+```{r}
+fns <- list.files('../data/shotgun_output/humann2/', pattern = '_lefse_ready.tsv')
+
+cmds <- tibble(
+  fns = fns
+) %>% 
+  transmute(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
+  gather() %>% 
+  select(value) %>% 
+  write_csv('../data/shotgun_output/humann2/lefse_cmd.sh', col_names = F)
+cmds
+
+```

### 2020-12-13 scripts/lefse_cmd.sh

@@ -1,0 +1,18 @@
+format_input.py I.cr_d30_lefse_ready.tsv  I.cr_d30_lefse_ready.tsv.in -c 1 -u 2 -o 1000000
+format_input.py I.toxicity_lefse_ready.tsv  I.toxicity_lefse_ready.tsv.in -c 1 -u 2 -o 1000000
+format_input.py M.cr_d30_lefse_ready.tsv  M.cr_d30_lefse_ready.tsv.in -c 1 -u 2 -o 1000000
+format_input.py M.toxicity_lefse_ready.tsv  M.toxicity_lefse_ready.tsv.in -c 1 -u 2 -o 1000000
+format_input.py P.cr_d30_lefse_ready.tsv  P.cr_d30_lefse_ready.tsv.in -c 1 -u 2 -o 1000000
+format_input.py P.toxicity_lefse_ready.tsv  P.toxicity_lefse_ready.tsv.in -c 1 -u 2 -o 1000000
+run_lefse.py I.cr_d30_lefse_ready.tsv.in  I.cr_d30_lefse_ready.tsv.res
+run_lefse.py I.toxicity_lefse_ready.tsv.in  I.toxicity_lefse_ready.tsv.res
+run_lefse.py M.cr_d30_lefse_ready.tsv.in  M.cr_d30_lefse_ready.tsv.res
+run_lefse.py M.toxicity_lefse_ready.tsv.in  M.toxicity_lefse_ready.tsv.res
+run_lefse.py P.cr_d30_lefse_ready.tsv.in  P.cr_d30_lefse_ready.tsv.res
+run_lefse.py P.toxicity_lefse_ready.tsv.in  P.toxicity_lefse_ready.tsv.res
+plot_res.py I.cr_d30_lefse_ready.tsv.res I.cr_d30_lefse_ready.tsv.png  --feature_font_size 4 --width 10 --dpi 300
+plot_res.py I.toxicity_lefse_ready.tsv.res I.toxicity_lefse_ready.tsv.png  --feature_font_size 4 --width 10 --dpi 300
+plot_res.py M.cr_d30_lefse_ready.tsv.res M.cr_d30_lefse_ready.tsv.png  --feature_font_size 4 --width 10 --dpi 300
+plot_res.py M.toxicity_lefse_ready.tsv.res M.toxicity_lefse_ready.tsv.png  --feature_font_size 4 --width 10 --dpi 300
[…]

### 2020-12-13 scripts/43_comparisons_3.Rmd
seriously wanna die

@@
 ```
+
+## the genefamilies regrouping
+
+```{r}
+fullgenefam <- read_tsv('../data/shotgun_output/humann2/humann2_genefamilies_CART.tsv') %>% 
+  select(!ends_with('_1'), MS_UPENN_1)
+
+gfcts <- sub_phenos %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann2/{.y}_gfcts.tsv'))) 
+```
+
+```{r}
+fns_ko <- list.files('../data/shotgun_output/humann2/', pattern = '_lefse_ready_ko.tsv')
+
+cmds_ko <- tibble(
+  fns = fns_ko
+) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
+  select(-fns) %>% 
+  gather() %>% 
+  select(value) %>% 
+  write_csv('../data/shotgun_output/humann2/lefse_cmd_ko.sh', col_names = F)
+```

### 2020-12-13 scripts/43_comparisons_3.Rmd
update

@@
 ```
+
+## run both of them with all of the samples pulled together
+
+```{r}
+all_sub_pheno <- pheno %>% 
+  split(., list(.$pheno)) %>% 
+  purrr::imap(~ filter(.data = ., value != 'not_assessed'))
+
+all_sub_pheno %>% 
+  imap(function(.x, .y){
+    select(.data = .x, value) %>% 
+      t() %>% 
+      write.table(str_glue('../data/shotgun_output/humann2/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
+  })
+
+all_pcts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann2/pull_{.y}_pcts.tsv')))
+
+all_gfcts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann2/pull_{.y}_gfcts.tsv'))) 
+```
+
+```{r}
+fns_pull <- list.files('../data/shotgun_output/humann2/', pattern = 'pull.+_lefse_ready.+$')
+
+cmds_pull <- tibble(
+  fns = fns_pull
+) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
+  select(-fns) %>% 
+  gather() %>% 
+  select(value) %>% 
+  write_csv('../data/shotgun_output/humann2/lefse_cmd_pull.sh', col_names = F)
+```
+
+```{r}
+# very quicly done some pcoa
+cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(pathway =  names(.)[1]) %>% 
+  spread('pathway', 'value') %>% 
+  column_to_rownames('sampleid')
+
+dist_bl <- vegan::vegdist(cts, method = 'bray')
+
+bl_bc <- cmdscale(dist_bl, k = 2) 
+pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
+  mutate(sampleid = fid)
+
+bl_bc %>% 
+    as.data.frame() %>% 
+    rownames_to_column('sampleid') %>% 
+    inner_join(pheno, by = 'sampleid') %>% 
+    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
+```
+```{r}
+empty_samp <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  group_by(sampleid) %>% 
+  summarise(Sum = sum(value)) %>% 
+  filter(Sum == 0) %>% 
+  pull(sampleid)
+
+cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(KO =  names(.)[1]) %>% 
+  filter(!sampleid %in% empty_samp) %>% 
+  spread('KO', 'value') %>% 
+  column_to_rownames('sampleid')
+
+dist_bl <- vegan::vegdist(cts, method = 'bray')
+
+bl_bc <- cmdscale(dist_bl, k = 2) 
+pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
+  mutate(sampleid = fid)
+
+bl_bc %>% 
+    as.data.frame() %>% 
+    rownames_to_column('sampleid') %>% 
+    inner_join(pheno, by = 'sampleid') %>% 
+    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
+```

### 2020-12-14 scripts/40_combine_all_samp_pheno.Rmd
going slowly but moving forward....
working hard this weekend I think I can actually take a break

@@
 ## the pheno label for the samples
+
+```{r}
+I2 <- read_csv('../data/shotgun_pheno/italy_all_later.csv')
+
+P <- read_csv('../data/shotgun_pheno/upenn_all_earlier.csv')
+
+P2 <- read_csv('../data/shotgun_pheno/upenn_all_later.csv')
+
+M <- read_csv('../data/shotgun_pheno/msk_all_pheno.csv')
+```
 
@@
   count(center, cr_d30)
+
+# output a final format with only the most essential columns
+all %>% 
+  select(fid, cr_d30, toxicity, center) %>% 
+  mutate(cr_d30 = str_to_lower(cr_d30)) %>% 
+  write_csv('../data/shotgun_pheno/final_simplified.csv')
+
+all %>% 
+  write_csv('../data/shotgun_pheno/final_comprehensive.csv')
+
+all %>% 
+  write_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/pheno_data_CR_toxicity_comprehensive.csv')
 ```

### 2020-12-15 scripts/40_combine_all_samp_pheno.Rmd
update pretty efficient with snakemake right now

@@
   write_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/pheno_data_CR_toxicity_comprehensive.csv')
+
+all_pheno <- read_csv('../data/shotgun_pheno/final_comprehensive.csv')
 ```
+
+```{r}
+# as the CR we currently want to look at the day 100 respone only
+all_pheno %>% 
+  select(fid, cr_d100, toxicity, center) %>% 
+  write_csv('../data/shotgun_pheno/final_simplified.csv')
+```

### 2020-12-15 scripts/43_comparisons_3.Rmd
update pretty efficient with snakemake right now

--- 2020-12-13 scripts/43_comparisons_3.Rmd
@@
 
-USING THE HUMANN2 RESULTS FOR NOW
 SUBSTITUTE WITH HUMANN3 RESULTS WHEN I CAN
@@
 ```{r}
+# a tally of the table
+all <- read_csv('../data/shotgun_pheno/final_simplified.csv')
+
+all %>% 
+  split(.$center) %>% 
+  map_dfr(~ count(x = ., cr_d100), .id = 'center')
+
+all %>% 
+  count(toxicity)
+
+```
+
+```{r}
 pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
-  gather('pheno', 'value', cr_d30:toxicity)
+  gather('pheno', 'value', cr_d100:toxicity)
 
@@
       t() %>% 
-      write.table(str_glue('../data/shotgun_output/humann2/{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
+      write.table(str_glue('../data/shotgun_output/humann3/{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
   })
@@
 # the pathway counts table
-full <- read_tsv('../data/shotgun_output/humann2/humann2_pathabundance_CART.tsv') %>% 
-  select(!ends_with('_1'), MS_UPENN_1)
+full <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_pathabundance_CART.tsv')  %>% 
+  rename_all(funs(str_replace(., '^CART_','')))  %>% 
+  rename_all(funs(str_replace(., '_humann3$','')))
 
@@
   imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann2/{.y}_pcts.tsv'))) 
+         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_pcts.tsv'))) 
 ```
 
+## the genefamilies regrouping
+
 ```{r}
-fns <- list.files('../data/shotgun_output/humann2/', pattern = '_lefse_ready.tsv')
+fullgenefam <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_genefamilies_CART.tsv')  %>% 
+  rename_all(funs(str_replace(., '^CART_','')))  %>% 
+  rename_all(funs(str_replace(., '_humann3$','')))
+
+gfcts <- sub_phenos %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_gfcts.tsv'))) 
+```
+
+## run both of them with all of the samples pulled together
+
+```{r}
+all_sub_pheno <- pheno %>% 
+  split(., list(.$pheno)) %>% 
+  purrr::imap(~ filter(.data = ., value != 'not_assessed'))
+
+all_sub_pheno %>% 
+  imap(function(.x, .y){
+    select(.data = .x, value) %>% 
+      t() %>% 
+      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
+  })
+
+all_pcts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
+
+all_gfcts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_gfcts.tsv'))) 
+```
+
+## generate command to run lefse in py2 environment
+
+cuz can't get them thru snakemake
+
+```{r}
+fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready')
 
@@
 ) %>% 
-  transmute(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
   mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
   mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
+  select(-fns) %>% 
   gather() %>% 
   select(value) %>% 
-  write_csv('../data/shotgun_output/humann2/lefse_cmd.sh', col_names = F)
-cmds
+  write_csv('../data/shotgun_output/humann3/lefse_run_cmd.sh', col_names = F)
+```
 
+## PCOA to see if batch effects
+
+```{r}
+# very quicly done some pcoa
+cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(pathway =  names(.)[1]) %>% 
+  spread('pathway', 'value') %>% 
+  column_to_rownames('sampleid')
+
+dist_bl <- vegan::vegdist(cts, method = 'bray')
+
+bl_bc <- cmdscale(dist_bl, k = 2) 
+pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
+  mutate(sampleid = fid)
+
+bl_bc %>% 
+    as.data.frame() %>% 
+    rownames_to_column('sampleid') %>% 
+    inner_join(pheno, by = 'sampleid') %>% 
+    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
 ```
+```{r}
+empty_samp <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  group_by(sampleid) %>% 
+  summarise(Sum = sum(value)) %>% 
+  filter(Sum == 0) %>% 
+  pull(sampleid)
+
+cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(KO =  names(.)[1]) %>% 
+  filter(!sampleid %in% empty_samp) %>% 
+  spread('KO', 'value') %>% 
+  column_to_rownames('sampleid')
+
+dist_bl <- vegan::vegdist(cts, method = 'bray')
+
+bl_bc <- cmdscale(dist_bl, k = 2) 
+pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
+  mutate(sampleid = fid)
+
+bl_bc %>% 
+    as.data.frame() %>% 
+    rownames_to_column('sampleid') %>% 
+    inner_join(pheno, by = 'sampleid') %>% 
+    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
+```

### 2020-12-16 scripts/40_combine_all_samp_pheno.Rmd
update

--- 2020-12-12 scripts/40_combine_all_samp_pheno.Rmd
@@
 ```{r}
+I2 <- read_csv('../data/shotgun_pheno/italy_all_later.csv')
+
+P <- read_csv('../data/shotgun_pheno/upenn_all_earlier.csv')
+
+P2 <- read_csv('../data/shotgun_pheno/upenn_all_later.csv')
+
+M <- read_csv('../data/shotgun_pheno/msk_all_pheno.csv')
+```
+
+```{r}
 I1 <- read_csv('../data/shotgun_pheno/italy_all_earlier.csv') %>% 
   filter(timepoint < 0) %>% 
-  ungroup() %>% 
-  select(fid, cr_d30:toxicity)
+  ungroup()
 
@@
   rename(pt_id = pid) %>% 
-  left_join(iptb) %>% 
-  select(fid, cr_d30:toxicity)
+  left_join(iptb)
 
 I_phe <- bind_rows(
-  I1, I2_pheno
+  I1 %>% 
+    select(fid, pt_id = pid, infusion_date:last_followup), 
+  I2_pheno %>% 
+    select(fid, pt_id, infusion_date:last_followup)
 ) %>% 
   mutate(center = 'I')
+```
 
+```{r}
 # upenn
@@
   left_join(P_ptb)  %>% 
-  select(fid, cr_d30:toxicity) %>% 
-  select(fid, cr_d30, cr_d100, cr_overall, toxicity)  %>% 
-  mutate(center = 'P')
+  mutate(center = 'P') %>% 
+  select(colnames(I_phe))
+```
 
+```{r}
 # MSK
 M_phe <- M %>% 
-  select(fid, cr_d30:toxicity) %>% 
-  mutate(center = 'M')
+  mutate(center = 'M',
+         pt_id = mrn) %>% 
+  select(colnames(I_phe))
+```
 
+```{r}
 # combined
@@
 all <- bind_rows(
-  M_phe, P_phe, I_phe
-)
+  M_phe %>% mutate(pt_id = as.character(pt_id),
+                   icansgrade = as.character(icansgrade)), 
+  P_phe %>% 
+    mutate(pt_id = as.character(pt_id)), 
+  I_phe %>% 
+    mutate(icansgrade = as.character(icansgrade))
+) %>% 
+  mutate(crsgrade = as.character(crsgrade))
 
-# a tally table
+# save a copy to the server for John
 all %>% 
-  count(center, toxicity)
+  write_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/pheno_data_comprehensive.csv')
 
+# save one for my own
 all %>% 
-  count(center, cr_d30)
+  write_csv('../data/shotgun_pheno/final_comprehensive.csv')
 ```
+
+## tally of the table
+
+```{r}
+# make proper columns factors as they should be
+
+should_be_factors <- all %>% 
+  select_if(is.character) %>% 
+  select(-fid, -pt_id) %>% 
+  colnames()
+
+ALL <- all %>% 
+  mutate_at(should_be_factors, as.factor)
+
+```
+
+```{r}
+ALL %>% 
+  split(.$center) %>% 
+  map_dfr(~ count(x = ., icans), .id = 'center')
+```
+```{r}
+ALL %>% 
+  split(.$center) %>% 
+  map_dfr(~ count(x = ., crs), .id = 'center')
+```
+
+```{r}
+outcome <- c('cr_d100','toxicity','icans','crs')
+
+ALL %>% 
+  select(outcome) %>% 
+  map(function(col){
+    ALL %>% 
+  split(.$center) %>% 
+  map_dfr(~ count(x = ., col), .id = 'center')
+  })
+
+for (i in 1:length(outcome)){
+  ALL %>% 
+    split(.$center) %>% 
+    map_dfr(~ count(x = ., outcome[1]), .id = 'center')
+}
+```

### 2020-12-17 scripts/40_combine_all_samp_pheno.Rmd
udpate

@@
 # make proper columns factors as they should be 
-
 should_be_factors <- all %>% 
@@
 
-```
-
-```{r}
-ALL %>% 
-  split(.$center) %>% 
-  map_dfr(~ count(x = ., icans), .id = 'center')
-```
-```{r}
-ALL %>% 
-  split(.$center) %>% 
-  map_dfr(~ count(x = ., crs), .id = 'center')
+ALL %>% summary
 ```
@@
 ALL %>% 
-  select(outcome) %>% 
-  map(function(col){
+  count(center, cr_d100)
+
 ALL %>% 
-  split(.$center) %>% 
-  map_dfr(~ count(x = ., col), .id = 'center')
-  })
+  count(center, toxicity)
 
-for (i in 1:length(outcome)){
 ALL %>% 
-    split(.$center) %>% 
-    map_dfr(~ count(x = ., outcome[1]), .id = 'center')
-}
+  count(center, icans)
+
+ALL %>% 
+  count(center, crs)
+
+tb <- ALL %>% 
+  select(center, outcome) 
 ```

### 2020-12-17 scripts/40_combine_all_samp_pheno.Rmd
was very stimulated and ..

@@
   write_csv('../data/shotgun_pheno/final_comprehensive.csv')
+
+all <- read_csv('../data/shotgun_pheno/final_comprehensive.csv')
 ```
@@
 ```
+
+```{r}
+tally <- ALL %>% 
+  select(center, outcome) %>% 
+  mutate_if(is.factor, as.character) %>% 
+  gather('metric', 'value', cr_d100 : crs) %>% 
+  count(center, metric, value) %>% 
+  filter(value != 'not_assessed') %>% 
+  spread('metric', 'n', fill = 0)
+```
+
+```{r}
+library(kableExtra)
+
+kbl(tally, caption = "") %>%
+  kable_classic_2( full_width = F) %>%
+  pack_rows("Group Italy", 1, 2) %>%
+  pack_rows("Group MSK", 3, 4) %>% 
+  pack_rows("Group UPENN", 5, 6)
+
+```

### 2020-12-17 scripts/43_comparisons_3.Rmd
udpate
was very stimulated and ..

@@
 ```{r}
-# a tally of the table
-all <- read_csv('../data/shotgun_pheno/final_simplified.csv')
-
-all %>% 
-  split(.$center) %>% 
-  map_dfr(~ count(x = ., cr_d100), .id = 'center')
-
-all %>% 
-  count(toxicity)
-
-```
-
-```{r}
-pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
-  gather('pheno', 'value', cr_d100:toxicity)
+pheno <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
+  select(fid, center, cr_d100, toxicity, icans, crs) %>% 
+  gather('pheno', 'value', cr_d100:crs)
 
@@
          write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
+```
 
+```{r}
 all_gfcts <- all_sub_pheno %>% 
@@
 ```{r}
-fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready')
+fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv$')
 
@@
   write_csv('../data/shotgun_output/humann3/lefse_run_cmd.sh', col_names = F)
+```
+
+## The taxa relative abundance from the shotgun data using metaphlan
+
+```{r}
+taxa <- read_tsv('../data/shotgun_output/metaphlan/merged_abundance_table.txt', skip = 1)  %>% 
+  rename_all(funs(str_replace(., '_profile$','')))
+
+tcts <- sub_phenos %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ taxa %>% select(clade_name, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_tcts.tsv')))
+
+all_tcts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ taxa %>% select(clade_name, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_tcts.tsv')))
+```
+
+```{r}
+# run lefse
+fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_tcts.tsv$')
+
+cmds <- tibble(
+  fns = fns
+) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
+  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.png  --label_font_size 4 --dpi 300 --format png --title {fns}')) %>% 
+  select(-fns) %>% 
+  gather() %>% 
+  select(value) %>% 
+  write_csv('../data/shotgun_output/humann3/lefse_run_cmd_taxa.sh', col_names = F)
 ```

### 2020-12-17 scripts/45_ggplot_venn_lefse.Rmd
update

--- 2020-12-15 scripts/43_comparisons_3.Rmd
@@
 ---
-title: "Comparisons in three center"
+title: "ggplot and the venn diagram for the lefse bars !"
 output: html_document
@@
 ```
-
-SUBSTITUTE WITH HUMANN3 RESULTS WHEN I CAN
-
-```{r}
-# a tally of the table
-all <- read_csv('../data/shotgun_pheno/final_simplified.csv')
-
-all %>% 
-  split(.$center) %>% 
-  map_dfr(~ count(x = ., cr_d100), .id = 'center')
-
-all %>% 
-  count(toxicity)
-
-```
-
-```{r}
-pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
-  gather('pheno', 'value', cr_d100:toxicity)
-
-sub_phenos <- pheno %>% 
-  split(., list(.$center, .$pheno)) %>% 
-  purrr::imap(~ filter(.data = ., value != 'not_assessed'))
-```
-
-```{r}
-# the pheno for the different center and tox/CR_d30
-sub_phenos %>% 
-  imap(function(.x, .y){
-    select(.data = .x, value) %>% 
-      t() %>% 
-      write.table(str_glue('../data/shotgun_output/humann3/{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
-  })
-```
-
-```{r}
-# the pathway counts table
-full <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_pathabundance_CART.tsv')  %>% 
-  rename_all(funs(str_replace(., '^CART_','')))  %>% 
-  rename_all(funs(str_replace(., '_humann3$','')))
-
-pcts <- sub_phenos %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_pcts.tsv'))) 
-```
-
-## the genefamilies regrouping
-
-```{r}
-fullgenefam <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_genefamilies_CART.tsv')  %>% 
-  rename_all(funs(str_replace(., '^CART_','')))  %>% 
-  rename_all(funs(str_replace(., '_humann3$','')))
-
-gfcts <- sub_phenos %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_gfcts.tsv'))) 
-```
-
-## run both of them with all of the samples pulled together
-
-```{r}
-all_sub_pheno <- pheno %>% 
-  split(., list(.$pheno)) %>% 
-  purrr::imap(~ filter(.data = ., value != 'not_assessed'))
-
-all_sub_pheno %>% 
-  imap(function(.x, .y){
-    select(.data = .x, value) %>% 
-      t() %>% 
-      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
-  })
-
-all_pcts <- all_sub_pheno %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
-
-all_gfcts <- all_sub_pheno %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_gfcts.tsv'))) 
-```
-
-## generate command to run lefse in py2 environment
-
-cuz can't get them thru snakemake
-
-```{r}
-fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready')
-
-cmds <- tibble(
-  fns = fns
-) %>% 
-  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
-  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
-  select(-fns) %>% 
-  gather() %>% 
-  select(value) %>% 
-  write_csv('../data/shotgun_output/humann3/lefse_run_cmd.sh', col_names = F)
-```
-
-## PCOA to see if batch effects
-
-```{r}
-# very quicly done some pcoa
-cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
-  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
-  rename(pathway =  names(.)[1]) %>% 
-  spread('pathway', 'value') %>% 
-  column_to_rownames('sampleid')
-
-dist_bl <- vegan::vegdist(cts, method = 'bray')
-
-bl_bc <- cmdscale(dist_bl, k = 2) 
-pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
-  mutate(sampleid = fid)
-
-bl_bc %>% 
-    as.data.frame() %>% 
-    rownames_to_column('sampleid') %>% 
-    inner_join(pheno, by = 'sampleid') %>% 
-    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
-```
-```{r}
-empty_samp <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
-  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
-  group_by(sampleid) %>% 
-  summarise(Sum = sum(value)) %>% 
-  filter(Sum == 0) %>% 
-  pull(sampleid)
-
-cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
-  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
-  rename(KO =  names(.)[1]) %>% 
-  filter(!sampleid %in% empty_samp) %>% 
-  spread('KO', 'value') %>% 
-  column_to_rownames('sampleid')
-
-dist_bl <- vegan::vegdist(cts, method = 'bray')
-
-bl_bc <- cmdscale(dist_bl, k = 2) 
-pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
-  mutate(sampleid = fid)
-
-bl_bc %>% 
-    as.data.frame() %>% 
-    rownames_to_column('sampleid') %>% 
-    inner_join(pheno, by = 'sampleid') %>% 
-    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
-```

### 2020-12-17 scripts/cat_combi_CT.sh

@@ -1,2 +1,0 @@
-cat ../data/lefse_input/pheno__CR.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__CR.tsv
-cat ../data/lefse_input/pheno__Toxicity.txt ../data/pheno_aligned_relab_table/humann2_pathabundance_cpm_unstratified.tsv > ../data/lefse_input/pathway_abundace_pheno__Toxicity.tsv

### 2020-12-17 scripts/cat_combi_genef.sh

@@ -1,2 +1,0 @@
-cat ../data/lefse_genefam/pheno__CR.txt ../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv > ../data/lefse_genefam/genefam_abundace_pheno__CR.tsv
-cat ../data/lefse_genefam/pheno__Toxicity.txt ../data/humann2_res/genefamily_split_KO/humann2_genefamilies_KO_regroup_relab_stratified_reordered.tsv > ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.tsv

### 2020-12-17 scripts/cat_combi_spp.sh

@@ -1,2 +1,0 @@
-cat ../data/lefse_input/pheno__CR.txt ../data/metaphlan2_result/merged_abundance_table_WHOLE_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__CR.tsv
-cat ../data/lefse_input/pheno__Toxicity.txt ../data/metaphlan2_result/merged_abundance_table_WHOLE_reordered.tsv > ../data/lefse_input/spp_abundace_pheno__Toxicity.tsv

### 2020-12-17 scripts/format_input_CT.sh

@@ -1,2 +1,0 @@
-format_input.py ../data/lefse_input/pathway_abundace_pheno__CR.tsv ../data/lefse_input/pathway_abundace_pheno__CR.in -c 1  -u 2 -o 1000000
-format_input.py ../data/lefse_input/pathway_abundace_pheno__Toxicity.tsv ../data/lefse_input/pathway_abundace_pheno__Toxicity.in -c 1  -u 2 -o 1000000

### 2020-12-17 scripts/format_input_genef.sh

@@ -1,2 +1,0 @@
-format_input.py ../data/lefse_genefam/genefam_abundace_pheno__CR.tsv ../data/lefse_genefam/genefam_abundace_pheno__CR.in -c 1 -u 2 -o 1000000
-format_input.py ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.tsv ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.in -c 1 -u 2 -o 1000000

### 2020-12-17 scripts/format_input_spp.sh

@@ -1,2 +1,0 @@
-format_input.py ../data/lefse_input/spp_abundace_pheno__CR.tsv ../data/lefse_input/spp_abundace_pheno__CR.in -c 1 -u 2 -o 1000000
-format_input.py ../data/lefse_input/spp_abundace_pheno__Toxicity.tsv ../data/lefse_input/spp_abundace_pheno__Toxicity.in -c 1 -u 2 -o 1000000

### 2020-12-17 scripts/humann2_bar_cmd.sh

@@ -1,23 +1,0 @@
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature ARGDEG-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/ARGDEG-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature HOMOSER-METSYN-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/HOMOSER-METSYN-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature MET-SAM-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/MET-SAM-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature METSYN-PWY --sort similarity sum metadata --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/METSYN-PWY_pw.png
[…]

### 2020-12-17 scripts/humann2_bar_cmd_no_unclassfified.sh

@@ -1,23 +1,0 @@
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature ARGDEG-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/ARGDEG-PWY_pw_no_unclassfified.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature HOMOSER-METSYN-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/HOMOSER-METSYN-PWY_pw_no_unclassfified.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature MET-SAM-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/MET-SAM-PWY_pw_no_unclassfified.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature METSYN-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/METSYN-PWY_pw_no_unclassfified.png
[…]

### 2020-12-17 scripts/humann2_bar_cmd_no_unclassfified_tox.sh

@@ -1,10 +1,0 @@
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature 7ALPHADEHYDROX-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/7ALPHADEHYDROX-PWY_pw_no_unclassfified.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-4242 --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/PWY-4242_pw_no_unclassfified.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature CITRULBIO-PWY --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/CITRULBIO-PWY_pw_no_unclassfified.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-5695 --sort similarity sum metadata --exclude-unclassified --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/without_unclassfified/Toxicity/PWY-5695_pw_no_unclassfified.png
[…]

### 2020-12-17 scripts/humann2_bar_cmd_tox.sh

@@ -1,10 +1,0 @@
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature 7ALPHADEHYDROX-PWY --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/7ALPHADEHYDROX-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-4242 --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/PWY-4242_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature CITRULBIO-PWY --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/CITRULBIO-PWY_pw.png
-humann2_barplot  --input ../data/humann2_res/humann2_final_out/humann2_pathabundance_relab_stratified_reordered_with_tox.tsv --focal-feature PWY-5695 --sort similarity sum metadata  --focal-metadatum Toxicity --last-metadatum Toxicity --scaling pseudolog --dimensions 10 5 -o /Volumes/castoricenter/CAR-Tcell_and_Microbiota/humann2_shotgun/figs/species_contribution_barplot_humann2/with_unclassfified/Toxicity/PWY-5695_pw.png
[…]

### 2020-12-17 scripts/plotting_cladogram_spp.sh

@@ -1,2 +1,0 @@
-plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__CR.res  ../figs/cladogram/spp_abundace_lefse_pheno__CR.png  --dpi 300 --format png --title pheno__CR. --label_font_size 4
-plot_cladogram.py ../data/lefse_input/spp_abundace_pheno__Toxicity.res  ../figs/cladogram/spp_abundace_lefse_pheno__Toxicity.png  --dpi 300 --format png --title pheno__Toxicity. --label_font_size 4

### 2020-12-17 scripts/plotting_de_feature.sh

@@ -1,6 +1,0 @@
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.res  ../figs/pathway_abundace_pheno__CR__CAR_Group.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__CAR_Group.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res  ../figs/pathway_abundace_pheno__CR__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Class.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.res  ../figs/pathway_abundace_pheno__CR__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR__Disease_Status_CAR.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.res  ../figs/pathway_abundace_pheno__Toxicity__CAR_Group.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__CAR_Group.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res  ../figs/pathway_abundace_pheno__Toxicity__Disease_Class.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Class.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.res  ../figs/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity__Disease_Status_CAR.

### 2020-12-17 scripts/plotting_de_feature_CT.sh

@@ -1,2 +1,0 @@
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__CR.res  ../figs/pathway_abundace_pheno__CR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR.
-plot_res.py ../data/lefse_input/pathway_abundace_pheno__Toxicity.res  ../figs/pathway_abundace_pheno__Toxicity.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity.

### 2020-12-17 scripts/plotting_de_feature_genef.sh

@@ -1,2 +1,0 @@
-plot_res.py ../data/lefse_genefam/genefam_abundace_pheno__CR.res  ../figs/genefam_abundace_pheno__CR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR.
-plot_res.py ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.res  ../figs/genefam_abundace_pheno__Toxicity.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity.

### 2020-12-17 scripts/plotting_de_feature_spp.sh

@@ -1,2 +1,0 @@
-plot_res.py ../data/lefse_input/spp_abundace_pheno__CR.res  ../figs/spp_abundace_pheno__CR.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__CR.
-plot_res.py ../data/lefse_input/spp_abundace_pheno__Toxicity.res  ../figs/spp_abundace_pheno__Toxicity.png --feature_font_size 4 --width 10 --dpi 300 --title pheno__Toxicity.

### 2020-12-17 scripts/run_lefse.sh

@@ -1,6 +1,0 @@
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.in ../data/lefse_input/pathway_abundace_pheno__CR__CAR_Group.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Class.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.in ../data/lefse_input/pathway_abundace_pheno__CR__Disease_Status_CAR.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__CAR_Group.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Class.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.in ../data/lefse_input/pathway_abundace_pheno__Toxicity__Disease_Status_CAR.res

### 2020-12-17 scripts/run_lefse_CT.sh

@@ -1,2 +1,0 @@
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__CR.in ../data/lefse_input/pathway_abundace_pheno__CR.res
-run_lefse.py ../data/lefse_input/pathway_abundace_pheno__Toxicity.in ../data/lefse_input/pathway_abundace_pheno__Toxicity.res

### 2020-12-17 scripts/run_lefse_genef.sh

@@ -1,2 +1,0 @@
-run_lefse.py ../data/lefse_genefam/genefam_abundace_pheno__CR.in ../data/lefse_genefam/genefam_abundace_pheno__CR.res
-run_lefse.py ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.in ../data/lefse_genefam/genefam_abundace_pheno__Toxicity.res

### 2020-12-17 scripts/run_lefse_spp.sh

@@ -1,2 +1,0 @@
-run_lefse.py ../data/lefse_input/spp_abundace_pheno__CR.in ../data/lefse_input/spp_abundace_pheno__CR.res
-run_lefse.py ../data/lefse_input/spp_abundace_pheno__Toxicity.in ../data/lefse_input/spp_abundace_pheno__Toxicity.res

### 2020-12-17 scripts/45_ggplot_venn_lefse.Rmd
ehh hard that i need to juggle between two different projecgs

@@
 ```
+
+Create a function of ggplot viewing the lefse res files and also venn to see the overlap and whatnot of the different comparisons
+
+```{r}
+pres <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv.res$', full.names = T)
+
+pres_all <- pres %>% 
+  set_names(pres) %>% 
+  map_dfr(~ read_tsv(., col_names = c('pathway','xx','direction','score','pval')) %>% 
+        filter(!is.na(score)), .id = 'pres') %>% 
+  mutate(pres = str_replace(pres, '^.+//',''),
+         pres = str_replace(pres, '_lefse.+$',''))  %>% 
+  rename(grp = pres) %>% 
+  mutate(grp = str_replace(grp, 'pull_','pull\\.')) %>% 
+  separate(grp, into = c('cohort','outcome'), sep = '\\.')
+
+pres_all_list <- pres_all %>% 
+  split(., list(.$outcome)) %>% 
+  split(~ split())
+  map(~ pull(.data = ., pathway))
+
+all_names <- names(pres_all_list) 
+icans <- all_names[str_detect(all_names, 'icans')]
+icans_list <- pres_all_list %>% 
+  pluck(icans)
+```
+
+```{r}
+library(ggvenn)
+
+ggvenn(pres_all_list[[icans]])
+```

### 2020-12-18 scripts/44_ggplot_venn_lefse.Rmd
gosh cant believe how much im learning purrr right now

--- 2020-12-17 scripts/45_ggplot_venn_lefse.Rmd
@@
 pres_all_list <- pres_all %>% 
-  split(., list(.$outcome)) %>% 
-  split(~ split())
-  map(~ pull(.data = ., pathway))
+  split(., list(.$cohort, .$outcome )) %>% 
+  map(~ pull(.data = ., pathway)) %>% 
+  compact()
 
-all_names <- names(pres_all_list) 
-icans <- all_names[str_detect(all_names, 'icans')]
-icans_list <- pres_all_list %>% 
-  pluck(icans)
+vars <- pres_all %>% 
+  distinct(outcome) %>% 
+  pull(outcome)
+
+pres_all_list_every <- vars %>% 
+  set_names(vars) %>% 
+  map(function(var){
+    pres_all_list %>% 
+        keep(.p = str_detect(names(.), var))
+  } )
 ```
@@
 
-ggvenn(pres_all_list[[icans]])
+pres_all_list_every %>% 
+  imap(~ ggvenn(.x) +
+         #labs(title = str_glue('{.y}')) +
+          ggsave(str_glue('../figs/venn_pathway_{.y}.pdf'), width = 7, height = 4))
 ```

### 2020-12-18 scripts/44_ggplot_venn_lefse.Rmd
update

--- 2020-12-17 scripts/45_ggplot_venn_lefse.Rmd
@@
 library(tidyverse)
+library(ggvenn)
 ```
+
+Create a function of ggplot viewing the lefse res files and also venn to see the overlap and whatnot of the different comparisons
+
+```{r}
+res <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_.cts.tsv.res$', full.names = T)
+
+res_all <- res %>% 
+  set_names(res) %>% 
+  map_dfr(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
+        filter(!is.na(score)), .id = 'res') %>% 
+  mutate(tbl = if_else(str_detect(res, 'pcts'), 'pathway', 'taxa')) %>% 
+  mutate(res = str_replace(res, '^.+//',''),
+         res = str_replace(res, '_lefse.+$',''))  %>% 
+  rename(grp = res) %>% 
+  mutate(grp = str_replace(grp, 'pull_','pull\\.')) %>% 
+  separate(grp, into = c('cohort','outcome'), sep = '\\.')
+
+res_all_list <- res_all %>% 
+  split(.$tbl) %>% 
+  map(~split(., list(.$cohort, .$outcome )) %>% 
+  map(~ pull(.data = ., feature)) %>% 
+  compact())
+
+vars <- res_all %>% 
+  distinct(outcome) %>% 
+  pull(outcome)
+
+tbs <- names(res_all_list)
+
+tbs %>% 
+  set_names(tbs) %>% 
+  map(function(tb){
+    vars %>% 
+      set_names(vars) %>% 
+      map(function(var){
+        res_all_list %>% 
+          pluck(tb) %>% 
+          keep(.p = str_detect(names(.), var))
+  } ) %>% 
+      imap(~ ggvenn(.x) +
+             labs(title = str_glue('{tb}')) +
+          ggsave(str_glue('../figs/{tb}_venn_{.y}.pdf'), width = 7, height = 4))
+  })
+```

### 2020-12-19 scripts/40_combine_all_samp_pheno.Rmd
update

--- 2020-12-17 scripts/40_combine_all_samp_pheno.Rmd
@@
 ```
+
+```{r}
+tally <- ALL %>% 
+  select(center, outcome) %>% 
+  mutate_if(is.factor, as.character) %>% 
+  gather('metric', 'value', cr_d100 : crs)
+
+individual <- tally %>% 
+  count(center, metric, value) %>% 
+  filter(value != 'not_assessed') %>% 
+  spread('metric', 'n', fill = 0)
+
+pull_ <- tally %>% 
+  count(metric, value)  %>% 
+  filter(value != 'not_assessed') %>% 
+  mutate(center = 'pull') %>% 
+  spread('metric', 'n', fill = 0) %>% 
+  select(colnames(individual))
+
+tally_all <- bind_rows(
+  individual, pull_
+)
+
+```
+
+```{r}
+library(kableExtra)
+
+kbl(tally_all, caption = "") %>%
+  kable_minimal( full_width = F) %>%
+  pack_rows("Group Italy", 1, 2) %>%
+  pack_rows("Group MSK", 3, 4) %>% 
+  pack_rows("Group UPENN", 5, 6)  %>% 
+  pack_rows("Group ALL", 7, 8)
+
+```

### 2020-12-19 scripts/43_comparisons_3.Rmd
update

@@
 
+```{bash}
+# regroup the whole gene family table here
+humann_regroup_table --input humann3_genefamilies_CART.tsv --groups uniref90_ko --output humann3_KO_CART.tsv
+```
+
 ```{r}
-fullgenefam <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_genefamilies_CART.tsv')  %>% 
+fullko <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_KO_CART.tsv')  %>% 
   rename_all(funs(str_replace(., '^CART_','')))  %>% 
@@
 
-gfcts <- sub_phenos %>% 
+kocts <- sub_phenos %>% 
   map(~ pull(.data = ., fid) ) %>% 
-  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_gfcts.tsv'))) 
+  imap(~ fullko %>% select(`# Gene Family`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_kocts.tsv'))) 
+```
+
+```{r}
+# the lefse for the ko only
+fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_kocts.tsv$')
+
+cmds <- tibble(
+  fns = fns
+) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
+  select(-fns) %>% 
+  gather() %>% 
+  select(value) %>% 
+  write_csv('../data/shotgun_output/humann3/lefse_run_cmd_KO.sh', col_names = F)
 ```
@@
          write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
+
+all_kocts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., fid) ) %>% 
+  imap(~ fullko %>% select(`# Gene Family`, matches(.x)) %>% 
+         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_kocts.tsv')))
 ```
@@
   mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
   select(-fns) %>% 
@@
   mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
   mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.png  --label_font_size 4 --dpi 300 --format png --title {fns}')) %>% 

### 2020-12-19 scripts/44_ggplot_venn_lefse.Rmd
update

@@
 Create a function of ggplot viewing the lefse res files and also venn to see the overlap and whatnot of the different comparisons
+
+## collect all of the res files
 
@@
   compact())
+```
 
+```{r}
 vars <- res_all %>% 
@@
 ```
+
+## a PCOA to see how the data
+
+```{r}
+cts <- read_tsv('../data/shotgun_output/humann3/') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(pathway =  names(.)[1]) %>% 
+  spread('pathway', 'value') %>% 
+  column_to_rownames('sampleid')
+dist_bl <- vegan::vegdist(cts, method = 'bray')
+bl_bc <- cmdscale(dist_bl, k = 2) 
+pheno <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
+  mutate(sampleid = fid)
+bl_bc %>% 
+    as.data.frame() %>% 
+    rownames_to_column('sampleid') %>% 
+    inner_join(pheno, by = 'sampleid') %>% 
+    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
+```

### 2020-12-20 scripts/44_ggplot_venn_lefse.Rmd
udpate

--- 2020-12-18 scripts/44_ggplot_venn_lefse.Rmd
@@
 
+## collect all of the res files
+
 ```{r}
-res <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_.cts.tsv.res$', full.names = T)
+res <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_.+cts.tsv.res$', full.names = T)
 
@@
   set_names(res) %>% 
-  map_dfr(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
-        filter(!is.na(score)), .id = 'res') %>% 
-  mutate(tbl = if_else(str_detect(res, 'pcts'), 'pathway', 'taxa')) %>% 
+  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
+        filter(!is.na(score))) %>% 
+  keep(~ nrow(.) > 0) %>% 
+  bind_rows(.id = 'res') %>% 
+  mutate(tbl = if_else(str_detect(res, 'pcts'), 'pathway', if_else(str_detect(res, 'kocts'), 'KO', 'taxa'))) %>% 
   mutate(res = str_replace(res, '^.+//',''),
@@
 
+```
+
+```{r}
 vars <- res_all %>% 
@@
 ```
+
+```{r}
+# for the KO table
+library(KEGGREST)
+res_all_list %>% pluck('KO') %>% 
+  keep(str_detect(names(.), 'toxicity')) %>% 
+  reduce(intersect) %>% 
+  map(~ list(KOID = keggGet(.)[[1]]$ENTRY,
+             KO_name = keggGet(.)[[1]]$NAME,
+             KO_def = keggGet(.)[[1]]$DEFINITION),
+             KO_pathway = keggGet(.)[[1]]$PATHWAY) %>%
+  bind_rows()
+```
+
+## a PCOA to see how the data
+
+```{r}
+cts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(pathway =  names(.)[1]) %>% 
+  spread('pathway', 'value') %>% 
+  column_to_rownames('sampleid')
+dist_bl <- vegan::vegdist(cts, method = 'bray')
+bl_bc <- cmdscale(dist_bl, k = 2) 
+pheno <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
+  mutate(sampleid = fid)
+bl_bc %>% 
+    as.data.frame() %>% 
+    rownames_to_column('sampleid') %>% 
+    inner_join(pheno, by = 'sampleid') %>% 
+    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
+```

### 2020-12-21 scripts/44_ggplot_venn_lefse.Rmd
such a busy day under immense pressure and stay remain calm

@@
 ```
+
+## output a pretty ggplot version of the lefse bar plot
+
+```{r}
+make <- res_all %>% 
+  filter(cohort == 'pull' & tbl == 'pathway') %>% 
+  mutate(score = if_else(direction == 'no', -score, score))
+make
+```
+
+```{r}
+all_title_fs <- 20
+axis_text_fs <- 16
+make %>% 
+  split(.$outcome) %>% 
+  imap(~   ggplot(data = .x, aes(x = reorder(feature, score), y = score, fill = direction)) +
+            geom_bar(stat = 'identity') +
+            coord_flip() +
+            #scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            #scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  axis.title.x = element_text(size=axis_text_fs),
+                  plot.title = element_text(size=all_title_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position="bottom") +
+            labs(title = str_glue('Response to {.y}') ,
+                 y = 'Score')+ 
+            ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/NEW_humann3_version/pathway/pull_{.y}_pathway.pdf'), device = 'pdf', width = 12))
+
+```
+
+```

### 2020-12-22 scripts/40_combine_all_samp_pheno.Rmd
gosh went back to the pheno table cleaning again

@@
 ```{r}
+# trying to get back the timepoint, sampleid, oligoid
+# starting from the patient and their earliest sample better have pre in their raw sampleid
+
 I2 <- read_csv('../data/shotgun_pheno/italy_all_later.csv')
+
+isamp <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/samples/opbg_sampleskey_with_oligoid_v2.xlsx', sheet = 1) %>% 
+  filter(pt_id %in% I2$pid) %>% 
+  filter(timepoint < 0) %>% 
+  distinct(sampleid, .keep_all = T) %>% 
+  full_join(I2 %>% select(pt_id = pid, fid))
 
@@
 iptb <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Italy/patients/italy_essential_data_v4.xlsx', sheet = 1)
-I2_pheno <- I2 %>% 
-  rename(pt_id = pid) %>% 
+I2_pheno <- isamp %>% 
   left_join(iptb)
@@
   I1 %>% 
-    select(fid, pt_id = pid, infusion_date:last_followup), 
+    select(fid, pt_id = pid, timepoint:last_followup), 
   I2_pheno %>% 
-    select(fid, pt_id, infusion_date:last_followup)
+    select(fid, pt_id, timepoint:last_followup)
 ) %>% 

### 2020-12-22 scripts/40_combine_all_samp_pheno.Rmd
update on the pheno cleaning again

@@
   I2_pheno %>% 
-    select(fid, pt_id, timepoint:last_followup)
+    select(fid, pt_id, timepoint:sampleid, oligoid:last_followup)
 ) %>% 
@@
   left_join(P_ptb)  %>% 
-  mutate(center = 'P') %>% 
+  mutate(center = 'P')
+
+psamp <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/samples/sample_mapping_UPenn.xlsx', sheet = 1) %>% 
+  mutate(pt_id = pt_id-8000) %>% 
+  filter(pt_id %in% P_phe$pt_id) %>% 
+  full_join(P_phe) %>% 
   select(colnames(I_phe))
@@
 # MSK
+msamp <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 1)
 M_phe <- M %>% 
+  mutate(timepoint = NA,
+         oligoid = NA) %>% 
   mutate(center = 'M',
@@
 all <- bind_rows(
-  M_phe %>% mutate(pt_id = as.character(pt_id),
-                   icansgrade = as.character(icansgrade)), 
-  P_phe %>% 
+  M_phe %>% mutate(pt_id = as.character(pt_id)), 
+  psamp %>% 
     mutate(pt_id = as.character(pt_id)), 
-  I_phe %>% 
-    mutate(icansgrade = as.character(icansgrade))
-) %>% 
-  mutate(crsgrade = as.character(crsgrade))
+  I_phe 
+)
 

### 2021-01-27 scripts/43_comparisons_3.Rmd
update

@@
   select(fid, center, cr_d100, toxicity, icans, crs) %>% 
+  filter(center != 'I') %>% # remove the Italy ones
   gather('pheno', 'value', cr_d100:crs)
@@
          write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
+```
 
+```{r}
+# KO
 all_kocts <- all_sub_pheno %>% 
@@
 
+## Need to go to snakemake for some steps related to humann before the lefse
+
 ## generate command to run lefse in py2 environment
@@
 ```{r}
+# PATHWAY
 fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv$')
@@
 ```{r}
-taxa <- read_tsv('../data/shotgun_output/metaphlan/merged_abundance_table.txt', skip = 1)  %>% 
+taxa <- read_tsv('../data/shotgun_output/metaphlan/merged_abundance_table.txt')  %>% 
   rename_all(funs(str_replace(., '_profile$','')))
@@
          write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_tcts.tsv')))
+
+# then just need to use the snakemake rule to cat the two
 ```
@@
   mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
-  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.png  --label_font_size 4 --dpi 300 --format png --title {fns}')) %>% 
+  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
   select(-fns) %>% 

### 2021-01-27 scripts/44_ggplot_venn_lefse.Rmd
update

--- 2020-12-20 scripts/44_ggplot_venn_lefse.Rmd
@@
 ```
+
+## output a pretty ggplot version of the lefse bar plot
+
+```{r}
+make <- res_all %>% 
+  filter(cohort == 'pull' & tbl == 'pathway') %>% 
+  mutate(score = if_else(direction == 'no', -score, score)) %>% 
+  mutate(feature = str_replace_all(feature, '_',' ')) %>% 
+  mutate(feature = str_wrap(feature, width = 55, indent = 0, exdent = 0))
+
+all_title_fs <- 20
+axis_text_fs <- 16
+make %>% 
+  split(.$outcome) %>% 
+  imap(~   ggplot(data = .x, aes(x = reorder(feature, score), y = score, fill = direction)) +
+            geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
+            coord_flip() +
+            #scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            #scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  axis.title.x = element_text(size=axis_text_fs),
+                  plot.title = element_text(size=all_title_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position="bottom") +
+         scale_fill_manual(values = c('#0099B4','#AD002A')) +
+            labs(title = str_glue('Response to {.y}') ,
+                 y = 'Score')+ 
+            ggsave(str_glue('../data/shotgun_output/humann3/pull_{.y}_pathway.pdf'), device = 'pdf', width = 14))
+            #ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/NEW_humann3_version/pathway/pull_{.y}_pathway.pdf'), device = 'pdf', width = 12))
+
+make %>% 
+  filter(outcome == 'cr_d100') %>% 
+  split(.$outcome) %>% 
+  imap(~   ggplot(data = .x, aes(x = reorder(feature, score), y = score, fill = direction)) +
+            geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
+            coord_flip() +
+            #scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            #scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  axis.title.x = element_text(size=axis_text_fs),
+                  plot.title = element_text(size=all_title_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position="bottom") +
+         scale_fill_manual(values = c('#ED0000','#00468B')) +
+            labs(title = str_glue('Response to {.y}') ,
+                 y = 'Score')+ 
+            ggsave(str_glue('../data/shotgun_output/humann3/pull_{.y}_pathway.pdf'), device = 'pdf', width = 14))
+
+# for above this is for the Italy excluded version 
+```

### 2021-02-11 scripts/44_ggplot_venn_lefse.Rmd
updte

--- 2020-12-20 scripts/44_ggplot_venn_lefse.Rmd
@@
 ```
+
+## output a pretty ggplot version of the lefse bar plot
+
+```{r}
+make <- res_all %>% 
+  filter(cohort == 'pull' & tbl == 'pathway') %>% 
+  mutate(score = if_else(direction == 'no', -score, score)) %>% 
+  mutate(feature = str_replace_all(feature, '_',' ')) 
+  #mutate(feature = str_wrap(feature, width = 55, indent = 0, exdent = 0))
+
+all_title_fs <- 20
+axis_text_fs <- 16
+make %>% 
+  split(.$outcome) %>% 
+  imap(~   ggplot(data = .x, aes(x = reorder(feature, score), y = score, fill = direction)) +
+            geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
+            coord_flip() +
+            #scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            #scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  axis.title.x = element_text(size=axis_text_fs),
+                  plot.title = element_text(size=all_title_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position="bottom") +
+         scale_fill_manual(values = c('#0099B4','#AD002A')) +
+            labs(title = str_glue('Response to {.y}') ,
+                 y = 'Score')+ 
+            ggsave(str_glue('../data/shotgun_output/humann3/pull_{.y}_pathway.pdf'), device = 'pdf', width = 8))
+            #ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/NEW_humann3_version/pathway/pull_{.y}_pathway.pdf'), device = 'pdf', width = 12))
+
+make %>% 
+  filter(outcome == 'cr_d100') %>% 
+  split(.$outcome) %>% 
+  imap(~   ggplot(data = .x, aes(x = reorder(feature, score), y = score, fill = direction)) +
+            geom_col(position = position_dodge2(width = 0.9, preserve = "single")) +
+            coord_flip() +
+            #scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            #scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  axis.title.x = element_text(size=axis_text_fs),
+                  plot.title = element_text(size=all_title_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position="bottom") +
+         scale_fill_manual(values = c('#ED0000','#00468B')) +
+            labs(title = str_glue('Response to {.y}') ,
+                 y = 'Score')+ 
+            ggsave(str_glue('../data/shotgun_output/humann3/pull_{.y}_pathway.pdf'), device = 'pdf', width = 8))
+
+# for above this is for the Italy excluded version 
+```

### 2021-02-15 scripts/45_16s_sorting_out.Rmd
update

---
title: "Sorting out the 16s"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readxl)
```

This script will be cleaning and organizing the 16s samples

```{r}
# upenn is all baseline samples 
ustb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/samples/sample_mapping_UPenn.xlsx')

uptb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 3) %>% 
  filter(!is.na(gender))

```

```{r}
# MSK contains baseline and following time points
mstb_ineli <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 2)
mstb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/sample_table_msk.xlsx', sheet = 1) %>% 
  filter(! sampleid %in% mstb_ineli$sampleid)

mptb_ineli <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 2)
mptb <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 1) %>% 
  mutate(mrn = as.numeric(mrn)) %>% 
  filter(! mrn %in% mptb_ineli$pt_id)

```

## baseline samples only meta

```{r}
m_meta_b <- mstb %>% 
  filter(str_detect(time_group, '^Baseline')) %>% 
  inner_join(mptb, by = "mrn")

u_meta_b <- ustb %>% 
  inner_join(uptb, by = "pt_id")

colnames(u_meta_b)
```

```{r}
colnames(m_meta_b)
```

## Look at the tally of the phenotypes

```{r}
m_meta_b %>% 
  count(cr_d30)

m_meta_b %>% 
  count(cr_d100)

m_meta_b %>% 
  count(toxicity)
```
```{r}
u_meta_b %>% 
  count(cr_d30)

u_meta_b %>% 
  count(cr_d100)

u_meta_b %>% 
  count(toxicity)
```

### 2021-02-15 scripts/46_16s_relative_abundance.Rmd
update

---
title: "16s relative abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(vdbR)
connect_database('~/dbConfig.txt')
library(tidyverse)
```

```{r}
get_table_from_database('asv_counts_ag')
get_table_from_database('asv_annotation_blast_ag')
```

```{r}
# the counts for the genus in MSK
mcts <- asv_counts_ag %>% 
  filter(sampleid %in% m_meta_b$sampleid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)

mannot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% mcts$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species))

mcts_all <- mcts %>% 
  full_join(mannot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, genus) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, genus, relab) 
```
```{r}
# the counts for the genus in Upenn
ucts <- asv_counts_ag %>% 
  filter(sampleid %in% u_meta_b$oligoid) %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)

uannot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% ucts$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species))

ucts_all <- ucts %>% 
  full_join(uannot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, genus) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(asv_counts_ag %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, genus, relab) 
```
```{r}
# combine them and save to a df
genus <- bind_rows(
  ucts_all %>% mutate(center = 'P'),
  mcts_all %>% mutate(center = 'M')
)

genus %>% 
  write_csv('../data/amplicon/stool/counts_genus_combined.csv')
```

### 2021-02-15 scripts/45_16s_sorting_out.Rmd
great slowly picking up the pace

@@
 ```
+## merge the two pheno tables and then write out
+
+```{r}
+m <- m_meta_b %>% 
+  select(-pt_id_old) %>% 
+  rename(pt_id = mrn) %>% 
+  select(-last_followup) %>% 
+  mutate(center = 'M') %>% 
+  mutate(icansgrade = as.character(icansgrade))
+
+p <- u_meta_b %>%   
+  mutate(time_group = 'Baseline') %>% 
+  select(-sampleid) %>% 
+  rename(sampleid = oligoid) %>% 
+  select(-last_followup_day) %>% 
+  mutate(center = 'P')  %>% 
+  select(colnames(m))
+
+combined <- bind_rows(m, p)
+```
+
+## also get the alpha diversity here
+
+```{r}
+get_table_from_database('asv_alpha_diversity_ag')
+
+combined_join <- combined %>% 
+  inner_join(asv_alpha_diversity_ag %>% 
+               select(sampleid, simpson_reciprocal))
+
+combined_join %>% 
+  write_csv('../data/amplicon/stool/combined_2_meta.csv')
+```

### 2021-02-15 scripts/47_genus_relab_and_phenotype.Rmd
good I made some progress

---
title: "16s genus relab and phenotype"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

Enteroccous relative abundance v CR D100

Enteroccous relative abundance v Toxicity

Toxicity grade (both crs and icans) in light of the relative abundance of bacteroidetes or enterococcus

## Enteroccous relative abundance v CR D100

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv')
```

```{r}
# get the relative abundance of bacteroidetes or enterococcus for those samples
target <- genus %>% 
  filter(genus %in% c('Enterococcus','Bacteroides')) %>% 
  select(-center) %>% 
  spread('genus', 'relab') %>% 
  mutate(Bacteroides_log = log10(Bacteroides + 2*10^-6),
         Enterococcus_log = log10(Enterococcus + 2*10^-6))

joined <- meta %>% 
  full_join(target) %>% 
  mutate(center = if_else(center == 'P', 'Upenn', "MSK"))
```

```{r}
# a tally of the yes and no and unknown for the CR D100 to put in the title
joined %>% 
  filter(cr_d100 %in% c('yes','no')) %>% 
  ggboxplot(x = 'cr_d100', y = 'Enterococcus_log', facet.by = 'center', add = 'jitter', color = 'cr_d100',
            xlab = 'CR Day 100', ylab = 'log10(Enterococcus relab)',
            title = 'CR D100 VS log10(Enterococcus relab)') +
  scale_color_manual(values = c('#ED0000','#00468B'))  +
  stat_compare_means(comparisons= list(c('yes', 'no')),
    label= "p.format",
    method= 'wilcox.test',
    exact=FALSE) +
  ggsave('../figs/amplicon/CR D100 VS log10(Enterococcus relab).pdf', width = 5, height = 5)
```

## Enteroccous relative abundance v Toxicity

```{r}
joined %>% 
  ggboxplot(x = 'toxicity', y = 'Enterococcus_log', facet.by = 'center', add = 'jitter', color = 'toxicity',
            xlab = 'toxicity', ylab = 'log10(Enterococcus relab)',
            title = 'toxicity VS log10(Enterococcus relab)') +
  scale_color_manual(values = c('#0099B4','#AD002A'))  +
   stat_compare_means(comparisons= list(c('yes', 'no')),
    label= "p.format",
    method= 'wilcox.test',
    exact=FALSE) +
  ggsave('../figs/amplicon/toxicity VS log10(Enterococcus relab).pdf', width = 5, height = 4.5)
```
## Toxicity grade (both crs and icans) VS the relative abundance of bacteroidetes or enterococcus

```{r}
joined %>% 
  count(center, icansgrade)

joined %>% 
  count(center, crsgrade)

# ignore the pending ones
df <- joined %>% 
  mutate(icansgrade = as.numeric(icansgrade)) %>% 
  filter(!is.na(icansgrade)) %>% 
  select(icansgrade, crsgrade, Bacteroides_log, Enterococcus_log, center)  %>% 
  gather('item', 'grade', icansgrade:crsgrade)

df %>% 
  ggboxplot(x = 'grade', y = 'Bacteroides_log',  add = 'jitter',
            xlab = 'Toxicity grade', ylab = 'log10(Bacteroides relab)',
            title = 'Toxicity grade VS log10(Bacteroides relab)') +
  facet_wrap(item ~ center, scales = 'free') +
  stat_compare_means(comparisons= list(c('0', '1'), c('0','2'), c('0','3')),label= "p.format", method= 'wilcox.test',exact=FALSE)  + 
  ggsave('../figs/amplicon/toxicity grade VS log10(Bacteroides relab).pdf', width = 7, height = 7)
```
```{r}
df %>% 
  ggboxplot(x = 'grade', y = 'Enterococcus_log',  add = 'jitter',
            xlab = 'Toxicity grade', ylab = 'log10(Enterococcus relab)',
            title = 'Toxicity grade VS log10(Enterococcus relab)') +
  facet_wrap(item ~ center, scales = 'free') +
  stat_compare_means(comparisons= list(c('0', '1'), c('0','2'), c('0','3')),label= "p.format", method= 'wilcox.test',exact=FALSE)  + 
  ggsave('../figs/amplicon/toxicity grade VS log10(Enterococcus relab).pdf', width = 7, height = 7)
```
 The problem is when they don't have all the comparison in a specific facet there is gonna be no comparison and p value gotta think about this

### 2021-02-17 scripts/48_16s_beta_diversity.Rmd
pretty sleep but managed to get this done its my love for the doing

---
title: "16s beta diversity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(vegan)
library(tidyverse)
set.seed(123)
```

Do a PcoA and do the anosim test

**Do M, P and pull separately, and then do for each of the 4 things CR_d100, Toxicity, ICANS and CRS**

```{r}
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv')
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') 
pheno <- meta %>% 
  select(sampleid, cr_d100, toxicity, crs, icans)

nsamp <- genus %>% distinct(sampleid) %>% nrow

# do some filtering 10% and 0.01%
keep <- genus %>% 
  filter(relab > 0.0001) %>% 
  count(genus) %>% 
  filter(n > floor(nsamp * 0.2)) %>% 
  filter(!is.na(genus)) %>% 
  pull(genus)

dat <- genus %>% 
  filter(genus %in% keep) %>% 
  spread(key = 'genus', value = 'relab') %>% 
  full_join(pheno)

```

```{r}
# a function to do the combinations of the anosim test
pheno4 <- colnames(pheno)[2:5]

combi <- data_frame(
  g1 = c(rep('M',4), rep('P',4), rep('M',4)),
  g2 = c(rep('M',4), rep('P',4), rep('P',4)),
  label = rep(pheno4, 3)
)

do_anosim_each <-  function(g1_, g2_, label_){
    df = dat %>% 
      filter(center %in% c(g1_, g2_)) %>% 
      filter(get(label_) %in% c('yes','no')) 
    cts = df %>% 
      select(sampleid, all_of(keep)) %>% 
      column_to_rownames('sampleid')
    phe = df %>% 
      select(matches(label_)) %>% pull(get(label_))

    # anosim run
    ret = anosim(cts, phe, permutations = 999, distance = "bray", strata = NULL, parallel = 4)

  return(list(pval = ret$signif, 
              statistic = ret$statistic))
}

all_anosim <- pmap(combi, function(g1, g2, label){
    do_anosim_each(g1, g2, label)
  }) %>% 
  set_names(str_glue('{combi$g1}_{combi$g2}__{combi$label}')) 
```

```{r}
# to organize the final results
anosim_res <- all_anosim %>% 
  bind_rows(.id = 'name') %>% 
  separate(name, into = c('group','pheno'), remove = T, sep = '__') %>% 
  mutate(cohort = case_when(group == 'M_M' ~ "MSK", group == 'P_P' ~ "Upenn", group == 'M_P' ~ "Combined"))
```

### 2021-02-17 scripts/48_16s_beta_diversity.Rmd
yeah I have added the pcoa ggplot to it

@@
       select(matches(label_)) %>% pull(get(label_)) 
+    phedf = df %>% 
+      select(sampleid, matches(label_))
 
@@
     ret = anosim(cts, phe, permutations = 999, distance = "bray", strata = NULL, parallel = 4)
+
+    # plot the pcoa
+    dist_ <- vegdist(cts, method = 'bray')
+    bc <- cmdscale(dist_, k = 2) 
+    bc %>% 
+      as.data.frame() %>% 
+      rownames_to_column('sampleid') %>% 
+      full_join(phedf, by = 'sampleid') %>% 
+      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
+      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
+      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))
 
@@
   mutate(cohort = case_when(group == 'M_M' ~ "MSK", group == 'P_P' ~ "Upenn", group == 'M_P' ~ "Combined"))
+
+anosim_res %>% 
+  write_csv('../data/amplicon/stool/anosim_res.csv')
 ```

### 2021-02-22 scripts/49_16s_stacked_barplot.Rmd
udpate

---
title: "16s taxa stacked bar plot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database("asv_annotation_blast_color_ag");
```

## load the tables and organize the data

```{r}
# my table of the CART stool cohort
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

# get the counts from database and also the color for the asv 
counts_data <- get_counts_subset(stb$sampleid)

counts_data %>% 
  distinct(sampleid)

dat <- counts_data %>% 
  select(asv_key:count_total, count_relative) %>% 
  left_join(asv_annotation_blast_color_ag %>% 
              select(asv_key,color_label_group_distinct), by = "asv_key")

# the color palette (inherited from Ying, used in lots of project in our lab, the palette used in the NEJM paper Fig 2D https://www.nejm.org/doi/full/10.1056/NEJMoa1900623)

asv_color_set <- asv_annotation_blast_color_ag %>% 
  distinct(color,color_label_group_distinct,color_label_group,color_base) %>% 
  select(color_label_group_distinct, color) %>% 
  deframe()
```

```{r}
# calculate the beta diversity between the samples which deicide the order of the samples in the plot
cbd <- compute_beta_diversity_and_tsne(sampleid = dat$sampleid, 
                                      taxonomy = dat$color_label_group_distinct,
                                      count = dat$count);
#compute beta diversity
cbd$compute_beta_diversity()
```

```{r}
#get beta diversity
d_beta <- cbd$get_betadiversity()

#compute hierarchical cluster
hc <- hclust(as.dist(d_beta), method = 'complete')

dend <- as.dendrogram(hc)

sample_dendogram_order <- labels(dend)

dat$sampleid = factor(dat$sampleid,levels = sample_dendogram_order)

ggplot(dat,aes(sampleid, count_relative, fill = color_label_group_distinct) ) +
  geom_bar(stat = "identity", position="fill", width = 1) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90),
        axis.text.y = element_blank(),
        legend.position = "none") +
  scale_fill_manual(values = asv_color_set)

```

### 2021-02-22 scripts/49_16s_stacked_barplot.Rmd
first time to work on the bars of the 16s

@@
   distinct(sampleid)
+```
 
+```{r}
 dat <- counts_data %>% 
@@
 
+# there are some ASVs that don't have a color with it, but can use the color for the genus level
+color_group <- dat %>% 
+  split(is.na(.$color_label_group_distinct))
+
+# find the genus for these asv
+get_table_from_database('asv_annotation_blast_ag')
+
+no_color <- color_group %>% 
+  pluck('TRUE') %>% 
+  distinct(asv_key) %>% 
+  inner_join(asv_annotation_blast_ag %>% 
+               select(asv_key, genus))
+
+# find the colors for these genera
+genera_colors <- no_color %>% 
+  distinct(genus) %>% 
+  inner_join(asv_annotation_blast_color_ag %>% 
+               distinct(genus, color_label_group_distinct))
+
+# the full df for the no color genera
+no_color_df <- no_color %>% 
+  left_join(genera_colors)
+
+no_color_df_full <- color_group %>% 
+  pluck('TRUE') %>% 
+  select(-color_label_group_distinct) %>% 
+  left_join(no_color_df %>% 
+              select(- genus))
+
+# the question is do we go one taxa level higher or make a new color base and shades for the new asv 
+```
+
+```{r}
 # the color palette (inherited from Ying, used in lots of project in our lab, the palette used in the NEJM paper Fig 2D https://www.nejm.org/doi/full/10.1056/NEJMoa1900623)
@@
 ```
+
+```{r}
+# the thing that is dividing the samples to lower and higher diversity 
+# or just order samples by their diversity
+div_order <- stb %>% 
+  arrange(simpson_reciprocal) %>% 
+  pull(sampleid)
+
+dat$sampleid = factor(dat$sampleid,levels = div_order)
+
+ggplot(dat,aes(sampleid, count_relative, fill = color_label_group_distinct) ) +
+  geom_bar(stat = "identity", position="fill", width = 1) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 90),
+        axis.text.y = element_blank(),
+        legend.position = "none") +
+  scale_fill_manual(values = asv_color_set)
+```

### 2021-03-02 scripts/49_16s_stacked_barplot.Rmd
did this centroid thing anyway

@@
 
-dat$sampleid = factor(dat$sampleid,levels = div_order)
+###
+# how about splitting the above dendrogram order into the low and higher diversity groups
+div_med <- median(stb$simpson_reciprocal)
+lower_samp <- stb %>% 
+  filter(simpson_reciprocal <= div_med) %>% 
+  pull(sampleid)
+lower_samp_o <- sample_dendogram_order[sample_dendogram_order %in% lower_samp]
+higher_samp_o <- sample_dendogram_order[!sample_dendogram_order %in% lower_samp]
+
+dat$sampleid = factor(dat$sampleid,levels = c(lower_samp_o, higher_samp_o))
 

### 2021-03-02 scripts/50_16s_diversity.Rmd
did this centroid thing anyway

---
title: "16s diversity"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
```
```{r}
# combining the institutions for the diversity boxplot colored by the toxicity and CR
dat <- stb %>% 
  select(sampleid, cr_d100, toxicity, simpson_reciprocal) %>% 
  gather('metric', 'value', cr_d100:toxicity)

# for coloring issues the two metric has to be done separately
dat %>% 
  filter(value %in% c('yes','no')) %>% 
  filter(metric == 'cr_d100') %>% 
  ggboxplot(x = 'value', y = 'simpson_reciprocal', add = 'jitter',
            title = 'cr_d100', ylab = 'Simpson Reciprocal', xlab = '', color = 'value',
            palette = c('#ED0000','#00468B')) +
            stat_compare_means(comparisons= list(c('yes', 'no')),
            label= "p.format",
            method= 'wilcox.test',
            correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/diversity/combined_cr_d100.pdf')

dat %>% 
  filter(value %in% c('yes','no')) %>% 
  filter(metric == 'toxicity') %>% 
  ggboxplot(x = 'value', y = 'simpson_reciprocal', add = 'jitter',
            title = 'toxicity', ylab = 'Simpson Reciprocal', xlab = '', color = 'value',
            palette = c('#0099B4','#AD002A'))  +
            stat_compare_means(comparisons= list(c('yes', 'no')),
            label= "p.format",
            method= 'wilcox.test',
            correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/diversity/combined_toxicity.pdf')

```

### 2021-03-02 scripts/51_distance_2_healthy.Rmd
did this centroid thing anyway

---
title: "16S distance to healthy people"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(vdbR)
connect_database('~/dbConfig.txt')
source('centroid_distance.R')
get_table_from_database("asv_alpha_diversity_ag")
get_table_from_database("asv_annotation_blast_ag")
get_table_from_database("healthy_volunteers_ag")
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
```

```{r}
healthy <- healthy_volunteers_ag %>% 
  inner_join(asv_alpha_diversity_ag, by = c("sampleid", "oligos_id"))

cts <- get_counts_subset(c(stb$sampleid, healthy %>% pull(sampleid)))

# get the genus of the asv_key
dat <- cts %>% 
  left_join(asv_annotation_blast_ag %>% 
              select(asv_key, genus))

# calculate the beta diversity
cbd <- compute_beta_diversity_and_tsne(sampleid = dat$sampleid,
                                      taxonomy = dat$genus,
                                      count =  dat$count);

cbd$compute_beta_diversity()

d_beta = cbd$get_betadiversity() 
```

```{r}
# assemble the df with sampleid  mrn and group
df <- bind_rows(
  tibble(
    sampleid = stb$sampleid,
    pid = as.character(stb$pt_id),
    center = stb$center,
    grp = 'CART'
  ),
  tibble(
    sampleid = healthy$sampleid,
    pid = healthy$volunteer_uid,
    center = 'healthy',
    grp = 'healthy'
  )
)
```

```{r}
# compute the distance to centroid combined and for individual centers
centroid_grp <- centroid_distance_selected_rows(
  M = d_beta, 
  selected_rows = df$sampleid,
  selected_rows_groups = df$grp, 
  ref_group = "healthy")

centroid_center <- centroid_distance_selected_rows(
  M = d_beta, 
  selected_rows = df$sampleid,
  selected_rows_groups = df$center, 
  ref_group = "healthy")
```

```{r}
# plotting
centroid_grp %>% 
  ggboxplot(x = 'groups', y = 'dist_to_centroid', add = 'jitter', xlab = '',
            ylab = 'Distance to healthy cohort', title = 'Per sampleid')  +
  stat_compare_means(comparisons= list(c('CART', 'healthy')),
            label= "p.format",
            method= 'wilcox.test',
						correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/distance_to_centroid/all_CART_compared_to_healthy.pdf')
```
```{r}
centroid_center %>% 
  ggboxplot(x = 'groups', y = 'dist_to_centroid', add = 'jitter', xlab = '',
            ylab = 'Distance to healthy cohort', title = 'Per sampleid')  +
  stat_compare_means(comparisons= list(c('M','P'), c('M', 'healthy'), c('P', 'healthy')),
            label= "p.format",
            method= 'wilcox.test',
						correct=FALSE) +
  ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/distance_to_centroid/centers_compared_to_healthy.pdf')
```

### 2021-03-02 scripts/centroid_distance.R

@@ -1,0 +1,60 @@
+#It computes the distance of points to `column-centroid` according to a reference `group` to compute centroid.
+#
+#
+
+library(data.table);
+
+#source('~/projects/general/library/antoniostats/Fishers_method.R'); #Use Fisher's method to aggregate p-value
+
+euclidean_norm <- function(x){
+  return( sqrt(sum(x^2)));
+}
+
+centroid_distance = function(M, groups, ref_group=NULL) {
+  
+  groups = factor(groups);
+  
+  if(is.null(ref_group)){
+    ref_group = levels(groups)[1];
+  }
+  
+  ind_ref = groups==ref_group;
+  centroid = apply(M[ind_ref,],2,mean);
+  
+  M_recentered = t(t(M) - centroid);
+  
+  dist_to_centroid = sqrt( apply((M_recentered)^2, 1, sum) );
+  
+  
+  d_set = data.table(dist_to_centroid=dist_to_centroid, id=names(dist_to_centroid), groups=groups, ref_group=ref_group);
+  return(d_set);
+  
+}
+
+centroid_distance_selected_rows = function(M, selected_rows, selected_rows_groups, ref_group=NULL) {
+  
+  if( ! all(selected_rows %in% row.names(M) ) ){
+    stop("all selected_row_groups must be in row.names of M");
+  }
+  
+  ind_matrix_selected = row.names(M) %in% selected_rows;
+  
+  M_selected = M[ind_matrix_selected,ind_matrix_selected];
+  
+  m_selected_order = merge( data.table(selected_rows_raw_order = rownames(M_selected)),
+                            data.table(selected_rows = selected_rows,
+                                       selected_rows_order = 1:length(selected_rows)),
+                            by.x="selected_rows_raw_order",
+                            by.y="selected_rows");
+  
+  M_selected =M_selected[order(m_selected_order$selected_rows_order),order(m_selected_order$selected_rows_order)];
+  #1
+  if(! all(row.names(M_selected) == selected_rows) ){
+    stop("I expected all row names in M_selected to match selected_rows in proper order.");
+  }
+  
+  d_set = centroid_distance(M_selected, selected_rows_groups, ref_group);
+  #d_set$M_selected = M_selected;
+  return(d_set);
+  
+}

### 2021-03-02 scripts/44_ggplot_venn_lefse.Rmd
this is actually pretty cool and it solves problems

@@
   bind_rows(.id = 'res') %>% 
-  mutate(tbl = if_else(str_detect(res, 'pcts'), 'pathway', if_else(str_detect(res, 'kocts'), 'KO', 'taxa'))) %>% 
+  mutate(tbl = if_else(str_detect(res, 'pcts'), 'pathway', 'taxa')) %>% 
   mutate(res = str_replace(res, '^.+//',''),
@@
   compact())
+
+res_all %>% 
+  write_csv('../data/shotgun_output/lefse_sig_results_taxa_and_pw.csv')
 

### 2021-03-02 scripts/52_shotgun_sig_pw_ontology.Rmd
this is actually pretty cool and it solves problems

---
title: "shotgun pathway ontology look"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
pw <- read_csv('../data/shotgun_output/lefse_sig_results_taxa_and_pw.csv') %>% 
  filter(tbl == 'pathway') %>% 
  mutate(pwid = str_extract(feature, '^.+_PWY|^PWY_\\d{3,4}'))  %>% 
  mutate(pwid = str_replace_all(pwid, '_', '-'))
```

```{r}
level <- read_csv('~/pipeline/scripts/shotgun_pipeline/data/metacyc_pathway_class_and_superclass_levels.csv', col_types = 'ccccccccccccccccc')

dat <- pw %>% 
  inner_join(level)

dat_list <- dat %>% 
  arrange(cohort, outcome , direction, -score) %>% 
  split(list(.$cohort, .$outcome , .$direction)) %>% 
  discard( ~nrow(.) == 0)

dat_list %>% 
  imap(function(.x, .y){
    .x %>% 
      write_csv(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/lefse_sig_pathways_ontology_tables/{.y}_result.csv'))
  })
```

### 2021-03-08 scripts/45_16s_sorting_out.Rmd
ok slowly picking up the work

@@
 uptb <- read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/U Penn/patients/upenn_essential_data_v3.xlsx', sheet = 3) %>% 
-  filter(!is.na(gender))
-
+  filter(sequenced == 'yes')
 ```
@@
   mutate(center = 'M') %>% 
-  mutate(icansgrade = as.character(icansgrade))
+  select(-cr_overall) # this column is now gone
 
@@
 ```{r}
+library(vdbR)
+connect_database('~/dbConfig.txt')
 get_table_from_database('asv_alpha_diversity_ag')

### 2021-03-08 scripts/48_16s_beta_diversity.Rmd
ok slowly picking up the work

@@
 
+# find the dominating genus in each sample
+domi <- genus %>% 
+  arrange(sampleid, -relab) %>% 
+  group_by(sampleid) %>% 
+  slice(1) %>% 
+  rename(domgenus = genus) %>% 
+  select(sampleid, domgenus)
+
 dat <- genus %>% 
@@
   spread(key = 'genus', value = 'relab') %>% 
-  full_join(pheno)
+  full_join(pheno) %>% 
+  full_join(domi)
 
@@
 # a function to do the combinations of the anosim test
-pheno4 <- colnames(pheno)[2:5]
+pheno_ <- c(colnames(pheno)[2:5] , 'domgenus')
+n <- length(pheno_)
 
 combi <- data_frame(
-  g1 = c(rep('M',4), rep('P',4), rep('M',4)),
-  g2 = c(rep('M',4), rep('P',4), rep('P',4)),
-  label = rep(pheno4, 3)
+  g1 = c(rep('M',n), rep('P',n), rep('M',n)),
+  g2 = c(rep('M',n), rep('P',n), rep('P',n)),
+  label = rep(pheno_, 3)
 )
@@
 do_anosim_each <-  function(g1_, g2_, label_){
+  g1_ = 'M'
+  g2_ = 'P'
+  label_ = 'domgenus'
     df = dat %>% 
@@
       select(sampleid, matches(label_))
-
-    # anosim run
-    ret = anosim(cts, phe, permutations = 999, distance = "bray", strata = NULL, parallel = 4)
-
-    # plot the pcoa
+    # 
+    # # anosim run
+    # ret = anosim(cts, phe, permutations = 999, distance = "bray", strata = NULL, parallel = 4)
+    # 
+    # # plot the pcoa
     dist_ <- vegdist(cts, method = 'bray')
     bc <- cmdscale(dist_, k = 2)
+    # bc %>% 
+    #   as.data.frame() %>% 
+    #   rownames_to_column('sampleid') %>% 
+    #   full_join(phedf, by = 'sampleid') %>% 
+    #   ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
+    #   labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
+    #   ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))
+
+    # the version without the anosim stats
     bc %>% 
@@
       ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
-      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
+      labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}')) +
       ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))
 
-  return(list(pval = ret$signif, 
-              statistic = ret$statistic))
+  # return(list(pval = ret$signif, 
+  #             statistic = ret$statistic))
 }
@@
 ```
+
+```{r}
+# do the pcoa colored by the dominating genus 
+do_domgenus_coloring_each <-  function(g1_, g2_, label_){
+  g1_ = 'M'
+  g2_ = 'P'
+  label_ = 'domgenus'
+    df = dat %>% 
+      filter(center %in% c(g1_, g2_)) 
+    cts = df %>% 
+      select(sampleid, all_of(keep)) %>% 
+      column_to_rownames('sampleid')
+    phe = df %>% 
+      select(matches(label_)) %>% pull(get(label_)) 
+    phedf = df %>% 
+      select(sampleid, matches(label_))
+
+    # # plot the pcoa
+    dist_ <- vegdist(cts, method = 'bray')
+    bc <- cmdscale(dist_, k = 2)
+
+    # the version without the anosim stats
+    bc %>% 
+      as.data.frame() %>% 
+      rownames_to_column('sampleid') %>% 
+      full_join(phedf, by = 'sampleid') %>% 
+      ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
+      labs(title = str_glue('PCoA  combined {label_}')) +
+      ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))
+
+}
+```

### 2021-03-08 scripts/48_16s_beta_diversity.Rmd
slowly but making progress

--- 2021-02-17 scripts/48_16s_beta_diversity.Rmd
@@
 # a function to do the combinations of the anosim test
-pheno4 <- colnames(pheno)[2:5]
+pheno_ <- colnames(pheno)[2:5] 
+n <- length(pheno_)
 
 combi <- data_frame(
-  g1 = c(rep('M',4), rep('P',4), rep('M',4)),
-  g2 = c(rep('M',4), rep('P',4), rep('P',4)),
-  label = rep(pheno4, 3)
+  g1 = c(rep('M',n), rep('P',n), rep('M',n)),
+  g2 = c(rep('M',n), rep('P',n), rep('P',n)),
+  label = rep(pheno_, 3)
 )
@@
 do_anosim_each <-  function(g1_, g2_, label_){
+
     df = dat %>% 
@@
       labs(title = str_glue('PCoA {g1_}_{g2_}__{label_}\npval:{ret$signif} statistic:{round(ret$statistic, 2)}')) +
+      theme_void() +
       ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))
@@
 ```
+
+```{r}
+# do the pcoa colored by the dominating genus 
+g1_ = 'M'  
+g2_ = 'P'
+label_ = 'domgenus'
+df = dat %>% 
+  filter(center %in% c(g1_, g2_)) 
+cts = df %>% 
+  select(sampleid, all_of(keep)) %>% 
+  column_to_rownames('sampleid')
+phe = df %>% 
+  select(matches(label_)) %>% pull(get(label_)) 
+phedf = df %>% 
+  select(sampleid, matches(label_))
+
+# # plot the pcoa
+dist_ <- vegdist(cts, method = 'bray')
+bc <- cmdscale(dist_, k = 2)
+
+# the version without the anosim stats
+bc %>% 
+  as.data.frame() %>% 
+  rownames_to_column('sampleid') %>% 
+  full_join(phedf, by = 'sampleid') %>% 
+  ggplot(aes_string(x = "V1", y = "V2", color =  label_)) + geom_point() +
+  labs(title = str_glue('Combined cohort colored by dominating genus')) +
+  theme_void() +
+  ggsave(str_glue('../figs/beta_{g1_}_{g2_}__{label_}.pdf'))
+```

### 2021-03-08 scripts/49_16s_stacked_barplot.Rmd
great slowly but making progress

@@
 
+# so if the genus is unknown then it's gonna be assigned "other" gray color  
 # the question is do we go one taxa level higher or make a new color base and shades for the new asv
+# after discussing with Tsoni, we decided that it's ok to assign gray to the unknown genus
+
+# merge the new no_color_df_full to the original df
+dat <- bind_rows(
+  no_color_df_full,
+  color_group %>% 
+    pluck('FALSE')
+)   
 ```
@@
   theme_classic() +
+  labs(title = 'stacked bar sorted with hclust') +
   theme(axis.text.x = element_text(angle = 90),
@@
         legend.position = "none") +
-  scale_fill_manual(values = asv_color_set)
+  scale_fill_manual(values = asv_color_set) +
+  ggsave('../figs/amplicon/stacked_bar_sorted_with_hclust.pdf', width = 7, height = 5)
 
@@
   theme_classic() +
+  labs(title = 'stacked bar sorted with hclust and divided to lower and higher diversity') +
   theme(axis.text.x = element_text(angle = 90),
@@
         legend.position = "none") +
-  scale_fill_manual(values = asv_color_set)
+  scale_fill_manual(values = asv_color_set) +
+  ggsave('../figs/amplicon/stacked_bar_sorted_with_hclust_lower_and_higher_diversity.pdf', width = 7, height = 5)
 ```

### 2021-03-09 scripts/53_bayesian_5fam.Rmd
finally did some real coding after 5 PM

---
title: "Bayesian 5 taxa family"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(brms)   
library(ggmcmc)
```

Model:
- CR_d100 ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

- Toxicity ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus

```{r}
# the data
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus')) %>% 
  select(-center) %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  select(-relab) %>% 
  spread(key = 'genus', value = 'relablog')

dat <- meta %>% 
  full_join(genus) %>% 
  mutate(toxicity = factor(toxicity, levels = c('no','yes')))

dat_cr <- dat %>% 
  filter(cr_d100 %in% c('yes','no')) %>% 
  mutate(cr_d100 = factor(cr_d100, levels = c('no','yes')))
```

```{r}
# the model
tox <- brm( toxicity  ~ 
              Akkermansia +
              Bacteroides +
              Enterococcus +
              Faecalibacterium +
              Ruminococcus +
               (1 | pt_id) ,  
              data = dat, 
              family = bernoulli(), 
              warmup = 1000, iter = 3000, 
              cores = 8, chains = 4, 
              seed = 123)
```

### 2021-03-09 scripts/44_ggplot_venn_lefse.Rmd
udpate

@@
                  y = 'Score')+ 
-            ggsave(str_glue('../data/shotgun_output/humann3/pull_{.y}_pathway.pdf'), device = 'pdf', width = 8))
+            ggsave(str_glue('../data/shotgun_output/humann3/pull_{.y}_pathway.pdf'), device = 'pdf', width = 12, height = 20))
             #ggsave(str_glue('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/shotgun_results/NEW_humann3_version/pathway/pull_{.y}_pathway.pdf'), device = 'pdf', width = 12))

### 2021-03-09 scripts/53_bayesian_5fam.Rmd
udpate

@@
 title: "Bayesian 5 taxa family"
-output: html_document
+author: "Anqi Dai"
+date: "`r Sys.Date()`"
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3 
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
 ```
@@
 Model:
+
 - CR_d100 ~ Faecalibacterium + Ruminococcus + Akkermansia + Bacteroides + Enterococcus
@@
               family = bernoulli(), 
-              warmup = 1000, iter = 3000, 
-              cores = 8, chains = 4, 
+              warmup = 5000, iter = 40000, 
+              cores = 8, chains = 2, 
               seed = 123)
+
+# It's giving errors: "Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
+# Running the chains for more iterations may help"
 ```
+
+```{r}
+# Sean suggested to use individual genus first, good idea. 
+# do 5 iteratively in a loop
+target <- c('Akkermansia','Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus')
+
+ret <- target %>% 
+  set_names(target) %>% 
+  map(function(genus) {
+      mod =  brm( as.formula(str_glue('toxicity  ~ 
+              {genus} +
+               (1 | pt_id)')),  
+             data = dat, 
+              family = bernoulli(), 
+              warmup = 3000, iter = 10000, 
+              cores = 8, chains = 2, 
+              seed = 123)
+
+      res = posterior_samples(mod) %>% 
+              select(starts_with('b')) %>% 
+              select(-b_Intercept)
+      return(res)
+  }) %>% 
+  bind_cols()
+
+```

### 2021-03-10 scripts/45_16s_sorting_out.Rmd
slowly but making progress
slowly but getting things going ...

@@
 mptb <-  read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/patients/msk_essential_data_v4.xlsx', sheet = 1) %>% 
-  mutate(mrn = as.numeric(mrn)) %>% 
-  filter(! mrn %in% mptb_ineli$pt_id)
+  filter(! pt_id %in% mptb_ineli$pt_id)
 
@@
   filter(str_detect(time_group, '^Baseline')) %>% 
-  inner_join(mptb, by = "mrn")
+  inner_join(mptb, by = c('mrn' = 'pt_id'))
 
@@
   mutate(center = 'M') %>% 
-  select(-cr_overall) # this column is now gone
+  select(pt_id:vital_status, center) # there are several more columns in the msk table now
 
@@
   write_csv('../data/amplicon/stool/combined_2_meta.csv')
+
+combined_join %>% 
+  write_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/16s_data_pheno_comprehensive.csv')
 ```

### 2021-03-10 scripts/53_bayesian_5fam.Rmd
slowly but getting things going ...

@@
 # the model 
-tox <- brm( toxicity  ~ 
-              Akkermansia +
-              Bacteroides +
-              Enterococcus +
-              Faecalibacterium +
-              Ruminococcus +
-               (1 | pt_id) ,  
-              data = dat, 
-              family = bernoulli(), 
-              warmup = 5000, iter = 40000, 
-              cores = 8, chains = 2, 
-              seed = 123)
+# tox <- brm( toxicity  ~ 
+#               Akkermansia +
+#               Bacteroides +
+#               Enterococcus +
+#               Faecalibacterium +
+#               Ruminococcus +
+#                (1 | pt_id) ,  
+#               data = dat, 
+#               family = bernoulli(), 
+#               warmup = 5000, iter = 40000, 
+#               cores = 8, chains = 2, 
+#               seed = 123)
 
@@
   }) %>%  
-  bind_cols()
+  bind_cols() %>% 
+  gather('genus','coeff')
+```
 
+```{r}
+# plotting in ridge plot
+df <- ret %>% 
+  group_by(genus) %>% 
+  mutate(meanpergenus = mean(coeff))%>%
+  ungroup()  %>% 
+  mutate(genus = str_replace(genus, 'b_',''))
+
+ggplot()+
+  ggridges::geom_density_ridges(data  = df, 
+                                aes(x      = coeff,
+                                    y      = reorder(as.factor(genus), meanpergenus),
+                                    height = ..density..),
+                                scale = 1, size = 0.25,
+                                alpha = 0.6) +
+  geom_vline(xintercept = 0, col = "red", linetype = 'dashed', size = 1) +
+  #scale_x_continuous(limits = c(-0.02,0.03)) +
+  labs(y = '',
+       title = 'Response: toxicity') +
+  ggthemes::theme_tufte() +
+  theme(axis.text.y = element_text(size=20),
+        axis.text.x = element_text(size=15)) +
+  ggsave('../figs/bayesian_genus_5_each_toxicity.jpg', width = 8)    
 ```
+
+```{r}
+# also when the response is cr_d100
+ret_CR <- target %>% 
+  set_names(target) %>% 
+  map(function(genus) {
+      mod =  brm( as.formula(str_glue('cr_d100  ~ 
+              {genus} +
+               (1 | pt_id)')),  
+             data = dat_cr, 
+              family = bernoulli(), 
+              warmup = 3000, iter = 10000, 
+              cores = 8, chains = 2, 
+              seed = 123)
+
+      res = posterior_samples(mod) %>% 
+              select(starts_with('b')) %>% 
+              select(-b_Intercept)
+      return(res)
+  }) %>%  
+  bind_cols() %>% 
+  gather('genus','coeff')
+```
+```{r}
+ret_CR %>% 
+  group_by(genus) %>% 
+  mutate(meanpergenus = mean(coeff))%>%
+  ungroup()  %>% 
+  mutate(genus = str_replace(genus, 'b_','')) %>% 
+  ggplot()+
+  ggridges::geom_density_ridges(aes(x      = coeff,
+                                    y      = reorder(as.factor(genus), meanpergenus),
+                                    height = ..density..),
+                                scale = 1, size = 0.25,
+                                alpha = 0.6) +
+  geom_vline(xintercept = 0, col = "red", linetype = 'dashed', size = 1) +
+  #scale_x_continuous(limits = c(-0.02,0.03)) +
+  labs(y = '',
+       title = 'Response: cr_d100') +
+  ggthemes::theme_tufte() +
+  theme(axis.text.y = element_text(size=20),
+        axis.text.x = element_text(size=15)) +
+  ggsave('../figs/bayesian_genus_5_each_CR.jpg', width = 8)   
+```

### 2021-03-11 scripts/54_rethinking_binary.Rmd
scary but a first model

---
title: "the rethinking way of doing the binary outcome modeling"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)  
```

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
  mutate(logdiv_s = scale(log(meta$simpson_reciprocal))) %>% 
  mutate(tox = if_else(toxicity == 'yes', 1, 0),
         cr100 = if_else(cr_d100 == 'yes', 1, 0),
         loca = if_else(center == 'M', 1, 2))
```

```{r}
# visualize the var distribution 
table(meta$cr_d100)  
table(meta$toxicity)
hist(meta$simpson_reciprocal)
hist(log(meta$simpson_reciprocal)) # I think I should use log transformed
table(meta$center)
```
```{r}
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    logdivs = meta$logdiv_s
    )

mtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- b*logdivs + a[location] ,
        b ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE )
```

### 2021-03-12 scripts/54_rethinking_binary.Rmd
it looks good then it is not

@@
 title: "the rethinking way of doing the binary outcome modeling"
-output: html_document
+author: "Anqi Dai"
+date: "`r Sys.Date()`"
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
 ---
@@
 library(tidyverse)  
+library(haven) #for reading sav data
+library(sjstats) #for calculating intra-class correlation (ICC)
+library(effects) #for plotting parameter effects
+library(jtools) #for transformaing model summaries
+library(ROCR)
 ```
+
+# CR/Toxicity ~ alpha diversity + center
+
+## Toxicity
 
@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
-  mutate(logdiv_s = scale(log(meta$simpson_reciprocal))) %>% 
+  mutate(div_std = scale(simpson_reciprocal)) %>%  # diversity is scaled 
   mutate(tox = if_else(toxicity == 'yes', 1, 0),
          cr100 = if_else(cr_d100 == 'yes', 1, 0),
-         loca = if_else(center == 'M', 1, 2))
+         loca = if_else(center == 'M', 1, 2)) %>%  # MSK 1 ; Upenn 2
+  mutate(center = factor(center),
+         toxicity = factor(toxicity, levels = c('no','yes')),
+         cr_d100 = factor(cr_d100, levels = c('no','yes')))
+
+meta %>% write_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')
 ```
@@
 table(meta$center)
+table(meta$toxicity, meta$center)
+table(meta$cr_d100, meta$center)   
+range(meta$div_std)
 ```
 ```{r}
+set.seed(123)
 dat_list <- list(
@@
     location = meta$loca,
-    logdivs = meta$logdiv_s
+    div_std = meta$div_std
     )
@@
         tox ~ dbinom( 1 , p ) ,
-        logit(p) <- b*logdivs + a[location] ,
-        b ~ dnorm( 0 , 1),
+        logit(p) <- b*div_std + a[location] ,
+        b ~ dnorm( 0 , 2), # the std of 2 is made by the later prior check
         a[location] ~ dnorm( 0 , 0.5 )
-    ) , data=dat_list , chains=4 , log_lik=TRUE )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+
+precis(mtox, depth = 2)
+plot(precis(mtox, depth = 2) ) 
+# there is baseline difference between penn and Msk and MSK has higher tox rage in general.
+# it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
 ```
+
+```{r}
+# doing a prior check
+prior <- extract.prior( mtox , n=1e4 ) 
+# don't know how to put in logdivs into this prior check
+divs_samp <- sample(meta$div_std, size = 1e4, replace = T)
+p <- sapply( 1:2 , function(k) inv_logit( prior$b * divs_samp + prior$a[,k] ) )
+mean( abs( p[,1] - p[,2] ) )
+dens(abs( p[,1] - p[,2] ), adj = 0.1) # seems the difference is a bit too high
+```
+
+```{r}
+# posterior prediction check
+
+# the raw proportion of yes toxicity in each place
+rawtox <- meta  %>% 
+  count(center, toxicity) %>% 
+  left_join(meta %>% 
+              count(center, name = 'total'), by = 'center') %>% 
+  group_by(center, toxicity) %>% 
+  summarise(frac = round(n/total, 2)) %>% 
+  filter(toxicity == 'yes')
+
+# don't know how to do the post check since the diversity is continuous variable
+
+```
+
+```{r}
+# I also want to do it in the glm way so that I can compare
+lm_tox <- glm(formula = toxicity ~ simpson_reciprocal + center,
+                    family = binomial(link = "logit"),
+                    data = meta)
+#summary(lm_tox)
+summ(lm_tox, exp = T)# this is the log odds
+
+```
+
+## CR_d100
+
+```{r}
+mcr <- ulam( 
+    alist( 
+        cr100 ~ dbinom( 1 , p ) ,
+        logit(p) <- b*div_std + a[location] ,
+        b ~ dnorm( 0 , 2), # the std of 2 is made by the later prior check
+        a[location] ~ dnorm( 0 , 0.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+
+precis(mcr, depth = 2)
+plot(precis(mcr, depth = 2) ) 
+# there isn't much difference in baseline level between the centers
+# diversity seems to be positively associating with CR
+
+# if switching to standardized diversity, the positive association is less sure.
+
+```
+
+```{r}
+# doing a prior check
+prior <- extract.prior( mcr , n=1e4 ) 
+# don't know how to put in logdivs into this prior check
+divs_samp <- sample(meta$div_std, size = 1e4, replace = T)
+p <- sapply( 1:2 , function(k) inv_logit( prior$b * divs_samp + prior$a[,k] ) )
+mean( abs( p[,1] - p[,2] ) )
+dens(abs( p[,1] - p[,2] ), adj = 0.1)
+```
+
+```{r}
+# I also want to do it in the glm way so that I can compare
+lm_cr <- glm(formula = cr_d100 ~ simpson_reciprocal + center,
+                    family = binomial(link = "logit"),
+                    data = meta)
+#summary(lm_tox)
+summ(lm_cr, exp = T)# this is the log odds
+```

### 2021-03-12 scripts/55_rethinking_abx.Rmd
it looks good then it is not

---
title: "rethinking antibiotics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)  
library(haven) #for reading sav data
library(sjstats) #for calculating intra-class correlation (ICC)
library(effects) #for plotting parameter effects
library(jtools) #for transformaing model summaries
library(ROCR)
```

```{r}
# none of the patients from Penn received the 3 anaerobe-targeting antibiotics in the 4 weeks prior to CAR T cell infusion
# msk table
m <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/antibiotics/sean_table/abx_table_msk_sean.xlsx') %>% 
  select(pt_id, anaerobe_targeting)

meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv') %>% 
  left_join(m)
```

### 2021-03-12 scripts/55_rethinking_abx.Rmd
ok the abx seems easier

@@
 
+# CR/Toxicity ~ abx + center
+
+## Toxicity
+
 ```{r}
@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv') %>% 
-  left_join(m)
+  left_join(m) %>% 
+  mutate(anaerobe_targeting = if_else(is.na(anaerobe_targeting), 'no', anaerobe_targeting)) %>% 
+  mutate(anaerobe = if_else(anaerobe_targeting == 'no', 0, 1))
+
+set.seed(123)
+dat_list <- list(
+    tox = meta$tox,
+    cr100 = meta$cr100,
+    location = meta$loca,
+    anaerobe = meta$anaerobe
+    )
 ```
+
+```{r}   
+atox <- ulam(
+    alist(
+        tox ~ dbinom( 1 , p ) ,
+        logit(p) <- b*div_std + a[location] ,
+        b ~ dnorm( 0 , 2), # the std of 2 is made by the later prior check
+        a[location] ~ dnorm( 0 , 0.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+precis(mtox, depth = 2)
+plot(precis(mtox, depth = 2) ) 
+```

### 2021-03-12 scripts/53_bayesian_5fam.Rmd
update

@@
   mutate(cr_d100 = factor(cr_d100, levels = c('no','yes')))
+
+# write out the sample and the log 10 genera 
+dat %>% 
+  select(sampleid, Akkermansia:Ruminococcus) %>% 
+  write_csv('../data/amplicon/stool/combined_5_genera.csv')
 ```

### 2021-03-12 scripts/55_rethinking_abx.Rmd
update

--- 2021-03-12 scripts/55_rethinking_abx.Rmd
@@
 title: "rethinking antibiotics"
-output: html_document
+author: "Anqi Dai"
+date: "`r Sys.Date()`"
+output:
+  html_document:
+    toc: true
+    toc_float: true
+    toc_depth: 3
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
 ```
@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv') %>% 
-  left_join(m)
+  left_join(m) %>% 
+  mutate(anaerobe_targeting = if_else(is.na(anaerobe_targeting), 'no', anaerobe_targeting)) %>% 
+  mutate(anaerobe = if_else(anaerobe_targeting == 'no', 1, 2))
+
+set.seed(123)
+dat_list <- list(
+    tox = meta$tox,
+    cr100 = meta$cr100,
+    location = meta$loca,
+    anaerobe = meta$anaerobe
+    )
+
+meta %>% 
+  count(center, anaerobe_targeting) 
 ```
+
+```{r}   
+atox <- ulam(
+    alist(
+        tox ~ dbinom( 1 , p ) ,
+        logit(p) <- b[anaerobe] + a[location] ,
+        b[anaerobe] ~ dnorm( 0 , 1), 
+        a[location] ~ dnorm( 0 , 0.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+precis(atox, depth = 2)
+plot(precis(atox, depth = 2) ) 
+# seems anaerobe abx is positively linked with toxicity 
+```
+
+## CR day 100
+
+```{r}
+acr <- ulam(  
+    alist(
+        cr100 ~ dbinom( 1 , p ) ,
+        logit(p) <- b[anaerobe] + a[location] ,
+        b[anaerobe] ~ dnorm( 0 , 1), 
+        a[location] ~ dnorm( 0 , 0.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+precis(acr, depth = 2)
+plot(precis(acr, depth = 2) ) 
+# nothing interesting 
+```

### 2021-03-12 scripts/56_rethinking_genera.Rmd
update

---
title: "rethinking and the 5 genera"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

```{r}
library(rethinking)
library(tidyverse)  
```

# CR/Toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center

## Toxicity

```{r}
genera <- read_csv('../data/amplicon/stool/combined_5_genera.csv')

meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
  full_join(genera) %>% 
  mutate(A_std = scale(Akkermansia),
         B_std = scale(Bacteroides),
         E_std = scale(Enterococcus),
         F_std = scale(Faecalibacterium),
         R_std = scale(Ruminococcus))

set.seed(123)   
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    A_std = meta$A_std,
    B_std = meta$B_std,
    E_std = meta$E_std,
    F_std = meta$F_std,
    R_std = meta$R_std
    ) 
```

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*A_std + bb*B_std + be*E_std + bf*F_std + br*R_std + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gtox, depth = 2)
plot(precis(gtox, depth = 2) )  
# indicative of the Bacteroides link with toxicity  
```

## CR

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*A_std + bb*B_std + be*E_std + bf*F_std + br*R_std + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, depth = 2)
plot(precis(gcr, depth = 2) ) 
# I think the Ruminococcus still holds for the CR
```

### 2021-03-12 scripts/54_rethinking_binary.Rmd
yeah finally am able to concentrate and make some progress

--- 2021-03-11 scripts/54_rethinking_binary.Rmd
@@
 
+# CR/Toxicity ~ alpha diversity + center
+
+## Toxicity
+
 ```{r}
 meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
-  mutate(logdiv_s = scale(log(meta$simpson_reciprocal))) %>% 
+  mutate(logdiv_s = scale(log(meta$simpson_reciprocal))) %>%  # diversity is scaled 
   mutate(tox = if_else(toxicity == 'yes', 1, 0),
          cr100 = if_else(cr_d100 == 'yes', 1, 0),
-         loca = if_else(center == 'M', 1, 2))
+         loca = if_else(center == 'M', 1, 2)) %>%  # MSK 1 ; Upenn 2
+  mutate(center = factor(center),
+         toxicity = factor(toxicity, levels = c('no','yes')),
+         cr_d100 = factor(cr_d100, levels = c('no','yes')))
+
+meta %>% write_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')
 ```
@@
 table(meta$toxicity)
-hist(meta$simpson_reciprocal)
-hist(log(meta$simpson_reciprocal)) # I think I should use log transformed
+#hist(meta$simpson_reciprocal)
+#hist(log(meta$simpson_reciprocal)) # I think I should use log transformed
 table(meta$center)
+table(meta$toxicity, meta$center)
+table(meta$cr_d100, meta$center)   
+range(meta$logdiv_s)
 ```
+
 ```{r}
+set.seed(123)
 dat_list <- list(
@@
     location = meta$loca,
-    logdivs = meta$logdiv_s
+    logdiv_s = meta$logdiv_s
     )
@@
         tox ~ dbinom( 1 , p ) ,
-        logit(p) <- b*logdivs + a[location] ,
-        b ~ dnorm( 0 , 1),
+        logit(p) <- b*logdiv_s + a[location] ,
+        b ~ dnorm( 0 , 0.25), 
         a[location] ~ dnorm( 0 , 0.5 )
-    ) , data=dat_list , chains=4 , log_lik=TRUE )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+
+precis(mtox, depth = 2)
+plot(precis(mtox, depth = 2) ) 
+# there is baseline difference between penn and Msk and MSK has higher tox rage in general.
+# it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
 ```
+
+### Prior
+
+```{r}
+# doing a prior check
+prior <- extract.prior( mtox , n=1e4 ) 
+# the diff in the probabilty of yes to toxicity when there are two std diff in diversity in each of the centers 
+# for msk  
+div_value <- c(-1, 1)
+p <- sapply( 1:2 , function(k) inv_logit( prior$b * div_value[k] + prior$a[,1] ) )
+mean( abs( p[,1] - p[,2] ) )
+# the probability difference is pretty big !  # so reduced the std of the b from 1.5 to 0.25
+
+# for upenn
+p <- sapply( 1:2 , function(k) inv_logit( prior$b * div_value[k] + prior$a[,2] ) )
+mean( abs( p[,1] - p[,2] ) )
+```
+
+### Posterior
+
+```{r}
+# posterior prediction check
+post <- extract.samples(mtox)
+post_tox <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(div_value)) %>% 
+      set_names(seq(1, length(div_value))) %>% 
+      map(function(idx){
+         inv_logit( post$b * div_value[idx] + post$a[, center_] ) 
+      })
+
+    ret = mean(abs( cols[[1]] -cols[[2]] ) )
+    return(ret)
+
+  })
+
+```
+
+```{r}
+# the raw proportion of yes toxicity in each place
+rawtox <- meta  %>% 
+  count(center, toxicity) %>% 
+  left_join(meta %>% 
+              count(center, name = 'total'), by = 'center') %>% 
+  group_by(center, toxicity) %>% 
+  summarise(frac = round(n/total, 2)) %>% 
+  filter(toxicity == 'yes')
+
+# don't know how to do the post check since the diversity is continuous variable
+
+```
+
+## CR_d100
+
+```{r}
+mcr <- ulam( 
+    alist( 
+        cr100 ~ dbinom( 1 , p ) ,
+        logit(p) <- b*logdiv_s + a[location] ,
+        b ~ dnorm( 0 , 1.5), 
+        a[location] ~ dnorm( 0 , 0.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+
+precis(mcr, depth = 2)
+plot(precis(mcr, depth = 2) ) 
+# there isn't much difference in baseline level between the centers
+# diversity seems to be positively associating with CR
+
+# if switching to standardized diversity, the positive association is less sure.
+
+```
+
+```{r}
+# doing a prior check
+prior <- extract.prior( mcr , n=1e4 ) 
+# don't know how to put in logdivs into this prior check
+divs_samp <- sample(meta$logdiv_s, size = 1e4, replace = T)
+p <- sapply( 1:2 , function(k) inv_logit( prior$b * divs_samp + prior$a[,k] ) )
+mean( abs( p[,1] - p[,2] ) )
+dens(abs( p[,1] - p[,2] ), adj = 0.1)
+```

### 2021-03-12 scripts/55_rethinking_abx.Rmd
yeah finally am able to concentrate and make some progress

--- 2021-03-12 scripts/55_rethinking_abx.Rmd
@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv') %>% 
-  left_join(m)
+  left_join(m) %>% 
+  mutate(anaerobe_targeting = if_else(is.na(anaerobe_targeting), 'no', anaerobe_targeting)) %>% 
+  mutate(anaerobe = if_else(anaerobe_targeting == 'no', 0, 1))
+
+set.seed(123)
+dat_list <- list(
+    tox = meta$tox,
+    cr100 = meta$cr100,
+    location = meta$loca,
+    anaerobe = meta$anaerobe
+    )
+
+meta %>% 
+  count(center, anaerobe_targeting) 
 ```
+
+```{r}   
+atox <- ulam(
+    alist(
+        tox ~ dbinom( 1 , p ) ,
+        logit(p) <- b*anaerobe + a[location] ,
+        b ~ dnorm( 0 , 1), 
+        a[location] ~ dnorm( 0 , 0.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+precis(atox, depth = 2)
+plot(precis(atox, depth = 2) ) 
+# seems anaerobe abx is positively linked with toxicity 
+```
+
+## CR day 100
+
+```{r}
+acr <- ulam(  
+    alist(
+        cr100 ~ dbinom( 1 , p ) ,
+        logit(p) <- b*anaerobe + a[location] ,
+        b ~ dnorm( 0 , 1), 
+        a[location] ~ dnorm( 0 , 0.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+precis(acr, depth = 2)
+plot(precis(acr, depth = 2) ) 
+# nothing interesting 
+```

### 2021-03-12 scripts/56_rethinking_genera.Rmd
yeah finally am able to concentrate and make some progress

@@
     location = meta$loca,
-    A_std = meta$A_std,
-    B_std = meta$B_std,
-    E_std = meta$E_std,
-    F_std = meta$F_std,
-    R_std = meta$R_std
+    Akkermansia = meta$Akkermansia,
+    Bacteroides = meta$Bacteroides,
+    Enterococcus = meta$Enterococcus,
+    Faecalibacterium = meta$Faecalibacterium,
+    Ruminococcus = meta$Ruminococcus
+
     ) 
@@
         tox ~ dbinom( 1 , p ) ,
-        logit(p) <- ba*A_std + bb*B_std + be*E_std + bf*F_std + br*R_std + a[location] ,
+        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + a[location] ,
         ba ~ dnorm( 0 , 1),
@@
         cr100 ~ dbinom( 1 , p ) ,
-        logit(p) <- ba*A_std + bb*B_std + be*E_std + bf*F_std + br*R_std + a[location] ,
+        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + a[location] ,
         ba ~ dnorm( 0 , 1),

### 2021-03-12 scripts/54_rethinking_binary.Rmd
do you know how hard is to make that plot

@@
 library(tidyverse) 
+library(ggpubr)
 ```
@@
 div_value <- c(-1, 1)
-p <- sapply( 1:2 , function(k) inv_logit( prior$b * div_value[k] + prior$a[,1] ) )
-mean( abs( p[,1] - p[,2] ) )
-# the probability difference is pretty big !  # so reduced the std of the b from 1.5 to 0.25
 
-# for upenn
-p <- sapply( 1:2 , function(k) inv_logit( prior$b * div_value[k] + prior$a[,2] ) )
-mean( abs( p[,1] - p[,2] ) )
+# make it into a function so that I can do the m and p the same time
+prior_tox <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(div_value)) %>% 
+      set_names(seq(1, length(div_value))) %>% 
+      map(function(idx){
+         inv_logit( prior$b * div_value[idx] + prior$a[, center_] ) 
+      })
+
+    diff = abs( cols[[1]] -cols[[2]] )
+    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(ret)
+  }) 
 ```
@@
 
-    ret = mean(abs( cols[[1]] -cols[[2]] ) )
+    diff = abs( cols[[1]] -cols[[2]] )
+    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
     return(ret)
+  }) 
+```
 
-  })
-
+```{r}
+# visualize the prior and post together
+tox <- bind_rows(
+  bind_cols(prior_tox, grp ='prior', quant = c(0.025, 0.5, 0.975)),
+  bind_cols(post_tox, grp ='post', quant = c(0.025, 0.5, 0.975))
+)  %>% 
+  gather('center', 'value', m:p) %>% 
+  spread('quant', 'value') %>% 
+  mutate(grp = factor(grp, levels = c('prior','post')))
+tox
+```
+```{r}
+tox %>% 
+  ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 std apart') +
+  geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') +
+  ylim(0,1) 
 ```
@@
   filter(toxicity == 'yes')
-
-# don't know how to do the post check since the diversity is continuous variable
 

### 2021-03-14 scripts/54_rethinking_binary.Rmd
udpate

@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
-  mutate(logdiv_s = scale(log(meta$simpson_reciprocal))) %>%  # diversity is scaled 
+  mutate(logdiv_s = scale(log(simpson_reciprocal))) %>%  # diversity is scaled 
   mutate(tox = if_else(toxicity == 'yes', 1, 0),
@@
         logit(p) <- b*logdiv_s + a[location] ,
-        b ~ dnorm( 0 , 0.25), 
+        b ~ dnorm( 0 , 2), 
         a[location] ~ dnorm( 0 , 0.5 )
@@
 
-precis(mtox, depth = 2)
-plot(precis(mtox, depth = 2) ) 
+# the std of b is 2, since we assume the microbiome could have big impact on immunology (based on our previous understanding and knowledge )
+#precis(mtox, depth = 2)
+plot(precis(mtox, depth = 2, prob = 0.95) ) 
 # there is baseline difference between penn and Msk and MSK has higher tox rage in general.
@@
     ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
-    return(ret)
-  }) 
+    return(cols)
+  }) # here I'm returning the cols instead of the diff 
+```
+
+```{r}
+# posterior check for tox
+# see the coeff distribution directly without going to the difference 
+c('m','p') %>% 
+  map(function(x){ tibble(
+           post_low = post_tox %>% pluck(x) %>% pluck(1),
+           post_high = post_tox %>% pluck(x) %>% pluck(2)
+          )  %>% gather() %>% 
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+      facet_grid(key ~ .)})
 ```
@@
         logit(p) <- b*logdiv_s + a[location] ,
-        b ~ dnorm( 0 , 1.5), 
+        b ~ dnorm( 0 , 2),  
         a[location] ~ dnorm( 0 , 0.5 )
@@
 
-precis(mcr, depth = 2)
-plot(precis(mcr, depth = 2) ) 
+#precis(mcr, depth = 2)
+plot(precis(mcr, depth = 2, prob = 0.95) ) 
 # there isn't much difference in baseline level between the centers
-# diversity seems to be positively associating with CR
-
-# if switching to standardized diversity, the positive association is less sure.
+# diversity seems to be positively associating with CR when the prior is too large sd
+# if using appropriate prior the association is less sure
 
@@
 prior <- extract.prior( mcr , n=1e4 )
-# don't know how to put in logdivs into this prior check
-divs_samp <- sample(meta$logdiv_s, size = 1e4, replace = T)
-p <- sapply( 1:2 , function(k) inv_logit( prior$b * divs_samp + prior$a[,k] ) )
-mean( abs( p[,1] - p[,2] ) )
-dens(abs( p[,1] - p[,2] ), adj = 0.1)
+
+prior_cr <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(div_value)) %>% 
+      set_names(seq(1, length(div_value))) %>% 
+      map(function(idx){
+         inv_logit( prior$b * div_value[idx] + prior$a[, center_] ) 
+      })
+
+    diff = abs( cols[[1]] -cols[[2]] )
+    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(cols)
+  })
 ```
+
+```{r}
+# posterior prediction check
+post <- extract.samples(mcr)
+post_cr <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(div_value)) %>% 
+      set_names(seq(1, length(div_value))) %>% 
+      map(function(idx){
+         inv_logit( post$b * div_value[idx] + post$a[, center_] ) 
+      })
+
+    diff = abs( cols[[1]] -cols[[2]] )
+    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(cols)
+  }) 
+```
+
+```{r}
+# the below is when the returned value is the difference between the low and high
+# bind_rows(
+#   bind_cols(prior_cr, grp ='prior', quant = c(0.025, 0.5, 0.975)),
+#   bind_cols(post_cr, grp ='post', quant = c(0.025, 0.5, 0.975))
+# )  %>% 
+#   gather('center', 'value', m:p) %>% 
+#   spread('quant', 'value') %>% 
+#   mutate(grp = factor(grp, levels = c('prior','post'))) %>% 
+#   ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 std apart') +
+#   geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') +
+#   ylim(0,1) 
+```
+```{r}
+# plot the histogram of the probs corresponding to low and high diversity
+# msk
+tibble(
+  prior_low = prior_cr %>% pluck('m') %>% pluck(1),
+  prior_high = prior_cr %>% pluck('m') %>% pluck(2)
+)%>% gather() %>% 
+  gghistogram(x = 'value')+
+  facet_grid(key ~ .)
+
+# posterior 
+# do the M and P in one function
+c('m','p') %>% 
+  map(function(x){ tibble(
+           post_low = post_cr %>% pluck(x) %>% pluck(1),
+           post_high = post_cr %>% pluck(x) %>% pluck(2)
+          )  %>% gather() %>% 
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+      facet_grid(key ~ .)})
+```

### 2021-03-14 scripts/55_rethinking_abx.Rmd
cool Im on the right track very happy with the results

@@
         logit(p) <- b*anaerobe + a[location] ,
-        b ~ dnorm( 0 , 1), 
+        b ~ dnorm( 0 , 1.5), 
         a[location] ~ dnorm( 0 , 0.5 )
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-precis(atox, depth = 2)
-plot(precis(atox, depth = 2) ) 
+#precis(atox, depth = 2)
+plot(precis(atox, depth = 2 ,prob = 0.95) ) 
 # seems anaerobe abx is positively linked with toxicity 
+```
+### Posterior check
+
+```{r}
+abx_val <- c(0, 1)
+post <- extract.samples(atox)
+post_tox <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(abx_val)) %>% 
+      set_names(seq(1, length(abx_val))) %>% 
+      map(function(idx){
+         inv_logit( post$b * abx_val[idx] + post$a[, center_] ) 
+      })
+
+    #diff = abs( cols[[1]] -cols[[2]] )
+    #ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(cols)
+  })
+
+c('m','p') %>% 
+  map(function(x){ tibble(
+           post_no = post_tox %>% pluck(x) %>% pluck(1),
+           post_yes = post_tox %>% pluck(x) %>% pluck(2)
+          )  %>% gather() %>% 
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+      facet_grid(key ~ .)})
+
+# to find out the mean for MSK
+df <- tibble(
+           post_no = post_tox %>% pluck('m') %>% pluck(1),
+           post_yes = post_tox %>% pluck('m') %>% pluck(2)
+          )  %>% gather()  %>% 
+  group_by(key) %>% 
+  summarise(meancoeff = mean(value))
 ```
@@
         logit(p) <- b*anaerobe + a[location] ,
-        b ~ dnorm( 0 , 1), 
+        b ~ dnorm( 0 , 1.5), 
         a[location] ~ dnorm( 0 , 0.5 )
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-precis(acr, depth = 2)
-plot(precis(acr, depth = 2) ) 
+#precis(acr, depth = 2)
+plot(precis(acr, depth = 2, prob = 0.95) ) 
 # nothing interesting 
 ```
+### Posterior check
+
+```{r}
+abx_val <- c(0, 1)
+post <- extract.samples(acr)
+post_cr <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(abx_val)) %>% 
+      set_names(seq(1, length(abx_val))) %>% 
+      map(function(idx){
+         inv_logit( post$b * abx_val[idx] + post$a[, center_] ) 
+      })
+
+    #diff = abs( cols[[1]] -cols[[2]] )
+    #ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(cols)
+  })
+
+c('m','p') %>% 
+  map(function(x){ tibble(
+           post_no = post_cr %>% pluck(x) %>% pluck(1),
+           post_yes = post_cr %>% pluck(x) %>% pluck(2)
+          )  %>% gather() %>% 
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+      facet_grid(key ~ .)})
+
+df <- tibble(
+           post_no = post_cr %>% pluck('m') %>% pluck(1),
+           post_yes = post_cr %>% pluck('m') %>% pluck(2)
+          )  %>% gather()  %>% 
+  group_by(key) %>% 
+  summarise(meancoeff = mean(value))
+```

### 2021-03-14 scripts/56_rethinking_genera.Rmd
the post check for genera and tox cr is not that easy

--- 2021-03-12 scripts/56_rethinking_genera.Rmd
@@
 library(tidyverse)  
+library(ggpubr)
 ```
@@
 
+# do not standardize the log transformed relative abundance 
 meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
-  full_join(genera) %>% 
-  mutate(A_std = scale(Akkermansia),
-         B_std = scale(Bacteroides),
-         E_std = scale(Enterococcus),
-         F_std = scale(Faecalibacterium),
-         R_std = scale(Ruminococcus))
+  full_join(genera)
 
+# wanna see the range of the 5 genera log 10 relab
+meta %>% select(Akkermansia:Ruminococcus) %>% 
+  summary
+```
+
+```{r}
 set.seed(123)   
@@
     location = meta$loca,
-    A_std = meta$A_std,
-    B_std = meta$B_std,
-    E_std = meta$E_std,
-    F_std = meta$F_std,
-    R_std = meta$R_std
+    Akkermansia = meta$Akkermansia,
+    Bacteroides = meta$Bacteroides,
+    Enterococcus = meta$Enterococcus,
+    Faecalibacterium = meta$Faecalibacterium,
+    Ruminococcus = meta$Ruminococcus
     ) 
@@
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-precis(gtox, depth = 2)
-plot(precis(gtox, depth = 2) )  
+plot(precis(gtox, depth = 2, prob = 0.95) )  
 # indicative of the Bacteroides link with toxicity  
+```
+
+### posterior check
+
+```{r}
+bac_value <- log10(c(0.05, 0.35))
+post <- extract.samples(gtox) 
+# if all the others are the same (10%) and only Bacteroides relative abundance could be 0 or  0.35
+post_tox <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(bac_value)) %>% 
+      set_names(seq(1, length(bac_value))) %>% 
+      map(function(idx){
+         inv_logit( post$ba * (-1) + post$bb * bac_value[idx] + post$be * (-1) +  post$bf * (-1) + post$br * (-1) + post$a[, center_] ) 
+      })
+
+    diff = abs( cols[[1]] -cols[[2]] )
+    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(cols)
+  })
+
+c('m','p') %>% 
+  map(function(x){ tibble(
+           post_low = post_tox %>% pluck(x) %>% pluck(1),
+           post_high = post_tox %>% pluck(x) %>% pluck(2)
+          )  %>% gather() %>% 
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+      facet_grid(key ~ .)}) 
 ```
@@
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-precis(gcr, depth = 2)
-plot(precis(gcr, depth = 2) ) 
+
+plot(precis(gcr, depth = 2, prob = 0.95) ) 
 # I think the Ruminococcus still holds for the CR
 ```
+
+### posterior check
+
+```{r}
+rum_value <- log10(c(0.05, 0.5))
+post <- extract.samples(gcr) 
+# if all the others are the same (10%) and only Ruminococcus relative abundance could be 0 or  0.5
+post_cr <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(rum_value)) %>% 
+      set_names(seq(1, length(rum_value))) %>% 
+      map(function(idx){
+         inv_logit( post$ba * (-1) + post$bb * rum_value[idx] + post$be * (-1) +  post$bf * (-1) + post$br * (-1) + post$a[, center_] ) 
+      })
+
+    diff = abs( cols[[1]] -cols[[2]] )
+    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(cols)
+  })
+
+c('m','p') %>% 
+  map(function(x){ tibble(
+           post_low = post_cr %>% pluck(x) %>% pluck(1),
+           post_high = post_cr %>% pluck(x) %>% pluck(2)
+          )  %>% gather() %>% 
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+      facet_grid(key ~ .)}) 
+```

### 2021-03-15 scripts/56_rethinking_genera.Rmd
good morning Monday on the right track of doing work

@@
 ```{r}
-bac_value <- log10(c(0.05, 0.35))
 post <- extract.samples(gtox) 
-# if all the others are the same (10%) and only Bacteroides relative abundance could be 0 or  0.35
-post_tox <- list(m = 1, p = 2) %>% 
-  map(function(center_){
-    cols = seq(1, length(bac_value)) %>% 
-      set_names(seq(1, length(bac_value))) %>% 
-      map(function(idx){
-         inv_logit( post$ba * (-1) + post$bb * bac_value[idx] + post$be * (-1) +  post$bf * (-1) + post$br * (-1) + post$a[, center_] ) 
-      })
+# using the patient with highest and lowest Bacteroides data to plug in
+# find the patients
+bl <- meta %>% 
+  arrange(Bacteroides) %>% 
+  slice(1) %>% 
+  select(Akkermansia:Ruminococcus)
 
-    diff = abs( cols[[1]] -cols[[2]] )
-    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
-    return(cols)
-  })
+bh <- meta %>% 
+  arrange(-Bacteroides) %>% 
+  slice(1) %>% 
+  select(Akkermansia:Ruminococcus)
+```
 
-c('m','p') %>% 
-  map(function(x){ tibble(
-           post_low = post_tox %>% pluck(x) %>% pluck(1),
-           post_high = post_tox %>% pluck(x) %>% pluck(2)
-          )  %>% gather() %>% 
-      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
-      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-      facet_grid(key ~ .)}) 
+```{r}
+pl <- inv_logit( post$ba * (-1.612806	) + post$bb * (-5.69897		) + post$be * (-5.69897	) +  post$bf * (-5.69897		) + post$br * (-0.5244935) + post$a[, 1] )
+
+ph <- inv_logit( post$ba * (-2.286752	) + post$bb * (-0.4575058) + post$be * (-1.745601	) +  post$bf * (-5.69897) + post$br * (-4.050354) + post$a[, 1] )
+
+bind_rows(
+  tibble(
+    value = pl,
+    grp = 'low'
+  ),
+  tibble( 
+    value = ph,
+    grp = 'high'
+  )
+) %>% 
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK')
+
 ```
@@
 ```{r}
-rum_value <- log10(c(0.05, 0.5))
 post <- extract.samples(gcr) 
-# if all the others are the same (10%) and only Ruminococcus relative abundance could be 0 or  0.5
-post_cr <- list(m = 1, p = 2) %>% 
-  map(function(center_){
-    cols = seq(1, length(rum_value)) %>% 
-      set_names(seq(1, length(rum_value))) %>% 
-      map(function(idx){
-         inv_logit( post$ba * (-1) + post$bb * rum_value[idx] + post$be * (-1) +  post$bf * (-1) + post$br * (-1) + post$a[, center_] ) 
-      })
+# using the patient with highest and lowest Ruminococcus data to plug in
+# find the patients
+rl <- meta %>% 
+  arrange(Ruminococcus) %>% 
+  slice(1) %>% 
+  select(Akkermansia:Ruminococcus)
 
-    diff = abs( cols[[1]] -cols[[2]] )
-    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
-    return(cols)
-  })
+rh <- meta %>% 
+  arrange(-Ruminococcus) %>% 
+  slice(1) %>% 
+  select(Akkermansia:Ruminococcus)
 
-c('m','p') %>% 
-  map(function(x){ tibble(
-           post_low = post_cr %>% pluck(x) %>% pluck(1),
-           post_high = post_cr %>% pluck(x) %>% pluck(2)
-          )  %>% gather() %>% 
-      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
-      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-      facet_grid(key ~ .)}) 
+pl <- inv_logit( post$ba * (-3.517128) + post$bb * (-2.447201	) + post$be * (-2.504205	) +  post$bf * (-5.69897	) + post$br * (-5.69897) + post$a[, 1] )
+
+ph <- inv_logit( post$ba * (-1.845325	) + post$bb * (-3.052386) + post$be * (-5.69897) +  post$bf * (-5.69897) + post$br * (-0.373353) + post$a[, 1] )
+
+bind_rows(
+  tibble(
+    value = pl,
+    grp = 'low'
+  ),
+  tibble( 
+    value = ph,
+    grp = 'high'
+  )
+) %>% 
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK')
+
 ```

### 2021-03-15 scripts/54_rethinking_binary.Rmd
updated the script and also the ppt good

@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
-  mutate(logdiv_s = scale(log(simpson_reciprocal))) %>%  # diversity is scaled 
+  mutate(logdiv_s = scale(log(simpson_reciprocal))) %>%  
   mutate(tox = if_else(toxicity == 'yes', 1, 0),
@@
       gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
-      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-      facet_grid(key ~ .)})
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) })
 ```
@@
 # visualize the prior and post together
-tox <- bind_rows(
-  bind_cols(prior_tox, grp ='prior', quant = c(0.025, 0.5, 0.975)),
-  bind_cols(post_tox, grp ='post', quant = c(0.025, 0.5, 0.975))
-)  %>% 
-  gather('center', 'value', m:p) %>% 
-  spread('quant', 'value') %>% 
-  mutate(grp = factor(grp, levels = c('prior','post')))
-tox
+# tox <- bind_rows(
+#   bind_cols(prior_tox, grp ='prior', quant = c(0.025, 0.5, 0.975)),
+#   bind_cols(post_tox, grp ='post', quant = c(0.025, 0.5, 0.975))
+# )  %>% 
+#   gather('center', 'value', m:p) %>% 
+#   spread('quant', 'value') %>% 
+#   mutate(grp = factor(grp, levels = c('prior','post')))
+# tox
 ```
 ```{r}
-tox %>% 
-  ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 std apart') +
-  geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') +
-  ylim(0,1) 
+# tox %>% 
+#   ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 std apart') +
+#   geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') +
+#   ylim(0,1) 
 ```

### 2021-03-15 scripts/55_rethinking_abx.Rmd
updated the script and also the ppt good

@@
 library(tidyverse)  
-library(haven) #for reading sav data
-library(sjstats) #for calculating intra-class correlation (ICC)
-library(effects) #for plotting parameter effects
-library(jtools) #for transformaing model summaries
-library(ROCR)
+library(ggpubr)
 ```
@@
       gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
-      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-      facet_grid(key ~ .)})
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) })
 
@@
       gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
-      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
-      facet_grid(key ~ .)})
+      scale_x_continuous(breaks = seq(0, 1, 0.1))})
 

### 2021-03-15 scripts/54_rethinking_binary.Rmd
ok now it is time to move to the shotgun

@@
     ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
-    return(cols)
+    return(ret)
   })
 ```
+### posterior
 
@@
       }) 
-
-    diff = abs( cols[[1]] -cols[[2]] )
-    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
     return(cols)
@@
 
-```{r}
-# the below is when the returned value is the difference between the low and high
-# bind_rows(
-#   bind_cols(prior_cr, grp ='prior', quant = c(0.025, 0.5, 0.975)),
-#   bind_cols(post_cr, grp ='post', quant = c(0.025, 0.5, 0.975))
-# )  %>% 
-#   gather('center', 'value', m:p) %>% 
-#   spread('quant', 'value') %>% 
-#   mutate(grp = factor(grp, levels = c('prior','post'))) %>% 
-#   ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 std apart') +
-#   geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') +
-#   ylim(0,1) 
-```
 ```{r}
@@
 ```
+
+```{r}
+# the difference of the probability 
+post_cr <- list(m = 1, p = 2) %>% 
+  map(function(center_){
+    cols = seq(1, length(div_value)) %>% 
+      set_names(seq(1, length(div_value))) %>% 
+      map(function(idx){
+         inv_logit( post$b * div_value[idx] + post$a[, center_] ) 
+      })
+
+    diff = abs( cols[[1]] -cols[[2]] )
+    ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
+    return(ret)
+  })
+```
+
+```{r}
+#the below is when the returned value is the difference between the low and high
+diff_prob_post <- bind_rows(
+  bind_cols(prior_cr, grp ='prior', quant = c(0.025, 0.5, 0.975)),
+  bind_cols(post_cr, grp ='post', quant = c(0.025, 0.5, 0.975))
+)  %>%
+  gather('center', 'value', m:p) %>%
+  spread('quant', 'value') %>%
+  mutate(grp = factor(grp, levels = c('prior','post'))) %>%
+  # only showing the post check value
+  filter(grp == 'post')
+# mean at both centers are 0.20 actually
+
+diff_prob_post  %>% 
+  ggscatter(x = 'center', y = '0.5', color = 'red', facet.by = 'grp', ylab = '', title = 'Difference in probability of responding yes when the diversity is 2 sd apart') +
+  geom_pointrange(aes( ymin  = `0.025`,ymax  = `0.975`), fatten = 5, col = 'red') 
+```
+
+```{r}
+#  The two patients with highest and lowest diversity
+# post only 
+# lowest appear in one patient in p; highest is in one patient in M  
+div_l <- min(meta$logdiv_s)
+pl <- inv_logit( post$b * div_l + post$a[, 2] )
+
+div_h <- max(meta$logdiv_s) 
+ph <- inv_logit( post$b * div_h + post$a[, 1] )
+
+bind_rows(
+  tibble(value = pl, grp = 'lowest'),
+  tibble(value = ph, grp = 'highest')
+) %>% 
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean', xlab = 'Predicted probability of CR d100', ylab = 'Probability density',
+              title = 'Predicted probability of CR d100 between the pt with highest div and the pt with lowest') 
+```

### 2021-03-16 scripts/43_comparisons_3.Rmd
finally did some real coding after 5 PM
good I tried the ordered categorical outcome not sure if i did it right tho

@@
 ```{r}
-pheno <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
+cur <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
   select(fid, center, cr_d100, toxicity, icans, crs) %>% 
-  filter(center != 'I') %>% # remove the Italy ones
+  filter(center != 'I') %>% # manually update the label for the "not_assessed"
+  mutate(cr_d100 = case_when(
+    fid == '2460A' ~ 'no',
+    fid == '2293A' ~ 'no',
+    fid == '2563A' ~ 'no',
+    TRUE ~ cr_d100
+  ))
+
+pheno <- cur %>% 
   gather('pheno', 'value', cr_d100:crs)

### 2021-03-16 scripts/54_rethinking_binary.Rmd
good I tried the ordered categorical outcome not sure if i did it right tho

@@
          toxicity = factor(toxicity, levels = c('no','yes')),
-         cr_d100 = factor(cr_d100, levels = c('no','yes')))
+         cr_d100 = factor(cr_d100, levels = c('no','yes'))) %>%
+  # have a 2 cutpoint for the CRS
+  mutate(crs3 = case_when(
+    crsgrade == 0 ~ 1,
+    crsgrade == 1 ~ 2,
+    crsgrade >= 2 ~ 3
+  ))
 
@@
 range(meta$logdiv_s)
+table(meta$crs3)
+table(meta$crsgrade, meta$center) 
 ```
@@
     tox = meta$tox,
+    crs3 = meta$crs3,
     cr100 = meta$cr100,
@@
 # it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
+```
+
+```{r}
+# another model with the crs grade 0,1,>=2 
+mcrs3 <- ulam(
+    alist(
+        crs3 ~ dordlogit( phi , cutpoints ),
+        phi <- b*logdiv_s + a[location] ,
+        b ~ dnorm( 0 , 1),  
+        a[location] ~ dnorm( 0 , 0.5 ),
+        cutpoints ~ dnorm( 0 , 1.5 )
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+plot(precis(mcrs3, depth = 2, prob = 0.95) ) 
+# the cutpoints are not of much interet, this suggests that the diversity really isn't having a correlation?
 ```
@@
 # the raw proportion of yes toxicity in each place
-rawtox <- meta  %>% 
-  count(center, toxicity) %>% 
-  left_join(meta %>% 
-              count(center, name = 'total'), by = 'center') %>% 
-  group_by(center, toxicity) %>% 
-  summarise(frac = round(n/total, 2)) %>% 
-  filter(toxicity == 'yes')
+# rawtox <- meta  %>% 
+#   count(center, toxicity) %>% 
+#   left_join(meta %>% 
+#               count(center, name = 'total'), by = 'center') %>% 
+#   group_by(center, toxicity) %>% 
+#   summarise(frac = round(n/total, 2)) %>% 
+#   filter(toxicity == 'yes')
 
@@
               title = 'Predicted probability of CR d100 between the pt with highest div and the pt with lowest')
+
+mean(pl)
+mean(ph)
 ```

### 2021-03-16 scripts/54_rethinking_binary.Rmd
ok slowly and getting into the mode ...

@@
           )  %>% gather() %>% 
-      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',
+                  xlab = 'Predicted probability of toxicity', ylab = 'Probability density',
+                  title = str_glue('{x}')) +
       scale_x_continuous(breaks = seq(0, 1, 0.1)) })
@@
   })
-```
 
-```{r}
 # plot the histogram of the probs corresponding to low and high diversity
@@
           )  %>% gather() %>% 
-      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',
+                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density',
+                  title = str_glue('{x}')) +
       scale_x_continuous(breaks = seq(0, 1, 0.1)) })

### 2021-03-16 scripts/55_rethinking_abx.Rmd
ok slowly and getting into the mode ...

@@
           )  %>% gather() %>% 
-      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',xlab = 'Predicted probability of toxicity', ylab = 'Probability density', title = str_glue('{x}')) +
       scale_x_continuous(breaks = seq(0, 1, 0.1)) })
@@
           )  %>% gather() %>% 
-      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',title = str_glue('{x}')) +
+      gghistogram(x = 'value',bins = 30, fill = 'key', palette = 'nejm', color = 'white',
+                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density', title = str_glue('{x}')) +
       scale_x_continuous(breaks = seq(0, 1, 0.1))})

### 2021-03-16 scripts/56_rethinking_genera.Rmd
feeling I didn something major cuz never did this before pmap within pmap

@@
   select(Akkermansia:Ruminococcus)
-```
 
-```{r}
 pl <- inv_logit( post$ba * (-1.612806	) + post$bb * (-5.69897		) + post$be * (-5.69897	) +  post$bf * (-5.69897		) + post$br * (-0.5244935) + post$a[, 1] )
@@
 ) %>% 
-  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK')
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK',
+              xlab = 'Predicted probability of toxicity', ylab = 'Probability density')
 
@@
 bind_rows(
-  tibble(
-    value = pl,
-    grp = 'low'
-  ),
-  tibble( 
-    value = ph,
-    grp = 'high'
-  )
+  tibble(value = pl,grp = 'low'),
+  tibble( value = ph,grp = 'high')
 ) %>% 
-  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK')
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK',
+              xlab = 'Predicted probability of CR d100', ylab = 'Probability density')
 
 ```
+
+```{r}
+# ruminococcus relab quantile
+meta_ <- meta %>% 
+  select(Akkermansia:Ruminococcus)
+
+# the post coeffs in df format
+postdf <- bind_cols(ba = post$ba, bb = post$bb, be = post$be, bf = post$bf, br = post$br, am = post$a[,1])
+
+# to responde to Sean's comments to not use max and min "extreme"patient data
+# use the 90% and 10% quantile patient data
+# Jonas: how do you know that the patient falling onto the 90%tile is not “exceptional” and therefore an “outlier” in another feature that your model includes, besides ruminococcus, he suggests to sample in that 0-10% quantile range patient data for example
+# It is 100 jointly sampled coefficients multiplied with the composition of bacteria in the current patient
+N <- 100
+# for the top and bottom quantile range of ruminococcus relab
+ru_top <- meta_ %>% 
+  filter(Ruminococcus >= quantile(meta$Ruminococcus, 0.9))
+ru_bot <- meta_ %>% 
+  filter(Ruminococcus <= quantile(meta$Ruminococcus, 0.1))
+
+# sample in the top and bot patients 100 times
+ru_top_samp <- ru_top %>% 
+    sample_n(size = N, replace = T)
+ru_bot_samp <- ru_bot %>% 
+    sample_n(size = N, replace = T)
+
+# a function for every patient data generate 100 post prob of CR (sampling from the post 100 times)
+
+per_pt_sample_post_prob <- function(lga_, lgb_, lge_, lgf_ , lgr_){
+  # the input is the log relab of the 5 genera for that patient
+  post_samp <- postdf %>% 
+    sample_n(size = N, replace = F)
+
+  ret = pmap(post_samp, function(ba, bb, be, bf, br, am) {
+    inv_logit( ba * lga_ + bb * lgb_ + be * lge_ +  bf * lgf_ + br * lgr_ + am ) 
+  }) %>% 
+    set_names(seq(1, N)) %>% 
+    bind_rows() %>% 
+    gather()
+  return(ret)
+}
+
+# for the top quantile patients 
+res_top <- pmap(ru_top_samp, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
+  per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
+}) %>% 
+  set_names(paste('P', seq(1, N), sep = '')) %>% 
+  bind_rows(.id = 'ptsamp')
+
+# for the bottom quantile patients
+res_bot <- pmap(ru_bot_samp, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
+  per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
+}) %>% 
+  set_names(paste('P', seq(1, N), sep = '')) %>% 
+  bind_rows(.id = 'ptsamp')
+```
+
+```{r}
+# plotting and visualizing
+res_top %>% 
+  gghistogram(x = 'value', fill = 'ptsamp', add_density = T,
+              title = '100 sampled patients that had top 10% Ruminococcus relab',
+              xlab = 'Probability of having CR d100', ylab = 'Probability density') +
+  theme(legend.position = 'none')
+```
+
+```{r}
+res_bot %>% 
+  gghistogram(x = 'value', fill = 'ptsamp', add_density = T,
+              title = '100 sampled patients that had bottom 10% Ruminococcus relab',
+              xlab = 'Probability of having CR d100', ylab = 'Probability density') +
+  theme(legend.position = 'none')
+```

### 2021-03-18 scripts/54_rethinking_binary.Rmd
update

@@
 # the std of b is 2, since we assume the microbiome could have big impact on immunology (based on our previous understanding and knowledge )
-#precis(mtox, depth = 2)
+precis(mtox,  prob = 0.95)
 plot(precis(mtox, depth = 2, prob = 0.95) ) 
@@
 
-#precis(mcr, depth = 2)
+precis(mcr,  prob = 0.95)
 plot(precis(mcr, depth = 2, prob = 0.95) ) 

### 2021-03-18 scripts/55_rethinking_abx.Rmd
update

@@
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-#precis(atox, depth = 2)
+precis(atox,prob = 0.95)
 plot(precis(atox, depth = 2 ,prob = 0.95) ) 
@@
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-#precis(acr, depth = 2)
+precis(acr,  prob = 0.95)
 plot(precis(acr, depth = 2, prob = 0.95) ) 

### 2021-03-18 scripts/56_rethinking_genera.Rmd
update

@@
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+precis(gtox, prob = 0.95)   
 plot(precis(gtox, depth = 2, prob = 0.95) )  
@@
 ) %>% 
-  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK',
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',add = 'mean',title = 'MSK',
               xlab = 'Predicted probability of toxicity', ylab = 'Probability density')
 
+bind_rows(
+  tibble(value = pl,grp = 'low'),
+  tibble( value = ph,grp = 'high')
+)   %>% 
+  group_by(grp) %>% 
+  summarise(m = mean(value))
 ```
@@
     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-
+precis(gcr, prob = 0.95) 
 plot(precis(gcr, depth = 2, prob = 0.95) ) 
@@
 ) %>% 
-  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK',
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK', add = 'mean',
               xlab = 'Predicted probability of CR d100', ylab = 'Probability density')
 
+bind_rows(
+  tibble(value = pl,grp = 'low'),
+  tibble( value = ph,grp = 'high')
+)   %>% 
+  group_by(grp) %>% 
+  summarise(m = mean(value))
 ```
@@
 # for the top quantile patients 
-res_top <- pmap(ru_top_samp, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
+res_top <- pmap(ru_top, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
   per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
 }) %>% 
-  set_names(paste('P', seq(1, N), sep = '')) %>% 
-  bind_rows(.id = 'ptsamp')
+  set_names(paste('P', seq(1, nrow(ru_top)), sep = '')) %>% 
+  bind_rows(.id = 'pt')
 
 # for the bottom quantile patients
-res_bot <- pmap(ru_bot_samp, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
+res_bot <- pmap(ru_bot, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
   per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
 }) %>% 
-  set_names(paste('P', seq(1, N), sep = '')) %>% 
-  bind_rows(.id = 'ptsamp')
+  set_names(paste('P', seq(1, nrow(ru_bot)), sep = '')) %>% 
+  bind_rows(.id = 'pt')
 ```
@@
 # plotting and visualizing
+# visualizing each patient
 res_top %>% 
-  gghistogram(x = 'value', fill = 'ptsamp', add_density = T,
+  gghistogram(x = 'value', fill = 'pt', add_density = T,
               title = '100 sampled patients that had top 10% Ruminococcus relab',
@@
   theme(legend.position = 'none')
+
+# averaging over all the patients
+res_top %>% 
+  gghistogram(x = 'value', 
+              #add_density = T,
+              title = '100 sampled patients that had top 10% Ruminococcus relab',
+              xlab = 'Probability of having CR d100', ylab = 'Probability density') +
+  theme(legend.position = 'none')
+
+mean(res_top$value)
 ```
@@
 ```{r}
+# visualizing each patient
 res_bot %>% 
-  gghistogram(x = 'value', fill = 'ptsamp', add_density = T,
+  gghistogram(x = 'value', fill = 'pt', add_density = T,
               title = '100 sampled patients that had bottom 10% Ruminococcus relab',
@@
   theme(legend.position = 'none')
+
+# averaging over all the patients
+res_bot %>% 
+  gghistogram(x = 'value', 
+              #add_density = T,
+              title = '100 sampled patients that had bottom 10% Ruminococcus relab',
+              xlab = 'Probability of having CR d100', ylab = 'Probability density') +
+  theme(legend.position = 'none')
+
+mean(res_bot$value)
 ```
+
+```{r}
+# plot them together in one 
+bind_rows(
+  res_top %>% mutate(grp = 'top'),
+  res_bot %>% mutate(grp = 'bot')
+) %>% 
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
+                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density')
+
+```

### 2021-03-18 scripts/57_16s_lefse.Rmd
update

---
title: "16s lefse"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# sort out the genera counts table and also do filtering (need to have all taxa levels)
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('asv_annotation_blast_ag')
get_table_from_database('asv_counts_ag')
cts <- get_counts_subset(meta$sampleid)

cts_ <- cts %>% 
  select(asv_key, sampleid, count) %>% 
  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
  arrange(asv_key)

annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts_$asv_key) %>% 
  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
  mutate(taxa_genus = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}'))

cts_genus_all <- cts_ %>% 
  full_join(annot %>%  select(asv_key, taxa_genus), by  = 'asv_key') %>% 
  select(-asv_key) %>% 
  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
  group_by(sampleid, taxa_genus) %>% 
  summarise(cnt = sum(count)) %>% 
  # get the total count from the db to calculate the relab
  left_join(cts %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
  mutate(relab = cnt/count_total) %>% 
  select(sampleid, taxa_genus, relab)

# the genera to keep
keepg <- cts_genus_all %>% 
  filter(relab > 0.0001) %>%
  ungroup() %>% 
  count(taxa_genus) %>% 
  filter(n > floor(nrow(meta) * 0.1)) %>% 
  filter(!str_detect(taxa_genus, 'NA$')) %>% 
  pull(taxa_genus)

cts_genus_fil <- cts_genus_all %>% 
  filter(taxa_genus %in% keepg) %>% 
  spread('sampleid', 'relab')

```

```{r}
pheno <- meta %>% 
  select(cr_d100:crs, )
  gather('pheno', 'value', cr_d100:crs)

all_sub_pheno <- pheno %>% 
  split(., list(.$pheno)) %>% 
  purrr::imap(~ filter(.data = ., value != 'not_assessed'))

all_sub_pheno %>% 
  imap(function(.x, .y){
    select(.data = .x, value) %>% 
      t() %>% 
      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
  })
```

### 2021-03-18 scripts/57_16s_lefse.Rmd
update

@@
 ```{r}
+library(tidyverse)
+```
+
+```{r}
 # sort out the genera counts table and also do filtering (need to have all taxa levels)
@@
 get_table_from_database('asv_annotation_blast_ag')
-get_table_from_database('asv_counts_ag')
 cts <- get_counts_subset(meta$sampleid)
@@
 ```{r}
+# the pheno label for the samples
 pheno <- meta %>% 
-  select(cr_d100:crs, )
-  gather('pheno', 'value', cr_d100:crs)
+  select(center, cr_d100:crs, icans, sampleid) %>% 
+  gather('pheno', 'value', cr_d100:icans)
 
@@
 
-all_sub_pheno %>% 
+tpheno <- all_sub_pheno %>% 
   imap(function(.x, .y){
-    select(.data = .x, value) %>% 
-      t() %>% 
-      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
+    select(.data = .x, value, center) %>% 
+      t() %>% write.table(str_glue('../data/amplicon/lefse/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
   })
+
+tcts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., sampleid) ) %>% 
+  imap(~ cts_genus_fil %>% select(taxa_genus, matches(.x)) %>%  write_tsv(str_glue('../data/amplicon/lefse/{.y}_tcts.tsv'))) 
 ```
+
+cat pull_toxicity.txt toxicity_tcts.tsv > pull_toxicity_tcts.tsv
+cat pull_cr_d100.txt cr_d100_tcts.tsv > pull_cr_d100_tcts.tsv 
+cat pull_crs.txt crs_tcts.tsv > pull_crs_tcts.tsv
+cat pull_icans.txt icans_tcts.tsv > pull_icans_tcts.tsv
+
+```{r}
+fns <- list.files('../data/amplicon/lefse/', pattern = 'pull.*_tcts.tsv$')
+
+cmds <- tibble(
+  fns = fns
+) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -s 2  -u 3  ')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
+  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
+  select(-fns) %>% 
+  gather() %>% 
+  select(value) %>% 
+  write_csv('../data/amplicon/lefse/lefse_run_cmd_taxa.sh', col_names = F)
+
+# No differentially abundant features found in the 4
+# also this is the case when using center as subclass 
+```

### 2021-03-18 scripts/43_comparisons_3.Rmd
feeling so bad and so upset

@@
 # PATHWAY
+# I already normalized the pcts so don't need to normalize again here 
 fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv$')
@@
 ) %>% 
-  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2')) %>% 
   mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 

### 2021-03-18 scripts/57_16s_lefse.Rmd
feeling so bad and so upset

@@
 # sort out the genera counts table and also do filtering (need to have all taxa levels)
-meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
+meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
+  filter(center == 'M')
 library(vdbR)

### 2021-03-19 scripts/45_16s_sorting_out.Rmd
uypdate

@@
 ```
+
+```{r}
+# check something about abx
+m <- readxl::read_excel('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/antibiotics/sean_table/abx_table_msk_sean.xlsx') %>% 
+  select(pt_id, anaerobe_targeting)
+meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv') %>% 
+  left_join(m) %>% 
+  mutate(anaerobe_targeting = if_else(is.na(anaerobe_targeting), 'no', anaerobe_targeting))
+
+m <- meta %>% 
+  filter(center == 'M')
+
+m %>% 
+  count(anaerobe_targeting, cr_d100) %>% 
+  ggbarplot(x = 'anaerobe_targeting', y= 'n', color = 'cr_d100', fill = 'cr_d100')
+```

### 2021-03-19 scripts/57_16s_lefse.Rmd
uypdate

--- 2021-03-18 scripts/57_16s_lefse.Rmd
@@
 ```{r}
+library(tidyverse)
+library(ggpubr)
+```
+
+```{r}
 # sort out the genera counts table and also do filtering (need to have all taxa levels)
 meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
+
 library(vdbR)
@@
 cts_ <- cts %>% 
-  select(asv_key, sampleid, count) %>% 
-  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
-  arrange(asv_key)
+  select(asv_key, sampleid, count_relative) %>% 
+  spread(key = 'sampleid', value = 'count_relative', fill = 0)
 
@@
          species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
-  mutate(taxa_genus = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}'))
+  mutate(taxa_asv = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}|a__{asv_key}'))
 
-cts_genus_all <- cts_ %>% 
-  full_join(annot %>%  select(asv_key, taxa_genus), by  = 'asv_key') %>% 
+cts_all <- cts_ %>% 
+  full_join(annot %>%  select(asv_key, taxa_asv), by  = 'asv_key') %>% 
   select(-asv_key)  %>% 
-  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
-  group_by(sampleid, taxa_genus) %>% 
-  summarise(cnt = sum(count)) %>% 
-  # get the total count from the db to calculate the relab
-  left_join(cts %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
-  mutate(relab = cnt/count_total) %>% 
-  select(sampleid, taxa_genus, relab)
+  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1])
 
-# the genera to keep
-keepg <- cts_genus_all %>% 
-  filter(relab > 0.0001) %>%
+# the asv to keep
+keepg <- cts_all %>% 
+  filter(relab > 0) %>%
   ungroup() %>% 
-  count(taxa_genus) %>% 
+  count(taxa_asv) %>% 
   filter(n > floor(nrow(meta) * 0.1)) %>% 
-  filter(!str_detect(taxa_genus, 'NA$')) %>% 
-  pull(taxa_genus)
+  pull(taxa_asv)
 
-cts_genus_fil <- cts_genus_all %>% 
-  filter(taxa_genus %in% keepg) %>% 
+cts_fil <- cts_all %>% 
+  filter(taxa_asv %in% keepg) %>% 
   spread('sampleid', 'relab')
+```
+
+```{r}
+dat <- genera %>% 
+  filter(genus == 'Ruminococcus') %>% 
+  gather('sampleid', 'relab', names(.)[2]:names(.)[ncol(.)]) %>% 
+  full_join(meta) %>% 
+  mutate(logrelab =  log10(relab + 2e-6))
+
+yes <- dat %>% 
+  filter(cr_d100 == 'yes')
+no <- dat %>% 
+  filter(cr_d100 == 'no')
+
+dat %>% 
+  #filter(center == 'M') %>% 
+  mutate(logrelab =  log10(relab + 2e-6)) %>% 
+  ggboxplot(x = 'cr_d100', y = 'logrelab', add = 'jitter') +
+stat_compare_means(comparisons= list(c('yes', 'no')),
+										label= "p.signif",
+										method= 'wilcox.test',
+										correct=FALSE)
+
+t.test(yes$logrelab, no$logrelab)
 
@@
 
-all_sub_pheno %>% 
+tpheno <- all_sub_pheno %>% 
   imap(function(.x, .y){
     select(.data = .x, value) %>% 
-      t() %>% 
-      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
+      t() %>% write.table(str_glue('../data/amplicon/lefse/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
   })
+
+tcts <- all_sub_pheno %>% 
+  map(~ pull(.data = ., sampleid) ) %>% 
+  imap(~ cts_fil %>% select(taxa_asv, matches(.x)) %>%  write_tsv(str_glue('../data/amplicon/lefse/{.y}_asv_tcts.tsv'))) 
 ```
+
+cat pull_toxicity.txt toxicity_asv_tcts.tsv > pull_toxicity_asv_tcts.tsv
+cat pull_cr_d100.txt cr_d100_asv_tcts.tsv > pull_cr_d100_asv_tcts.tsv 
+cat pull_crs.txt crs_asv_tcts.tsv > pull_crs_asv_tcts.tsv
+cat pull_icans.txt icans_asv_tcts.tsv > pull_icans_asv_tcts.tsv
+
+```{r}
+fns <- list.files('../data/amplicon/lefse/', pattern = 'pull.*_asv_tcts.tsv$')
+
+cmds <- tibble(
+  fns = fns
+) %>% 
+  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1  -u 2 -o 1000000   ')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
+  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
+  select(-fns) %>% 
+  gather() %>% 
+  select(value) %>% 
+  write_csv('../data/amplicon/lefse/lefse_run_cmd_taxa.sh', col_names = F)
+
+# No differentially abundant features found in the 4
+# also this is the case when using center as subclass 
+```

### 2021-03-19 scripts/58_shotgun_lefse_and_picked_target.Rmd
uypdate

---
title: "lefse and the picked targets"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
all_pw <- read_tsv('~/pipeline/scripts/shotgun_pipeline/data/metacyc_pathway_name_and_ID.tsv')
pick <- read_csv('../data/shotgun_output/metacyc_pathway_name_and_ID_MSv2.csv') %>% 
  select(Metabolite:cluster)

lefse <- read_csv('../data/shotgun_output/pull_pathway_lefse.csv') %>% 
  mutate(PWID = str_extract(feature, '^.+_PWY|^PWY.*_\\d{3,4}'))  %>% 
  mutate(
    PWID = case_when(
      str_detect(feature, 'NAD_BIOSYNTHESIS_II') ~ 'NAD_BIOSYNTHESIS_II',
      str_detect(feature, 'ARG_POLYAMINE_SYN') ~ 'ARG+POLYAMINE_SYN',
      str_detect(feature, 'PWY5F9') ~ 'PWY5F9',
      TRUE ~ PWID
    )
  ) %>% 
  mutate(PWID = str_replace_all(PWID, '_', '-'))

res <- lefse %>% 
  left_join(pick)

res %>% 
  write_csv('../data/shotgun_output/pull_lefse_results_pathway_join_with_picked_ones.csv')
```

### 2021-03-23 scripts/49_16s_stacked_barplot.Rmd
ud[ate

@@
 )
+
+dat %>%  write_csv('../data/the_data_to_make_panel_B.csv')
 ```

### 2021-03-23 scripts/54_rethinking_binary.Rmd
ud[ate

@@
 # it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
+div_value <- c(-1, 1)
 ```
@@
       scale_x_continuous(breaks = seq(0, 1, 0.1)) })
+```
+
+```{r}
+# combine M and P since they look very similar and also save some space
+low_M <- post_tox %>% pluck('m') %>% pluck(1)
+low_P <- post_tox %>% pluck('p') %>% pluck(1)
+high_M <- post_tox %>% pluck('m') %>% pluck(2)
+high_P <- post_tox %>% pluck('p') %>% pluck(2)
+low_all <- bind_rows(
+  tibble(value = low_M),
+  tibble(value = low_P)
+) %>% mutate(grp = 'low')
+
+high_all <- bind_rows(
+  tibble(value = high_M),
+  tibble(value = high_P)
+)%>% mutate(grp = 'high')
+
+tox_all <- bind_rows(
+  low_all,
+  high_all
+)
+
+tox_all  %>% 
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',
+                  xlab = 'Predicted probability of toxicity', ylab = 'Probability density',
+                  title = '') +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  ggsave('../figs/bayesican_tox_combined.pdf', width = 6, height = 4)
 ```

### 2021-03-23 scripts/59_peruse_16s_lefse.Rmd
ud[ate

---
title: "Detailed look at the 16s lefse results"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggpubr)
```

```{r}
res <- list.files('../data/amplicon/lefse/', pattern = 'asv_tcts.tsv.res$', full.names = T)

res_all <- res %>% 
  set_names(res) %>% 
  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  keep(~ nrow(.) > 0) %>% 
  bind_rows(.id = 'res') %>% 
  mutate(res = str_replace(res, '^.+//',''),
         res = str_replace(res, '_asv.+$',''))  %>% 
  rename(grp = res)
res_all$feature

```

### 2021-03-24 scripts/54_rethinking_binary.Rmd
good I even got some work done after back

@@
 ```{r}
+# combine M and P
+low_all <- tibble(
+  value = c(post_cr %>% pluck('m') %>% pluck(1), post_cr %>% pluck('p') %>% pluck(1))
+) %>% mutate(grp = 'low')
+
+high_all <- tibble(
+  value = c(post_cr %>% pluck('m') %>% pluck(2), post_cr %>% pluck('p') %>% pluck(2))
+) %>% mutate(grp = 'high')
+
+cr_all <- bind_rows(
+  low_all,
+  high_all
+)
+
+cr_all  %>% 
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',
+                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density',
+                  title = '') +
+      scale_x_continuous(breaks = seq(0, 1, 0.1)) +
+  ggsave('../figs/bayesican_cr_combined.pdf', width = 6, height = 4)
+```
+
+```{r}
 # the difference of the probability 

### 2021-03-24 scripts/60_rethinking_g_clostridium.Rmd
update

---
title: "replacing bacteroides with Clostridium genus bayesian"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)  
library(ggpubr)
```

# CR/Toxicity ~ Akkermansia + Clostridium + Enterococcus + Faecalibacterium + Ruminococcus + center

## Toxicity

```{r}
# first I need to get the Clostridium genus abundance
clos <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus == 'Clostridium') %>% 
  mutate(relablog = log10(relab + 2*10^-6)) %>% 
  select(sampleid, Clostridium = relablog)

g5 <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
  full_join(clos)

g5 %>% 
  write_csv('../data/amplicon/stool/combined_6_genera.csv')

meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
  full_join(g5)

set.seed(123)   
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    Akkermansia = meta$Akkermansia,
    Clostridium = meta$Clostridium,
    Enterococcus = meta$Enterococcus,
    Faecalibacterium = meta$Faecalibacterium,
    Ruminococcus = meta$Ruminococcus
    ) 
```

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bc*Clostridium + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + a[location] ,
        ba ~ dnorm( 0 , 1),
        bc ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
#precis(gtox, prob = 0.95)   
plot(precis(gtox, depth = 2, prob = 0.95) )  
```

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bc*Clostridium + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + a[location] ,
        ba ~ dnorm( 0 , 1),
        bc ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, prob = 0.95) 
plot(precis(gcr, depth = 2, prob = 0.95) ) 
```

### 2021-03-25 scripts/43_comparisons_3.Rmd
update

@@
   mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
   mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 

### 2021-03-25 scripts/56_rethinking_genera.Rmd
update

@@
 # indicative of the Bacteroides link with toxicity  
+```
+
+```{r}
+post <- extract.samples(gtox) 
+meta_ <- meta %>% 
+  select(Akkermansia:Ruminococcus)
+
+# the post coeffs in df format
+postdf <- bind_cols(ba = post$ba, bb = post$bb, be = post$be, bf = post$bf, br = post$br, am = post$a[,1])
+
+N <- 100
+# for the top and bottom quantile range of Bacteroides relab
+ba_top <- meta_ %>% 
+  filter(Bacteroides >= quantile(meta$Bacteroides, 0.9))
+ba_bot <- meta_ %>% 
+  filter(Bacteroides <= quantile(meta$Bacteroides, 0.1))
+
+per_pt_sample_post_prob <- function(lga_, lgb_, lge_, lgf_ , lgr_){
+  # the input is the log relab of the 5 genera for that patient
+  post_samp <- postdf %>% 
+    sample_n(size = N, replace = F)
+
+  ret = pmap(post_samp, function(ba, bb, be, bf, br, am) {
+    inv_logit( ba * lga_ + bb * lgb_ + be * lge_ +  bf * lgf_ + br * lgr_ + am ) 
+  }) %>% 
+    set_names(seq(1, N)) %>% 
+    bind_rows() %>% 
+    gather()
+  return(ret)
+}
+
+# for the top quantile patients 
+res_top <- pmap(ba_top, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
+  per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
+}) %>% 
+  set_names(paste('P', seq(1, nrow(ba_top)), sep = '')) %>% 
+  bind_rows(.id = 'pt')
+
+# for the bottom quantile patients  
+res_bot <- pmap(ba_bot, function(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus){
+  per_pt_sample_post_prob(Akkermansia, Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus)
+}) %>% 
+  set_names(paste('P', seq(1, nrow(ba_bot)), sep = '')) %>% 
+  bind_rows(.id = 'pt')
+
+# plot them together in one 
+bind_rows(
+  res_top %>% mutate(grp = 'top'),
+  res_bot %>% mutate(grp = 'bot')
+) %>% 
+  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
+                  xlab = 'Predicted probability of Toxicity', ylab = 'Probability density') +
+  ggsave('../figs/predicted_tox_bacteroides_top_bottom_10.pdf', width = 7, height = 5)
 ```
@@
   gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
-                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density')
+                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density')+
+  ggsave('../figs/predicted_CR_Ruminococcus_top_bottom_10.pdf', width = 7, height = 5)
 

### 2021-03-25 scripts/57_16s_lefse.Rmd
update

@@
   mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
   mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 

### 2021-04-14 scripts/43_comparisons_3.Rmd
although slow but Im progressing

@@
   purrr::imap(~ filter(.data = ., value != 'not_assessed'))
+
+cur %>% 
+  write_csv('../data/shotgun_pheno/final_comprehensive_updated_simple.csv')
 ```
@@
   mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
+  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
   select(-fns) %>% 

### 2021-04-14 scripts/45_16s_sorting_out.Rmd
ud[ate
although slow but Im progressing

@@
 library(readxl)
+library(ggpubr)
 ```
@@
   rename(sampleid = oligoid) %>% 
-  select(-last_followup_day) %>% 
+  select(-last_followup) %>% 
   mutate(center = 'P')  %>% 
@@
   count(anaerobe_targeting, cr_d100) %>% 
-  ggbarplot(x = 'anaerobe_targeting', y= 'n', color = 'cr_d100', fill = 'cr_d100')
+  ggbarplot(x = 'anaerobe_targeting', y= 'n', color = 'cr_d100', fill = 'cr_d100', label = TRUE)
 ```

### 2021-04-14 scripts/51_distance_2_healthy.Rmd
although slow but Im progressing

@@
   ggsave('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/Angel_amplicon/distance_to_centroid/centers_compared_to_healthy.pdf')
+
+centroid_center %>% 
+  filter(groups == 'healthy')
 ```
+```{r}
+# dominance in healthy volunteers 
+h <- get_counts_subset(healthy$sampleid)
+
+cts <- h %>% 
+  select(asv_key, sampleid, count) %>% 
+  spread(key = 'sampleid', value = 'count', fill = 0) %>% 
+  arrange(asv_key)
+
+mannot <- asv_annotation_blast_ag %>% 
+  filter(asv_key %in% cts$asv_key) %>% 
+  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
+         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
+         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
+         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) 
+mcts_all <- cts %>% 
+  full_join(mannot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
+  select(-asv_key) %>% 
+  gather(key = 'sampleid', value = 'count', names(.)[1]:names(.)[ncol(.) - 1]) %>% 
+  group_by(sampleid, genus) %>% 
+  summarise(cnt = sum(count)) %>% 
+  # get the total count from the db to calculate the relab
+  left_join(h %>% distinct(sampleid,count_total ), by = 'sampleid') %>% 
+  mutate(relab = cnt/count_total) %>% 
+  select(sampleid, genus, relab)
+
+mcts_all %>% 
+  write_csv('../data/amplicon/healthy_volunteer_genus_relab.csv')
+
+h %>% 
+  write_csv('../data/amplicon/healthy_volunteer_asv_relab.csv')
+```

### 2021-04-14 scripts/56_rethinking_genera.Rmd
although slow but Im progressing

@@
   ggsave('../figs/predicted_tox_bacteroides_top_bottom_10.pdf', width = 7, height = 5)
+
+bind_rows(
+  res_top %>% mutate(grp = 'top'),
+  res_bot %>% mutate(grp = 'bot')
+) %>% 
+  group_by(grp) %>% 
+  summarise(ave = mean(value))
+
 ```
@@
 
+bind_rows(
+  res_top %>% mutate(grp = 'top'),
+  res_bot %>% mutate(grp = 'bot')
+)  %>% 
+  group_by(grp) %>% 
+  summarise(ave = mean(value))
 ```

### 2021-04-14 scripts/56_rethinking_genera.Rmd
update

@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
-  full_join(genera)
+  inner_join(genera)
 

### 2021-04-14 scripts/57_16s_lefse.Rmd
update

@@
 ```{r}
-dat <- genera %>% 
-  filter(genus == 'Ruminococcus') %>% 
-  gather('sampleid', 'relab', names(.)[2]:names(.)[ncol(.)]) %>% 
-  full_join(meta) %>% 
-  mutate(logrelab =  log10(relab + 2e-6))
-
-yes <- dat %>% 
-  filter(cr_d100 == 'yes')
-no <- dat %>% 
-  filter(cr_d100 == 'no')
-
-dat %>% 
-  #filter(center == 'M') %>% 
-  mutate(logrelab =  log10(relab + 2e-6)) %>% 
-  ggboxplot(x = 'cr_d100', y = 'logrelab', add = 'jitter') +
-stat_compare_means(comparisons= list(c('yes', 'no')),
-										label= "p.signif",
-										method= 'wilcox.test',
-										correct=FALSE)
-
-t.test(yes$logrelab, no$logrelab)
+# dat <- genera %>% 
+#   filter(genus == 'Ruminococcus') %>% 
+#   gather('sampleid', 'relab', names(.)[2]:names(.)[ncol(.)]) %>% 
+#   full_join(meta) %>% 
+#   mutate(logrelab =  log10(relab + 2e-6)) 
+# 
+# yes <- dat %>% 
+#   filter(cr_d100 == 'yes')
+# no <- dat %>% 
+#   filter(cr_d100 == 'no')
+# 
+# 
+# dat %>% 
+#   #filter(center == 'M') %>% 
+#   mutate(logrelab =  log10(relab + 2e-6)) %>% 
+#   ggboxplot(x = 'cr_d100', y = 'logrelab', add = 'jitter') +
+# stat_compare_means(comparisons= list(c('yes', 'no')),
+# 										label= "p.signif",
+# 										method= 'wilcox.test',
+# 										correct=FALSE)
+# 
+# t.test(yes$logrelab, no$logrelab)
 

### 2021-04-15 scripts/43_comparisons_3.Rmd
update
very hard to concentrate but damn I am getting things done ..

@@
 ```{r}
+# UPDATE THIS TABLE BABY
 cur <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
@@
 
+# Removing the patients that the data we can't use
+# upated 2021-4-12
+now <- read_csv('/Volumes/castoricenter/CAR-Tcell_and_Microbiota/MSK/samples/eligible_sampleslist_JS.csv') %>% 
+  filter(eligible == 'yes' & shotgun == 'yes') %>% 
+  pull(pt_id)
+
+final <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
+  filter(center != 'I') %>% 
+  mutate(pt_id = as.numeric(pt_id)) %>% 
+  mutate(pt_id = if_else(center == 'P', pt_id + 8000, pt_id)) %>% 
+  filter(pt_id %in% now) # keep to the most updated list
+
+final %>% write_csv('../data/shotgun_pheno/final_comprehensive_UPDATED.csv')
+
+# the final simple list with most up to date pheno label
+simple <- cur %>% 
+  filter(fid %in% final$fid)
+
+simple %>% write_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')
+```
+
+```{r}
 pheno <- cur %>% 

### 2021-04-19 scripts/43_comparisons_3.Rmd
great got this shit done
udpate

@@
 simple %>% write_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')
+
+pheno <- simple %>% 
+  gather('pheno', 'value', cr_d100:crs) 
 ```
@@
 ```{r}
-pheno <- cur %>% 
-  gather('pheno', 'value', cr_d100:crs)
 
@@
 
-cur %>% 
-  write_csv('../data/shotgun_pheno/final_comprehensive_updated_simple.csv')
 ```

### 2021-04-19 scripts/61_shotgun_taxa_lefse.Rmd
udpate

--- 2021-04-14 scripts/43_comparisons_3.Rmd
@@
 
-SUBSTITUTE WITH HUMANN3 RESULTS WHEN I CAN
+For Marcel's slides , including the 9 that were removed and do taxa lefse (2021-4-19)
 
 ```{r}
+# UPDATE THIS TABLE BABY
 cur <- read_csv('../data/shotgun_pheno/final_comprehensive.csv') %>% 
@@
   gather('pheno', 'value', cr_d100:crs) 
-
-sub_phenos <- pheno %>% 
-  split(., list(.$center, .$pheno)) %>% 
-  purrr::imap(~ filter(.data = ., value != 'not_assessed'))
-
-cur %>% 
-  write_csv('../data/shotgun_pheno/final_comprehensive_updated_simple.csv')
-```
-
-```{r}
-# the pheno for the different center and tox/CR_d30
-sub_phenos %>% 
-  imap(function(.x, .y){
-    select(.data = .x, value) %>% 
-      t() %>% 
-      write.table(str_glue('../data/shotgun_output/humann3/{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
-  })
-```
-
-```{r}
-# the pathway counts table
-full <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_pathabundance_CART.tsv')  %>% 
-  rename_all(funs(str_replace(., '^CART_','')))  %>% 
-  rename_all(funs(str_replace(., '_humann3$','')))
-
-pcts <- sub_phenos %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_pcts.tsv'))) 
-```
-
-## the genefamilies regrouping
-
-```{bash}
-# regroup the whole gene family table here
-humann_regroup_table --input humann3_genefamilies_CART.tsv --groups uniref90_ko --output humann3_KO_CART.tsv
-```
-
-```{r}
-fullko <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_KO_CART.tsv')  %>% 
-  rename_all(funs(str_replace(., '^CART_','')))  %>% 
-  rename_all(funs(str_replace(., '_humann3$','')))
-
-kocts <- sub_phenos %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ fullko %>% select(`# Gene Family`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_kocts.tsv'))) 
-```
-
-```{r}
-# the lefse for the ko only
-fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_kocts.tsv$')
-
-cmds <- tibble(
-  fns = fns
-) %>% 
-  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
-  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.png  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
-  select(-fns) %>% 
-  gather() %>% 
-  select(value) %>% 
-  write_csv('../data/shotgun_output/humann3/lefse_run_cmd_KO.sh', col_names = F)
 ```
@@
 
-all_pcts <- all_sub_pheno %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
-```
-
-```{r}
-# KO
-all_kocts <- all_sub_pheno %>% 
-  map(~ pull(.data = ., fid) ) %>%  
-  imap(~ fullko %>% select(`# Gene Family`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_kocts.tsv')))
-```
-
-```{r}
-all_gfcts <- all_sub_pheno %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ fullgenefam %>% select(`# Gene Family`, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_gfcts.tsv'))) 
-```
-
-## Need to go to snakemake for some steps related to humann before the lefse
-
-## generate command to run lefse in py2 environment
-
-cuz can't get them thru snakemake
-
-```{r}
-# PATHWAY
-# I already normalized the pcts so don't need to normalize again here 
-fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv$')
-
-cmds <- tibble(
-  fns = fns
-) %>% 
-  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2')) %>% 
-  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
-  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
-  select(-fns) %>% 
-  gather() %>% 
-  select(value) %>% 
-  write_csv('../data/shotgun_output/humann3/lefse_run_cmd.sh', col_names = F)
 ```
@@
   rename_all(funs(str_replace(., '_profile$','')))
-
-tcts <- sub_phenos %>% 
-  map(~ pull(.data = ., fid) ) %>% 
-  imap(~ taxa %>% select(clade_name, matches(.x)) %>% 
-         write_tsv(str_glue('../data/shotgun_output/humann3/{.y}_tcts.tsv')))
 
@@
 ```
-
-## PCOA to see if batch effects
-
-```{r}
-# very quicly done some pcoa
-cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
-  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
-  rename(pathway =  names(.)[1]) %>% 
-  spread('pathway', 'value') %>% 
-  column_to_rownames('sampleid')
-
-dist_bl <- vegan::vegdist(cts, method = 'bray')
-
-bl_bc <- cmdscale(dist_bl, k = 2) 
-pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
-  mutate(sampleid = fid)
-
-bl_bc %>% 
-    as.data.frame() %>% 
-    rownames_to_column('sampleid') %>% 
-    inner_join(pheno, by = 'sampleid') %>% 
-    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
-```
-```{r}
-empty_samp <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
-  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
-  group_by(sampleid) %>% 
-  summarise(Sum = sum(value)) %>% 
-  filter(Sum == 0) %>% 
-  pull(sampleid)
-
-cts <- read_tsv('../data/shotgun_output/humann2/humann_final_out/pull_toxicity_kocts_cpm_unstratified.tsv') %>% 
-  gather('sampleid', 'value', names(.)[2]:names(.)[ncol(.)]) %>% 
-  rename(KO =  names(.)[1]) %>% 
-  filter(!sampleid %in% empty_samp) %>% 
-  spread('KO', 'value') %>% 
-  column_to_rownames('sampleid')
-
-dist_bl <- vegan::vegdist(cts, method = 'bray')
-
-bl_bc <- cmdscale(dist_bl, k = 2) 
-pheno <- read_csv('../data/shotgun_pheno/final_simplified.csv') %>% 
-  mutate(sampleid = fid)
-
-bl_bc %>% 
-    as.data.frame() %>% 
-    rownames_to_column('sampleid') %>% 
-    inner_join(pheno, by = 'sampleid') %>% 
-    ggpubr::ggscatter(x = 'V1', y = 'V2', color =  'center',  shape = 'toxicity') 
-```

### 2021-06-28 scripts/57_16s_lefse.Rmd
update

@@
 ```{r}
-# sort out the genera counts table and also do filtering (need to have all taxa levels)
+# sort out the asv counts table and also do filtering (need to have all taxa levels)
 meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
@@
   select(-asv_key)  %>% 
-  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1])
+  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1]) %>% 
+  left_join(meta %>% select(sampleid, cr_d100, toxicity), by = 'sampleid')
 
 # the asv to keep
+# keep the asvs that show up in at least 10% of the samples
 keepg <- cts_all %>% 
-  filter(relab > 0) %>%
+  filter(relab > 0.0001) %>%
   ungroup() %>% 
@@
   filter(taxa_asv %in% keepg) %>% 
-  spread('sampleid', 'relab')
+  spread('sampleid', 'relab', fill = 0)
 ```
@@
 
-cat pull_toxicity.txt toxicity_asv_tcts.tsv > pull_toxicity_asv_tcts.tsv
-cat pull_cr_d100.txt cr_d100_asv_tcts.tsv > pull_cr_d100_asv_tcts.tsv 
-cat pull_crs.txt crs_asv_tcts.tsv > pull_crs_asv_tcts.tsv
-cat pull_icans.txt icans_asv_tcts.tsv > pull_icans_asv_tcts.tsv
+```{bash}
+cat ../data/amplicon/lefse/pull_toxicity.txt ../data/amplicon/lefse/toxicity_asv_tcts.tsv > ../data/amplicon/lefse/pull_toxicity_asv_tcts.tsv
+cat ../data/amplicon/lefse/pull_cr_d100.txt ../data/amplicon/lefse/cr_d100_asv_tcts.tsv > ../data/amplicon/lefse/pull_cr_d100_asv_tcts.tsv 
+cat ../data/amplicon/lefse/pull_crs.txt ../data/amplicon/lefse/crs_asv_tcts.tsv > ../data/amplicon/lefse/pull_crs_asv_tcts.tsv
+cat ../data/amplicon/lefse/pull_icans.txt ../data/amplicon/lefse/icans_asv_tcts.tsv > ../data/amplicon/lefse/pull_icans_asv_tcts.tsv
+```
 
@@
 
-# No differentially abundant features found in the 4
-# also this is the case when using center as subclass 
 ```
+
+```{r}
+# to see the relative abundance of those taxa 
+# to get the top and bottom three taxa of the lefse results
+res <- list.files('../data/amplicon/lefse/', pattern = 'asv_tcts.tsv.res$', full.names = T)
+
+# gather the species level taxa in the lefse significant results
+res_all <- res %>% 
+  set_names(res) %>% 
+  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
+        filter(!is.na(score))) %>% 
+  keep(~ nrow(.) > 0) %>% 
+  bind_rows(.id = 'res') %>% 
+  mutate(res = str_replace(res, '^.+//',''),
+         res = str_replace(res, '_asv.+$',''))  %>% 
+  rename(grp = res) %>% 
+  filter(grp %in% c('pull_cr_d100','pull_toxicity')) %>% 
+  mutate(feature = str_replace_all(feature, '\\.','\\|')) %>% 
+  #split(., list(.$grp, .$direction)) %>% 
+  #map_dfr(~ top_n(x = ., n = 4, wt = score) %>% arrange(-score)) %>% 
+  filter(str_detect(feature, 's__.+$')) %>% 
+  filter(!str_detect(feature, 'a__.+$')) %>% 
+  mutate(feature = str_replace(feature, '^.+s__','')) %>% 
+  mutate(feature = str_replace(feature, '_Clostridium_', '[Clostridium]')) %>% 
+  ungroup()
+
+# plot the relab of those taxa (at species level) in boxplot
+# get the species counts of the sampels
+cts_spp <- cts_ %>% 
+  full_join(annot %>%  select(asv_key, species), by  = 'asv_key') %>% 
+  select(-asv_key)  %>% 
+  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1]) %>% 
+  group_by(sampleid, species) %>% 
+  summarise(Relab = sum(relab)) %>% 
+  select(sampleid, species, Relab) %>% 
+  left_join(meta %>% select(sampleid, cr_d100, toxicity), by = 'sampleid') %>% 
+  ungroup()
+
+joined <- cts_spp %>% 
+  inner_join(res_all, by = c('species' = 'feature'))
+
+# finally I can do the plotting
+
+joined  %>%
+  filter(grp == 'pull_cr_d100') %>% 
+  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter') +
+  facet_wrap(direction ~ species)
+
+```
+
+```{r}
+joined  %>%
+  filter(grp == 'pull_toxicity') %>% 
+  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter') +
+  facet_wrap(direction ~ species)
+```

### 2021-06-28 scripts/48_16s_beta_diversity.Rmd
finally getting something done basically finishing my list

@@
 ```
+
+# do a PCOA between the healthy and the CART baseline at asv level
+
+```{r}
+library(vdbR)
+connect_database('~/dbConfig.txt')
+get_table_from_database("healthy_volunteers_ag")
+get_table_from_database("asv_alpha_diversity_ag")
+stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
+
+healthy <- healthy_volunteers_ag %>% 
+  inner_join(asv_alpha_diversity_ag, by = c("sampleid", "oligos_id")) 
+cts <- get_counts_subset(c(stb$sampleid, healthy %>% pull(sampleid)))
+
+# a total of 75 samples counts there are 3 healthy samples I don't have count ???
+nsamp <- cts %>% 
+  distinct(sampleid) %>% 
+  nrow
+
+all_pheno <- bind_rows(healthy %>% 
+  select(sampleid) %>% 
+  mutate(grp = 'healthy', center = 'healthy'),
+  stb %>% select(sampleid, center)  %>% 
+    mutate(grp = 'CART') %>% 
+    select(sampleid, grp, center)
+  ) %>% 
+  ungroup
+
+# filter >0.01% in more than 10% samples
+keepa <- cts %>% 
+  filter(count_relative > 0.0001) %>% 
+  count(asv_key) %>% 
+  filter(n > floor(nsamp * 0.1)) %>% 
+  pull(asv_key)
+
+cts_fil <- cts %>% 
+  filter(asv_key %in% keepa) %>% 
+  select(sampleid, asv_key,count_relative ) %>% 
+  spread(key = 'asv_key', value = 'count_relative', fill = 0) %>% 
+  column_to_rownames('sampleid')
+
+dist_ <- vegdist(cts_fil, method = 'bray')
+bc <- cmdscale(dist_, k = 2)
+
+bc %>%
+  as.data.frame() %>%
+  rownames_to_column('sampleid') %>% 
+  ungroup() %>% 
+  inner_join(all_pheno) %>% 
+  distinct() %>% 
+  ggscatter(x = 'V1', y = 'V2', color =  'grp') +
+  labs(title = 'PCOA of healthy and CART patients') +
+  #theme_void() +
+  ggsave('../figs/PCOA of healthy and CART patients.pdf')
+
+bc %>%
+  as.data.frame() %>%
+  rownames_to_column('sampleid') %>% 
+  ungroup() %>% 
+  inner_join(all_pheno) %>% 
+  distinct() %>% 
+  ggscatter(x = 'V1', y = 'V2', color =  'center') +
+  labs(title = 'PCOA of healthy and CART patients center') +
+  #theme_void() +
+  ggsave('../figs/PCOA of healthy and CART patients centetr.pdf')
+
+```

### 2021-06-28 scripts/57_16s_lefse.Rmd
finally getting something done basically finishing my list

@@
   filter(grp == 'pull_cr_d100') %>% 
-  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter') +
-  facet_wrap(direction ~ species)
+  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
+  facet_wrap(direction ~ species) +
+  ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)
 
@@
   filter(grp == 'pull_toxicity') %>% 
-  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter') +
-  facet_wrap(direction ~ species)
+  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
+  facet_wrap(direction ~ species) +
+  ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 15, height = 13)
 ```

### 2021-06-30 scripts/62_logistic_regression_version.Rmd
every movement is so hard

---
title: "logistic regression"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(magrittr)
library(ggeffects)
library(tidyverse)
library(lme4)
```

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
  mutate(center = factor(center),
         toxicity = factor(toxicity, levels = c('no','yes')),
         cr_d100 = factor(cr_d100, levels = c('no','yes')))

dat <- meta %>% 
  select(toxicity,cr_d100, simpson_reciprocal, center)

dat %>% 
  head
```

# Alpha diversity as predictor

## Toxicity ~ alpha diversity + center

```{r}
m_tox <- glm(
  toxicity ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)
summary(m_tox)

ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal','center')) %>% plot()
```

## CR ~ alpha diversity + center

```{r}
m_CR <- glm(
  cr_d100 ~ simpson_reciprocal + center, 
  data = dat, 
  family = binomial(link = "logit")
)
summary(m_CR)

ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal','center')) %>% plot()
```

# 5 genera as predictor

## Toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center

```{r}
genera <- read_csv('../data/amplicon/stool/combined_5_genera.csv')

gmeta <- meta  %>% 
  inner_join(genera) %>% 
  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)
```

```{r}
g_tox <- glm(
  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta, 
  family = binomial(link = "logit")
)
summary(g_tox)

ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides','center')) %>% plot()
ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus','center')) %>% plot()
ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus','center')) %>% plot()
ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia','center')) %>% plot()
ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium','center')) %>% plot()
```

## CR ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center

```{r}
g_CR <- glm(
  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta, 
  family = binomial(link = "logit")
)
summary(g_CR)

ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides','center')) %>% plot()
ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus','center')) %>% plot()
ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus','center')) %>% plot()
ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia','center')) %>% plot()
ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium','center')) %>% plot()
```

### 2021-07-01 scripts/57_16s_lefse.Rmd
update

@@
 # the asv to keep
-# keep the asvs that show up in at least 10% of the samples
+# keep the asvs that show up in at least 25% of the samples
 keepg <- cts_all %>% 
@@
   count(taxa_asv) %>% 
-  filter(n > floor(nrow(meta) * 0.1)) %>% 
+  filter(n > floor(nrow(meta) * 0.25)) %>% 
   pull(taxa_asv)
@@
   spread('sampleid', 'relab', fill = 0)
-```
-
-```{r}
-# dat <- genera %>% 
-#   filter(genus == 'Ruminococcus') %>% 
-#   gather('sampleid', 'relab', names(.)[2]:names(.)[ncol(.)]) %>% 
-#   full_join(meta) %>% 
-#   mutate(logrelab =  log10(relab + 2e-6)) 
-# 
-# yes <- dat %>% 
-#   filter(cr_d100 == 'yes')
-# no <- dat %>% 
-#   filter(cr_d100 == 'no')
-# 
-# 
-# dat %>% 
-#   #filter(center == 'M') %>% 
-#   mutate(logrelab =  log10(relab + 2e-6)) %>% 
-#   ggboxplot(x = 'cr_d100', y = 'logrelab', add = 'jitter') +
-# stat_compare_means(comparisons= list(c('yes', 'no')),
-# 										label= "p.signif",
-# 										method= 'wilcox.test',
-# 										correct=FALSE)
-# 
-# t.test(yes$logrelab, no$logrelab)
-
 ```

### 2021-07-01 scripts/62_logistic_regression_version.Rmd
update

@@
 library(lme4)
+library(cowplot)
+library(grid)
+library(ggpubr)
 ```
@@
 
+**The ggeffects package computes estimated marginal means (predicted values) for the response, at the margin of specific values or levels from certain model terms, i.e.
+it generates predictions by a model by holding the non-focal variables constant and varying the focal variable(s).
+ggpredict() uses predict() for generating predictions**
+
 ```{r}
@@
 
-ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal','center')) %>% plot()
+ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
 ```
@@
 
-ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal','center')) %>% plot()
+ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
 ```
@@
 
+The genera's relative abundance is log10 transformed.
+
 ```{r}
@@
 
+# the CLR transformed genera abundance 
+clr <- read_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv')  %>% 
+  spread(key = 'genus', value = 'clr')
+
 gmeta <- meta  %>% 
-  inner_join(genera) %>% 
+  inner_join(clr) %>% 
   select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)
+
+gmeta %>% 
+  head
 ```
@@
 
-ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides','center')) %>% plot()
-ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus','center')) %>% plot()
-ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus','center')) %>% plot()
-ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia','center')) %>% plot()
-ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium','center')) %>% plot()
+gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
+gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
+ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
+ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
+gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
+
+gtox <- ggarrange(gb, gr, ge, ga, gf,
+    ncol = 1
+    #o = 'hv',
+    #labels = 'AUTO',
+    #axis = 'bltr'
+    ) +
+    ggsave(str_glue('../figs/amplicon/gtox_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
 ```
@@
 
-ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides','center')) %>% plot()
-ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus','center')) %>% plot()
-ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus','center')) %>% plot()
-ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia','center')) %>% plot()
-ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium','center')) %>% plot()
+gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
+gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
+ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
+ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
+gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
+
+gCR <- ggarrange(gb, gr, ge, ga, gf,
+    ncol = 1
+    #o = 'hv',
+    #labels = 'AUTO',
+    #axis = 'bltr'
+    ) +
+    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
 ```

### 2021-07-01 scripts/63_CLR_transformation.Rmd
update

---
title: "CLR transformation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(compositions)
library(tidyverse)
```

```{r}
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  select(-center) %>% 
  spread(key = 'sampleid', value = 'relab', fill = 0) %>% 
  mutate(genus = if_else(is.na(genus), 'NA',genus)) %>% 
  column_to_rownames('genus')

clr_res <- clr(genus) %>% 
  as.data.frame()  %>% 
  rownames_to_column('genus') %>% 
  gather('sampleid','clr', names(.)[2]:names(.)[ncol(.)])

clr_5 <- clr_res %>% 
  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus'))

clr_5 %>% 
  write_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv')
```

### 2021-07-01 scripts/48_16s_beta_diversity.Rmd
every movement is so hard
update
udpate

@@
   ) %>% 
-  ungroup
+  ungroup %>% 
+  inner_join(asv_alpha_diversity_ag %>% 
+               select(path_pool, sampleid))
 
@@
   #theme_void() +
-  ggsave('../figs/PCOA of healthy and CART patients.pdf')
+  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients.pdf')
 
@@
   #theme_void() +
-  ggsave('../figs/PCOA of healthy and CART patients centetr.pdf')
+  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients centetr.pdf')
 
 ```
+
+```{r}
+# a pcoa at asv level for only M and P
+cts <- get_counts_subset(c(stb$sampleid))
+
+keepa <- cts %>% 
+  filter(count_relative > 0.0001) %>% 
+  count(asv_key) %>% 
+  filter(n > floor(nsamp * 0.1)) %>% 
+  pull(asv_key)
+
+cts_fil <- cts %>% 
+  filter(asv_key %in% keepa) %>% 
+  select(sampleid, asv_key,count_relative ) %>% 
+  spread(key = 'asv_key', value = 'count_relative', fill = 0) %>% 
+  column_to_rownames('sampleid')
+
+dist_ <- vegdist(cts_fil, method = 'bray')
+bc <- cmdscale(dist_, k = 2)
+
+mp <- bc %>%
+  as.data.frame() %>%
+  rownames_to_column('sampleid') %>% 
+  ungroup() %>% 
+  inner_join(all_pheno) %>% 
+  distinct(sampleid, .keep_all = T)  %>% 
+  mutate(pool = str_extract(path_pool, 'Sample.+/'))
+
+mp %>% 
+  ggscatter(x = 'V1', y = 'V2', color =  'pool', shape = 'center') +
+  labs(title = 'PCOA of healthy and CART patients') +
+  #theme_void() +
+  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 5, height = 6)
+```

### 2021-07-01 scripts/57_16s_lefse.Rmd
udpate

@@
   #map_dfr(~ top_n(x = ., n = 4, wt = score) %>% arrange(-score)) %>% 
-  filter(str_detect(feature, 's__.+$')) %>% 
-  filter(!str_detect(feature, 'a__.+$')) %>% 
-  mutate(feature = str_replace(feature, '^.+s__','')) %>% 
+  # filter(str_detect(feature, 's__.+$')) %>% 
+  # filter(!str_detect(feature, 'a__.+$')) %>% 
+  filter(str_detect(feature, 'g__.+$')) %>% 
+  filter(!str_detect(feature, 's__.+$')) %>% 
+  mutate(feature = str_replace(feature, '^.+g__','')) %>% 
   mutate(feature = str_replace(feature, '_Clostridium_', '[Clostridium]')) %>% 
@@
 
-joined <- cts_spp %>% 
-  inner_join(res_all, by = c('species' = 'feature'))
+cts_genus <- cts_ %>% 
+  full_join(annot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
+  select(-asv_key)  %>% 
+  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1]) %>% 
+  group_by(sampleid, genus) %>% 
+  summarise(Relab = sum(relab)) %>% 
+  select(sampleid, genus, Relab) %>% 
+  left_join(meta %>% select(sampleid, cr_d100, toxicity), by = 'sampleid') %>% 
+  ungroup()
+
+joined <- cts_genus %>% 
+  inner_join(res_all, by = c('genus' = 'feature'))
 
@@
   ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
-  facet_wrap(direction ~ species) +
+  facet_wrap(direction ~ genus) +
   ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)
@@
   ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
-  facet_wrap(direction ~ species) +
+  facet_wrap(direction ~ genus) +
   ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 15, height = 13)

### 2021-07-01 scripts/62_logistic_regression_version.Rmd
udpate

@@
 
-gtox <- ggarrange(gb, gr, ge, ga, gf,
+gtox_clr <- ggarrange(gb, gr, ge, ga, gf,
     ncol = 1
@@
     #axis = 'bltr'
-    ) +
-    ggsave(str_glue('../figs/amplicon/gtox_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
+    ) 
 ```
@@
 
-gCR <- ggarrange(gb, gr, ge, ga, gf,
+gCR_clr <- ggarrange(gb, gr, ge, ga, gf,
     ncol = 1
@@
 ```
+
+# log 10 transformed version of the genera relab
+
+```{r}
+gmeta_log10 <- meta  %>% 
+  inner_join(genera) %>% 
+  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)
+
+```
+
+```{r}
+g_tox <- glm(
+  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
+  data = gmeta_log10, 
+  family = binomial(link = "logit")
+)
+summary(g_tox)
+gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
+gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
+ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
+ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
+gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
+gtox_log <- ggarrange(gb, gr, ge, ga, gf,
+    ncol = 1
+    #o = 'hv',
+    #labels = 'AUTO',
+    #axis = 'bltr'
+    ) +
+    ggsave(str_glue('../figs/amplicon/gtox_log10_5.pdf'), width = 3, height = 8, dpi = 300)
+```
+
+## CR ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center
+
+```{r}
+g_CR <- glm(
+  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
+  data = gmeta_log10, 
+  family = binomial(link = "logit")
+)
+summary(g_CR)
+gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
+gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
+ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
+ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
+gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
+gCR_log <- ggarrange(gb, gr, ge, ga, gf,
+    ncol = 1
+    #o = 'hv',
+    #labels = 'AUTO',
+    #axis = 'bltr'
+    ) 
+```
+
+# side-by-side comparison of the CLR and log10 transformed
+
+```{r}
+tox_all <- ggarrange(gtox_log, gtox_clr, nrow = 1) +
+    ggsave(str_glue('../figs/amplicon/gtox_all_5.pdf'), width = 6, height = 8, dpi = 300)
+
+cr_all <- ggarrange(gCR_log, gCR_clr, nrow = 1) +
+    ggsave(str_glue('../figs/amplicon/gCR_all_5.pdf'), width = 6, height = 8, dpi = 300)
+```

### 2021-07-13 scripts/43_comparisons_3.Rmd
update

@@
 simple %>% write_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')
+
+simple <- read_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')
 
@@
          write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
+```
+
+```{r}
+# add a filtering step here 
+# 50 and 25% 
+pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(pw = `# Pathway`)
+
+keeppw <- pcts %>% 
+  filter(cpm > 50) %>%
+  ungroup() %>% 
+  count(pw) %>% 
+  filter(n > floor(nrow(simple) * 0.25)) %>% 
+  pull(pw)
+
+cts_fil <- pcts %>% 
+  filter(pw %in% keeppw) %>% 
+  spread('sampleid', 'cpm', fill = 0)
+
+cts_fil %>% 
+  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified_fil.tsv')
+```
+
+```{r}
+# tox
+pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
+  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
+  rename(pw = `# Pathway`)
+
+keeppw <- pcts %>% 
+  filter(cpm > 50) %>%
+  ungroup() %>% 
+  count(pw) %>% 
+  filter(n > floor(nrow(simple) * 0.25)) %>% 
+  pull(pw)
+
+cts_fil <- pcts %>% 
+  filter(pw %in% keeppw) %>% 
+  spread('sampleid', 'cpm', fill = 0)
+
+cts_fil %>% 
+  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified_fil.tsv')
 ```

### 2021-07-13 scripts/48_16s_beta_diversity.Rmd
update

@@
   distinct(sampleid, .keep_all = T)  %>% 
-  mutate(pool = str_extract(path_pool, 'Sample.+/'))
+  mutate(pool = str_extract(path_pool, 'Sample.+/')) %>% 
+  mutate(pool = str_replace(pool, 'Sample_','')) %>% 
+  mutate(pool = if_else(str_detect(pool, 'IGO'), str_extract(pool, 'IGO.+$'), pool)) %>% 
+  mutate(pool = str_replace(pool, '_1/|_comple.+$',''))
 
 mp %>% 
-  ggscatter(x = 'V1', y = 'V2', color =  'pool', shape = 'center') +
-  labs(title = 'PCOA of healthy and CART patients') +
+  ggscatter(x = 'V1', y = 'V2', color =  'pool') +
+  labs(title = 'PCOA of CART patients') +
   #theme_void() +

### 2021-07-13 scripts/54_rethinking_binary.Rmd
back from my wonderful vacation
update

@@
 library(ggpubr)
+
 ```
@@
 precis(mtox,  prob = 0.95)
-plot(precis(mtox, depth = 2, prob = 0.95) ) 
+plot(precis(mtox, depth = 2, prob = 0.95 ) , pars = ('b'), labels = c('simpson_reciprocal'))
 # there is baseline difference between penn and Msk and MSK has higher tox rage in general.
@@
 div_value <- c(-1, 1)
+
 ```
@@
 precis(mcr,  prob = 0.95)
-plot(precis(mcr, depth = 2, prob = 0.95) ) 
+plot(precis(mcr, depth = 2, prob = 0.95 ) , pars = ('b'), labels = c('simpson_reciprocal'))
 # there isn't much difference in baseline level between the centers

### 2021-07-13 scripts/57_16s_lefse.Rmd
update

@@
   ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
-  facet_wrap(direction ~ genus) +
+  facet_wrap(direction ~ genus, scales="free_y") +
   ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)
@@
   ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
-  facet_wrap(direction ~ genus) +
+  facet_wrap(direction ~ genus, scales="free_y") +
   ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 15, height = 13)

### 2021-07-13 scripts/62_logistic_regression_version.Rmd
update

--- 2021-07-01 scripts/62_logistic_regression_version.Rmd
@@
 library(ggpubr)
+library(sjPlot)
+theme_set(theme_sjplot())
 ```
@@
 )
+
 summary(m_tox)
 
-ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
+#ggpredict(m_tox,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
+
+# plot_model(m_tox, transform = NULL) +
+#   ggsave('../figs/amplicon/diversity_tox_logi.pdf', height = 3, width = 4)
+
+or_CI <- round(cbind(coef(m_tox), confint(m_tox)), digits=3) %>% 
+  as.data.frame()
+
+or_CI <- or_CI %>% 
+  mutate(variable=rownames(or_CI)) %>% 
+  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)
+
+# Now plot them
+or_CI[2,] %>%  #remove row number 1 (The intercept)
+  ggplot(aes(x = reorder(variable, V1), y = V1)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = lower_bound,
+                    ymax  = upper_bound),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-0.5, 1)) +
+  #geom_hline(yintercept = 1, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'Toxicity') +
+  ggsave('../figs/amplicon/logis_div_tox.pdf', width = 5, height = 4)
+
 ```
@@
 )
-summary(m_CR)
 
-ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
+or_CI <- round(cbind(coef(m_CR), confint(m_CR)), digits=3) %>% 
+  as.data.frame()
+
+or_CI <- or_CI %>% 
+  mutate(variable=rownames(or_CI)) %>% 
+  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)
+
+#Now plot them
+or_CI[2,] %>%  #remove row number 1 (The intercept)
+  ggplot(aes(x = reorder(variable, V1), y = V1)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = lower_bound,
+                    ymax  = upper_bound),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-0.5, 1)) +
+  #geom_hline(yintercept = 1, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'CR') +
+  ggsave('../figs/amplicon/logis_div_cr.pdf', width = 5, height = 4)
+```
+```{r}
+# plot_model(m_CR, transform = NULL, terms = c('simpson_reciprocal')) +
+#   ggsave('../figs/amplicon/diversity_cr_logi.pdf', height = 3, width = 4)
+# 
+# get_model_data(m_CR, type = "est")
+# log(get_model_data(m_CR, type = "est")$estimate)
+```
+
+```{r}
+#ggpredict(m_CR,se=TRUE,interactive=TRUE,digits=3, c('simpson_reciprocal')) %>% plot()
 ```
@@
 
-gtox <- ggarrange(gb, gr, ge, ga, gf,
+plot_model(g_tox)+
+  ggsave('../figs/amplicon/genera_tox_logi_CLR.pdf', height = 3, width = 4)
+```
+
+```{r}
+gtox_clr <- ggarrange(gb, gr, ge, ga, gf,
     ncol = 1
@@
     #axis = 'bltr'
-    ) +
-    ggsave(str_glue('../figs/amplicon/gtox_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
+    )  
 ```
@@
 
-gCR <- ggarrange(gb, gr, ge, ga, gf,
+plot_model(g_CR)+
+  ggsave('../figs/amplicon/genera_cr_logi_CLR.pdf', height = 3, width = 4)
+```
+
+```{r}
+gCR_clr <- ggarrange(gb, gr, ge, ga, gf,
     ncol = 1
@@
 ```
+
+# log 10 transformed version of the genera relab
+
+```{r}
+gmeta_log10 <- meta  %>% 
+  inner_join(genera) %>% 
+  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)
+
+```
+
+```{r}
+g_tox <- glm(
+  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
+  data = gmeta_log10, 
+  family = binomial(link = "logit")
+)
+summary(g_tox)
+gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
+gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
+ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
+ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
+gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
+gtox_log <- ggarrange(gb, gr, ge, ga, gf,
+    ncol = 1
+    #o = 'hv',
+    #labels = 'AUTO',
+    #axis = 'bltr'
+    ) +
+    ggsave(str_glue('../figs/amplicon/gtox_log10_5.pdf'), width = 3, height = 8, dpi = 300)
+
+plot_model(g_tox) +
+  ggsave('../figs/amplicon/genera_tox_logi_log10.pdf', height = 3, width = 4)
+```
+
+## CR ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center
+
+```{r}
+g_CR <- glm(
+  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
+  data = gmeta_log10, 
+  family = binomial(link = "logit")
+)
+summary(g_CR)
+gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
+gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
+ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
+ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
+gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
+gCR_log <- ggarrange(gb, gr, ge, ga, gf,
+    ncol = 1
+    #o = 'hv',
+    #labels = 'AUTO',
+    #axis = 'bltr'
+    )
+
+plot_model(g_CR)+
+  ggsave('../figs/amplicon/genera_cr_logi_log10.pdf', height = 3, width = 4)
+```
+
+# side-by-side comparison of the CLR and log10 transformed
+
+```{r}
+tox_all <- ggarrange(gtox_log, gtox_clr, nrow = 1) +
+    ggsave(str_glue('../figs/amplicon/gtox_all_5.pdf'), width = 6, height = 8, dpi = 300)
+
+cr_all <- ggarrange(gCR_log, gCR_clr, nrow = 1) +
+    ggsave(str_glue('../figs/amplicon/gCR_all_5.pdf'), width = 6, height = 8, dpi = 300)
+```

### 2021-07-13 scripts/64_Bayesian_with_bifido.Rmd
update

---
title: "bayesian with Bifido"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(rethinking)
library(tidyverse)  
library(ggpubr)
set.seed(123)
```

# CR/Toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + Bifidobacterium + center

```{r}
bifido <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
  filter(genus == 'Bifidobacterium') %>% 
  select(-center) %>% 
  spread(key = 'genus', value = 'relab') %>% 
  mutate(Bifidobacterium = log10(Bifidobacterium + 2*10^-6))

genera <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
  inner_join(bifido)

# do not standardize the log transformed relative abundance 
meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
  inner_join(genera) %>% 
  mutate(diseaseID = if_else(disease == 'NHL', 1, 2))

# wanna see the range of the 5 genera log 10 relab
meta %>% select(Akkermansia:Bifidobacterium) %>% 
  summary
```

```{r}
set.seed(123)   
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    Akkermansia = meta$Akkermansia,
    Bacteroides = meta$Bacteroides,
    Enterococcus = meta$Enterococcus,
    Faecalibacterium = meta$Faecalibacterium,
    Ruminococcus = meta$Ruminococcus,
    Bifidobacterium = meta$Bifidobacterium,
    did = meta$diseaseID
    ) 
```

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + bd*Bifidobacterium + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        bd ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gtox, prob = 0.95)   
plot(precis(gtox, depth = 2, prob = 0.95) )  
# the  Ruminococcus looks more significant now
```

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + bd*Bifidobacterium + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        bd ~ dnorm( 0 , 1),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, prob = 0.95) 
plot(precis(gcr, depth = 2, prob = 0.95) ) 
# no posterior prediction 
# genera from literature , 
```

# add disease to the original model

```{r}
gtox_d <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + d[did] + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        d[did] ~ dnorm( 0 , 0.5 ),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gtox_d, prob = 0.95)   
plot(precis(gtox_d, depth = 2, prob = 0.95) )  
# the disease effect is not strong as center 
```

```{r}
gcr_d <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + d[did] + a[location] ,
        ba ~ dnorm( 0 , 1),
        bb ~ dnorm( 0 , 1),
        be ~ dnorm( 0 , 1),
        bf ~ dnorm( 0 , 1),
        br ~ dnorm( 0 , 1),
        d[did] ~ dnorm( 0 , 0.5 ),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr_d, prob = 0.95) 
plot(precis(gcr_d, depth = 2, prob = 0.95) ) 
```

```{r}
# the distribution of the outcome in each of the disease
table(meta$disease, meta$cr_d100)
table(meta$disease, meta$toxicity)
```

### 2021-07-13 scripts/65_bayesian_with_CLR.Rmd
update

---
title: "Bayesian with CLR"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(rethinking)
library(tidyverse)  
library(ggpubr)
set.seed(123)

```

```{r}
clr_5 <- read_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv') %>% 
  spread(key = 'genus', value = 'clr')

meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
  inner_join(clr_5) 
```

```{r}
dat_list <- list(
    tox = meta$tox,
    cr100 = meta$cr100,
    location = meta$loca,
    Akkermansia = meta$Akkermansia,
    Bacteroides = meta$Bacteroides,
    Enterococcus = meta$Enterococcus,
    Faecalibacterium = meta$Faecalibacterium,
    Ruminococcus = meta$Ruminococcus
    ) 
```

```{r}
gtox <- ulam(
    alist(
        tox ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus  + a[location] ,
        ba ~ dnorm( 0 , 0.75),
        bb ~ dnorm( 0 , 0.75),
        be ~ dnorm( 0 , 0.75),
        bf ~ dnorm( 0 , 0.75),
        br ~ dnorm( 0 , 0.75),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 16)
precis(gtox, prob = 0.95)   
plot(precis(gtox, depth = 2, prob = 0.95) )  
# the  Ruminococcus looks more significant now
```

```{r}
# maybe I can do a prior check with one real sample value for a test
prior <- extract.prior( gtox , n=1e4, cores = 16 )

# test sample values
test_dat <- clr_5 %>% 
  slice_sample(n = 1)

# make it into a function so that I can do the m and p the same time
# prior_tox <- list(m = 1, p = 2) %>% 
#   map(function(center_){
#     ret = inv_logit( prior$ba * test_dat$Akkermansia + prior$bb * test_dat$Bacteroides + prior$be * test_dat$Enterococcus +  prior$bf * test_dat$Faecalibacterium + prior$br * test_dat$Ruminococcus + prior$a[, center_] ) 
#     return(ret)
#   })

ret_m = inv_logit( prior$ba * test_dat$Akkermansia + prior$bb * test_dat$Bacteroides + prior$be * test_dat$Enterococcus +  prior$bf * test_dat$Faecalibacterium + prior$br * test_dat$Ruminococcus + prior$a[, 1] )

ret_p = inv_logit( prior$ba * test_dat$Akkermansia + prior$bb * test_dat$Bacteroides + prior$be * test_dat$Enterococcus +  prior$bf * test_dat$Faecalibacterium + prior$br * test_dat$Ruminococcus + prior$a[, 2] )

# the distribution of probability
tibble(m = ret_m, p = ret_p) %>% 
  gather() %>% 
  gghistogram(x  = 'value', fill = 'key')
# a little uneven 
```

```{r}
gcr <- ulam(
    alist(
        cr100 ~ dbinom( 1 , p ) ,
        logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus  + a[location] ,
        ba ~ dnorm( 0 , 0.25),
        bb ~ dnorm( 0 , 0.25),
        be ~ dnorm( 0 , 0.25),
        bf ~ dnorm( 0 , 0.25),
        br ~ dnorm( 0 , 0.25),
        a[location] ~ dnorm( 0 , 0.5 )
    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
precis(gcr, prob = 0.95)
plot(precis(gcr, depth = 2, prob = 0.95) )
```

### 2021-07-13 scripts/54_rethinking_binary.Rmd
udpate

@@
 plot(precis(mtox, depth = 2, prob = 0.95 ) , pars = ('b'), labels = c('simpson_reciprocal'))
+
+# make this plot myself with a pre-defined range in the axis
+df <- precis(mtox,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = 'simpson_reciprocal')
+
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = reorder(variable, mean), y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-0.7, 1)) +
+  #geom_hline(yintercept = 1, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'Toxicity') +
+  ggsave('../figs/amplicon/bayesian_div_tox.pdf', width = 5, height = 4)
+
+```
+
+```{r}
 # there is baseline difference between penn and Msk and MSK has higher tox rage in general.
@@
 ```{r}
+# make this plot myself with a pre-defined range in the axis
+df <- precis(mcr,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = 'simpson_reciprocal')
+
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = reorder(variable, mean), y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-0.7, 1)) +
+  #geom_hline(yintercept = 1, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'CR') +
+  ggsave('../figs/amplicon/bayesian_div_cr.pdf', width = 5, height = 4)
+```
+
+```{r}
 # doing a prior check

### 2021-07-13 scripts/56_rethinking_genera.Rmd
back from my wonderful vacation
udpate

@@
 library(ggpubr)
+set.seed(123)
 ```
@@
         a[location] ~ dnorm( 0 , 0.5 )
-    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 16)
 precis(gtox, prob = 0.95)   
-plot(precis(gtox, depth = 2, prob = 0.95) )  
+plot(precis(gtox, depth = 2, prob = 0.95) , pars = c('ba','bb','be','bf','br'), labels = c('Akkermansia', 'Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus'))  
 # indicative of the Bacteroides link with toxicity  
+```
+
+```{r}
+# make this plot myself with a pre-defined range in the axis
+df <- precis(gtox,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = c('Akkermansia', 'Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus'))
+
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = variable, y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-1.5, 1.5)) +
+  geom_hline(yintercept = 0, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'Toxicity') +
+  ggsave('../figs/amplicon/bayesian_genera_tox.pdf', width = 5, height = 4)
 ```
@@
 # I think the Ruminococcus still holds for the CR
+```
+
+```{r}
+# make this plot myself with a pre-defined range in the axis
+df <- precis(gcr,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = c('Akkermansia', 'Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus'))
+
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = variable, y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-1.5, 1.5)) +
+  geom_hline(yintercept = 0, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'CR') +
+  ggsave('../figs/amplicon/bayesian_genera_cr.pdf', width = 5, height = 4)
 ```

### 2021-07-13 scripts/62_logistic_regression_version.Rmd
update

@@
 
-# the CLR transformed genera abundance 
-clr <- read_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv')  %>% 
-  spread(key = 'genus', value = 'clr')
-
 gmeta <- meta  %>% 
@@
 
-```{r}
-g_tox <- glm(
-  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
-  data = gmeta, 
-  family = binomial(link = "logit")
-)
-summary(g_tox)
-
-gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
-gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
-ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
-ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
-gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
-
-plot_model(g_tox)+
-  ggsave('../figs/amplicon/genera_tox_logi_CLR.pdf', height = 3, width = 4)
-```
-
-```{r}
-gtox_clr <- ggarrange(gb, gr, ge, ga, gf,
-    ncol = 1
-    #o = 'hv',
-    #labels = 'AUTO',
-    #axis = 'bltr'
-    )  
-```
-
 ## CR ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center
-
-```{r}
-g_CR <- glm(
-  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
-  data = gmeta, 
-  family = binomial(link = "logit")
-)
-summary(g_CR)
-
-gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
-gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
-ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
-ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
-gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
-
-plot_model(g_CR)+
-  ggsave('../figs/amplicon/genera_cr_logi_CLR.pdf', height = 3, width = 4)
-```
-
-```{r}
-gCR_clr <- ggarrange(gb, gr, ge, ga, gf,
-    ncol = 1
-    #o = 'hv',
-    #labels = 'AUTO',
-    #axis = 'bltr'
-    ) +
-    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
-```
 
@@
 
-plot_model(g_tox) +
-  ggsave('../figs/amplicon/genera_tox_logi_log10.pdf', height = 3, width = 4)
+# plot_model(g_tox) +
+#   ggsave('../figs/amplicon/genera_tox_logi_log10.pdf', height = 3, width = 4)
+```
+
+```{r}
+or_CI <- round(cbind(coef(g_tox), confint(g_tox)), digits=3) %>% 
+  as.data.frame()
+or_CI <- or_CI %>% 
+  mutate(variable=rownames(or_CI)) %>% 
+  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`) 
+# Now plot them
+or_CI[-c(1,7),] %>%  #remove row number 1 (The intercept)
+  ggplot(aes(x = variable, y = V1)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = lower_bound,
+                    ymax  = upper_bound),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-1.5, 1.5)) +
+  geom_hline(yintercept = 0, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'Toxicity') +
+  ggsave('../figs/amplicon/logis_genera_tox.pdf', width = 5, height = 4)
 ```
@@
 
+```{r}
+or_CI <- round(cbind(coef(g_CR), confint(g_CR)), digits=3) %>% 
+  as.data.frame()
+or_CI <- or_CI %>% 
+  mutate(variable=rownames(or_CI)) %>% 
+  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`) 
+# Now plot them
+or_CI[-c(1,7),] %>%  #remove row number 1 (The intercept)
+  ggplot(aes(x = variable, y = V1)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = lower_bound,
+                    ymax  = upper_bound),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-1.5, 1.5)) +
+  geom_hline(yintercept = 0, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'CR') +
+  ggsave('../figs/amplicon/logis_genera_cr.pdf', width = 5, height = 4)
+```
+
 # side-by-side comparison of the CLR and log10 transformed

### 2021-07-13 scripts/66_logistic_with_CLR.Rmd
update

---
title: "logistic with CLR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(magrittr) 
library(ggeffects)
library(tidyverse)
library(lme4)
library(cowplot)
library(grid)
library(ggpubr)
library(sjPlot)
theme_set(theme_sjplot())
```

```{r}
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
  mutate(center = factor(center),
         toxicity = factor(toxicity, levels = c('no','yes')),
         cr_d100 = factor(cr_d100, levels = c('no','yes')))

# the CLR transformed genera abundance 
clr <- read_csv('../data/amplicon/stool/counts_genus_combined_CLR.csv')  %>% 
  spread(key = 'genus', value = 'clr')

gmeta <- meta  %>% 
  inner_join(clr) %>% 
  select(toxicity,cr_d100, Akkermansia,Bacteroides, Enterococcus, Faecalibacterium, Ruminococcus , center)
```

```{r}
g_tox <- glm(
  toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta, 
  family = binomial(link = "logit")
)
summary(g_tox)

gb <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_tox,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()

# plot_model(g_tox)+
#   ggsave('../figs/amplicon/genera_tox_logi_CLR.pdf', height = 3, width = 4)
```

```{r}
or_CI <- round(cbind(coef(g_tox), confint(g_tox)), digits=3) %>% 
  as.data.frame()

or_CI <- or_CI %>% 
  mutate(variable=rownames(or_CI)) %>% 
  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)

# Now plot them
or_CI[-c(1,7),] %>%  #remove row number 1 (The intercept)
  ggplot(aes(x = variable, y = V1)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-1.5, 1.5)) +
  geom_hline(yintercept = 0, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'Toxicity') +
  ggsave('../figs/amplicon/logis_genera_tox_CLR.pdf', width = 5, height = 4)
```

```{r}
g_CR <- glm(
  cr_d100 ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + center, 
  data = gmeta, 
  family = binomial(link = "logit")
)
summary(g_CR)

gb <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Bacteroides')) %>% plot()
gr <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Ruminococcus')) %>% plot()
ge <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Enterococcus')) %>% plot()
ga <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Akkermansia')) %>% plot()
gf <- ggpredict(g_CR,se=TRUE,interactive=TRUE,digits=3, c('Faecalibacterium')) %>% plot()
# 
# plot_model(g_CR)+
#   ggsave('../figs/amplicon/genera_cr_logi_CLR.pdf', height = 3, width = 4)
```

```{r}
or_CI <- round(cbind(coef(g_CR), confint(g_CR)), digits=3) %>% 
  as.data.frame()

or_CI <- or_CI %>% 
  mutate(variable=rownames(or_CI)) %>% 
  rename(lower_bound = `2.5 %` , upper_bound = `97.5 %`)

# Now plot them
or_CI[-c(1,7),] %>%  #remove row number 1 (The intercept)
  ggplot(aes(x = variable, y = V1)) +
  geom_point(shape = 15,
             size  = 4, width = 0.1,
             position = "dodge", color="black") +
  geom_errorbar(aes(ymin  = lower_bound,
                    ymax  = upper_bound),
                width = 0.1,
                size  = 0.7,
                position = "dodge", color="turquoise4") +
  theme(axis.title = element_text(face = "bold")) +
  xlab("Variables") + ylab("Coefficient with 95% CI") +
  coord_flip(ylim = c(-1.5, 1.5)) +
  geom_hline(yintercept = 0, color = "red", size = 1) +
   theme(axis.title = element_text(size = 0)) +
  theme(axis.text = element_text(size = 14)) +
  labs(title = 'CR') +
  ggsave('../figs/amplicon/logis_genera_cr_CLR.pdf', width = 5, height = 4)
```

```{r}
gtox_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    )  
```
```{r}
gCR_clr <- ggarrange(gb, gr, ge, ga, gf,
    ncol = 1
    #o = 'hv',
    #labels = 'AUTO',
    #axis = 'bltr'
    ) +
    ggsave(str_glue('../figs/amplicon/gcr_CLR_5.pdf'), width = 3, height = 8, dpi = 300)
```

### 2021-07-19 scripts/48_16s_beta_diversity.Rmd
update

@@
 
+ttt <- all_pheno %>% 
+  ungroup() %>% 
+  filter(grp == 'healthy') %>% 
+  distinct(sampleid)
+
 # filter >0.01% in more than 10% samples
@@
   column_to_rownames('sampleid')
+library(vegan)
+dist_ <- vegdist(cts_fil, method = 'bray')
+eigen <- pcoa(dist_)$values$Eigenvalues
+percent_var <- signif(eigen/sum(eigen), 3)*100
 
-dist_ <- vegdist(cts_fil, method = 'bray')
 bc <- cmdscale(dist_, k = 2)
@@
   labs(title = 'PCOA of healthy and CART patients') +
+  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
+  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
   #theme_void() +
   ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients.pdf')
-
-bc %>%
-  as.data.frame() %>%
-  rownames_to_column('sampleid') %>% 
-  ungroup() %>% 
-  inner_join(all_pheno) %>% 
-  distinct() %>% 
-  ggscatter(x = 'V1', y = 'V2', color =  'center') +
-  labs(title = 'PCOA of healthy and CART patients center') +
-  #theme_void() +
-  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients centetr.pdf')
 
@@
 dist_ <- vegdist(cts_fil, method = 'bray')
+eigen <- pcoa(dist_)$values$Eigenvalues
+percent_var <- signif(eigen/sum(eigen), 3)*100
+
 bc <- cmdscale(dist_, k = 2)
@@
 mp %>% 
-  ggscatter(x = 'V1', y = 'V2', color =  'pool') +
+  ggscatter(x = 'V1', y = 'V2', color =  'pool', size = 3) +
   labs(title = 'PCOA of CART patients') +
+  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
+  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
   #theme_void() +
-  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 5, height = 6)
+  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 9, height = 9)
 ```

### 2021-07-19 scripts/54_rethinking_binary.Rmd
update

@@
 # another model with the crs grade 0,1,>=2 
-mcrs3 <- ulam(
-    alist(
-        crs3 ~ dordlogit( phi , cutpoints ),
-        phi <- b*logdiv_s + a[location] ,
-        b ~ dnorm( 0 , 1),  
-        a[location] ~ dnorm( 0 , 0.5 ),
-        cutpoints ~ dnorm( 0 , 1.5 )
-    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-plot(precis(mcrs3, depth = 2, prob = 0.95) ) 
+# mcrs3 <- ulam(
+#     alist(
+#         crs3 ~ dordlogit( phi , cutpoints ),
+#         phi <- b*logdiv_s + a[location] ,
+#         b ~ dnorm( 0 , 1),  
+#         a[location] ~ dnorm( 0 , 0.5 ),
+#         cutpoints ~ dnorm( 0 , 1.5 )
+#     ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+# plot(precis(mcrs3, depth = 2, prob = 0.95) ) 
 # the cutpoints are not of much interet, this suggests that the diversity really isn't having a correlation?
@@
   }) 
+hist(prior_tox$p)
 ```

### 2021-07-19 scripts/57_16s_lefse.Rmd
update

@@
   filter(asv_key %in% cts_$asv_key) %>% 
-  mutate(ordr =  if_else(ordr == '', str_glue('of_class_{class}'), ordr),
-         family =  if_else(family == '', str_glue('of_order_{ordr}'), family),
-         genus =  if_else(genus == '', str_glue('of_family_{family}'), genus),
-         species =  if_else(species == '', str_glue('of_genus_{genus}'), species)) %>% 
+  mutate(ordr =  if_else(is.na(ordr), str_glue('unknown_of_class_{class}'), ordr),
+         family =  if_else(is.na(family), str_glue('unknown_of_order_{ordr}'), family),
+         genus =  if_else(is.na(genus) , str_glue('unknown_of_family_{family}'), genus),
+         species =  if_else(is.na(species) , str_glue('unknown_of_genus_{genus}'), species)) %>% 
   mutate(taxa_asv = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}|a__{asv_key}'))
@@
   mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1  -u 2 -o 1000000   ')) %>% 
-  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
+  mutate(run_cmd = str_glue('run_lefse.py -l  4 {fns}.in  {fns}.res')) %>% 
   mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
@@
 
+```
+
+```{r}
+# run in terminal:
+# bash /Users/daia1/projects/CART_microbiome/data/amplicon/lefse/lefse_run_cmd_taxa.sh
 ```

### 2021-07-19 scripts/63_CLR_transformation.Rmd
update

@@
 library(tidyverse)
+library(ggpubr)
+```
+CLR works with counts, not relab. And also need to add a pseudocount
+
+```{r}
+meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
+
+library(vdbR)
+connect_database('~/dbConfig.txt')
+get_table_from_database('asv_annotation_blast_ag')
+cts <- get_counts_subset(meta$sampleid)
+cts_ <- cts %>% 
+  select(asv_key, sampleid, count)
+
+annot <- asv_annotation_blast_ag %>% 
+  filter(asv_key %in% cts_$asv_key)
+
+cts_all <- cts_ %>% 
+  left_join(annot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
+  group_by(sampleid, genus) %>% 
+  summarise(cnt = sum(count)) %>% 
+  spread('sampleid', 'cnt', fill = 0) %>% 
+  filter(!is.na(genus)) %>% 
+  column_to_rownames('genus')
 ```
@@
 ```{r}
-genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv') %>% 
-  select(-center) %>% 
-  spread(key = 'sampleid', value = 'relab', fill = 0) %>% 
-  mutate(genus = if_else(is.na(genus), 'NA',genus)) %>% 
-  column_to_rownames('genus')
-
-clr_res <- clr(genus) %>% 
+clr_res <- clr(cts_all + 0.5) %>% 
   as.data.frame()  %>% 
@@
 ```
+
+```{r}
+log10 <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
+  gather('genus', 'log10', Akkermansia:Ruminococcus)
+
+two <- log10 %>% 
+  inner_join(clr_5)
+
+two %>% 
+  arrange(log10) %>% 
+  #filter(log10 > -5) %>% 
+  ggscatter(x = 'log10', y = 'clr', facet.by = 'genus', alpha = 0.4,
+             add = "reg.line",  # Add regressin line
+           add.params = list(color = "blue", fill = "lightgray"), # Customize line
+           conf.int = TRUE, # Add confidence interval
+           cor.coef = TRUE, # Add correlation coefficient.
+           cor.coeff.args = list(method = "pearson",  label.sep = "\n") )
+```

### 2021-07-19 scripts/65_bayesian_with_CLR.Rmd
update

@@
         logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus  + a[location] ,
-        ba ~ dnorm( 0 , 0.75),
-        bb ~ dnorm( 0 , 0.75),
-        be ~ dnorm( 0 , 0.75),
-        bf ~ dnorm( 0 , 0.75),
-        br ~ dnorm( 0 , 0.75),
+        ba ~ dnorm( 0 , 0.25),
+        bb ~ dnorm( 0 , 0.25),
+        be ~ dnorm( 0 , 0.25),
+        bf ~ dnorm( 0 , 0.25),
+        br ~ dnorm( 0 , 0.25),
         a[location] ~ dnorm( 0 , 0.5 )
@@
 # the  Ruminococcus looks more significant now
+```
+
+```{r}
+# make this plot myself with a pre-defined range in the axis
+df <- precis(gtox,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = c('Akkermansia', 'Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus'))
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = variable, y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-1.5, 1.5)) +
+  geom_hline(yintercept = 0, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'Toxicity') +
+  ggsave('../figs/amplicon/bayesian_genera_tox_CLR.pdf', width = 5, height = 4)
 ```
@@
 ```
+```{r}
+# make this plot myself with a pre-defined range in the axis
+df <- precis(gcr,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = c('Akkermansia', 'Bacteroides','Enterococcus','Faecalibacterium','Ruminococcus'))
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = variable, y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-1.5, 1.5)) +
+  geom_hline(yintercept = 0, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'CR') +
+  ggsave('../figs/amplicon/bayesian_genera_cr_CLR.pdf', width = 5, height = 4)
+
+get_table_from_database('counts_ag')
+```

### 2021-07-19 scripts/67_lefse_shotgun_taxa.Rmd
update

---
title: "lefse with shotgun taxa"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
simple <- read_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')

pheno <- simple %>% 
  gather('pheno', 'value', cr_d100:crs)

all_sub_pheno <- pheno %>% 
  split(., list(.$pheno)) %>% 
  purrr::imap(~ filter(.data = ., value != 'not_assessed'))

all_sub_pheno %>% 
  imap(function(.x, .y){
    select(.data = .x, value) %>% 
      t() %>% 
      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
  })
```

```{r}
taxa <- read_tsv('../data/shotgun_output/metaphlan/merged_abundance_table.txt')  %>% 
  rename_all(funs(str_replace(., '_profile$',''))) %>% 
  filter(str_detect(clade_name, 's__.+$')) %>% 
  select(- NCBI_tax_id) %>% 
  gather('sampleid', 'relab', names(.)[2]:names(.)[ncol(.)]) %>% 
  filter(sampleid %in% simple$fid)

keeptaxa <- taxa %>% 
  filter(relab > 0.05) %>%
  ungroup() %>% 
  count(clade_name) %>% 
  filter(n > floor(nrow(simple) * 0.25)) %>% 
  pull(clade_name)

cts_fil <- taxa %>% 
  filter(clade_name %in% keeptaxa) %>% 
  spread('sampleid', 'relab', fill = 0)

all_tcts <- all_sub_pheno %>% 
  map(~ pull(.data = ., fid) ) %>% 
  imap(~ cts_fil %>% select(clade_name, matches(.x)) %>% 
         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_tcts.tsv')))

```

```{bash}
cat ~/projects/CART_microbiome/data/shotgun_output/humann3/pull_toxicity.txt ~/projects/CART_microbiome/data/shotgun_output/humann3/pull_toxicity_tcts.tsv > ~/projects/CART_microbiome/data/shotgun_output/humann3/pull_toxicity_lefse_ready_tcts.tsv
cat ~/projects/CART_microbiome/data/shotgun_output/humann3/pull_cr_d100.txt ~/projects/CART_microbiome/data/shotgun_output/humann3/pull_cr_d100_tcts.tsv > ~/projects/CART_microbiome/data/shotgun_output/humann3/pull_cr_d100_lefse_ready_tcts.tsv
```

```{r}
# run lefse
fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_tcts.tsv$')

cmds <- tibble(
  fns = fns
) %>% 
  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2 -o 1000000')) %>% 
  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
  select(-fns) %>% 
  gather() %>% 
  select(value) %>% 
  write_csv('../data/shotgun_output/humann3/lefse_run_cmd_taxa.sh', col_names = F)

connect_database('~/dbConfig.txt')
get_table_from_database('asv_alpha_diversity_ag')
list_table_from_database('asv')
```

### 2021-07-21 scripts/68_rebuttal.Rmd
update

---
title: "CART rebuttal"
author: "Anqi Dai"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F)
```

```{r}
library(vegan)
library(ggpubr)
library(tidyverse)
library(grid)
library(cowplot)
```

# Reivewer 1

## Major point 2

**LEfSe is biased towards the identification of rare taxa and these rare taxa may be present in only one or two individuals. For this reason, please define parameters for minimum relative abundance and minimum number of patients in which taxa are seen. Please provide the relative abundance of the taxa of interest and the number of patients in which they are observed that are differentially abundant between groups.**

```{r}
# sort out the asv counts table and also do filtering (need to have all taxa levels)
meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('asv_annotation_blast_ag')
cts <- get_counts_subset(meta$sampleid)

cts_ <- cts %>% 
  select(asv_key, sampleid, count_relative) %>% 
  spread(key = 'sampleid', value = 'count_relative', fill = 0)

annot <- asv_annotation_blast_ag %>% 
  filter(asv_key %in% cts_$asv_key) %>% 
  mutate(ordr =  if_else(is.na(ordr), str_glue('unknown_of_class_{class}'), ordr),
         family =  if_else(is.na(family), str_glue('unknown_of_order_{ordr}'), family),
         genus =  if_else(is.na(genus) , str_glue('unknown_of_family_{family}'), genus),
         species =  if_else(is.na(species) , str_glue('unknown_of_genus_{genus}'), species)) %>% 
  mutate(taxa_asv = str_glue('k__{kingdom}|p__{phylum}|c__{class}|o__{ordr}|f__{family}|g__{genus}|s__{species}|a__{asv_key}'))

cts_all <- cts_ %>% 
  full_join(annot %>%  select(asv_key, taxa_asv), by  = 'asv_key') %>% 
  select(-asv_key)  %>% 
  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1]) %>% 
  left_join(meta %>% select(sampleid, cr_d100, toxicity), by = 'sampleid')

# the asv to keep
# keep the asvs that show up in at least 25% of the samples
keepg <- cts_all %>% 
  filter(relab > 0.0001) %>%
  ungroup() %>% 
  count(taxa_asv) %>% 
  filter(n > floor(nrow(meta) * 0.25)) %>% 
  pull(taxa_asv)

cts_fil <- cts_all %>% 
  filter(taxa_asv %in% keepg) %>% 
  spread('sampleid', 'relab', fill = 0)
```

```{r}
# the pheno label for the samples
pheno <- meta %>% 
  select(center, cr_d100:crs, icans, sampleid) %>% 
  gather('pheno', 'value', cr_d100:icans)

all_sub_pheno <- pheno %>% 
  split(., list(.$pheno)) %>% 
  purrr::imap(~ filter(.data = ., value != 'not_assessed'))

tpheno <- all_sub_pheno %>% 
  imap(function(.x, .y){
    select(.data = .x, value) %>% 
      t() %>% write.table(str_glue('../data/amplicon/lefse/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
  })

tcts <- all_sub_pheno %>% 
  map(~ pull(.data = ., sampleid) ) %>% 
  imap(~ cts_fil %>% select(taxa_asv, matches(.x)) %>%  write_tsv(str_glue('../data/amplicon/lefse/{.y}_asv_tcts.tsv'))) 
```

```{bash}
cat ../data/amplicon/lefse/pull_toxicity.txt ../data/amplicon/lefse/toxicity_asv_tcts.tsv > ../data/amplicon/lefse/pull_toxicity_asv_tcts.tsv
cat ../data/amplicon/lefse/pull_cr_d100.txt ../data/amplicon/lefse/cr_d100_asv_tcts.tsv > ../data/amplicon/lefse/pull_cr_d100_asv_tcts.tsv 
```

```{r}
fns <- list.files('../data/amplicon/lefse/', pattern = 'pull.*_asv_tcts.tsv$')

cmds <- tibble(
  fns = fns
) %>% 
  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1  -u 2 -o 1000000   ')) %>% 
  mutate(run_cmd = str_glue('run_lefse.py -l  4 {fns}.in  {fns}.res')) %>% 
  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
  mutate(clado_cmd = str_glue('plot_cladogram.py {fns}.res {fns}_clado.pdf  --label_font_size 4 --dpi 300 --format pdf --title {fns}')) %>% 
  select(-fns) %>% 
  gather() %>% 
  select(value) %>% 
  write_csv('../data/amplicon/lefse/lefse_run_cmd_taxa.sh', col_names = F)

```

```{r}
# run in terminal:
# bash /Users/daia1/projects/CART_microbiome/data/amplicon/lefse/lefse_run_cmd_taxa.sh
```

```{r}
# redo the resulted bar plot of the lefse in ggplot
# the input : lefse res files
fns <- list.files('../data/amplicon/lefse/', pattern = 'pull.*_asv_tcts.tsv.res$', full.names = T)

# join all of the tables feature together
feature <- fns %>% 
  set_names(fns) %>% 
  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  bind_rows(.id = 'group') %>% 
  mutate(group = str_replace(group, '../data/amplicon/lefse//pull_','')) %>% 
  mutate(group = str_replace(group, '_asv_tcts.tsv.res$',''))

# change the "N" direction to be minus score
feature <- bind_rows(
  feature %>% 
    split(.$direction) %>% 
    pluck('no') %>% 
    mutate(score = -score),
  feature %>% 
    split(.$direction) %>% 
    pluck('yes')  
) %>% 
  arrange(group, feature, score) %>% 
  mutate(feature = str_replace_all(feature, '^.+\\.', ''))
```

```{r}
all_title_fs <- 20
axis_text_fs <- 16

CR <- feature %>% 
  filter(group == 'cr_d100') %>% 
  ggplot(aes(x = reorder(feature, score), y = score, fill = direction)) +
            geom_bar( stat = 'identity') +
            coord_flip() +
            scale_color_manual(values = c('#925E9F', '#42B540'))  +
            scale_fill_manual(values = c('#925E9F', '#42B540')) +
            theme_classic() +
            theme(axis.title.y  = element_blank(), 
                  plot.title = element_text(size=all_title_fs),
                  axis.title.x = element_text(size=axis_text_fs),
                  axis.text.x = element_text(size=axis_text_fs),
                  legend.position='bottom') +
            labs(title = str_glue('Response to CR') ,
                 y = 'Score')

tox <- feature %>% 
  filter(group == 'toxicity') %>% 
  ggplot(aes(x = reorder(feature, score), y = score, fill = direction)) +
            geom_bar(stat = 'identity') +
            coord_flip() +
            scale_color_manual(values = c('#0099B4', '#AD002A'))  +
            scale_fill_manual(values = c('#0099B4', '#AD002A')) +
            theme_classic() +
            theme(axis.title.y  = element_blank(), 
                  axis.title.x = element_text(size=axis_text_fs),
                  plot.title = element_text(size=all_title_fs),
                  axis.text.x = element_text(size=axis_text_fs),
                  legend.position="bottom") +
            labs(title = str_glue('Response to Toxicity') ,
                 y = 'Score')

g <- plot_grid(CR,tox,  
          nrow = 2, 
          align = 'hv',
          axis = 'b')  +
  ggsave('../figs/amplicon/16s_lefse_combined.pdf', device = 'pdf', height = 15, width = 15)

g
```

```{r}
# to see the relative abundance of those taxa 
# to get the top and bottom three taxa of the lefse results
res <- list.files('../data/amplicon/lefse/', pattern = 'asv_tcts.tsv.res$', full.names = T)

# gather the species level taxa in the lefse significant results
res_all <- res %>% 
  set_names(res) %>% 
  map(~ read_tsv(., col_names = c('feature','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  keep(~ nrow(.) > 0) %>% 
  bind_rows(.id = 'res') %>% 
  mutate(res = str_replace(res, '^.+//',''),
         res = str_replace(res, '_asv.+$',''))  %>% 
  rename(grp = res) %>% 
  filter(grp %in% c('pull_cr_d100','pull_toxicity')) %>% 
  mutate(feature = str_replace_all(feature, '\\.','\\|')) %>% 
  #split(., list(.$grp, .$direction)) %>% 
  #map_dfr(~ top_n(x = ., n = 4, wt = score) %>% arrange(-score)) %>% 
  # filter(str_detect(feature, 's__.+$')) %>% 
  # filter(!str_detect(feature, 'a__.+$')) %>% 
  filter(str_detect(feature, 'g__.+$')) %>% 
  filter(!str_detect(feature, 's__.+$')) %>% 
  mutate(feature = str_replace(feature, '^.+g__','')) %>% 
  mutate(feature = str_replace(feature, '_Clostridium_', '[Clostridium]')) %>% 
  ungroup()

# plot the relab of those taxa (at species level) in boxplot
# get the species counts of the sampels
cts_spp <- cts_ %>% 
  full_join(annot %>%  select(asv_key, species), by  = 'asv_key') %>% 
  select(-asv_key)  %>% 
  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1]) %>% 
  group_by(sampleid, species) %>% 
  summarise(Relab = sum(relab)) %>% 
  select(sampleid, species, Relab) %>% 
  left_join(meta %>% select(sampleid, cr_d100, toxicity), by = 'sampleid') %>% 
  ungroup()

cts_genus <- cts_ %>% 
  full_join(annot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
  select(-asv_key)  %>% 
  gather('sampleid', 'relab', names(.)[1]:names(.)[ncol(.)-1]) %>% 
  group_by(sampleid, genus) %>% 
  summarise(Relab = sum(relab)) %>% 
  select(sampleid, genus, Relab) %>% 
  left_join(meta %>% select(sampleid, cr_d100, toxicity), by = 'sampleid') %>% 
  ungroup()

joined <- cts_genus %>% 
  inner_join(res_all, by = c('genus' = 'feature'))

# finally I can do the plotting

joined  %>%
  filter(grp == 'pull_cr_d100') %>% 
  ggboxplot(x = 'cr_d100', y = 'Relab', add = 'jitter', title = 'Outcome: cr_d100') +
  facet_wrap(direction ~ genus, scales="free_y") +
  ggsave('../figs/amplicon/lefse_taxa_crd100.pdf', width = 15, height = 13)

joined  %>%
  filter(grp == 'pull_toxicity') %>% 
  ggboxplot(x = 'toxicity', y = 'Relab', add = 'jitter', title = 'Outcome: toxicity') +
  facet_wrap(direction ~ genus, scales="free_y") +
  ggsave('../figs/amplicon/lefse_taxa_toxicity.pdf', width = 10, height = 13)
```

- Thank you reivewer for making this point. Previously the abundance filtering threshold for the ASV was set at 0 and prevalence threshold at 10%. The abundance threshold was set at 0 since I assumed I could be more lenient as the LEfSe was done at ASV level. Thus very rare taxa showed up in the results. Per the reviewer's comment, I've changed the abundance threshold to be at 0.01% and prevalence threshold at 25% (This is a small cohort with 45 samples).

- A lot of more commenly known taxa appeared in the updated results, and for display's purpose, bar plot only showed taxa with LDA score above 4. Moreover, as represented in the boxplot, the significant taxa (at genus level) look quite abundant.

- A big portion of the results are the same in the two outcomes. It doesn't quite help with making association conclusion, but it points us to the next step of selecting important genera for modeling.

## Major point 5

**Were controls used in this study--positive controls with known microbial consortia and negative reagent controls in order to account for potential contamination and to normalize abundance across batches? This is particularly important because several of the taxa identified using LEfSe analsyis could have represented contamination, yet we do not know their relative abundance and whether they were present in small numbers of reads or not. Were samples processed in batches and were batch effects taken into account?**

```{r}
get_table_from_database("asv_alpha_diversity_ag")
# a pcoa at asv level for only M and P
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

all_pheno <-  stb %>% select(sampleid, center)  %>% 
    mutate(grp = 'CART') %>% 
    select(sampleid, grp, center) %>% 
  ungroup %>% 
  inner_join(asv_alpha_diversity_ag %>% 
               select(path_pool, sampleid))

cts <- get_counts_subset(c(stb$sampleid))

nsamp <- nrow(stb)

keepa <- cts %>% 
  filter(count_relative > 0.0001) %>% 
  count(asv_key) %>% 
  filter(n > floor(nsamp * 0.25)) %>% 
  pull(asv_key)

cts_fil <- cts %>% 
  filter(asv_key %in% keepa) %>% 
  select(sampleid, asv_key,count_relative ) %>% 
  spread(key = 'asv_key', value = 'count_relative', fill = 0) %>% 
  column_to_rownames('sampleid')

dist_ <- vegdist(cts_fil, method = 'bray')
eigen <- pcoa(dist_)$values$Eigenvalues
percent_var <- signif(eigen/sum(eigen), 3)*100

bc <- cmdscale(dist_, k = 2)

mp <- bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct(sampleid, .keep_all = T)  %>% 
  mutate(pool = str_extract(path_pool, 'Sample.+/')) %>% 
  mutate(pool = str_replace(pool, 'Sample_','')) %>% 
  mutate(pool = if_else(str_detect(pool, 'IGO'), str_extract(pool, 'IGO.+$'), pool)) %>% 
  mutate(pool = str_replace(pool, '_1/|_comple.+$',''))

mp %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'pool', size = 3) +
  labs(title = 'PCOA of CART patients') +
  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) (ASV level)of CART patients_pool.pdf', width = 9, height = 9)
``` 
Controls were not used in the study. Several important taxa identified by LefSe could be statistical artefact due to the lenient filtering threshold. In terms of batch effects, as shown in the above scatter plot, 45 samples came from 13 different sequencing batches, and they look well mixed. Therefore I would conclude that batch effects wouldn't be a concern. Plus, in the following modeling step, center from which the sample was collected is incorporated as a covariet.

## Major point 13

**I would prefer to see an MDS/PCOA visualization instead of Figure 2C boxplots for beta-diversity. This helps much more with understanding how different communities are and what variability exists within communities**

```{r}
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database("healthy_volunteers_ag")
get_table_from_database("asv_alpha_diversity_ag")
stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')

healthy <- healthy_volunteers_ag %>% 
  inner_join(asv_alpha_diversity_ag, by = c("sampleid", "oligos_id")) 
cts <- get_counts_subset(c(stb$sampleid, healthy %>% pull(sampleid)))

# a total of 75 samples  
nsamp <- cts %>% 
  distinct(sampleid) %>% 
  nrow

all_pheno <- bind_rows(healthy %>% 
  select(sampleid) %>% 
  mutate(grp = 'healthy', center = 'healthy'),
  stb %>% select(sampleid, center)  %>% 
    mutate(grp = 'CART') %>% 
    select(sampleid, grp, center)
  ) %>% 
  ungroup %>% 
  inner_join(asv_alpha_diversity_ag %>% 
               select(path_pool, sampleid))

# filter >0.01% in more than 25% samples
keepa <- cts %>% 
  filter(count_relative > 0.0001) %>% 
  count(asv_key) %>% 
  filter(n > floor(nsamp * 0.25)) %>% 
  pull(asv_key)

cts_fil <- cts %>% 
  filter(asv_key %in% keepa) %>% 
  select(sampleid, asv_key,count_relative ) %>% 
  spread(key = 'asv_key', value = 'count_relative', fill = 0) %>% 
  column_to_rownames('sampleid')

library(vegan)
dist_ <- vegdist(cts_fil, method = 'bray')
eigen <- pcoa(dist_)$values$Eigenvalues
percent_var <- signif(eigen/sum(eigen), 3)*100

bc <- cmdscale(dist_, k = 2)

bc %>%
  as.data.frame() %>%
  rownames_to_column('sampleid') %>% 
  ungroup() %>% 
  inner_join(all_pheno) %>% 
  distinct() %>% 
  ggscatter(x = 'V1', y = 'V2', color =  'grp') +
  labs(title = 'PCOA of healthy and CART patients') +
  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
  #theme_void() +
  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients.pdf')
```

We agree a PCOA is a better visualization of the community difference between the healthy and the CART cohort. The PCOA is done at ASV level with same filtering threshold as above in the LEfSe and the Bray-Curtis distance was used. As illustrated in the plot, the CART and healthy cohort's microbiome is quite different.

## Major point 14

**Why was a Bayesian statistical approach used? I would like to either see “standard” statistical approaches or have a strong justification for why this is more appropriate and the other methods are not**

- Using Bayesian approach, we are able to obtain the posterior distribution of the coefficients, which retains the uncertainty of the coefficient estimates, and also do posterior predictive check to quantify the impact of the coefficients (as shown in the Fig 2H & J). The odds ratio in logistic regression is hard to interpretate.

## Major point 15

**Why was Ruminococcus chosen but not the Clostridium or Fusobacterium genera chosen for analysis in association with outcomes? The results seem disjointed and do not create a story to explain why the prior results are being ignored here. Why was Ruminococcus not observed in LEfSe for example?**

As mentioned in the above reply, rare taxa's apprearance such as Fusobacterium may just be statistical artefact due to the lenient filtering threshold. When the thresholds have been updated, Ruminococcus does show up and turn out to be quite significant with LDA score above 4 with taxa level spanning from family to species.

## Major point 17

**Please indicate when 16S sequencing results were used in analyses versus when shotgun sequencing was used. It is unclear why 16S was used for taxonomic results when shotgun sequencing was available for all of the samples as well. In addition, please include the median read sizes for each and the range of sizes. More data about the sequencing reads are needed.**

- 16s sequencing data was used in looking at the compositions, alpha, beta diversity, LefSe analysis and the Bayesian modeling, while shotgun data was used in LefSe analysis with the taxa and pathway abundance data.

- During the progession of the project, the 16s data was obtained sooner, which was why we did the major analysis with 16s data.

```{r}
# for the read count of the 16s samples after preprocessing
cts <- get_counts_subset(c(stb$sampleid))

sample_counts <- cts %>% 
  distinct(sampleid, count_total)

summary(sample_counts$count_total)
```

Thank reviewer for reminding us to add more details of the sequencing reads. For 16s data, the median read size is 50788 and the range is from 7041 to 93953. For the shotgun data,

## Minor point 17

**Figure 2: Are Ruminococcus/Bacteroides “low” and “high” relative abundance determined according to the median abundance? Or some other way?**

# Reviewer 2

## Major point 1

**Code availability statement does list names of software packages that have been used but does not provide any information on the code that implements the analyses included in this manuscript. It also seems unlikely that the actual code would have used as few packages than what has been mentioned in the Methods. The source code for the actual analyses of this manuscript has not been provided despite generally mentioning some external open source algorithms and software by name. The standard way of providing reproducible analysis source code is either as a supplementary zip file or as an online repository (such as Github), associated with permanent DOI (e.g. through Zenodo). Without the code it is not possible to assess the workflow details or assess the robustness of the workflows to varying methodological choices. This is concern in particular due to the other questions related to methodology (see below).**

Thank reviewer for bringing up this point. We've created a github page to display the code and the resulted figures. The address is: https://vdblab.github.io/CART_and_microbiome/

## Major point 2

**520-521: beta diversity has been calculated at genus level; the beta diversity is often calculated using the finest (ASV) resolution in order to use maximal information in the data, and in many cases this can have remarkable impact on the results. Are the results and conclusions robust to this choice?**

```{r}
# beta diversity at genus level but a PCOA between healthy and the CART
genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv')

keep <- genus %>% 
  filter(relab > 0.0001) %>% 
  count(genus) %>% 
  filter(n > floor(nsamp * 0.2)) %>% 
  filter(!is.na(genus)) %>% 
  pull(genus) 
```

### 2021-07-21 scripts/48_16s_beta_diversity.Rmd
update

@@
   count(asv_key) %>% 
-  filter(n > floor(nsamp * 0.1)) %>% 
+  filter(n > floor(nsamp * 0.25)) %>% 
   pull(asv_key)

### 2021-07-21 scripts/68_rebuttal.Rmd
update

@@
 output:
+  word_document:
+    toc: yes
+    toc_depth: '3'
   html_document:
-    toc: true
-    toc_float: true
+    toc: yes
+    toc_float: yes
     toc_depth: 3
@@
 
-bc %>%
+asv_pcoa <- bc %>%
   as.data.frame() %>%
@@
   ggscatter(x = 'V1', y = 'V2', color =  'grp') +
-  labs(title = 'PCOA of healthy and CART patients') +
+  labs(title = 'PCOA of healthy and CART patients at ASV level') +
   xlab(paste0("PC 1 [",percent_var[1],"%]")) +
@@
   #theme_void() +
-  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients.pdf')
+  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients_asv_level.pdf')
 ```
@@
 # beta diversity at genus level but a PCOA between healthy and the CART
-genus <- read_csv('../data/amplicon/stool/counts_genus_combined.csv')
+library(vdbR)
+connect_database('~/dbConfig.txt')
+get_table_from_database("healthy_volunteers_ag")
+get_table_from_database("asv_alpha_diversity_ag")
+stb <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
 
-keep <- genus %>% 
+healthy <- healthy_volunteers_ag %>% 
+  inner_join(asv_alpha_diversity_ag, by = c("sampleid", "oligos_id")) 
+cts <- get_counts_subset(c(stb$sampleid, healthy %>% pull(sampleid)))
+
+# a total of 75 samples  
+nsamp <- cts %>% 
+  distinct(sampleid) %>% 
+  nrow
+
+all_pheno <- bind_rows(healthy %>% 
+  select(sampleid) %>% 
+  mutate(grp = 'healthy', center = 'healthy'),
+  stb %>% select(sampleid, center)  %>% 
+    mutate(grp = 'CART') %>% 
+    select(sampleid, grp, center)
+  ) %>% 
+  ungroup %>% 
+  inner_join(asv_alpha_diversity_ag %>% 
+               select(path_pool, sampleid))
+
+joined <- cts %>% 
+  left_join(asv_annotation_blast_ag %>% select(asv_key, genus)) %>% 
+  group_by(sampleid, genus) %>% 
+  summarise(relab = sum(count_relative)) %>% 
+  ungroup()
+
+keepg <- joined %>% 
   filter(relab > 0.0001) %>% 
   count(genus) %>% 
-  filter(n > floor(nsamp * 0.2)) %>% 
+  filter(n > floor(nsamp * 0.25)) %>% 
   filter(!is.na(genus)) %>% 
   pull(genus)
+
+cts_fil <- joined %>% 
+  filter(genus %in% keepg) %>% 
+  spread('genus', 'relab', fill = 0) %>% 
+  column_to_rownames('sampleid')
+
+library(vegan)
+dist_ <- vegdist(cts_fil, method = 'bray')
+eigen <- pcoa(dist_)$values$Eigenvalues
+percent_var <- signif(eigen/sum(eigen), 3)*100
+bc <- cmdscale(dist_, k = 2)
+bc %>%
+  as.data.frame() %>%
+  rownames_to_column('sampleid') %>% 
+  ungroup() %>% 
+  inner_join(all_pheno) %>% 
+  distinct(sampleid, .keep_all = T) %>% 
+  ggscatter(x = 'V1', y = 'V2', color =  'grp') +
+  labs(title = 'PCOA of healthy and CART patients at genus level') +
+  xlab(paste0("PC 1 [",percent_var[1],"%]")) +
+  ylab(paste0("PC 2 [",percent_var[2],"%]")) +
+  #theme_void() +
+  ggsave('../figs/PCOA(bray-curtis) of healthy and CART patients_genus_level.pdf')
+
+asv_pcoa
+
 ```
+- Thank reviewer for bringing up this point.
+As shown by the above two PCOA, one at genus level, one at ASV level, with same filtering threshold, the healthy and the CART cohort look about the same level of separation.
+The result is robust to this choice.
+
+## Major point 3
+
+**l.
+580-581 log relative abundances have been use to evaluate associations; this ignores the compositionality effect, which is commonly treated by using Aitchison transformations (e.g.
+CLR) in microbiome studies.
+Ignoring compositionality has potentially notable impact on the analysis outcomes.
+It would be essential to check that the overall conclusions are not affected by this.
+Using clr transformed abundances would be more in line with the current practices than using log relative abundance.**
+
+```{r}
+# compare between the transformed counts of CLR and log10
+meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv')
+
+library(vdbR)
+connect_database('~/dbConfig.txt')
+cts <- get_counts_subset(meta$sampleid)
+cts_ <- cts %>% 
+  select(asv_key, sampleid, count)
+
+annot <- asv_annotation_blast_ag %>% 
+  filter(asv_key %in% cts_$asv_key)
+
+# CLR transformation works at the counts level
+cts_all <- cts_ %>% 
+  left_join(annot %>%  select(asv_key, genus), by  = 'asv_key') %>% 
+  group_by(sampleid, genus) %>% 
+  summarise(cnt = sum(count)) %>% 
+  spread('sampleid', 'cnt', fill = 0) %>% 
+  filter(!is.na(genus)) %>% 
+  column_to_rownames('genus')
+
+clr_res <- compositions::clr(cts_all + 0.5) %>% 
+  as.data.frame()  %>% 
+  rownames_to_column('genus') %>% 
+  gather('sampleid','clr', names(.)[2]:names(.)[ncol(.)])
+
+clr_5 <- clr_res %>% 
+  filter(genus %in% c('Faecalibacterium','Ruminococcus','Akkermansia','Bacteroides','Enterococcus'))
+
+# my log10 tranformation results
+log10 <- read_csv('../data/amplicon/stool/combined_5_genera.csv') %>% 
+  gather('genus', 'log10', Akkermansia:Ruminococcus)
+
+# compare the two
+two <- log10 %>% 
+  inner_join(clr_5)
+
+two %>% 
+  arrange(log10) %>% 
+  #filter(log10 > -5) %>% 
+  ggscatter(x = 'log10', y = 'clr', facet.by = 'genus', alpha = 0.4,
+            xlab = 'log10 transformation counts', ylab = 'CLR transformation counts',
+            title = 'Scatter plot and Pearson correlation' ,
+             add = "reg.line",  # Add regressin line
+           add.params = list(color = "blue", fill = "lightgray"), # Customize line
+           conf.int = TRUE, # Add confidence interval
+           cor.coef = TRUE, # Add correlation coefficient.
+           cor.coeff.args = list(method = "pearson",  label.sep = "\n") )
+```
+
+- Thank reviewer for pointing this out.
+We visualized the Pearson correlation between the CLR and the log10 transformed counts for the 5 genera that we put into the model.
+As depicted in the above plot, the correlation is about 1 for all of them.
+Therefore the log10 transformation is fine with out purpose.
+
+## Minor point 1
+
+**The Bayesian analysis is interesting and it is good to see that this methodology is starting to be adopted also in this research area but the analysis does not mention covariates.
+It the assumption that the groups are sufficiently homogeneous so that no covariates need to be considered in the analysis?
+**

### 2021-07-28 scripts/68_rebuttal.Rmd
udpate

@@
 
-```{r}
+```{r fig.height = 13, fig.width = 8, fig.align = "center"}
 all_title_fs <- 20
@@
 
-g <- plot_grid(CR,tox,  
+g <- cowplot::plot_grid(CR,tox,  
           nrow = 2, 
           align = 'hv',
+          rel_heights = c(1.8,1.2),
           axis = 'b')  +
@@
 ```
+
 Controls were not used in the study.
@@
 
-summary(sample_counts$count_total)
+#summary(sample_counts$count_total)
 ```
 
-Thank reviewer for reminding us to add more details of the sequencing reads. For 16s data, the median read size is 50788 and the range is from 7041 to 93953. For the shotgun data,
+```{r}
+# read depth of the shotgun reads 
+kneaddata_logs <- list.files('../data/log/', '_knead.log$', full.names = T)
+
+knead <- kneaddata_logs %>% 
+  set_names(kneaddata_logs) %>% 
+  map(~ suppressWarnings(read_table(., col_names = F, col_types = 'c')) %>% 
+        rename(content = names(.)[1]) %>% 
+        filter(str_detect(content, 'Total reads after merging results from multiple databases')) %>% 
+        transmute(knead_num = str_extract(content, 'Total reads after merging results from multiple databases.+$')) %>% 
+        mutate(knead_num = as.numeric(str_replace(knead_num,'^.+: ',''))) %>% 
+        head(1)) %>% 
+  bind_rows(.id = 'sampleid') %>% 
+  mutate(sampleid = str_replace(sampleid, '../data/log//',''),
+         sampleid = str_replace(sampleid, '_knead.log',''))
+
+#summary(knead$knead_num)
+```
+
+Thank reviewer for reminding us to add more details of the sequencing reads.
+For 16s data, the median read size is 50788 and the range is from 7041 to 93953.
+For the shotgun data, after preprocessing and decontamination, the median read depth is 19476595.
+The range is from 6106173 to 48289024.
 
@@
 **Figure 2: Are Ruminococcus/Bacteroides “low” and “high” relative abundance determined according to the median abundance? Or some other way?**
+
+```{r}
+genera <- read_csv('../data/amplicon/stool/combined_5_genera.csv')
+
+# do not standardize the log transformed relative abundance 
+meta <- read_csv('../data/amplicon/stool/combined_2_meta_expanded.csv')  %>% 
+  inner_join(genera)
+meta_ <- meta %>% 
+  select(Akkermansia:Ruminococcus)
+
+ba_top <- meta_ %>% 
+  filter(Bacteroides >= quantile(meta$Bacteroides, 0.9))
+
+ba_bot <- meta_ %>% 
+  filter(Bacteroides <= quantile(meta$Bacteroides, 0.1))
+```
+
+We thank reviewer for asking the clarification.
+The "high" represents the samples that have the top 10% Ruminococcus/Bacteroides relative abundance among all the samples, while the "low" for the bottom 10%.
 
@@
            cor.coef = TRUE, # Add correlation coefficient.
-           cor.coeff.args = list(method = "pearson",  label.sep = "\n") )
+           cor.coeff.args = list(method = "pearson",  label.sep = "\n") ) +
+  ggsave('../figs/amplicon/scatter_clr_log10.pdf')
 ```
@@
 **The Bayesian analysis is interesting and it is good to see that this methodology is starting to be adopted also in this research area but the analysis does not mention covariates.
-It the assumption that the groups are sufficiently homogeneous so that no covariates need to be considered in the analysis?
-**
+It the assumption that the groups are sufficiently homogeneous so that no covariates need to be considered in the analysis?**
+
+We thank reviewer for raising this question.
+The covariats included in the model will be more clear if the code is available, which we will provide alongside this rebuttal.
+Simply speaking, the models are actually:
+
+- CR/Toxicity ~ alpha diversity + Center
+
+- CR/Toxicity ~ Akkermansia + Bacteroides + Enterococcus + Faecalibacterium + Ruminococcus + Center
+
+Center is the most important covariate that could confound this analysis, so we added it. We also considered the disease type each patient had, but it turned out to be of little importance.
+
+## Minor point 4
+
+**PCoA is a standard visualization in this context and it would be helpful to include this at least in the supplementary material so that we can see how the healthy and treated cases map on the familiar population landscape characterized by PCoA.**
+
+```{r}
+asv_pcoa
+```
+We thank reviewer for mentioning this. The above is a PCOA of Bray-Curtis distance done at ASV level.
+
+## Minor point 5
+
+**The methods could provide some justification for the choice of the specific methods, including Bray-Curtis, inverse Simpson, and LefSe since there is a vast number of available alternatives, with potentially differing outcomes.
+It has been recently reported, for instance, that LefSe provides a larger number of positive hits (and hence potentially more false positives) than many other DA methods (https://www.biorxiv.org/content/10.1101/2021.05.10.443486v1).
+The choices are acceptable but the methods could benefit from brief motivation of the choices.**
+
+- According to [their paper](https://genomebiology.biomedcentral.com/articles/10.1186/gb-2011-12-6-r60), "LEfSe determines the features (organisms, clades, operational taxonomic units, genes, or functions) most likely to explain differences between classes by coupling standard tests for statistical significance with additional tests encoding biological consistency and effect relevance."
+
+# Reviewer 3
+
+## Major point 1
+
+**The “hypothesis-driven exploratory analysis” using five selected bacterial genera to identify the association of taxa with response seems biased to this reviewer.
+These five genera were previously identified in other studies as being associated with ICB and HCT, but only Bacteroides was found to be differentially abundant in the untargeted analysis.
+The strength of this data as possible biomarkers of response or toxicity is weak, and seems unfairly biased towards giving an association with previously-identified taxa in other disease states.
+The taxa should be confirmed using quantitative PCR and the data re-analyzed using machine learning.**
+
+After adjusting the filtering threshold, Ruminococcus, Bacteroides and Faecalibacterium of the 5 genera we chose appeared in the significant results in the untargeted analysis.
+And we picked a total of 5 genera since we aim to limit the number of covariants in the model due to the small sample size in this cohort.
+While machine learning is a good approach, the Bayesian method used in the current analysis is also valid, with additional benefits such as retaining the uncertainty of the coefficient estimates, and enabling posterior predictive check to quantify the impact of the coefficients.
+
+## Major point 2
+
+**Is there any association between pathway enrichment and therapeutic toxicity when examining the shotgun metagenomics dataset?
+This is a major objective of the investigation, and it seems odd not to address this, and only to look at response.
+Even if there is no association, this should be stated by the authors.**
+
+```{r fig.height = 10, fig.width = 8, fig.align = "center"}
+# look at the pathway results
+library(vdbR)
+connect_database('~/dbConfig.txt')
+get_table_from_database('metacyc_pathway_name')
+get_table_from_database('metacyc_pathway_ontology')
+
+fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv.res$', full.names = T)
+
+feature <- fns %>% 
+  set_names(fns) %>% 
+  map(~ read_tsv(., col_names = c('pathway','xx','direction','score','pval')) %>% 
+        filter(!is.na(score))) %>% 
+  bind_rows(.id = 'group') %>% 
+  mutate(group = str_replace(group, '../data/shotgun_output/humann3//','')) %>% 
+  mutate(group = str_replace(group, '_lefse_ready_pcts.tsv.res$',''))
+
+# change the "N" direction to be minus score
+feature <- bind_rows(
+  feature %>% 
+    split(.$direction) %>% 
+    pluck('no') %>% 
+    mutate(score = -score),
+  feature %>% 
+    split(.$direction) %>% 
+    pluck('yes')  
+) %>% 
+  arrange(group, pathway, score) %>% 
+  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY.*_\\d{3,4}')) %>% 
+  mutate(pwid = str_replace_all(pwid, '_', '-')) %>% 
+  mutate(pwid = if_else(str_detect(pathway, '^TCA'), 'TCA', pwid)) %>% 
+  mutate(pwid = if_else(str_detect(pathway, '^NAD'), 'NAD-BIOSYNTHESIS-II', pwid)) %>% 
+  inner_join(metacyc_pathway_name %>% select(pwid, pw_name)) %>% 
+  inner_join(metacyc_pathway_ontology %>% select(pwid, l4:l9))
+
+all_title_fs <- 20
+axis_text_fs <- 16
+CR <- feature %>% 
+  filter(group == 'pull_cr_d100') %>% 
+  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
+            geom_bar( stat = 'identity') +
+            coord_flip() +
+            scale_color_manual(values = c('#925E9F', '#42B540'))  +
+            scale_fill_manual(values = c('#925E9F', '#42B540')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  plot.title = element_text(size=all_title_fs),
+                  axis.title.x = element_text(size=axis_text_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position='bottom') +
+            labs(title = str_glue('Response to CR') ,
+                 y = 'Score')
+
+tox <- feature %>% 
+  filter(group == 'pull_toxicity') %>% 
+  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
+            geom_bar(stat = 'identity') +
+            coord_flip() +
+            scale_color_manual(values = c('#0099B4', '#AD002A'))  +
+            scale_fill_manual(values = c('#0099B4', '#AD002A')) +
+            theme_classic() +
+            theme(axis.title.y  = element_blank(), 
+                  axis.title.x = element_text(size=axis_text_fs),
+                  plot.title = element_text(size=all_title_fs),
+                  axis.text.x = element_text(size=axis_text_fs),
+                  legend.position="bottom") +
+            labs(title = str_glue('Response to Toxicity') ,
+                 y = 'Score')
+
+cowplot::plot_grid(CR,tox,  
+          nrow = 2, 
+          rel_heights = c(1,3),
+          align = 'hv',
+          axis = 'b')
+```
+
+## Major point 4
+
+**Altered microbial abundance itself may be in fact a source of difference between CAR T patients and healthy controls, or between toxicity/response outcomes, and this possibility was not examined by the authors.
+Relative abundance assessment can mask the real community dynamics of the microbiota and as such a method of complete quantitative profiling should be performed (PMID: 29143816).**
+
+(Need help in answering this.)

### 2021-07-28 scripts/43_comparisons_3.Rmd
update
udpate

@@
 # the pathway counts table
-full <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_pathabundance_CART.tsv')  %>% 
-  rename_all(funs(str_replace(., '^CART_','')))  %>% 
+full <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_pathabundance_cpm_joined_unstratified.tsv')  %>% 
   rename_all(funs(str_replace(., '_humann3$','')))
@@
 ```{r}
-# add a filtering step here 
+# add a filtering step here after the normalization in the snakemake 
 # 50 and 25% 
@@
   write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified_fil.tsv')
+
+# and then bind with snakemake
 ```
@@
   rename_all(funs(str_replace(., '_profile$','')))
+
+keeptaxa <- taxa %>% 
+  filter(cpm > 50) %>%
+  ungroup() %>% 
+  count(pw) %>% 
+  filter(n > floor(nrow(simple) * 0.25)) %>% 
+  pull(pw)
+
+cts_fil <- pcts %>% 
+  filter(pw %in% keeppw) %>% 
+  spread('sampleid', 'cpm', fill = 0)
 

### 2021-07-28 scripts/56_rethinking_genera.Rmd
udpate

@@
   labs(title = 'Toxicity') +
-  ggsave('../figs/amplicon/bayesian_genera_tox.pdf', width = 5, height = 4)
+  ggsave('../figs/amplicon/bayesian_genera_tox_log10.pdf', width = 5, height = 4)
 ```
@@
   labs(title = 'CR') +
-  ggsave('../figs/amplicon/bayesian_genera_cr.pdf', width = 5, height = 4)
+  ggsave('../figs/amplicon/bayesian_genera_cr_log10.pdf', width = 5, height = 4)
 ```

### 2021-07-28 scripts/67_lefse_shotgun_taxa.Rmd
udpate

@@
 ```{r}
-taxa <- read_tsv('../data/shotgun_output/metaphlan/merged_abundance_table.txt')  %>% 
-  rename_all(funs(str_replace(., '_profile$',''))) %>% 
+taxa <- read_delim('../data/shotgun_output/metaphlan/merged_abundance_table_metaphlan3.txt', delim = '\t', skip = 1)  %>% 
+  rename_all(funs(str_replace(., '_metaphlan3_profile$',''))) %>% 
   filter(str_detect(clade_name, 's__.+$')) %>% 
@@
 
+cts <- read_delim('../data/shotgun_output/metaphlan/merged_abundance_table_metaphlan3.txt', delim = '\t', skip = 1)  %>% 
+  rename_all(funs(str_replace(., '_metaphlan3_profile$',''))) %>% 
+  #filter(str_detect(clade_name, 's__.+$')) %>% 
+  select(- NCBI_tax_id)
+
 keeptaxa <- taxa %>% 
-  filter(relab > 0.05) %>%
+  filter(relab > 0.0001) %>%
   ungroup() %>% 
@@
 cts_fil <- taxa %>% 
-  filter(clade_name %in% keeptaxa) %>% 
+  #filter(clade_name %in% keeptaxa) %>% 
   spread('sampleid', 'relab', fill = 0)
@@
   map(~ pull(.data = ., fid) ) %>% 
-  imap(~ cts_fil %>% select(clade_name, matches(.x)) %>% 
+  imap(~ cts %>% select(clade_name, matches(.x)) %>% 
          write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_tcts.tsv')))
@@
 
-connect_database('~/dbConfig.txt')
-get_table_from_database('asv_alpha_diversity_ag')
-list_table_from_database('asv')
 ```

### 2021-07-28 scripts/69_lefse_shotgun_pathway.Rmd
udpate

---
title: "lefse shotgun pathway"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
simple <- read_csv('../data/shotgun_pheno/final_comprehensive_UPDATED_simple.csv')
pheno <- simple %>% 
  gather('pheno', 'value', cr_d100:crs) 
```

```{r}
# the pathway counts table
full <- read_tsv('../data/shotgun_output/humann3_tsv/humann3_pathabundance_CART.tsv')  %>% 
  rename_all(funs(str_replace(., '^CART_','')))  %>% 
  rename_all(funs(str_replace(., '_humann3$','')))

all_sub_pheno <- pheno %>% 
  split(., list(.$pheno)) %>% 
  purrr::imap(~ filter(.data = ., value != 'not_assessed'))

all_sub_pheno %>% 
  imap(function(.x, .y){
    select(.data = .x, value) %>% 
      t() %>% 
      write.table(str_glue('../data/shotgun_output/humann3/pull_{.y}.txt'), sep = '\t', quote = F, row.names = T, col.names = F)
  })

all_pcts <- all_sub_pheno %>% 
  map(~ pull(.data = ., fid) ) %>% 
  imap(~ full %>% select(`# Pathway`, matches(.x)) %>% 
         write_tsv(str_glue('../data/shotgun_output/humann3/pull_{.y}_pcts.tsv')))
```

Snakemake do the normalization and the split

```{r}
# add a filtering step here 
# 50 and 25% 
pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified.tsv') %>% 
  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
  rename(pw = `# Pathway`)

keeppw <- pcts %>% 
  filter(cpm > 50) %>%
  ungroup() %>% 
  count(pw) %>% 
  filter(n > floor(nrow(simple) * 0.25)) %>% 
  pull(pw)

cts_fil <- pcts %>% 
  filter(pw %in% keeppw) %>% 
  spread('sampleid', 'cpm', fill = 0)

cts_fil %>% 
  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_cr_d100_pcts_cpm_unstratified_fil.tsv')
```

```{r}
# tox
pcts <- read_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified.tsv') %>% 
  gather('sampleid', 'cpm', names(.)[2]:names(.)[ncol(.)]) %>% 
  rename(pw = `# Pathway`)

keeppw <- pcts %>% 
  filter(cpm > 50) %>%
  ungroup() %>% 
  count(pw) %>% 
  filter(n > floor(nrow(simple) * 0.25)) %>% 
  pull(pw)

cts_fil <- pcts %>% 
  filter(pw %in% keeppw) %>% 
  spread('sampleid', 'cpm', fill = 0)

cts_fil %>% 
  write_tsv('../data/shotgun_output/humann3/humann_final_out/pull_toxicity_pcts_cpm_unstratified_fil.tsv')
```

Snakemake do the binding

```{r}
# PATHWAY
# I already normalized the pcts so don't need to normalize again here 
fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv$')

cmds <- tibble(
  fns = fns
) %>% 
  mutate(format_cmd = str_glue('format_input.py {fns}  {fns}.in -c 1 -u 2')) %>% 
  mutate(run_cmd = str_glue('run_lefse.py {fns}.in  {fns}.res')) %>% 
  mutate(plot_cmd = str_glue('plot_res.py {fns}.res {fns}.pdf --format pdf  --feature_font_size 4 --width 10 --dpi 300 --title {fns}')) %>% 
  select(-fns) %>% 
  gather() %>% 
  select(value) %>% 
  write_csv('../data/shotgun_output/humann3/lefse_run_cmd.sh', col_names = F)
```

```{r}
# look at the pathway results
library(vdbR)
connect_database('~/dbConfig.txt')
get_table_from_database('metacyc_pathway_name')
get_table_from_database('metacyc_pathway_ontology')

fns <- list.files('../data/shotgun_output/humann3/', pattern = 'lefse_ready_pcts.tsv.res$', full.names = T)

feature <- fns %>% 
  set_names(fns) %>% 
  map(~ read_tsv(., col_names = c('pathway','xx','direction','score','pval')) %>% 
        filter(!is.na(score))) %>% 
  bind_rows(.id = 'group') %>% 
  mutate(group = str_replace(group, '../data/shotgun_output/humann3//','')) %>% 
  mutate(group = str_replace(group, '_lefse_ready_pcts.tsv.res$',''))

# change the "N" direction to be minus score
feature <- bind_rows(
  feature %>% 
    split(.$direction) %>% 
    pluck('no') %>% 
    mutate(score = -score),
  feature %>% 
    split(.$direction) %>% 
    pluck('yes')  
) %>% 
  arrange(group, pathway, score) %>% 
  mutate(pwid = str_extract(pathway, '^.+_PWY|^PWY.*_\\d{3,4}')) %>% 
  mutate(pwid = str_replace_all(pwid, '_', '-')) %>% 
  mutate(pwid = if_else(str_detect(pathway, '^TCA'), 'TCA', pwid)) %>% 
  mutate(pwid = if_else(str_detect(pathway, '^NAD'), 'NAD-BIOSYNTHESIS-II', pwid)) %>% 
  inner_join(metacyc_pathway_name %>% select(pwid, pw_name)) %>% 
  inner_join(metacyc_pathway_ontology %>% select(pwid, l4:l9))

all_title_fs <- 20
axis_text_fs <- 16
CR <- feature %>% 
  filter(group == 'pull_cr_d100') %>% 
  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
            geom_bar( stat = 'identity') +
            coord_flip() +
            scale_color_manual(values = c('#925E9F', '#42B540'))  +
            scale_fill_manual(values = c('#925E9F', '#42B540')) +
            theme_classic() +
            theme(axis.title.y  = element_blank(), 
                  plot.title = element_text(size=all_title_fs),
                  axis.title.x = element_text(size=axis_text_fs),
                  axis.text.x = element_text(size=axis_text_fs),
                  legend.position='bottom') +
            labs(title = str_glue('Response to CR') ,
                 y = 'Score')

tox <- feature %>% 
  filter(group == 'pull_toxicity') %>% 
  ggplot(aes(x = reorder(pathway, score), y = score, fill = direction)) +
            geom_bar(stat = 'identity') +
            coord_flip() +
            scale_color_manual(values = c('#0099B4', '#AD002A'))  +
            scale_fill_manual(values = c('#0099B4', '#AD002A')) +
            theme_classic() +
            theme(axis.title.y  = element_blank(), 
                  axis.title.x = element_text(size=axis_text_fs),
                  plot.title = element_text(size=all_title_fs),
                  axis.text.x = element_text(size=axis_text_fs),
                  legend.position="bottom") +
            labs(title = str_glue('Response to Toxicity') ,
                 y = 'Score')

cowplot::plot_grid(CR,tox,  
          nrow = 2, 
          rel_heights = c(1,3),
          align = 'hv',
          axis = 'b')
```

### 2021-09-27 scripts/54_rethinking_binary.Rmd
finally donne with this yeaaaaaaaaaaaah

--- 2021-07-13 scripts/54_rethinking_binary.Rmd
@@
 library(ggpubr)
-
+options(mc.cores = parallel::detectCores())
+rstan_options(auto_write = TRUE)
+library(tidybayes)
 ```
@@
 meta <- read_csv('../data/amplicon/stool/combined_2_meta.csv') %>% 
-  mutate(logdiv_s = scale(log(simpson_reciprocal))) %>%  
+  mutate(logdiv_s = as.numeric(scale(log(simpson_reciprocal)))) %>%  
   mutate(tox = if_else(toxicity == 'yes', 1, 0),
@@
          toxicity = factor(toxicity, levels = c('no','yes')),
-         cr_d100 = factor(cr_d100, levels = c('no','yes'))) %>%
-  # have a 2 cutpoint for the CRS
-  mutate(crs3 = case_when(
-    crsgrade == 0 ~ 1,
-    crsgrade == 1 ~ 2,
-    crsgrade >= 2 ~ 3
-  ))
+         cr_d100 = factor(cr_d100, levels = c('no','yes')))
 
@@
         a[location] ~ dnorm( 0 , 0.5 )
-    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 16)
 
@@
 plot(precis(mtox, depth = 2, prob = 0.95 ) , pars = ('b'), labels = c('simpson_reciprocal'))
-# there is baseline difference between penn and Msk and MSK has higher tox rage in general.
-# it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
-div_value <- c(-1, 1)
 
+# make this plot myself with a pre-defined range in the axis
+df <- precis(mtox,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = 'simpson_reciprocal')
+
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = reorder(variable, mean), y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-0.7, 1)) +
+  #geom_hline(yintercept = 1, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'Toxicity')
+
+ggsave('../figs/amplicon/bayesian_div_tox.pdf', width = 5, height = 4)
 ```
-
-```{r}
-# another model with the crs grade 0,1,>=2 
-mcrs3 <- ulam(
-    alist(
-        crs3 ~ dordlogit( phi , cutpoints ),
-        phi <- b*logdiv_s + a[location] ,
-        b ~ dnorm( 0 , 1),  
-        a[location] ~ dnorm( 0 , 0.5 ),
-        cutpoints ~ dnorm( 0 , 1.5 )
-    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
-plot(precis(mcrs3, depth = 2, prob = 0.95) ) 
-# the cutpoints are not of much interet, this suggests that the diversity really isn't having a correlation?
-```
+There is baseline difference between penn and Msk and MSK has higher tox rage in general.
+it doesn't seem diversity has a clear association with tox rate after correcting for center wise difference
 
@@
 
-    diff = abs( cols[[1]] -cols[[2]] )
+    diff =  cols[[1]] -cols[[2]] 
     ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
-    return(ret)
+    return(diff)
   }) 
+boxplot(prior_tox$p)
+hist(c(prior_tox$m$lowdiv, prior_tox$p$lowdiv)) 
+hist(c(prior_tox$m$highdiv, prior_tox$p$highdiv))
 ```
@@
 ```{r}
+# make this plot myself with a pre-defined range in the axis
+df <- precis(mcr,  prob = 0.95)  %>% as.tibble()
+df <- df %>%
+  mutate(variable = 'simpson_reciprocal')
+
+#remove row number 1 (The intercept)
+df %>% 
+  ggplot(aes(x = reorder(variable, mean), y = mean)) +
+  geom_point(shape = 15,
+             size  = 4, width = 0.1,
+             position = "dodge", color="black") +
+  geom_errorbar(aes(ymin  = `2.5%`,
+                    ymax  = `97.5%`),
+                width = 0.1,
+                size  = 0.7,
+                position = "dodge", color="turquoise4") +
+  theme(axis.title = element_text(face = "bold")) +
+  xlab("Variables") + ylab("Coefficient with 95% CI") +
+  coord_flip(ylim = c(-0.7, 1)) +
+  #geom_hline(yintercept = 1, color = "red", size = 1) +
+   theme(axis.title = element_text(size = 0)) +
+  theme(axis.text = element_text(size = 14)) +
+  labs(title = 'CR') 
+ggsave('../figs/amplicon/bayesian_div_cr.pdf', width = 5, height = 4)
+```
+
+```{r}
 # doing a prior check
-prior <- extract.prior( mcr , n=1e4 )
+prior <- extract.prior( mcr , n=1e4, cores =16 )
 
@@
 
-    diff = abs( cols[[1]] -cols[[2]] )
+    diff = cols[[1]] -cols[[2]] 
     ret = quantile(diff , probs = c(0.025, 0.5, 0.975))
-    return(ret)
+    return(diff)
   })
+
+boxplot(prior_cr$p)
+hist(c(prior_cr$m$lowdiv, prior_cr$p$lowdiv)) 
+hist(c(prior_cr$m$highdiv, prior_cr$p$highdiv)) 
 ```
@@
       }) 
-    return(cols)
+    diff =  cols[[1]] -cols[[2]] 
+    return(diff)
   })
 
+hist(c(post_cr$m, post_cr$p)) 
+hist(c(post_cr$m$highdiv, post_cr$p$highdiv))
+```
+
+```{r}
+# if putting them together (the prior and the post)
+both <- bind_rows(
+  data_frame(
+      coeff = c(prior_cr$m, prior_cr$p),
+      grp = 'prior'
+    ),
+  data_frame(
+      coeff = c(post_cr$m, post_cr$p),
+      grp = 'post'
+    )
+)
+
+both %>% 
+    ggplot(aes(x = coeff, y = grp, color = grp)) +
+    stat_pointinterval(.width = c(.66, .95))
+```
+
+```{r}
 # plot the histogram of the probs corresponding to low and high diversity

### 2021-09-27 scripts/56_rethinking_genera.Rmd
finally donne with this yeaaaaaaaaaaaah

--- 2021-07-13 scripts/56_rethinking_genera.Rmd
@@
   theme(axis.text = element_text(size = 14)) +
-  labs(title = 'Toxicity') +
-  ggsave('../figs/amplicon/bayesian_genera_tox.pdf', width = 5, height = 4)
+  labs(title = 'Toxicity') 
+ggsave('../figs/amplicon/bayesian_genera_tox_log10.pdf', width = 5, height = 4)
 ```
 
+### Prior check
+
+I will just do something similar to the nutrition project : the prior annd post coeff forest plot
+
 ```{r}
-post <- extract.samples(gtox) 
+theme_set(theme_tidybayes() + cowplot::panel_border())
+# logit(p) <- ba*Akkermansia + bb*Bacteroides + be*Enterococcus + bf*Faecalibacterium + br*Ruminococcus + a[location]
+prior <- extract.prior( gtox , n=4000 ) %>% as.tibble()  %>% select(-a) %>% mutate(grp = 'prior')
+post <- extract.samples(gtox) %>% as.tibble()  %>% select(-a) %>% mutate(grp = 'post')
+# comparing prior and post 
+tox <- bind_rows(prior, post) %>% 
+  gather('term', 'coeff', ba:br) %>% 
+  mutate(genus = case_when(
+    term == 'ba' ~ 'Akkermansia',
+    term == 'bb' ~ 'Bacteroides',
+    term == 'be' ~ 'Enterococcus',
+    term == 'bf' ~ 'Faecalibacterium',
+    term == 'br' ~ 'Ruminococcus'
+  ))
+
+post_order <- tox %>% 
+  filter(grp == 'post') %>% 
+  group_by(term, genus) %>% 
+  summarise(q50 = median(coeff)) %>% 
+  arrange(-q50) %>% 
+  pull(genus)
+
+tox %>%
+  mutate(genus = factor(genus, levels = post_order)) %>% 
+  ggplot(aes(x = coeff, y = grp, color = grp)) +
+  stat_pointinterval(.width = c(.66, .95)) +
+  scale_color_manual(values = c('#EC0000','#00468B')) +
+  geom_vline(xintercept = 0, col = 'gray', linetype = 'dashed') +
+  facet_grid(genus ~ .) +
+  labs(x = 'Regression coefficients',
+       y = '') +
+  theme(legend.position = 'none')
+```
+
+### posterior check
+
+```{r}
+
 meta_ <- meta %>% 
@@
 # plot them together in one 
-bind_rows(
+bacteroides_top_bottom_10 <- bind_rows(
   res_top %>% mutate(grp = 'top'),
@@
   gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
-                  xlab = 'Predicted probability of Toxicity', ylab = 'Probability density') +
+                  xlab = 'Predicted probability of Toxicity', ylab = 'Probability density')
+
 ggsave('../figs/predicted_tox_bacteroides_top_bottom_10.pdf', width = 7, height = 5)
@@
 
-```
-
-### posterior check
-
-```{r}
-post <- extract.samples(gtox) 
-# using the patient with highest and lowest Bacteroides data to plug in
-# find the patients
-bl <- meta %>% 
-  arrange(Bacteroides) %>% 
-  slice(1) %>% 
-  select(Akkermansia:Ruminococcus)
-
-bh <- meta %>% 
-  arrange(-Bacteroides) %>% 
-  slice(1) %>% 
-  select(Akkermansia:Ruminococcus)
-
-pl <- inv_logit( post$ba * (-1.612806	) + post$bb * (-5.69897		) + post$be * (-5.69897	) +  post$bf * (-5.69897		) + post$br * (-0.5244935) + post$a[, 1] )
-
-ph <- inv_logit( post$ba * (-2.286752	) + post$bb * (-0.4575058) + post$be * (-1.745601	) +  post$bf * (-5.69897) + post$br * (-4.050354) + post$a[, 1] )
-
-bind_rows(
-  tibble(
-    value = pl,
-    grp = 'low'
-  ),
-  tibble( 
-    value = ph,
-    grp = 'high'
-  )
-) %>% 
-  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',add = 'mean',title = 'MSK',
-              xlab = 'Predicted probability of toxicity', ylab = 'Probability density')
-
-bind_rows(
-  tibble(value = pl,grp = 'low'),
-  tibble( value = ph,grp = 'high')
-)   %>% 
-  group_by(grp) %>% 
-  summarise(m = mean(value))
 ```
@@
         a[location] ~ dnorm( 0 , 0.5 )
-    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 8)
+    ) , data=dat_list , chains=4 , log_lik=TRUE , cores = 16)
 precis(gcr, prob = 0.95) 
@@
   theme(axis.text = element_text(size = 14)) +
-  labs(title = 'CR') +
-  ggsave('../figs/amplicon/bayesian_genera_cr.pdf', width = 5, height = 4)
+  labs(title = 'CR') 
+ggsave('../figs/amplicon/bayesian_genera_cr_log10.pdf', width = 5, height = 4)
 ```
@@
 ### posterior check
-
-```{r}
-post <- extract.samples(gcr) 
-# using the patient with highest and lowest Ruminococcus data to plug in
-# find the patients
-rl <- meta %>% 
-  arrange(Ruminococcus) %>% 
-  slice(1) %>% 
-  select(Akkermansia:Ruminococcus)
-
-rh <- meta %>% 
-  arrange(-Ruminococcus) %>% 
-  slice(1) %>% 
-  select(Akkermansia:Ruminococcus)
-
-pl <- inv_logit( post$ba * (-3.517128) + post$bb * (-2.447201	) + post$be * (-2.504205	) +  post$bf * (-5.69897	) + post$br * (-5.69897) + post$a[, 1] )
-
-ph <- inv_logit( post$ba * (-1.845325	) + post$bb * (-3.052386) + post$be * (-5.69897) +  post$bf * (-5.69897) + post$br * (-0.373353) + post$a[, 1] )
-
-bind_rows(
-  tibble(value = pl,grp = 'low'),
-  tibble( value = ph,grp = 'high')
-) %>% 
-  gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white',title = 'MSK', add = 'mean',
-              xlab = 'Predicted probability of CR d100', ylab = 'Probability density')
-
-bind_rows(
-  tibble(value = pl,grp = 'low'),
-  tibble( value = ph,grp = 'high')
-)   %>% 
-  group_by(grp) %>% 
-  summarise(m = mean(value))
-```
 
@@
   gghistogram(x = 'value',bins = 30, fill = 'grp', palette = 'nejm', color = 'white', add = 'mean',
-                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density')+
+                  xlab = 'Predicted probability of CR d100', ylab = 'Probability density')
   ggsave('../figs/predicted_CR_Ruminococcus_top_bottom_10.pdf', width = 7, height = 5)
